---
title: "Run VISION in human snRNAseq objects"
output: html_document
date: "2024-12-02"
---

# Introduction

This notebook prepares selected GMT signature files and runs VISION on
preprocessed Seurat objects.

The workflow is organized into three parts:

1. Define paths and helper functions.
2. Build GMT signatures needed downstream.
3. Run VISION on human and mouse Seurat objects.

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(Seurat)
library(VISION)
library(here)
library(msigdbr)
library(readr)
library(openxlsx)
library(tidyverse)

# -------------------------------------------------------------------------
# Define inputs and outputs
# -------------------------------------------------------------------------
today <- format(Sys.Date(), "%Y_%m_%d")

# External signature source
BATLAS_EXCEL <- here("data/BATLAS_signature.xlsx")

# Signature files
HEAT_SIGNATURE <- here("output/3.heat_signature/data/heat_signature.gmt")
MOUSE_HEAT_GMT <- here("output/3.heat_signature/data/mouse_heat.gmt")
UCP1_CORR_SIGNATURE <- here("output/3.heat_signature/data/non_HEAT_UCP1_corr_signature.gmt")

BATLAS_SIGNATURE <- here("data/8.gmt_signatures/batlas.gmt" ) 
HALLMARK_SIGNATURE <- here("data/8.gmt_signatures/hallmark_subset.gmt")
GOBP_SIGNATURE <- here("data/8.gmt_signatures/gobp_subset.gmt")

# Seurat inputs
HUMAN_SEURAT_DIR <- here("output/4.downstream_pipelines/vision/data/cleaned_datasets_human") 
CAMARAH_RDS <- file.path(HUMAN_SEURAT_DIR, "CamaraH_Dataset.rds") 
SHAMSI_RDS <- here("output/4.downstream_pipelines/vision/data/cleaned_datasets_mouse/ShamsiF_Dataset.rds") 
ZHANG_BAT_RDS_YOUNG <- "~/Projects/Joy_2025/data/Zhang_2025/GSE247719_PanSci_07_BAT_adata_young_wt.rds"

# Output directories
VIS_OBJ_OUTPUT_DIR <- here("output", "4.downstream_pipelines", "vision", "data", "vision_objects")
GMT_SIGNATURES_OUTPUT_DIR <- here("output", "4.downstream_pipelines", "vision", "data", "vision_gmt_signatures")

dir.create(VIS_OBJ_OUTPUT_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(GMT_SIGNATURES_OUTPUT_DIR, recursive = TRUE, showWarnings = FALSE)
```

# Helper functions

```{r}
# Write one GMT line from a gene vector
make_gmt_line <- function(signature_name, description, genes) {
  paste(signature_name, description, paste(genes, collapse = "\t"), sep = "\t")
}

# Build GMT lines from selected MSigDB collections
build_msigdb_gmt <- function(msig_table, selected_sets) {
  gmt_lines <- vector("character", length(selected_sets))

  for (i in seq_along(selected_sets)) {
    gs_name <- selected_sets[i]

    sig_data <- msig_table %>%
      filter(gs_name == !!gs_name)

    if (nrow(sig_data) == 0) {
      stop("No genes found for signature: ", gs_name)
    }

    description <- sig_data %>%
      mutate(description = paste(gs_name, gs_id, db_version, sep = "_")) %>%
      distinct(description) %>%
      pull(description)

    gmt_lines[i] <- make_gmt_line(
      signature_name = gs_name,
      description = description,
      genes = sig_data$gene_symbol
    )
  }

  gmt_lines
}

# Run VISION on one Seurat object and save the resulting VISION object
runAndSaveVISION <- function(seurat_file,
                             Signatures,
                             subset_adipocytes = FALSE,
                             vis_output_dir = VIS_OBJ_OUTPUT_DIR,
                             skip_existing = TRUE,
                             subset_large_seurat = TRUE,
                             n_pool = 100000,
                             seed = 1234) {

  dir.create(vis_output_dir, recursive = TRUE, showWarnings = FALSE)

  file_basename <- tools::file_path_sans_ext(basename(seurat_file))
  output_file <- file.path(vis_output_dir, paste(file_basename, "VISION.rds", sep = "_"))
  log_file <- file.path(vis_output_dir, paste(file_basename, "VISION.log", sep = "_"))

  message("Analyzing ", basename(seurat_file))

  if (skip_existing && file.exists(output_file)) {
    message("Skipping ", basename(seurat_file), " (output already exists)")
    return(NULL)
  }

  tryCatch(
    {
      s.object <- readRDS(seurat_file)

      if (subset_adipocytes) {
        s.object <- subset(s.object, idents = c("White Adipocytes", "Brown Adipocytes"))
      }

      if (subset_large_seurat && ncol(s.object) > n_pool) {
        set.seed(seed)
        s.object <- subset(s.object, cells = sample(colnames(s.object), size = n_pool))
      }

      sink(log_file, append = TRUE)
      message("Started at ", Sys.time())
      sink()

      vis <- Vision(
        s.object,
        signatures = Signatures,
        dimRed = "umap",
        dimRedComponents = 2
      )
      vis <- analyze(vis)

      saveRDS(vis, output_file)
      message("Saved: ", output_file)
      
      vision_run_summary <- tibble::tibble(
        dataset = file_basename,
        n_cells_input = ncol(readRDS(seurat_file)),
        n_cells_analyzed = ncol(s.object),
        subset_adipocytes = subset_adipocytes,
        subset_large_seurat = subset_large_seurat,
        n_pool = n_pool,
        signatures = paste(basename(Signatures), collapse = ";"),
        output_file = output_file
      )
      
      readr::write_csv(
        vision_run_summary,
        file.path(vis_output_dir, paste(file_basename, "VISION_run_summary.csv", sep = "_"))
      )
      
    },
    error = function(e) {
      err_msg <- paste("Error occurred with", basename(seurat_file), ":", conditionMessage(e))
      message(err_msg)

      sink(log_file, append = TRUE)
      message(err_msg)
      sink()
    }
  )
}
```

# Build signature files

## BATLAS signature

```{r}
BATLAS <- read.xlsx(BATLAS_EXCEL) %>%
  select(1, 2)

names(BATLAS) <- c("white_ad_genes", "brown_ad_genes")

batlas_description <- paste(
  "BATLAS",
  "97_thermo_associated_genes",
  "https://www.cell.com/cell-reports/fulltext/S2211-1247(18)31489-X",
  sep = "_"
)

batlas_line <- make_gmt_line(
  signature_name = "BATLAS",
  description = batlas_description,
  genes = BATLAS$brown_ad_genes
)

write_lines(batlas_line, BATLAS_SIGNATURE)
```

## Hallmark signatures from MSigDB

```{r}
msig_hallmark <- msigdbr(
  species = "Homo sapiens",
  collection = "H"
)

hallmark_subcollections <- c(
  "HALLMARK_ADIPOGENESIS",
  "HALLMARK_FATTY_ACID_METABOLISM",
  "HALLMARK_G2M_CHECKPOINT",
  "HALLMARK_GLYCOLYSIS",
  "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
  "HALLMARK_REACTIVE_OXYGEN_SPECIES_PATHWAY",
  "HALLMARK_UNFOLDED_PROTEIN_RESPONSE"
)

hallmark_gmt <- build_msigdb_gmt(
  msig_table = msig_hallmark,
  selected_sets = hallmark_subcollections
)

write_lines(hallmark_gmt, HALLMARK_SIGNATURE)
```

## GOBP signatures from MSigDB

```{r}
msig_gobp <- msigdbr(
  species = "Homo sapiens",
  collection = "C5"
)

gobp_subcollections <- c(
  "GOBP_RESPONSE_TO_TOPOLOGICALLY_INCORRECT_PROTEIN",
  "GOBP_RESPONSE_TO_HEAT",
  "GOBP_PROTEIN_FOLDING",
  "GOBP_CELLULAR_RESPIRATION",
  "GOBP_AEROBIC_RESPIRATION",
  "GOBP_CELLULAR_RESPONSE_TO_PEPTIDE_HORMONE_STIMULUS",
  "GOBP_ADAPTIVE_THERMOGENESIS",
  "GOBP_NEGATIVE_REGULATION_OF_COLD_INDUCED_THERMOGENESIS",
  "GOBP_POSITIVE_REGULATION_OF_COLD_INDUCED_THERMOGENESIS"
)

gobp_gmt <- build_msigdb_gmt(
  msig_table = msig_gobp,
  selected_sets = gobp_subcollections
)

write_lines(gobp_gmt, GOBP_SIGNATURE)
```

# Run VISION in human Seurat objects

## Discover input Seurat files

```{r}
seurat_files <- list.files(HUMAN_SEURAT_DIR, full.names = TRUE)
```

## Run HEAT only for non-CamaraH human datasets

```{r}
Signatures <- c(HEAT_SIGNATURE)

seurat_files_wo_camarah <- seurat_files[
  !grepl("CamaraH", basename(seurat_files), ignore.case = TRUE)
]

for (f in seurat_files_wo_camarah) {
  runAndSaveVISION(
    seurat_file = f,
    Signatures = Signatures,
    skip_existing = FALSE
  )
}
```

## Run the full signature panel for CamaraH

```{r}
Signatures <- c(
  HEAT_SIGNATURE,
  UCP1_CORR_SIGNATURE,
  BATLAS_SIGNATURE,
  HALLMARK_SIGNATURE,
  GOBP_SIGNATURE
)

runAndSaveVISION(
  seurat_file = CAMARAH_RDS,
  Signatures = Signatures,
  vis_output_dir = VIS_OBJ_OUTPUT_DIR,
  skip_existing = FALSE
)

sapply(Signatures, readLines)
```

# Run VISION in mouse Seurat objects

## Shamsi 2021

```{r}
runAndSaveVISION(
  seurat_file = SHAMSI_RDS,
  Signatures = MOUSE_HEAT_GMT,
  subset_adipocytes = FALSE,
  vis_output_dir = VIS_OBJ_OUTPUT_DIR, skip_existing = TRUE
)
```

# Session Info
```{r}
writeLines(
  capture.output(sessionInfo()),
  file.path(VIS_OBJ_OUTPUT_DIR, "sessionInfo_vision_analysis.txt")
)
```

