---
title: "Emont HEAT Signature"
output: html_document
date: "2024-12-02"
---

## Loading Packages

The first part of the script is to load the packages that we're going to use.

```{r setup, include=FALSE}
pacman::p_load(SingleCellExperiment,
               metap,
               tradeSeq,
               Seurat,
               tidyverse, 
               Matrix,
               scales,
               cowplot,
               viridis,
               VISION,
               DESeq2,
               patchwork,
               DGEobj.utils)
```

The main purpose of this script is to see how the data from Emont et al. (2022) compares to our Neck Adipose Tissue dataset. In this paper, Emont et al. highlight hAd6 as a more thermogenic cluster of adipocytes. Here we'll use VISION to determine signature scores between clusters to see whether this HEAT signature is a good marker for thermogenic adipocytes broadly or whether it would be more specific to neck adipocytes.

## Loading Data

The first part of this script reads in the emont dataset and subsets the dataset to look specifically at adipocytes.

```{r read-data}

emont <- readRDS('/Volumes/Ben/Joslin/Data/Published Dataset/PublishedSingleCellDatasets/Emont 2022/modified/human_all.rds')

UMAPPlot(emont)

Idents(emont) = "cell_type2"
emont.adipo <- subset(emont, idents = "adipocyte")

```

### Running VISION

The next part is to rescale the data in accordance to what is recommended for the VISION workflow. Here the data is put into a separate object that is normalised called "scaled_counts". The UMAP projections are put into an object called "projection" and the associated metadata is put into "meta".

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- emont@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- emont@reductions$umap@cell.embeddings

# Read in meta data (Cells x Vars)
meta = emont@meta.data
```

The next part is to create the Vision object and to analyse that object.

```{r}

# Check for NAs in expression data
sum(is.na(scaled_counts))

Signatures="/Volumes/Ben/Joslin/Output/2.heat_score/5.HEAT_scRNA_human/HEAT_BATLAS_Hallmark.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)

mc.cores = 10
vis <- analyze(vis)

# saveRDS(vis, file = "../Output/15.Emont_dataset_analysis/VISION_emont_HEAT.rds")
# vis <- readRDS(file = "../Output/15.Emont_dataset_analysis/VISION_emont_HEAT.rds")

```

### Creating Object with VISION Data

From this vision analysis we create a new dataset called "graph.data" that is joining the UMAP embeddings witht he signature scores, the sample type (type of adipose tissue probed) and the cell type that was previously identified in the object. The graph produced below is

```{r}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "UCP1_CORR_SUE_GAVRILA_TSENG"]
BATLAS <- getSignatureScores(vis)[, "BATLAS"]
Adipo <- getSignatureScores(vis)[, "HALLMARK_ADIPOGENESIS"]
FA <- getSignatureScores(vis)[, "HALLMARK_FATTY_ACID_METABOLISM"]
OxPhos <- getSignatureScores(vis)[, "HALLMARK_OXIDATIVE_PHOSPHORYLATION"]
sampletype <- emont.adipo$depot
celltype <- emont.adipo$cell_type
individual <- emont.adipo$individual

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(BATLAS) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Adipo) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(FA) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(OxPhos) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(sampletype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(individual) %>% 
              rownames_to_column(var = "cells"), 
            by = "cells") %>% 
 gather(4:8, key = "Hallmark", value = "sigScores")

ggplot(graph.data) + 
  aes(x=UMAP_1, y=UMAP_2, color=sigScores) + 
  geom_point(size = 0.5) +
  scale_color_viridis() +
  facet_grid(sampletype~Hallmark)

```

Following this, we get a graph where we plot the signature scores by the adipocyte subtype described in this paper.

```{r}

ggplot(graph.data, aes(x = celltype, y = sigScores, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) + # Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", vjust = -1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "Boxplot of Signature Scores by Cell Type" # Plot title
  ) +
  facet_grid(~Hallmark)

```

```{r}

VlnPlot(emont.adipo, features = c("PPARGC1A", "EBF2"), split.by = "cell_type", stack = TRUE, flip = TRUE)

```

```{r}

mean.sig.scores <- graph.data %>% 
  filter(celltype == "hAd6") %>% 
  group_by(individual, Hallmark) %>%
  summarise(meansigscore = mean(sigScores))
  

ggplot(mean.sig.scores, aes(x = individual, y = meansigscore, fill = individual)) +
  geom_jitter(aes(color = individual), width = 0.2, alpha = 0.6, size = 0.8) + # Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", vjust = -1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "Boxplot of Signature Scores by Cell Type" # Plot title
  ) +
  facet_grid(~Hallmark)


```

```{r}

# created a norm.sig.scores object that is filtering for just the HEAT signature
norm.sig.scores <- graph.data %>% 
  filter(Hallmark == "HEAT")

mean_sigScore <- mean(norm.sig.scores$sigScore[norm.sig.scores$celltype == "hAd6"])

norm.sig.scores$normalized_sigScore <- norm.sig.scores$sigScore / mean_sigScore

head(norm.sig.scores)

ggplot(norm.sig.scores, aes(x = celltype, y = normalized_sigScore, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) +
  geom_hline(yintercept = 1, color = "grey40", linetype = "dashed") +# Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", vjust = -1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "HEAT Signature Scores by Adipocyte Subcluster" # Plot title
  ) 
  
```

```{r}

# created a norm.sig.scores object that is filtering for just the HEAT signature
norm.sig.scores <- graph.data %>% 
  filter(Hallmark == "BATLAS")

mean_sigScore <- mean(norm.sig.scores$sigScore[norm.sig.scores$celltype == "hAd6"])

norm.sig.scores$normalized_sigScore <- norm.sig.scores$sigScore / mean_sigScore

head(norm.sig.scores)

ggplot(norm.sig.scores, aes(x = celltype, y = normalized_sigScore, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) +
  geom_hline(yintercept = 1, color = "grey40", linetype = "dashed") +# Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", vjust = -1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "BATLAS Signature Scores by Adipocyte Subcluster" # Plot title
  ) 
  
```

#### Figure for publication comparing the HEAT signature scores vs. the clustering done in the original paper.

```{r fig.height = 6, fig.width = 12}

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=UMAP_1, y=UMAP_2, color=sigScores) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(option = "plasma",
                      limits = c(-0.3,2.6)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "Emont et al 2022",
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) 

c <- DimPlot(emont.adipo, reduction = "umap", group.by = "cell_type", label = TRUE, repel = TRUE) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "Emont et al 2022"
  ) + 
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  NoLegend()

c + b


```

### Comparison of SAT vs VAT

This section is intended to do a pseudobulk analysis to see how the different depots (SAT vs. VAT) compare in gene expression, particularly when pairing individuals. This analysis is done primarily with just adipocytes and with ASPCs. The main goal is to ask the question - are there generally a higher expression of thermogenic transcription factors (similar to what was observed in the deep neck adipose tissue) that could imply one depot over the other is more prone to thermogenesis?

```{r}

emont.adipo <- subset(emont, idents = c("adipocyte", "ASPC"))

```

```{r}

cell_level = "cell_type2" 

cell_level_sym = sym(cell_level)
cts = AggregateExpression(emont.adipo, assays = "RNA", 
                          slot = "counts", #DESeq2 uses raw counts
                          return.seurat = F, 
                          group.by = c("individual", cell_level, "depot"))
cts = cts$RNA %>% 
  t() %>% 
  as.data.frame()

```

```{r}

split_names = strsplit(rownames(cts), "_")
df <- t(data.frame(split_names)) %>% 
  as.data.frame()
# Replace all "-" with "_" in every value of the dataframe

colnames(df) = c("donor_id", cell_level, "depot")
for (col in colnames(df)){
  df[[col]] = gsub("-", "_", df[[col]])}
df_rownames = paste(df$donor_id, df[[cell_level]], df$depot, sep = "_")
rownames(df) = df_rownames

```

```{r}
t.cts = t(cts)
CPM_cts = convertCounts(
  t.cts,
  unit = "CPM",
  log = FALSE,
  normalize = "none",
  prior.count = NULL
)
rm(t.cts)
# Assuming 'df' is your dataframe
column_sums <- colSums(CPM_cts, na.rm = TRUE)

```

```{r}

CPMcts.filtered <- CPM_cts %>% 
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Comparison") %>%
  gather(key = Gene, value = Expression, -1) %>% 
    dplyr::left_join(df %>% 
                     rownames_to_column(var = "Comparison"), by = "Comparison") %>% 
  mutate(cell_type2 = factor(cell_type2, levels = c("ASPC", "adipocyte")))

```

##### Thermogenic Transcription Factors

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("EBF2", "MECOM", "PRDM16", "PPARGC1A")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)") +
           facet_wrap(~cell_type2))

}

EBF2 + MECOM + PRDM16 + PPARGC1A + plot_layout(ncol = 4)
rm(EBF2, MECOM, PRDM16, PPARGC1A)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 12}

genes <- c("EBF2", "MECOM", "PRDM16", "PPARGC1A")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id, cell_type2) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)") +
    facet_grid(~cell_type2))
}

EBF2 + MECOM + PRDM16 + PPARGC1A + plot_layout(ncol = 4)
rm(EBF2, MECOM, PRDM16, PPARGC1A)

```

##### Thermogenic Machinery Genes

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("CKMT2", "GK", "CA12", "BDH1")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)") +
           facet_wrap(~cell_type2))

}

CKMT2 + GK + CA12 + BDH1 + plot_layout(ncol = 4)
rm(CKMT2, GK, CA12, BDH1)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 12}

genes <- c("CKMT2", "GK", "CA12", "BDH1")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id, cell_type2) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)") +
    facet_grid(~cell_type2))
}

CKMT2 + GK + CA12 + BDH1 + plot_layout(ncol = 4)
rm(CKMT2, GK, CA12, BDH1)

```


##### Adipocyte Transcript Factor Genes

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("TWIST1", "CRNDE", "SIM1", "DMRT2")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)") +
           facet_wrap(~cell_type2))

}

TWIST1 + CRNDE + SIM1 + DMRT2 + plot_layout(ncol = 4)
rm(TWIST1, CRNDE, SIM1, DMRT2)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 12}

genes <- c("TWIST1", "CRNDE", "SIM1", "DMRT2")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id, cell_type2) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)") +
    facet_grid(~cell_type2))
}

TWIST1 + CRNDE + SIM1 + DMRT2 + plot_layout(ncol = 4)
rm(TWIST1, CRNDE, SIM1, DMRT2)

```

##### Adipocyte Machinery Genes

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("ADIPOQ", "PLIN1", "PLIN4", "FABP4")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)") +
           facet_wrap(~cell_type2))

}

ADIPOQ + PLIN1 + PLIN4 + FABP4 + plot_layout(ncol = 4)
rm(ADIPOQ, PLIN1, PLIN4, FABP4)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 12}

genes <- c("ADIPOQ", "PLIN1", "PLIN4", "FABP4")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id, cell_type2) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)") +
    facet_grid(~cell_type2))
}

ADIPOQ + PLIN1 + PLIN4 + FABP4 + plot_layout(ncol = 4)
rm(ADIPOQ, PLIN1, PLIN4, FABP4)

```

##### Developmental Genes

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 18}

# View the result

genes <- c("PAX3", "HOXD4", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)") +
           facet_wrap(~cell_type2))

}

PAX3 + HOXD4 + HOXA3 + HOXA4 + HOXA5 + HOXA6 + plot_layout(ncol = 6)
rm(PAX3, HOXD4, HOXA3, HOXA4, HOXA5, HOXA6)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 18}

genes <- c("PAX3", "HOXD4", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id, cell_type2) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)") +
    facet_grid(~cell_type2))
}

PAX3 + HOXD4 + HOXA3 + HOXA4 + HOXA5 + HOXA6 + plot_layout(ncol = 6)
rm(PAX3, HOXD4, HOXA3, HOXA4, HOXA5, HOXA6)

```

### Comparison of SAT vs VAT (excluding hAd6)

```{r}

Idents(emont) = "cell_type2"
emont.adipo.2 <- subset(emont, idents = c("adipocyte"))

Idents(emont.adipo.2) = "cell_type"
emont.adipo.2 <- subset(emont.adipo.2, idents = c("hAd1", "hAd2", "hAd3", "hAd4", "hAd5", "hAd7"))

```

```{r}

emont.adipo@meta.data %>% 
  dplyr::filter(is.na(compartment)) %>% 
  group_by(sample) %>% 
  summarise(n = n())

```

```{r}

cell_level = "cell_type2" 

cell_level_sym = sym(cell_level)
cts = AggregateExpression(emont.adipo.2, assays = "RNA", 
                          slot = "counts", #DESeq2 uses raw counts
                          return.seurat = F, 
                          group.by = c("individual", "depot"))
cts = cts$RNA %>% 
  t() %>% 
  as.data.frame()

```

```{r}

split_names = strsplit(rownames(cts), "_")
df <- t(data.frame(split_names)) %>% 
  as.data.frame()
# Replace all "-" with "_" in every value of the dataframe

colnames(df) = c("donor_id", "depot")
for (col in colnames(df)){
  df[[col]] = gsub("-", "_", df[[col]])}
df_rownames = paste(df$donor_id, df$depot, sep = "_")
rownames(df) = df_rownames

```

```{r}
t.cts = t(cts)
CPM_cts = convertCounts(
  t.cts,
  unit = "CPM",
  log = FALSE,
  normalize = "none",
  prior.count = NULL
)
rm(t.cts)
# Assuming 'df' is your dataframe
column_sums <- colSums(CPM_cts, na.rm = TRUE)

```

```{r}

CPMcts.filtered <- CPM_cts %>% 
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Comparison") %>%
  gather(key = Gene, value = Expression, -1) %>% 
    dplyr::left_join(df %>% 
                     rownames_to_column(var = "Comparison"), by = "Comparison")

```

##### Thermogenic Transcription Factors

The purpose of this section is to plot the thermogenic transcription factors both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 8}

# View the result

genes <- c("EBF2", "MECOM", "PRDM16", "PPARGC1A")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)"))
}

EBF2 + MECOM + PRDM16 + PPARGC1A + plot_layout(ncol = 4)
rm(EBF2, MECOM, PRDM16, PPARGC1A)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 8}

genes <- c("EBF2", "MECOM", "PRDM16", "PPARGC1A")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)"))
}

EBF2 + MECOM + PRDM16 + PPARGC1A + plot_layout(ncol = 4)
rm(EBF2, MECOM, PRDM16, PPARGC1A)

```

##### Thermogenic Machinery Genes

The purpose of this section is to plot the gene Adipocyte Machinery genes both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 8}

# View the result

genes <- c("CKMT2", "GK", "CA12", "BDH1")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)"))
}

CKMT2 + GK + CA12 + BDH1 + plot_layout(ncol = 4)
rm(CKMT2, GK, CA12, BDH1)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 8}

genes <- c("CKMT2", "GK", "CA12", "BDH1")

for (gene in genes) {
  
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)"))
}

CKMT2 + GK + CA12 + BDH1 + plot_layout(ncol = 4)
rm(CKMT2, GK, CA12, BDH1)

```

##### Adipocyte Transcript Factor Genes

The purpose of this section is to plot the gene Adipocyte Machinery genes both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 8}

# View the result

genes <- c("TWIST1", "CRNDE", "SIM1", "DMRT2")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)"))
}

TWIST1 + CRNDE + SIM1 + DMRT2 + plot_layout(ncol = 4)
rm(TWIST1, CRNDE, SIM1, DMRT2)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 8}

genes <- c("TWIST1", "CRNDE", "SIM1", "DMRT2")

for (gene in genes) {
  
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)"))
}

TWIST1 + CRNDE + SIM1 + DMRT2 + plot_layout(ncol = 4)
rm(TWIST1, CRNDE, SIM1, DMRT2)

```

##### Adipocyte Machinery Genes

The purpose of this section is to plot the gene Adipocyte Machinery genes both as boxplots (without pairing) and as line plots. This first set are the boxplots:

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("ADIPOQ", "PLIN1", "PLIN4", "FABP4")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)"))
}

ADIPOQ + PLIN1 + PLIN4 + FABP4 + plot_layout(ncol = 4)
rm(ADIPOQ, PLIN1, PLIN4, FABP4)

```

This second set are the line plots:

```{r, fig.height = 4, fig.width = 8}

genes <- c("ADIPOQ", "PLIN1", "PLIN4", "FABP4")

for (gene in genes) {
  # gene = "EBF2"
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)"))
}

ADIPOQ + PLIN1 + PLIN4 + FABP4 + plot_layout(ncol = 4)
rm(ADIPOQ, PLIN1, PLIN4, FABP4)

```

##### Developmental Genes

```{r, fig.height = 4, fig.width = 12}

# View the result

genes <- c("PAX3", "HOXD4", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

for(gene in genes){
  
  assign(gene,
         ggboxplot(CPMcts.filtered %>% 
                  dplyr::filter(Gene == gene), 
                x = "depot", 
                y = "Expression", 
                fill = "depot", 
                add = "jitter") +
           stat_compare_means(method = "anova") +
           scale_fill_manual(values = c("#ffe29e", "#ffe29e")) +
           theme(plot.title = element_text(hjust = 0.5),
                 axis.title.x = element_blank(),
                 axis.text.x = element_text(angle = 45, hjust = 1)) +
           labs(title = gene,
                y = "Expression (CPM)"))
}

PAX3 + HOXD4 + HOXA3 + HOXA4 + HOXA5 + HOXA6 + plot_layout(ncol = 6)
rm(PAX3, HOXD4, HOXA3, HOXA4, HOXA5, HOXA6)

```

```{r, fig.height = 4, fig.width = 12}

genes <- c("PAX3", "HOXD4", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

for (gene in genes) {
  
  data <- CPMcts.filtered %>%
    dplyr::filter(Gene == gene) %>%  
    group_by(donor_id) %>%  
    dplyr::filter(all(c("VAT", "SAT") %in% depot)) %>% 
    ungroup()

  assign(gene, ggline(data,
                x = "depot",
                y = "Expression",
                group = "donor_id") +
    coord_cartesian(ylim = c(0, max(data$Expression) * 1.2)) +
    stat_compare_means(method = "t.test", 
                       label.y = max(data$Expression) * 1.15,
                       aes(label = paste0("p = ", after_stat(p.format))),
                       paired = TRUE) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(title = gene, y = "Expression (CPM)"))
}

PAX3 + HOXD4 + HOXA3 + HOXA4 + HOXA5 + HOXA6 + plot_layout(ncol = 6)
rm(PAX3, HOXD4, HOXA3, HOXA4, HOXA5, HOXA6)

```
