---
title: "18.Wang_dataset_analysis"
output: html_document
date: "2024-12-04"
---

```{r setup, include=FALSE}
pacman::p_load(SingleCellExperiment,
               metap,
               tradeSeq,
               Seurat,
               tidyverse, 
               Matrix,
               scales,
               cowplot,
               viridis,
               VISION)
```

This script is for doing any analysis of the subcutaneous adipose tissue dataset described in Wang et al 2024 from Christian Wolfrum's lab where they are characterising a set of thermogenic adipocytes that have a characterisitic ATP-dependent futile cycle. This dataset is produced from 4 human samples. The purpose of this analysis is to see how our HEAT signature maps onto their data. This first part just reads in their data.

```{r}

wang <- readRDS("/Volumes/Ben/Joslin/Data/Published Dataset/PublishedSingleCellDatasets/Wang2024/original/GSE184687_human_subcutaneous_AT.rds")

DimPlot(wang, reduction = "umap", group.by = "seurat_clusters", label = TRUE)

```

Next, cells are labelled based on the descriptions given in their original paper. On looking at the markers in their clusters, it seemed like they had some odd annotations - particular that one of their skeletal muscle clusters seemed likely to be smooth muscle and other seemed likely to be lymphatic endothelial cells. The relabelled identity is provided as "new.clust.ident".

```{r}

metadata <- wang@meta.data

metadata$clust.ident[which(metadata$seurat_clusters==0)] <- "Preadipocytes"
metadata$clust.ident[which(metadata$seurat_clusters==1)] <- "Adipocytes"
metadata$clust.ident[which(metadata$seurat_clusters==2)] <- "Endothelial cells 1"
metadata$clust.ident[which(metadata$seurat_clusters==3)] <- "Endothelial cells 2"
metadata$clust.ident[which(metadata$seurat_clusters==4)] <- "Endothelial cells 3"
metadata$clust.ident[which(metadata$seurat_clusters==5)] <- "Macrophages"
metadata$clust.ident[which(metadata$seurat_clusters==6)] <- "Skeletal muscle 1"
metadata$clust.ident[which(metadata$seurat_clusters==7)] <- "Skeletal muscle 2"
metadata$clust.ident[which(metadata$seurat_clusters==8)] <- "Skeletal muscle 3"
metadata$clust.ident[which(metadata$seurat_clusters==9)] <- "Cells unknown"
metadata$clust.ident[which(metadata$seurat_clusters==10)] <- "CD4+ Tcm"
metadata$clust.ident[which(metadata$seurat_clusters==11)] <- "Skeletal muscle 4"
metadata$clust.ident[which(metadata$seurat_clusters==12)] <- "Endothelial cells 4"

metadata$new.clust.ident[which(metadata$seurat_clusters==0)] <- "Preadipocytes"
metadata$new.clust.ident[which(metadata$seurat_clusters==1)] <- "Adipocytes"
metadata$new.clust.ident[which(metadata$seurat_clusters==2)] <- "Endothelial cells 1"
metadata$new.clust.ident[which(metadata$seurat_clusters==3)] <- "Endothelial cells 2"
metadata$new.clust.ident[which(metadata$seurat_clusters==4)] <- "Endothelial cells 3"
metadata$new.clust.ident[which(metadata$seurat_clusters==5)] <- "Macrophages"
metadata$new.clust.ident[which(metadata$seurat_clusters==6)] <- "Smooth Muscle Cell"
metadata$new.clust.ident[which(metadata$seurat_clusters==7)] <- "Skeletal muscle 1"
metadata$new.clust.ident[which(metadata$seurat_clusters==8)] <- "Lymphatic EC"
metadata$new.clust.ident[which(metadata$seurat_clusters==9)] <- "Cells unknown"
metadata$new.clust.ident[which(metadata$seurat_clusters==10)] <- "CD4+ Tcm"
metadata$new.clust.ident[which(metadata$seurat_clusters==11)] <- "Skeletal muscle 2"
metadata$new.clust.ident[which(metadata$seurat_clusters==12)] <- "Endothelial cells 4"

wang@meta.data <- metadata

a <- DimPlot(wang, reduction = "umap", group.by = "clust.ident", label = TRUE) + NoLegend()
b <- DimPlot(wang, reduction = "umap", group.by = "new.clust.ident", label = TRUE) + NoLegend()

a + b

```

This next chunk looks at the expression of some of the standard markers that we use for describing populations within our neck adipose tissue dataset.

```{r}

DefaultAssay(wang) = "RNA"

features = c("ADIPOQ","PPARGC1A", "PDGFRA", "DCN", "MYOCD", "PDE4D",  "ADGRF5", "JAM2", "STEAP4", "PKHD1L1", "CDH19", "MPZ",  "CARD11", "CD163", "CPA3", "CSF3R", "CNTNAP2")

VlnPlot(wang, 
        features = features, 
        stack = TRUE, 
        flip = TRUE, 
        group.by = "clust.ident") + 
  NoLegend() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 1.5), "cm"))

DefaultAssay(wang) = "integrated"

```

Next, the dataset is prepared according to the VISION workflow. In VISION, we will characterise the HEAT signature in addition to the general adipogenesis signature and the BATLAS signature produced by the Wolfrum laboratory. In this first part, the object is split into multiple parts that will be re-inserted into a vision object in the next step.

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- wang@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- wang@reductions$umap@cell.embeddings

# Read in meta data (Cells x Vars)
meta = wang@meta.data
```

This next chunk loads the VISION signatures, loads the data that was split in the previous chunk, then runs the algorithm. The resulting output is default saved as an object (since this process takes 10-20 minutes, this can expedite future analyses.

```{r}

Signatures='/Volumes/Ben/Joslin/Output/2.heat_score/5.HEAT_scRNA_human/HEAT_BATLAS_Hallmark.gmt'

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)

# saveRDS(vis, file = "../Output/19.Wang_dataset_analysis/VISION_wang_HEAT.rds")
# vis <- readRDS(file = "../Output/19.Wang_dataset_analysis/VISION_wang_HEAT.rds")

```

Here the output scores are pasted onto a table with the UMAP embeddings along with the cluster annotations above. The purpose of this is to be able to plot the signature scores either on the UMAP embeddings or by cell type. In this first chunk, the different signature scores are mapped on the UMAP embeddings.

```{r}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "UCP1_CORR_SUE_GAVRILA_TSENG"]
BATLAS <- getSignatureScores(vis)[, "BATLAS"]
Adipo <- getSignatureScores(vis)[, "HALLMARK_ADIPOGENESIS"]
FA <- getSignatureScores(vis)[, "HALLMARK_FATTY_ACID_METABOLISM"]
OxPhos <- getSignatureScores(vis)[, "HALLMARK_OXIDATIVE_PHOSPHORYLATION"]
sampletype <- wang$orig.ident
celltype <- wang$new.clust.ident

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(BATLAS) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Adipo) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(FA) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(OxPhos) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(sampletype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  gather(4:8, key = "Hallmark", value = "sigScores")


ggplot(graph.data) + 
  aes(x = UMAP_1, y = UMAP_2, color = sigScores) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(option = "plasma") +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "UMAP Plots for Different Signatures",
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
  ) +
  facet_wrap(~Hallmark)

ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x = UMAP_1, y = UMAP_2, color = sigScores) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(option = "plasma") +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "UMAP Plot of HEAT Signature",
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
  )



```

Note in these graphs that the Adipogenesis, BATLAS, and Fatty acid gene signatures strongly mark the adipocyte cluster. By comparison, the HEAT signature has a relatively strong score for the skeletal muscle which is consistent with a more metabolically active gene signature.

In the below chunk, these signature scores (HEAT, BATLAS, and adipogenesis) are further investigated split by cluster.

```{r}


ggplot((graph.data %>% dplyr::filter(Hallmark == "HEAT")), aes(x = celltype, y = sigScores, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) + # Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", angle = 45, hjust = 1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "Boxplot of HEAT Signature Scores by Cell Type" # Plot title
  ) 

ggplot((graph.data %>% dplyr::filter(Hallmark == "BATLAS")), aes(x = celltype, y = sigScores, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) + # Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", angle = 45, hjust = 1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "Boxplot of BATLAS Signature Scores by Cell Type" # Plot title
  ) 

ggplot((graph.data %>% dplyr::filter(Hallmark == "Adipo")), aes(x = celltype, y = sigScores, fill = celltype)) +
  geom_jitter(aes(color = celltype), width = 0.2, alpha = 0.6, size = 0.8) + # Add jittered points
  geom_boxplot(outlier.shape = NA,  color = "black", width = 0.6) + # Boxplot aesthetics
  scale_fill_brewer(palette = "Set3") +
  scale_color_brewer(palette = "Set3") + # Optional: Nice color palette for points
  theme_minimal() + # Clean theme
  theme(
    axis.text.x = element_text(size = 14, face = "bold", angle = 45, hjust = 1), # Rotate x-axis labels for better readability
    axis.text.y = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 16, face = "bold"),
    panel.grid.major.y = element_line(color = "grey90", size = 0.6), # Light horizontal gridlines
    panel.grid.major.x = element_blank(), # Remove vertical gridlines
    panel.grid.minor = element_blank(), # Remove minor gridlines
    panel.border = element_blank(), # Remove border
    axis.line = element_line(color = "black", size = 1.2), # Add solid x and y axes
    axis.ticks = element_line(color = "black", size = 1.2) # Add solid axis ticks
  ) +
  labs(
    x = NULL, # "Cell Type", # Label for x-axis
    y = "Signature Scores", # Label for y-axis
    title = "Boxplot of Adipo Signature Scores by Cell Type" # Plot title
  ) 

```

