---
title: "Angueira analysis"
author: "Henrique"
date: "2025-11-17"
output: html_document
---

# Introduction

This notebook analyzes the Angueira dataset with two main goals:

1. Relate HEAT signature scores to UCP1 expression in adipocytes.
2. Recluster Angueira adipocytes and quantify the fraction of HEAT-positive
   cells per adipocyte cluster.

The retained outputs are:

- `Angueira_HEATvUCP1.csv`
- `AngueiraA_Dataset_adipo_recluster.rds`
- `heat_pos_adipo_anguiera.csv`

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(tidyverse)
library(Seurat)
library(here)

# -------------------------------------------------------------------------
# Define inputs and outputs
# -------------------------------------------------------------------------
ANGUEIRA_RDS <- here(
"output/4.downstream_pipelines/vision/data/cleaned_datasets_human/AngueiraA_Dataset.rds"
)

ANGUEIRA_VISION_SIGSCORE <- here(
  "output/4.downstream_pipelines/vision/data/vision_signature_scores/AngueiraA_Dataset_VISION_sigScore.csv"
)

OUTPUT_DATA_DIR <- here("output/3.heat_signature/3.public_snrnaseq/data")
OUTPUT_FIGURE_DIR <- here("output/3.heat_signature/3.public_snrnaseq/figures")
CLEANED_DATASET_DIR <- here("output/1.processing/data/cleaned_objects")

dir.create(OUTPUT_DATA_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(OUTPUT_FIGURE_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(CLEANED_DATASET_DIR, showWarnings = FALSE, recursive = TRUE)

# -------------------------------------------------------------------------
# Analysis settings
# -------------------------------------------------------------------------
wad_HEAT_percentile <- 0.989345 ## CONSIDERING READING IT FROM A GENERATED FILE
```

# Load dataset and signature scores

```{r}
s.object <- readRDS(ANGUEIRA_RDS)
sigscores_df <- read.csv(ANGUEIRA_VISION_SIGSCORE)
```

# Relate HEAT signature score to UCP1 expression in adipocytes

```{r}
Idents(s.object) <- "cell_type"
s.object.small <- subset(s.object, idents = "Adipocytes")

ucp1_df <- FetchData(s.object.small, vars = "UCP1", layer = "data")

ucp1_v_sigscores <- sigscores_df |>
  filter(Hallmark == "HEAT") |>
  left_join(ucp1_df |> rownames_to_column("barcode"), by = "barcode") |>
  filter(cell_type == "Adipocytes")

write_csv(
  ucp1_v_sigscores,
  file.path(OUTPUT_DATA_DIR, "Angueira_HEATvUCP1.csv")
)
```

# Recluster adipocytes

```{r}
s.object.new <- subset(s.object, idents = "Adipocytes")

s.object.new <- NormalizeData(object = s.object.new)
s.object.new <- FindVariableFeatures(object = s.object.new)
s.object.new <- ScaleData(object = s.object.new)
s.object.new <- SCTransform(object = s.object.new)

s.object.new <- RunPCA(object = s.object.new)
s.object.new <- FindNeighbors(object = s.object.new, dims = 1:30)
s.object.new <- FindClusters(
  object = s.object.new,
  resolution = 0.5,
  cluster.name = "unintegrated_clusters"
)
s.object.new <- RunUMAP(object = s.object.new, dims = 1:30)
```

# Add HEAT score and define HEAT-positive cells

```{r}
meta <- s.object.new@meta.data

if (!"sigScores" %in% colnames(meta)) {
  meta <- meta |>
    rownames_to_column("barcode") |>
    left_join(
      sigscores_df |>
        filter(Hallmark == "HEAT") |>
        select(barcode, HEAT = sigScores),
      by = "barcode",
      unmatched = "drop"
    ) |>
    column_to_rownames("barcode")
}

meta <- meta |>
  mutate(heat_positive = HEAT >= wad_HEAT_percentile)

s.object.new@meta.data <- meta
```

# Save reclustered Seurat object

```{r}
saveRDS(
  s.object.new,
  file.path(CLEANED_DATASET_DIR, "AngueiraA_Dataset_adipo_recluster.rds")
)
```

# Count HEAT-positive adipocytes by cluster

```{r}
heat_pos_anguiera <- meta |>
  group_by(unintegrated_clusters) |>
  count(heat_positive) |>
  mutate(heat_positive_pct = n / sum(n) * 100)

write_csv(
  heat_pos_anguiera,
  file.path(OUTPUT_DATA_DIR, "heat_pos_adipo_anguiera.csv")
)
```
