---
title: "33.MouseSingleCellDatasets"
output: html_document
date: "2025-03-11"
---

```{r setup, include=FALSE}

pacman::p_load(Seurat,
               tidyverse,
               SoupX,
               biomaRt,
               reticulate,
               anndata, 
               schard,
               VISION)

```

## Shamsi 2022

``` {r}

shamsi <- readRDS("/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Shamsi 2022 - scRNAseq with immune cells/scBAT_scanpy_noCold2_1_reprocess_paga_assign_cluster_celltype_updated.rds")

```

``` {r}

a <- DimPlot(shamsi, label = TRUE) + NoLegend()

shamsi.noimmun <- subset(shamsi, idents = c("VSM", "EC", "Pericytes", "Lymph_EC", "Pdgfra_APC", "NMSC", "MSC", "Adipocytes", "Macrophages"))

b <- DimPlot(shamsi.noimmun, label = TRUE) + NoLegend()

a + b

rm(shamsi)

```

#### Vision Analysis

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- shamsi.noimmun@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- shamsi.noimmun@reductions$umap@cell.embeddings

# Read in meta data (Cells x Vars)
meta = shamsi.noimmun@meta.data
```

The next part is to create the Vision object and to analyse that object.

``` {r}

Signatures="/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Output/VISION/Gene Signatures/mouse_heat.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)

# saveRDS(vis, file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Shamsi 2022 - scRNAseq with immune cells/VISION_Shamsi_HEAT.rds")
# vis <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Shamsi 2022 - scRNAseq with immune cells/VISION_Shamsi_HEAT.rds")

```

``` {r fig.width = 8, fig.height = 4}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "mouse_heat"]
celltype <- shamsi.noimmun$cell_type
cond <- shamsi.noimmun$cond
Ucp1 <- FetchData(shamsi.noimmun, vars = "Ucp1")

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Ucp1) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(cond) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
 gather(4, key = "Hallmark", value = "sigScores") 

a <- DimPlot(shamsi.noimmun, 
             reduction = "umap", 
             label = TRUE, 
             pt.size = 0.5,
             repel = TRUE) + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank()) +
    NoLegend()

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=umap_1, y=umap_2, color=as.numeric(sigScores)) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(limits = c(-0.3,3.5)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(~Hallmark)

c <- FeaturePlot(shamsi.noimmun, reduction = "umap", feature = "Ucp1") + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
        ) +
  NoLegend()

a + b + c

```


``` {r, fig.height = 3.5, fig.width = 3.5}

adipo.genes <- c("Prdm16", "Ebf2", "Adipoq", "Plin1")

graph.data <- graph.data %>% 
  left_join(as.data.frame(FetchData(shamsi.noimmun, vars = adipo.genes)) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") 

adipo.data <- graph.data %>% 
  dplyr::filter(celltype == "Adipocytes") 

ggplot(adipo.data) +
  geom_point(aes(x = sigScores, y = Prdm16), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Prdm16 in Adipocytes")

ggplot(adipo.data,
       aes(x = sigScores, 
           y = log10(Ucp1))) +
  geom_point(, size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ucp1 in Adipocytes") + 
  xlab("HEAT Signature Score") +
  ylab(expression(log[10] * " UCP1 Expression")) +
  geom_smooth(method = "lm", 
              linetype = "dashed",
              color = "black",
              se = FALSE) + 
  stat_poly_eq(aes(label = paste(..rr.label..,
                                 ..p.value.label.., sep = "~~~")),
               formula = y ~ x,
               parse = TRUE)

ggplot(adipo.data) +
  geom_point(aes(x = sigScores, y = Ebf2), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ebf2 in Adipocytes")

```

``` {r}

ggplot(graph.data %>% 
         dplyr::filter(celltype == "Adipocytes"),
       aes(x = cond,
           y = sigScores)) +
  geom_jitter() +
  geom_violin() + 
  theme_pubr()

ggplot(graph.data %>% 
         dplyr::filter(celltype == "VSM"),
       aes(x = cond,
           y = sigScores)) +
  geom_jitter() +
  geom_violin() + 
  theme_pubr()

ggplot(graph.data, 
         aes(x = celltype,
           y = sigScores)) +
  geom_violin() + 
  theme_pubr()

ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT") %>% 
         dplyr::filter(celltype == "VSM")) + 
  aes(x=umap_1, y=umap_2, color=as.numeric(sigScores)) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(limits = c(-0.3,1.5)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(~cond)

```

``` {r, fig.height = 18, fig.width = 4.5}

VlnPlot(shamsi.noimmun,
        features = mouse.heat,
        stack = TRUE,
        flip = TRUE) +
  NoLegend()

```

``` {r, fig.height = 18, fig.width = 3}

vsm <- subset(shamsi.noimmun, idents = "VSM")
Idents(vsm) = "cond"
VlnPlot(vsm,
        features = mouse.heat,
        stack = TRUE,
        flip = TRUE,
        split.by = "cond") +
  NoLegend()

```

## Yu 2024 - Involution in mouse BAT

``` {r}

young <- Read10X(data.dir = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Yu 2024/GSM7920451")
mid <- Read10X(data.dir = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Yu 2024/GSM7920452")
old <- Read10X(data.dir = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Yu 2024/GSM7920453")

young = CreateSeuratObject(young, project = "young")
mid = CreateSeuratObject(mid, project = "mid")
old = CreateSeuratObject(old, project = "old")

```

#### Merge Datasets

Mouse datasets

``` {r}

yu.mouse = merge(x = young,
                 y = c(mid, 
                       old),
                 add.cell.ids = c("young", 
                                  "mid", 
                                  "old"))

```

``` {r}

yu.mouse=NormalizeData(yu.mouse)
yu.mouse=FindVariableFeatures(yu.mouse)
yu.mouse=ScaleData(yu.mouse)
yu.mouse=RunPCA(yu.mouse)

# The below is plotting how many PCs are required. 

ElbowPlot(yu.mouse,ndims = 30)

```

#### Filtering by mitochondrial RNA

``` {r}

# Infer UMAP based on number of PCs selected above

yu.mouse = RunUMAP(yu.mouse,
                   dims = 1:30,
                   reduction.name = "umap.unintegrated")
DimPlot(yu.mouse,
        reduction = "umap.unintegrated")

# Visualize the mitochondrial content and ribosomal gene content. If there are specific clusters that have a number of mito genes, we can simply remove the whole cluster.

yu.mouse[["percent.mt"]] <- PercentageFeatureSet(yu.mouse, pattern = "^mt-")
yu.mouse[["percent.rpl"]] <- PercentageFeatureSet(yu.mouse, pattern = "^rpl")

FeaturePlot(yu.mouse,"percent.rpl")
FeaturePlot(yu.mouse,"percent.mt")

VlnPlot(yu.mouse,"percent.mt")

yu.mouse <- subset(yu.mouse, subset = percent.mt < 1)

FeaturePlot(yu.mouse, features = "Plin1")
FeaturePlot(yu.mouse, features = "Ucp1")

```

#### Integrating Datasets

``` {r}

yu.mouse <- IntegrateLayers(object = yu.mouse, 
                            method = CCAIntegration, 
                            orig.reduction = "pca", 
                            new.reduction = "integrated.cca",
                            verbose = FALSE)

# re-join layers after integration
yu.mouse[["RNA"]] <- JoinLayers(yu.mouse[["RNA"]])

yu.mouse <- FindNeighbors(yu.mouse, reduction = "integrated.cca", dims = 1:30)
yu.mouse <- FindClusters(yu.mouse, resolution = 0.8)
yu.mouse <- RunUMAP(yu.mouse, dims = 1:30, reduction = "integrated.cca")

a <- DimPlot(yu.mouse, reduction = "umap.unintegrated")
b <- DimPlot(yu.mouse, reduction = "umap")

a + b

```

#### Visualising Markers and annotating clusters

``` {r, fig.height = 10, fig.width = 5}

FeaturePlot(yu.mouse,
            features = c("Ucp1", "Adipoq", "Prdm16", "Ebf2"))

VlnPlot(yu.mouse,
        features = c("Ucp1", "Plin1", "Adipoq", "Prdm16", "Ebf2", "Pdgfra", 
                     "Jam2", "Shank3", "Myocd", "Prox1", "Steap4",
                     "Cd163", "Cpa3", "Il7r", "Csf3r", "Cdh19"),
        stack = TRUE,
        flip = TRUE)

```

``` {r}

new.cluster.ids <- c("Adipocyte", "Adipocyte", "Adipocyte", "Adipocyte", "Adipocyte", 
                     "Adipocyte", "Adipocyte", "Adipocyte", "Adipocyte Progenitor", "Pericytes",
                     "Endothelial Cell", "Unknown 1", "Adipocyte", "Endothelial Cell",
                     "Unknown 1", "Unknown 2")
names(new.cluster.ids) <- levels(yu.mouse)
yu.mouse <- RenameIdents(yu.mouse, new.cluster.ids)
yu.mouse$cell_type <- Idents(yu.mouse)
DimPlot(yu.mouse, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()

```

<!-- ``` {r} -->

<!-- # Convert Seurat object into SoupX format -->
<!-- sc <- SoupChannel(toc = yu.mouse@assays$RNA$data) -->
<!-- # sc$tod <- yu.mouse@assays$RNA$counts -->

<!-- # Add metadata-based cell type annotation -->
<!-- metaData <- yu.mouse@meta.data -->
<!-- cellTypes <- setNames(metaData$cell_type, rownames(metaData)) -->

<!-- # Estimate contamination fraction -->
<!-- sc <- setClusters(sc, cellTypes)  # Assign clusters based on 'cell_types' -->
<!-- sc <- autoEstCont(sc)  # Automatically estimate contamination fraction -->

<!-- # Apply SoupX correction -->
<!-- corrected_counts <- adjustCounts(sc) -->

<!-- # Create a new Seurat object with corrected counts (optional) -->
<!-- s.object[["RNA_corrected"]] <- CreateAssayObject(corrected_counts) -->

<!-- # Set default assay and normalize -->
<!-- DefaultAssay(s.object) <- "RNA_corrected" -->
<!-- s.object <- NormalizeData(s.object) -->

<!-- # Check metadata -->
<!-- head(s.object@meta.data) -->

<!-- all(rownames(yu.mouse@meta.data) %in% colnames(sc$tod))  # Should return TRUE -->

<!-- sum(rowSums(sc$tod) == 0)  # Count genes with zero expression -->
<!-- sum(colSums(sc$tod) == 0)  # Count cells with zero expression -->

<!-- ``` -->


<!-- ``` {r} -->

<!-- yu.mouse = autoEstCont(yu.mouse) -->
<!-- out = adjustCounts(yu.mouse) -->

<!-- ``` -->



#### Heat Markers


``` {r}

mouse_human_genes = read.csv("https://www.informatics.jax.org/downloads/reports/HOM_MouseHumanSequence.rpt",sep="\t")

HEAT <- read.csv("../Data/Published Bulk Datasets/GSE82155_RAW - Neonates vs infancy epicardial/HEAT_Gene_Signature.csv", header = FALSE)

heat_conversion <- mouse_human_genes %>% 
  dplyr::select(DB.Class.Key, Common.Organism.Name, Symbol) %>% 
  distinct(DB.Class.Key, Common.Organism.Name, .keep_all = TRUE) %>% 
  tidyr::spread(key = "Common.Organism.Name", value = "Symbol") %>% 
  dplyr::filter(human %in% HEAT$V1)

mouse.heat <- heat_conversion$'mouse, laboratory'
rm(heat_conversion)

```

``` {r}

yu.mouse$orig.ident <- factor(yu.mouse$orig.ident, levels = c("young", "mid", "old"))
FeaturePlot(yu.mouse, features = "Ppargc1a")
FeaturePlot(yu.mouse, features = "Ucp1", split.by = "orig.ident")

```

#### Vision Analysis


```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- yu.mouse@assays$RNA@layers$counts
rownames(counts) = rownames(yu.mouse@assays$RNA@features)
colnames(counts) = rownames(yu.mouse@assays$RNA@cells)

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- yu.mouse@reductions$umap@cell.embeddings

# Read in meta data (Cells x Vars)
meta = yu.mouse@meta.data
```

The next part is to create the Vision object and to analyse that object.

``` {r}

Signatures="/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Output/VISION/Gene Signatures/mouse_heat.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)

# saveRDS(vis, file = "../Output/18.NeckAdipose_dataset_analysis/VISION_neck_HEAT.rds")
# vis <- readRDS(file = "../Output/18.NeckAdipose_dataset_analysis/VISION_neck_HEAT.rds")

```

``` {r fig.width = 8, fig.height = 4}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "mouse_heat"]
celltype <- yu.mouse$cell_type
Ucp1 <- FetchData(yu.mouse, vars = "Ucp1")

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Ucp1) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
 gather(4, key = "Hallmark", value = "sigScores") 

a <- DimPlot(yu.mouse, 
             reduction = "umap", 
             label = TRUE, 
             pt.size = 0.5,
             repel = TRUE) + NoLegend()

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=umap_1, y=umap_2, color=as.numeric(sigScores)) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(limits = c(-0.3,2.6)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(~Hallmark)

c <- FeaturePlot(yu.mouse, reduction = "umap", feature = "Ucp1") + NoLegend()

a + b + c

```

``` {r}

graph.data <- graph.data %>% 
  left_join(as.data.frame(FetchData(yu.mouse, vars = "Prdm16")) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(FetchData(yu.mouse, vars = "Ebf2")) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") 

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = Prdm16), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Prdm16")

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = Ucp1), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ucp1")

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = Ebf2), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ebf2")

```

### Old dataset only

``` {r}

old=NormalizeData(old)
old=FindVariableFeatures(old)
old=ScaleData(old)
old=RunPCA(old)

# The below is plotting how many PCs are required. 

ElbowPlot(old,ndims = 30)

old = RunUMAP(old,dims = 1:30)
old <- FindNeighbors(old, reduction = "pca", dims = 1:30)
old <- FindClusters(old, resolution = 0.2)
Idents(old) = "RNA_snn_res.0.2"

```

``` {r}

old[["percent.mt"]] <- PercentageFeatureSet(old, pattern = "^mt-")
a <- DimPlot(old)
FeaturePlot(old,"percent.mt")
VlnPlot(old,"percent.mt")

old <- subset(old, subset = percent.mt < 1)

b <- DimPlot(old)

a + b

```

``` {r}

UMAPPlot(old)

adipo.compare <- FindMarkers(old, ident.1 = 0, ident.2 = 1)

View(adipo.compare %>% dplyr::filter(pct.2 > 0.5))
```

``` {r}

feature = "Prdm16"

FeaturePlot(yu.mouse, reduction = "umap.unintegrated", split.by = "orig.ident", feature = feature)

```

``` {r}

VlnPlot(old, 
        features = c("Ucp1", "Cd36", "Fabp4", "Prdm16", "Mgll", "Zfp423"),
        stack = TRUE,
        flip = TRUE)

```

## Zhang 2025 Pansc

``` {r}

bat <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/GSE247719_PanSci_07_BAT_adata (1).h5ad")
gwat <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/GSE247719_PanSci_09_gWAT_adata.h5ad")
iwat <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/GSE247719_PanSci_08_iWAT_adata.h5ad")

```

## Zhang 2025

### BAT

This dataset is from Zhang et al. 2025 and is a single cell RNAseq dataset of mouse BAT. There has been some preparation of the data -

``` {r}

bat = schard::h5ad2seurat('/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/bat_adipo.h5ad')

```

``` {r}

custom_colors <- c("Brown Adipocytes" = "#97600f",  
                   "White Adipocytes" = "#ffe65a")

```

``` {r}

bat[["percent.mt"]] <- PercentageFeatureSet(bat, pattern = "^mt-")

meta <- bat@meta.data %>% 
  mutate(nCount_RNA = ifelse(nCount_RNA > 800, 1, 0),
         nCount_RNA_high = ifelse(nCount_RNA < 3000, 1, 0),
         percent_mt_cutoff = ifelse(percent.mt < 5, 1, 0),
         both_cutoffs = ifelse(nCount_RNA == 1 & percent_mt_cutoff == 1 & Genotype == "WT", 1, 0))

print("nCount_RNA")
table(meta$nCount_RNA)
print("nCount_RNA_high")
table(meta$nCount_RNA_high)
print("percent_mt_cutoff")
table(meta$percent_mt_cutoff)
print("both_cutoffs")
table(meta$both_cutoffs)

VlnPlot(bat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

``` {r}

bat <- subset(bat, subset = Genotype == "WT" & nCount_RNA > 800 & nCount_RNA < 3000 & percent.mt < 5)

```

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- bat@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- bat@reductions$Xumap_@cell.embeddings

# Read in meta data (Cells x Vars)
meta = bat@meta.data
```

``` {r}

Signatures="/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Output/VISION/Gene Signatures/mouse_heat.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)


# saveRDS(vis, file = "../Data/PublishedSingleCellDatasets/Zhang 2025/BAT_neck_HEAT.rds")
# vis <- readRDS(file = "../Output/18.NeckAdipose_dataset_analysis/VISION_neck_HEAT.rds")

```

``` {r fig.width = 8, fig.height = 4}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "mouse_heat"]
celltype <- bat$Sub_cell_type
cond <- bat$Age_group
Ucp1 <- FetchData(bat, vars = "Ucp1")

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Ucp1) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(cond) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
 gather(4, key = "Hallmark", value = "sigScores") 

a <- DimPlot(bat, 
             reduction = "Xumap_", 
             label = TRUE, 
             pt.size = 0.5,
             repel = TRUE) + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank()) +
    NoLegend()

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=Xumap_1, y=Xumap_2, color=as.numeric(sigScores)) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(limits = c(-0.3,3.5)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(~Hallmark)

c <- FeaturePlot(bat, reduction = "Xumap_", feature = "Ucp1") + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
        ) +
  NoLegend()

a + b + c

```

``` {r}

graph.data <- graph.data %>% 
  left_join(as.data.frame(FetchData(bat, vars = "Prdm16")) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(FetchData(bat, vars = "Ebf2")) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") 

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = Prdm16), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Prdm16")

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = log10(Ucp1)), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ucp1")

ggplot(graph.data) +
  geom_point(aes(x = sigScores, y = Ebf2), size = 1) + 
  theme_pubr() +
  ggtitle("HEAT Score vs Ebf2")

```

``` {r}

ggplot(graph.data, 
         aes(x = cond,
           y = sigScores)) +
  geom_jitter() +
  geom_violin() + 
  theme_pubr()

```


### iWAT

``` {r}

iwat = schard::h5ad2seurat('/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/iwat_adipo.h5ad')

```

``` {r}

iwat[["percent.mt"]] <- PercentageFeatureSet(iwat, pattern = "^mt-")

meta <- iwat@meta.data %>% 
  mutate(nCount_RNA = ifelse(nCount_RNA > 800, 1, 0),
         nCount_RNA_high = ifelse(nCount_RNA < 3000, 1, 0),
         percent_mt_cutoff = ifelse(percent.mt < 5, 1, 0),
         both_cutoffs = ifelse(nCount_RNA == 1 & percent_mt_cutoff == 1 & Genotype == "WT", 1, 0))

print("nCount_RNA")
table(meta$nCount_RNA)
print("nCount_RNA_high")
table(meta$nCount_RNA_high)
print("percent_mt_cutoff")
table(meta$percent_mt_cutoff)
print("both_cutoffs")
table(meta$both_cutoffs)

VlnPlot(iwat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

``` {r}

iwat <- subset(iwat, subset = Genotype == "WT" & nCount_RNA > 800 & nCount_RNA < 3000 & percent.mt < 5)

```

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- iwat@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- iwat@reductions$Xumap_@cell.embeddings

# Read in meta data (Cells x Vars)
meta = iwat@meta.data
```

``` {r}

Signatures="/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Output/VISION/Gene Signatures/mouse_heat.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)


# saveRDS(vis, file = "../Data/PublishedSingleCellDatasets/Zhang 2025/iwat_neck_HEAT.rds")
# vis <- readRDS(file = "../Output/18.NeckAdipose_dataset_analysis/VISION_neck_HEAT.rds")

```

``` {r fig.width = 8, fig.height = 4}

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "mouse_heat"]
celltype <- iwat$Sub_cell_type
cond <- iwat$Age_group
Ucp1 <- FetchData(iwat, vars = "Ucp1")

graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(celltype) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Ucp1) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(cond) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
 gather(4, key = "Hallmark", value = "sigScores") 

a <- DimPlot(iwat, 
             reduction = "Xumap_", 
             label = TRUE, 
             pt.size = 0.5,
             repel = TRUE) + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank()) +
    NoLegend()

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=Xumap_1, y=Xumap_2, color=as.numeric(sigScores)) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(limits = c(-0.3,3.5)) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  ) +
  facet_grid(~Hallmark)

c <- FeaturePlot(iwat, reduction = "Xumap_", feature = "Ucp1") + 
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold")
        ) +
  NoLegend()

a + b + c

```



### gWAT

``` {r}

gWAT = schard::h5ad2seurat('/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/gwat_adipo.h5ad')

```

``` {r}

gWAT[["percent.mt"]] <- PercentageFeatureSet(gWAT, pattern = "^mt-")

meta <- gWAT@meta.data %>% 
  mutate(nCount_RNA = ifelse(nCount_RNA > 800, 1, 0),
         nCount_RNA_high = ifelse(nCount_RNA < 3000, 1, 0),
         percent_mt_cutoff = ifelse(percent.mt < 5, 1, 0),
         both_cutoffs = ifelse(nCount_RNA == 1 & percent_mt_cutoff == 1 & Genotype == "WT", 1, 0))

print("nCount_RNA")
table(meta$nCount_RNA)
print("nCount_RNA_high")
table(meta$nCount_RNA_high)
print("percent_mt_cutoff")
table(meta$percent_mt_cutoff)
print("both_cutoffs")
table(meta$both_cutoffs)

VlnPlot(gWAT, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

```

``` {r}

gWAT <- subset(gWAT, subset = Genotype == "WT" & nCount_RNA > 800 & nCount_RNA < 3000 & percent.mt < 5)

```

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- gWAT@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- gWAT@reductions$Xumap_@cell.embeddings

# Read in meta data (Cells x Vars)
meta = gWAT@meta.data
```

``` {r}

Signatures="/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Output/VISION/Gene Signatures/mouse_heat.gmt"

vis <- Vision(scaled_counts, 
               signatures = Signatures, 
               meta = meta, pool = FALSE)
vis <- addProjection(vis3, "UMAP", projection)
vis <- analyze(vis3)


# saveRDS(vis, file = "../Data/PublishedSingleCellDatasets/Zhang 2025/gWAT_neck_HEAT.rds")
# vis <- readRDS(file = "../Output/18.NeckAdipose_dataset_analysis/VISION_neck_HEAT.rds")

```

### Aggregated Analysis

``` {r}

gWAT_vis <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/gWAT_neck_HEAT.rds")
iWAT_vis <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/iWAT_neck_HEAT.rds")
BAT_vis <- readRDS(file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data/PublishedSingleCellDatasets/Zhang 2025/BAT_neck_HEAT.rds")

```

``` {r, fig.height = 10, fig.width = 6}

tissue_types <- c("BAT", "iWAT", "gWAT")
seurat_objects <- list(BAT_vis, iWAT_vis, gWAT_vis) 
names(seurat_objects) <- tissue_types

## Create a more re-usable function
process_adipose_data <- function(vision_object, tissue_type_name) {
  
  umap_projection <- getProjections(vision_object)[["UMAP"]]
  signature_score <- getSignatureScores(vision_object)[, "mouse_heat"]
  celltype <- vision_object@metaData %>% dplyr::select("Age_group", "Sex","ID")
  
  combined_data <- cbind(as.data.frame(umap_projection),
                           HEAT = signature_score,
                           celltype) %>% 
    mutate(type = tissue_type_name, 
           Age_group = factor(Age_group, 
                              levels = c("03_months", 
                                         "06_months", 
                                         "12_months", 
                                         "16_months", 
                                         "23_months"))) %>%
    rownames_to_column(var = "cells")
}

all_data <- purrr::imap(seurat_objects, ~process_adipose_data(.x, .y)) %>%  #Apply function
  bind_rows() %>% 
  mutate(adipocyte_type = ifelse(type == "BAT", "Brown Adipocytes", "White Adipocytes")) 

ggplot(all_data, 
         aes(x = Age_group,
           y = HEAT, 
           color = adipocyte_type)) +
  geom_jitter() +
  geom_boxplot() +
  scale_color_manual(values = c("#97600f","#ffe65a","#ffe65a")) + 
  theme_pubr() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
  ) +
  facet_grid(Sex~type)

```

``` {r}

by_sample_data <- all_data %>% 
  group_by(ID, type) %>% 
  summarise(mean_HEAT = mean(HEAT),
            sd_HEAT = sd(HEAT),
            n = n()) %>% 
  left_join(all_data %>% dplyr::select(ID, Age_group, Sex, type) %>% distinct(),
            by = c("ID", "type")) %>% 
  ungroup()

ggplot(by_sample_data, 
         aes(x = Age_group,
           y = mean_HEAT, 
           color = type)) +
  geom_jitter() +
  geom_boxplot() +
  scale_color_manual(values = c("#97600f","#ffe65a","#ffe65a")) + 
  theme_pubr() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
  ) +
  facet_grid(Sex~type)



```
