---
title: "Sun_dataset_analysis"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}

pacman::p_load(SingleCellExperiment,
               metap,
               tradeSeq,
               Seurat,
               tidyverse, 
               Matrix,
               scales,
               cowplot,
               viridis,
               VISION)

```


```{r }

setwd("/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject/Data")

load("../Data/PublishedSingleCellDatasets/Sun 2020/seurat_object.rdata")

```

``` {r}

UMAPPlot(integrated, label = TRUE)
FeaturePlot(integrated, features = "PLIN1")
FeaturePlot(integrated, features = "EBF2")
FeaturePlot(integrated, features = "ADIPOQ")
FeaturePlot(integrated, features = "PDGFRA")
FeaturePlot(integrated, features = "CKMT2")
FeaturePlot(integrated, features = "PPARGC1A")
FeaturePlot(integrated, features = "percent.mt", max.cutoff = 10)

features = c("ADIPOQ","PDGFRA", "DCN", "MYOCD", "ADGRF5", "JAM2", "STEAP4", "PKHD1L1", "CDH19", "MPZ",  "CARD11", "CD163", "CPA3", "CSF3R", "CNTNAP2", "MSLN", "PRLR", "NRXN1", "SHANK3", "PROX1", "IL7R", "MS4A1", "CDH19", "PTH", "COL1A1")

VlnPlot(integrated, 
        features = features, 
        stack = TRUE, 
        flip = TRUE, 
        group.by = "seurat_clusters") + 
  NoLegend() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 1.5), "cm"))

```

``` {r}

a = c("GHR", "ACACB", "SLC19A3", "PCDH9", "DMD")
b = c("ABCA10", "NEGR1", "LAMA2", "ABCA9", "DCN")
c = c("FAM155A", "LDB2", "PTPRB", "TMTC1", "VWF")
d = c("PKHD1L1", "FLRT2", "NRG3", "PDE1A", "PPFIBP1")
e = c("STAT4", "TPT1", "RPS12", "RPL12", "RPL32")
f = c("KIAA1217", "PRKG1", "DLG2", "FRMD4B", "KCNMA1")
g = c("FRMD4B", "KCNMA1", "CD163L1", "LGMN", "RBM47")
h = c("LSAMP", "LPAR1", "TMEM178B", "CALD1", "ARI3BP")
i = c("BCL11A", "MEF2C", "ADAM28")
j = c("MYH11", "KCNAB1", "RGS6", "ITGA8", "MYL9")
k = c("GPC5", "LAMP3", "WNT5B", "FLT3")

sets = c(a, b, c, d, e, f)

for (set in sets) {
  VlnPlot(integrated, 
          features = g, 
          stack = TRUE, 
          flip = TRUE, 
          group.by = "seurat_clusters") + 
    NoLegend() +
    theme(plot.margin = unit(c(0.1, 0.1, 0.1, 1.5), "cm"))
}

```

0 = CD4+ T Cells
1 = CD4+ T Cells
2 = Adipocytes
3 = Preads
4 = B Cells
6 = Dendritic Cells
7 = Macrophages
9 = Endothelial Cells 2
10 = Endothelial Cells 1
11 = Fibroblasts
12 = CD4+ T Cells


``` {r}

integrated@meta.data <- integrated@meta.data %>% 
  mutate(cell.id = ifelse(seurat_clusters %in% c(0, 6), "T Cells",
                          ifelse(seurat_clusters %in% c(1, 8), "Lymphoid Cells",
                                 ifelse(seurat_clusters == 2, "Adipocytes",
                                        ifelse(seurat_clusters == 3, "Adipocyte Progenitor Cells",
                                               ifelse(seurat_clusters == 4, "B Cells",
                                                      ifelse(seurat_clusters == 7, "Neutrophils",
                                                             ifelse(seurat_clusters == 9, "Lymphatic Endothelial Cells",
                                                                    ifelse(seurat_clusters == 10, "Endothelial Cells", "Unknown")))))))))


```

0 - Lymphoid Cells (T Cells)
1 - Lymphoid Cells (T or B Cells)
2 - Adipocytes
3 - Adipocyte Progenitor Cells
4 - Lymphoid Cells (B Cells)
5
6 - Lymphoid Cells (T Cells)
7 - Neutrophils
8 - Lymphoid Cells (T or B Cells)
9 - Lymphatic Endothelial Cells
10 - Endothelial Cells
11
12

The next part is to rescale the data in accordance to what is recommended for the VISION workflow. Here the data is put into a separate object that is normalised called "scaled_counts". The UMAP projections are put into an object called "projection" and the associated metadata is put into "meta".

```{r normalise-data, echo=FALSE}

# Read in expression counts (Genes X Cells)
counts <- integrated@assays$RNA@counts

# Scale counts within a sample
n.umi <- colSums(counts)
scaled_counts <- t(t(counts) / n.umi) * median(n.umi)

# Adding UMAP
projection <- integrated@reductions$umap@cell.embeddings

# Read in meta data (Cells x Vars)
meta = integrated@meta.data
```

The next part is to create the Vision object and to analyse that object.

```{r}

Signatures="../Output/VISION/Gene Signatures/HEAT_BATLAS_Hallmark.gmt"

vis <- Vision(scaled_counts, signatures = Signatures, meta = meta)
vis <- addProjection(vis, "UMAP", projection)
vis <- analyze(vis)

# saveRDS(vis, file = "/Users/seankodani/Library/CloudStorage/GoogleDrive-seankodani@yahoo.com/My Drive/NeckAdiposeProject//Output/Sun_Dataset_analysis/VISION_Sun_HEAT.rds")
# vis <- readRDS(file = "../Output/17.Efthymiou_dataset_analysis/VISION_efthymiou_HEAT.rds")

```


```{r }

UMAP <- getProjections(vis)[["UMAP"]]
HEAT <- getSignatureScores(vis)[, "UCP1_CORR_SUE_GAVRILA_TSENG"]
BATLAS <- getSignatureScores(vis)[, "BATLAS"]
Adipo <- getSignatureScores(vis)[, "HALLMARK_ADIPOGENESIS"]
FA <- getSignatureScores(vis)[, "HALLMARK_FATTY_ACID_METABOLISM"]
OxPhos <- getSignatureScores(vis)[, "HALLMARK_OXIDATIVE_PHOSPHORYLATION"]


graph.data <- as.data.frame(UMAP) %>% 
  rownames_to_column(var = "cells") %>% 
  left_join(as.data.frame(HEAT) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(BATLAS) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(Adipo) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(FA) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
  left_join(as.data.frame(OxPhos) %>% 
              rownames_to_column(var = "cells"),
            by = "cells") %>% 
 gather(4:8, key = "Hallmark", value = "sigScores")

```

```{r fig.height = 6, fig.width = 12}

b <- ggplot(graph.data %>% 
         dplyr::filter(Hallmark == "HEAT")) + 
  aes(x=umap_1, y=umap_2, color=sigScores) + 
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_gradient(low = "#FFF5F0", high = "#CB181D") +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "Sun et al 2020",
    color = "Signature Scores"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) 

c <- DimPlot(integrated, reduction = "umap", group.by = "cell.id", label = TRUE, repel = TRUE) +
  labs(
    x = "UMAP Dimension 1", 
    y = "UMAP Dimension 2", 
    title = "Sun et al 2020",
    color = "Signature Scores"
  ) + 
  theme_minimal() + 
  theme(
    legend.position = "right",
    legend.title = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  NoLegend()

c + b

```
```
