---
title: "CIBERSORTx sampling of snRNA-seq reference"
output: html_document
date: "2026-04-22"
---

# Introduction

This notebook generates 10 downsampled versions of the snRNA-seq neck
adipose reference for repeated CIBERSORTx deconvolution.

In each iteration, a random sampling fraction between 20% and 25% is
drawn, and cells are sampled proportionally within each cell type. The
resulting sampled Seurat object and sampled reference matrix are saved in
a separate `sample_N` folder.

## Important notes

- This notebook prepares the **reference inputs** for CIBERSORTx.
- After these `sample_N` folders are created, the sampled reference files
  are uploaded manually to CIBERSORTx together with the bulk RNA-seq
  matrices.
- The resulting CIBERSORTx outputs are later downloaded and stored back
  into the corresponding dataset/sample folders for downstream
  aggregation.


```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
library(Seurat)
library(viridis)
library(DGEobj.utils)
library(MASS)
library(Matrix)
library(data.table)
library(here)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------
SEURAT_INPUT <- here(
  "data",
  "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds"
)

OUTPUT_DATA_DIR <- here(
  "output",
  "4.downstream_pipelines",
  "cibersort",
  "data",
  "samples"
)

dir.create(OUTPUT_DATA_DIR, recursive = TRUE, showWarnings = FALSE)

# -------------------------------------------------------------------------
# Analysis settings
# -------------------------------------------------------------------------
CELL_TYPE_COLUMN <- "cell_type"
N_SAMPLES <- 10
MIN_PERCENT <- 20
MAX_PERCENT <- 25
BATCH_SIZE <- 10000
```

# Load inputs

```{r}
s.object <- readRDS(SEURAT_INPUT)
```

# Helper function: check cell type distribution

This helper is retained from the original script for quick inspection of
cell type proportions in the Seurat object.

```{r}
check_cell_type_distribution <- function(seurat_object,
                                         cell_type_column = "cell_type") {
  metadata <- seurat_object@meta.data

  cell_type_summary <- metadata %>%
    group_by(!!sym(cell_type_column)) %>%
    summarise(
      count = n(),
      fraction = count / nrow(metadata),
      .groups = "drop"
    ) %>%
    arrange(desc(count))

  print(cell_type_summary)

  ggplot(
    cell_type_summary,
    aes(x = !!sym(cell_type_column), y = fraction, fill = !!sym(cell_type_column))
  ) +
    geom_bar(stat = "identity") +
    labs(
      title = "Cell Type Distribution",
      x = "Cell Type",
      y = "Fraction of Total Cells"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    guides(fill = "none")
}
```

# Helper function: write large data frames in batches

This helper is used to write the sampled reference matrix as a tab-delimited
text file without holding the full export in memory at once.

```{r}
write_in_batches <- function(data, file_path, batch_size = 10000) {
  total_rows <- nrow(data)

  for (i in seq(1, total_rows, by = batch_size)) {
    end_row <- min(i + batch_size - 1, total_rows)
    batch <- data[i:end_row, ]
    batch_dt <- as.data.table(batch)

    fwrite(
      batch_dt,
      file = file_path,
      sep = "\t",
      col.names = (i == 1),
      append = (i != 1)
    )

    message(paste("Written rows", i, "to", end_row))
  }
}
```

# Generate sampled references for CIBERSORTx

```{r}
for (i in seq_len(N_SAMPLES)) {

  # -----------------------------------------------------------------------
  # Sampling fraction for this sample_N folder
  # -----------------------------------------------------------------------

  x <- sample(MIN_PERCENT:MAX_PERCENT, 1)
  fraction <- x / 100

  cat("--------------------------------------------------------------------------------\n")
  cat("Iteration:", i, "- Sampling fraction:", fraction, "\n")
  cat("--------------------------------------------------------------------------------\n")

  # -----------------------------------------------------------------------
  # Copy the original Seurat object so the full input is not overwritten
  # -----------------------------------------------------------------------
  s.object.sample <- s.object
  metadata <- s.object.sample@meta.data
  metadata$cell_names <- Cells(s.object.sample)

  # -----------------------------------------------------------------------
  # Stratified sampling within each cell type
  # -----------------------------------------------------------------------
  sampled_cells <- metadata %>%
    group_by(!!sym(CELL_TYPE_COLUMN)) %>%
    sample_frac(fraction) %>%
    pull(cell_names)

  s.object.sample <- subset(s.object.sample, cells = sampled_cells)

  # -----------------------------------------------------------------------
  # Create the sampled reference expression matrix
  # -----------------------------------------------------------------------
  s.object.deep <- s.object.sample

  cts.deep <- GetAssayData(object = s.object.deep, layer = "counts")
  meta.deep <- s.object.deep@meta.data

  # Preserve the original collaborator logic that reassigns endothelial cells
  endo_pos <- grepl("Endothelial", meta.deep$cell_type_group)
  meta.deep$cell_type_group[endo_pos] <- meta.deep$cell_type_res.1.6[endo_pos]
  s.object.deep@meta.data <- meta.deep

  final_mtx.deep <- cts.deep
  colnames(final_mtx.deep) <- meta.deep$cell_type_group

  final_mtx.deep_dense <- as.matrix(final_mtx.deep)
  final_mtx.deep_df <- as.data.frame(final_mtx.deep_dense)
  final_mtx.deep_df <- cbind(
    "Gene name" = rownames(final_mtx.deep_df),
    final_mtx.deep_df
  )

  # -----------------------------------------------------------------------
  # Create a subfolder for this sampled reference
  # -----------------------------------------------------------------------
  sample_dir <- file.path(OUTPUT_DATA_DIR, paste0("sample_", i))
  if (!dir.exists(sample_dir)) {
    dir.create(sample_dir, recursive = TRUE)
  }

  # -----------------------------------------------------------------------
  # Save sampled reference matrix for manual CIBERSORTx upload
  # -----------------------------------------------------------------------
  output_file_ref <- file.path(sample_dir, "deep_cervical_ref_snRNAseq_sampled.txt")

  if (file.exists(output_file_ref)) {
    file.remove(output_file_ref)
  }

  write_in_batches(
    final_mtx.deep_df,
    output_file_ref,
    batch_size = BATCH_SIZE
  )

  # -----------------------------------------------------------------------
  # Save sampled Seurat object for provenance and reproducibility
  # -----------------------------------------------------------------------
  output_file_rds <- file.path(sample_dir, "sampled_seurat_object.rds")
  saveRDS(s.object.deep, output_file_rds)

  cat("Saved sample", i, "with sampling fraction =", fraction, "\n")
}

cat("All sampled references have been processed and saved.\n")
```
