---
title: "Process CIBERSORTx deconvolution results"
author: "Henrique"
date: "2025-09-10"
output: html_document
---

# Introduction

This notebook processes CIBERSORTx estimated cellular proportions from
multiple repeated runs, averages the bootstrap estimates for each sample
and cell type, adds simple sample-level metadata, and exports the final
deconvolution table.

The retained final output is:

- `deconvolution_cellular_proportions_table.csv`

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggpubr)
library(stringr)
library(here)

# -------------------------------------------------------------------------
# Define inputs and outputs
# -------------------------------------------------------------------------

# Primary directory containing CIBERSORTx result folders
INPUT_DIR <- here("output", "4.downstream_pipelines", "cibersort", "data", "bootstrap_results")

OUTPUT_DATA_DIR <- here("output", "4.downstream_pipelines", "cibersort", "data")

dir.create(OUTPUT_DATA_DIR, showWarnings = FALSE, recursive = TRUE)

OUTPUT_TABLE <- file.path(
  OUTPUT_DATA_DIR,
  "deconvolution_cellular_proportions_table.csv"
)
```

# Load CIBERSORTx results

```{r}
datasets <- list.files(
  INPUT_DIR,
  pattern = ".*Job\\d+_Results.*\\.csv$",
  recursive = TRUE,
  full.names = TRUE
)

combined_data <- lapply(datasets, function(file) {
  data <- read.csv(file)
  data$File <- file
  data$Sample <- basename(dirname(file))

  pivot_longer(
    data,
    cols = 2:23,
    names_to = "Cell_type",
    values_to = "Proportions"
  )
}) %>%
  bind_rows() %>%
  mutate(
    Proportions = Proportions * 100,
    File = str_replace_all(File, "Resultsz", "Results"),
    Dataset = str_extract(File, "(?<=Results_)[^_]+"),
    Dataset = str_to_sentence(Dataset)
  )

combined_data %>% distinct(File, Dataset)
combined_data %>% filter(is.na(Dataset))
```

# Average repeated runs and add sample metadata

```{r}
combined_data_with_meta <- combined_data %>%
  group_by(Mixture, Cell_type, Dataset) %>%
  summarise(
    Proportions = mean(Proportions),
    .groups = "drop"
  ) %>%
  mutate(
    Depot = case_when(
      str_detect(Mixture, "Deep|BAT") ~ "Deep",
      str_detect(Mixture, "WAT|SC") ~ "Superficial",
      str_detect(Mixture, "^C") ~ "Control",
      str_detect(Mixture, "^F") ~ "Pheo",
      TRUE ~ Mixture
    ),
    Donor_ID = case_when(
      str_detect(Mixture, "\\d") ~ str_extract(Mixture, "\\d+"),
      TRUE ~ str_remove(Mixture, "[BW]AT")
    )
  )
```

# Save final deconvolution table

```{r}
write.csv(combined_data_with_meta, OUTPUT_TABLE, row.names = FALSE)
```
