---
title: "HEAT Signature Generation"
output: html_document
date: "2023-10-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(openxlsx)
library(Seurat)
library(tidyverse)
library(here)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------

# Bulk RNA-seq correlation source
INPUT_SALEJ_LOGCPM <- here(
"data", "6.internal_datasets", "Gavrila Lab Bulk RNAseq Data", "logCPM_for_correlations.csv"
)

# Microarray correlation source
INPUT_LYNES_MICROARRAY <- here(
"data/6.internal_datasets/Tseng_Datasets/Ruidan Xue's Microarray data/20 arrays for hBAT hWAT tissues/Supplementary_Table_cervical_at_microarray.xlsx"
)

# Cleaned Seurat object from the preprocessing pipeline
INPUT_SEURAT_RDS <- here(
  "data",
  "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds"
)

OUTPUT_DATA_DIR <- here("output", "3.heat_signature", "data")
OUTPUT_FIG_DIR <- here("output", "3.heat_signature", "figures")

dir.create(OUTPUT_DATA_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(OUTPUT_FIG_DIR, showWarnings = FALSE, recursive = TRUE)

# -------------------------------------------------------------------------
# Check required inputs
# -------------------------------------------------------------------------
required_inputs <- c(
  INPUT_SALEJ_LOGCPM,
  INPUT_LYNES_MICROARRAY,
  INPUT_SEURAT_RDS
)

missing_inputs <- required_inputs[!file.exists(required_inputs)]

if (length(missing_inputs) > 0) {
  stop(
    "Missing input file(s):\n",
    paste(missing_inputs, collapse = "\n")
  )
}

# -------------------------------------------------------------------------
# Helper functions
# -------------------------------------------------------------------------

# Write a gene set collection in GMT format
write_gmt <- function(gene_sets, file) {
  lines <- lapply(names(gene_sets), function(set_name) {
    paste(c(set_name, "NA", gene_sets[[set_name]]), collapse = "\t")
  })
  writeLines(unlist(lines), con = file)
}

# Compute Pearson correlation and p-value against a reference vector
cor_with_reference <- function(gene_expr, ref_expr) {
  cor_result <- cor.test(gene_expr, ref_expr, method = "pearson")
  c(cor = cor_result$estimate, pval = cor_result$p.value)
}
```

# Introduction

This notebook generates the HEAT signature by intersecting three sources of
evidence:

1. Genes positively correlated with UCP1 in deep neck bulk RNA-seq data
   from Salej et al.
2. Genes positively correlated with UCP1 in deep neck microarray data from
   Lynes et al.
3. Genes enriched in brown adipocytes in the snRNA-seq neck adipose object.

The final outputs are saved as tables and gene set files for downstream use.

# Salej et al. bulk RNA-seq UCP1 correlation

```{r}
# Load TPM/logCPM data
log2_cpm <- read_csv(INPUT_SALEJ_LOGCPM, show_col_types = FALSE)

# Remove rows without gene symbols
log2_cpm <- log2_cpm %>%
  filter(!is.na(gene))

# Prepare deep neck expression matrix
log2_cpm_clean <- log2_cpm %>%
  mutate(gene = make.unique(gene)) %>%
  column_to_rownames("X") %>%
  select(contains("Deep"))

cpm_deep_t <- as.data.frame(t(log2_cpm_clean))
cpm_deep_t <- data.frame(lapply(cpm_deep_t, function(x) as.numeric(as.character(x))))

# Identify the Ensembl ID corresponding to UCP1
ucp1_ensembl <- log2_cpm %>%
  filter(gene == "UCP1") %>%
  pull("X")

ucp1_idx <- which(colnames(cpm_deep_t) == ucp1_ensembl)
ucp1_exp <- cpm_deep_t[, ucp1_idx]

# Correlate all genes against UCP1
gene_names <- colnames(cpm_deep_t)

cor_results <- t(sapply(
  gene_names,
  function(gene_id) cor_with_reference(cpm_deep_t[[gene_id]], ucp1_exp)
))

salej_cor_df <- data.frame(
  Ensembl = gene_names,
  estimate = cor_results[, "cor.cor"],
  PValue = cor_results[, "pval"]
) %>%
  left_join(
    log2_cpm %>% select(X, gene),
    by = c("Ensembl" = "X")
  ) %>%
  rename(Gene = gene) %>%
  mutate(FDR = p.adjust(PValue, method = "fdr")) %>%
  arrange(FDR)

write_csv(
  salej_cor_df,
  file.path(OUTPUT_DATA_DIR, "Salej_et_al_logCPM_correlation.csv")
)
```

# Lynes et al. microarray UCP1 correlation

```{r}
# Load microarray data
lynes_cts <- read.xlsx(
  INPUT_LYNES_MICROARRAY,
  sheet = "All Data by Individual"
)

# Standardize column names
names(lynes_cts) <- gsub("[ \\.]", "_", names(lynes_cts))

# Keep deep neck (CS) samples and create unique probe-level identifiers
data_lynes <- lynes_cts %>%
  dplyr::select(grep("CS", names(lynes_cts), value = TRUE), Gene_Symbol, Probe_Set_ID) %>%
  mutate(Gene_ID = paste0(Gene_Symbol, "_(", Probe_Set_ID, ")")) %>%
  column_to_rownames("Gene_ID") %>%
  select(-c(Gene_Symbol, Probe_Set_ID))

# Locate UCP1 probe(s)
reference_gene <- grep("UCP1", rownames(data_lynes))
data_lynes <- as.matrix(data_lynes)
ref_values <- data_lynes[reference_gene, ]

# Remove non-informative rows
data_lynes <- data_lynes[str_detect(rownames(data_lynes), "^NA_", negate = TRUE), ]
data_lynes <- data_lynes[str_detect(rownames(data_lynes), "---", negate = TRUE), ]

# Correlate every probe against UCP1
lynes_results <- sapply(seq_len(nrow(data_lynes)), function(i) {
  cor.test(data_lynes[i, ], ref_values, method = "pearson")
})

lynes_results <- t(lynes_results)
lynes_results <- as.data.frame(lynes_results)

lynes_cor_df <- data.frame(
  Gene = rownames(data_lynes),
  p_value = sapply(lynes_results$p.value, function(x) x[1]),
  estimate = sapply(lynes_results$estimate, function(x) x[1])
) %>%
  mutate(FDR = p.adjust(p_value, method = "fdr"))

# Expand multi-gene probe annotations into one row per gene
expanded_lynes_cor_df <- lynes_cor_df %>%
  mutate(Gene_Probe = Gene) %>%
  separate_rows(Gene, sep = " /// ") %>%
  mutate(Gene = str_remove(Gene, "_\\(.*"))

write_csv(
  expanded_lynes_cor_df,
  file.path(OUTPUT_DATA_DIR, "Lynes_et_al_correlation.csv")
)
```

# Filter positively correlated genes in Salej and Lynes

```{r}
# Correlation filters used to define UCP1-associated genes
estimate_th <- 0
fdr_th <- 0.1
pval_th <- 1

salej_ucp1_pos_corr <- salej_cor_df %>%
  filter(
    estimate >= estimate_th,
    FDR <= fdr_th,
    PValue <= pval_th
  ) %>%
  distinct(Gene) %>%
  filter(Gene != "UCP1") %>%
  pull(Gene)

lynes_ucp1_pos_corr <- expanded_lynes_cor_df %>%
  filter(
    estimate >= estimate_th,
    FDR <= fdr_th,
    p_value <= pval_th
  ) %>%
  distinct(Gene) %>%
  filter(Gene != "UCP1") %>%
  pull(Gene)
```

# Brown adipocyte-enriched genes from snRNA-seq

```{r}
s.object <- readRDS(INPUT_SEURAT_RDS)

bad_mrk <- FindMarkers(
  s.object,
  ident.1 = "Brown Adipocytes",
  only.pos = TRUE,
  test.use = "wilcox"
)

filtered_bad_mrk <- bad_mrk %>%
  rownames_to_column("Gene") %>%
  filter(
    avg_log2FC > 1,
    pct.1 > 0.3,
    p_val_adj <= 0.05
  ) %>%
  pull(Gene)
```

# Generate the HEAT signature

```{r}
HEAT_signature <- intersect(
  filtered_bad_mrk,
  intersect(salej_ucp1_pos_corr, lynes_ucp1_pos_corr)
)

HEAT_signature

write.csv(
  HEAT_signature,
  file = file.path(OUTPUT_DATA_DIR, "heat_signature.csv"),
  row.names = FALSE
)

save(
  HEAT_signature,
  file = file.path(OUTPUT_DATA_DIR, "heat_signature.Rdata")
)

write_gmt(
  list("HEAT" = HEAT_signature),
  file.path(OUTPUT_DATA_DIR, "heat_signature.gmt")
)
```

# Save supporting gene lists

```{r}
genes_for_venn <- list(
  "Salej et al\nBulk RNAseq" = salej_ucp1_pos_corr,
  "Lynes et al\nMicroarray" = lynes_ucp1_pos_corr,
  "Brown Adipocyte Enriched" = filtered_bad_mrk
)

save(
  genes_for_venn,
  file = file.path(OUTPUT_DATA_DIR, "genes_for_venn.Rdata")
)
```

# Generate related UCP1-correlated signatures

```{r}
ucp1_corr <- intersect(salej_ucp1_pos_corr, lynes_ucp1_pos_corr)
ucp1_corr <- setdiff(ucp1_corr, "UCP1")

non_heat_ucp1_corr <- setdiff(ucp1_corr, HEAT_signature)
non_heat_ucp1_corr <- setdiff(non_heat_ucp1_corr, "UCP1")

write_gmt(
  list(
    "UCP1 correlated" = ucp1_corr,
    "Non_HEAT UCP1 correlated" = non_heat_ucp1_corr
  ),
  file.path(OUTPUT_DATA_DIR, "non_HEAT_UCP1_corr_signature.gmt")
)
```

# Save summary

```{r}
signature_summary <- tibble::tibble(
  metric = c(
    "n_salej_deep_samples",
    "n_salej_ucp1_correlated_genes",
    "n_lynes_deep_samples",
    "n_lynes_ucp1_correlated_genes",
    "n_brown_adipocyte_nuclei",
    "n_total_nuclei",
    "n_brown_adipocyte_enriched_genes",
    "n_heat_genes",
    "n_ucp1_correlated_genes_intersection",
    "n_non_heat_ucp1_correlated_genes"
  ),
  value = c(
    nrow(cpm_deep_t),
    length(salej_ucp1_pos_corr),
    ncol(data_lynes),
    length(lynes_ucp1_pos_corr),
    as.numeric(table(Idents(s.object))["Brown Adipocytes"]),
    ncol(s.object),
    length(filtered_bad_mrk),
    length(HEAT_signature),
    length(ucp1_corr),
    length(non_heat_ucp1_corr)
  )
)

readr::write_csv(
  signature_summary,
  file.path(OUTPUT_DATA_DIR, "heat_signature_generation_summary.csv")
)

write_csv(
  bad_mrk %>% rownames_to_column("Gene"),
  file.path(OUTPUT_DATA_DIR, "brown_adipocyte_FindMarkers_all_genes.csv")
)

write_csv(
  bad_mrk %>%
    rownames_to_column("Gene") %>%
    filter(avg_log2FC > 1, pct.1 > 0.3, p_val_adj <= 0.05),
  file.path(OUTPUT_DATA_DIR, "brown_adipocyte_enriched_genes_filtered.csv")
)
```


# Save as a single Excel
```{r}
library(openxlsx)

# Create workbook
wb <- createWorkbook()

# 1. Summary
addWorksheet(wb, "Summary")
writeData(wb, "Summary", signature_summary)

# 2. Salej correlation
addWorksheet(wb, "Salej corr")
writeData(wb, "Salej corr", salej_cor_df)

# 3. Lynes correlation
addWorksheet(wb, "Lynes corr")
writeData(wb, "Lynes corr", expanded_lynes_cor_df)

# 4. Brown adipocyte FindMarkers
addWorksheet(wb, "BAds FindMarkers")
writeData(wb, "BAds FindMarkers", bad_mrk %>% rownames_to_column("Gene"))

# 4. Brown adipocyte FindMarkers
addWorksheet(wb, "BAds FindMarkers Filt")
writeData(wb, "BAds FindMarkers Filt",   bad_mrk %>%
    rownames_to_column("Gene") %>%
    filter(avg_log2FC > 1, pct.1 > 0.3, p_val_adj <= 0.05))

# Save Excel file
saveWorkbook(
  wb,
  file = file.path(OUTPUT_DATA_DIR, "Supplemental_Table_6_heat_signature_generation_summary.xlsx"),
  overwrite = TRUE
)
```

# Session Info
```{r}
writeLines(
  capture.output(sessionInfo()),
  file.path(OUTPUT_DATA_DIR, "sessionInfo_heat_signature_generation.txt")
)
```

