---
title: "36.ReferenceMappingSubtypes"
output: html_document
date: "2025-05-22"
---

# Purpose

This notebook uses reference mapping to help label adipocyte and adipocyte
progenitor subtypes in the neck adipose dataset by projecting query cells onto
a published reference dataset from Emont et al.

Two approaches are explored:

1. Symphony reference mapping
2. Seurat reference mapping

The main biological goal is to use transferred subtype labels from the
published reference to better interpret the adipocyte and ASPC subclusters in
the study dataset.

# Reference notes from Emont et al.

For context, the Emont human adipocyte / ASPC reference includes several
subclusters of interest:

- hAd1, hAd3, hAd4, and hAd7 were primarily in SAT
- hAd2 and hAd6 were primarily in VAT
- hAd5 was more evenly distributed
- hAd3 was associated with larger adipocytes and triglyceride synthesis
- hAd4 and hAd7 were negatively associated with BMI
- hAd6 expressed thermogenesis-related genes such as EBF2, PPARGC1A, and ESRRG
- hASPC1 appeared closer to the more differentiated arm of ASPCs
- hASPC2 resembled early multipotent progenitors
- hASPC3 appeared committed but not yet fully differentiated
- hASPC4 resembled anti-adipogenic AREG-like cells

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(Seurat)
library(symphony)
library(tidyverse)
library(here)

# -------------------------------------------------------------------------
# Define inputs and outputs
# -------------------------------------------------------------------------
INPUT_EMONT_RDS <- HUMAN_SEURAT_DIR <- here("output", "4.downstream_pipelines", "vision", "data", "cleaned_datasets_human", "Emont_Dataset.rds") 

INPUT_QUERY_RDS <- here(
  "data",
  "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds"
)

OUTPUT_DATA_DIR <- here("output", "4.downstream_pipelines", "reference_mapping", "data")
OUTPUT_FIG_DIR  <- here("output", "4.downstream_pipelines", "reference_mapping", "figures")

dir.create(OUTPUT_DATA_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUTPUT_FIG_DIR, recursive = TRUE, showWarnings = FALSE)

# -------------------------------------------------------------------------
# Helper loaders
# -------------------------------------------------------------------------
load_first_existing_rds <- function(primary_path, alternative_paths = NULL, object_label = "RDS file") {
  if (file.exists(primary_path)) {
    message("Loading ", object_label, " from primary path.")
    return(readRDS(primary_path))
  }

  for (path in alternative_paths) {
    if (file.exists(path)) {
      message("Loading ", object_label, " from fallback path: ", path)
      return(readRDS(path))
    }
  }

  stop("Could not find ", object_label, ".")
}
```

# Load the reference and query datasets

```{r}
# Load Emont reference dataset
emont <- load_first_existing_rds(
  primary_path = INPUT_EMONT_RDS,
  alternative_paths = INPUT_EMONT_ALT,
  object_label = "Emont reference dataset"
)

# Split the reference into ASPC and adipocyte compartments
Idents(emont) <- "cell_type"
ref.aspc  <- subset(emont, idents = c("ASPC"))
ref.adipo <- subset(emont, idents = c("adipocyte"))
rm(emont)

# Load the main neck adipose dataset
s.object <- load_first_existing_rds(
  primary_path = INPUT_QUERY_RDS,
  alternative_paths = INPUT_QUERY_ALT,
  object_label = "main neck adipose dataset"
)

# Split the query into adipocyte and ASPC compartments
Idents(s.object) <- "cell_type"
adipocytes <- subset(s.object, idents = c("White Adipocytes"))
aspc       <- subset(s.object, idents = c("Adipocyte Progenitor Cells"))
rm(s.object)
```

# Reference mapping with Symphony

This first mapping attempt uses Symphony, which extends Harmony-based
integration for reference mapping.

## Build the adipocyte reference object

```{r}
reference <- symphony::buildReference(
  ref.adipo@assays$RNA["counts"],   # reference expression matrix, genes x cells
  ref.adipo@meta.data,              # reference metadata, cells x attributes
  vars = c("sample"),               # batch variable to harmonize over
  K = 100,                          # number of Harmony soft clusters
  verbose = TRUE,
  do_umap = TRUE,                   # build a reference UMAP model
  do_normalize = TRUE,              # log(CP10k) normalization of reference
  vargenes_method = "vst",
  vargenes_groups = "cell_type",    # group-aware variable gene selection
  topn = 2000,
  theta = 2,
  d = 20,
  additional_genes = NULL,
  save_uwot_path = file.path(OUTPUT_DATA_DIR, "./testing_uwot_model_2")
)

umap_labels <- cbind(reference[["umap"]][["embedding"]], reference[["meta_data"]])

ggplot(umap_labels, aes(x = UMAP1, y = UMAP2, color = cell_type)) +
  geom_point()
```

## Map query adipocytes onto the adipocyte reference

```{r}
query <- mapQuery(
  adipocytes@assays$RNA["counts"],  # query expression matrix, genes x cells
  adipocytes@meta.data,             # query metadata
  reference,                        # Symphony reference object
  do_normalize = FALSE,             # query already normalized upstream
  do_umap = TRUE                    # project query cells into reference UMAP
)
```

## Transfer labels and store the projected UMAP in the query Seurat object

```{r}
query <- knnPredict(query, reference, reference$meta_data$cell_type, k = 5)

adipocytes$cell_type_pred_knn <- query$meta_data$cell_type_pred_knn
adipocytes$cell_type_pred_knn_prob <- query$meta_data$cell_type_pred_knn_prob

query.reduction <- CreateDimReducObject(
  embeddings = query$umap,
  key = "ref.umap",
  assay = DefaultAssay(adipocytes)
)

adipocytes[["ref_umap"]] <- query.reduction
```

## Visualize transferred adipocyte subtype labels

```{r}
DimPlot(
  adipocytes,
  reduction = "ref_umap",
  group.by = "cell_type_pred_knn",
  label = TRUE,
  label.size = 6,
  repel = TRUE
)

FeaturePlot(
  adipocytes,
  features = "ADIPOQ",
  reduction = "ref_umap",
  label = TRUE,
  label.size = 3,
  repel = TRUE
) + NoLegend()

VlnPlot(
  adipocytes,
  features = "PLIN4",
  group.by = "cell_type_pred_knn"
)

query_labels <- cbind(query$meta_data, query$umap) %>%
  mutate(dataset = ifelse(cell_type == "White Adipocytes", "query", "reference"))

ggplot(query_labels, aes(x = UMAP1, y = UMAP2, color = cell_type_pred_knn)) +
  geom_point() +
  facet_wrap(~neck_region)
```

## Compare transferred subtype proportions across samples and neck regions

```{r}
table(query$meta_data$sample)

neck_region <- distinct(query$meta_data[c("sample", "neck_region")])

temp <- prop.table(table(query$meta_data$cell_type_pred_knn, query$meta_data$sample), margin = 2) %>%
  as.data.frame() %>%
  left_join(neck_region, by = c("Var2" = "sample"))

ggplot(temp, aes(x = Var1, y = Freq, color = neck_region)) +
  geom_boxplot(
    position = position_dodge(width = 0.75),
    width = 0.6,
    fill = "transparent",
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(x = Var1, y = Freq, color = neck_region),
    position = position_dodge(width = 0.75),
    alpha = 0.7,
    size = 1.5
  ) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73")) +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  labs(
    color = "Neck Region",
    x = "ASPC Subtype",
    y = "Proportion"
  )

compare_means(
  Freq ~ neck_region,
  data = temp %>% filter(neck_region != "Intermediate"),
  group.by = "Var1"
)
```

## Stacked barplot of transferred subtype composition by neck region

```{r fig.width = 4, fig.height = 6}
stack_barplot <- temp %>%
  group_by(neck_region, Var1) %>%
  summarise(Freq = sum(Freq), .groups = "drop") %>%
  ungroup()

p1 <- ggplot(stack_barplot, aes(x = neck_region, y = Freq, fill = Var1)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Neck Region",
    y = "Proportion within Region",
    title = "Distribution of Adipocyte Subtypes",
    fill = "Subtypes"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(size = 12, angle = 45, vjust = 1.25, hjust = 1.25),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank()
  )

print(p1)
```

## Donut plot representation of transferred subtype composition

```{r}
donut_data <- temp %>%
  group_by(neck_region, Var1) %>%
  summarise(Freq_sum = sum(Freq), .groups = "drop_last") %>%
  mutate(proportion = Freq_sum / sum(Freq_sum)) %>%
  ungroup() %>%
  group_by(neck_region) %>%
  mutate(
    cumulative_proportion = cumsum(proportion),
    mid_point_proportion = cumulative_proportion - 0.5 * proportion
  ) %>%
  ungroup() %>%
  mutate(
    ring_pos = case_when(
      neck_region == "Deep" ~ 1,
      neck_region == "Intermediate" ~ 2,
      neck_region == "Superficial" ~ 3
    ),
    ring_width = 0.9
  )

p2 <- ggplot(donut_data, aes(x = ring_pos, y = proportion, fill = Var1)) +
  geom_col(width = donut_data$ring_width, color = "white") +
  coord_polar(theta = "y") +
  labs(
    title = "Distribution of Adipocyte Subtypes",
    fill = "Subtypes"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "right"
  )

print(p2)
```

# Reference mapping with Seurat

This section keeps the second reference mapping strategy using Seurat.
It follows the same general idea as the Symphony mapping above, but using
Seurat’s anchor-based transfer framework.

## Prepare reference and query objects

```{r}
# Ensure the reference and query objects have normalized data
ref.adipo <- NormalizeData(ref.adipo)
ref.adipo <- FindVariableFeatures(ref.adipo)

adipocytes <- NormalizeData(adipocytes)
adipocytes <- FindVariableFeatures(adipocytes)
```

## Find transfer anchors and map adipocyte query cells

```{r}
anchors.adipo <- FindTransferAnchors(
  reference = ref.adipo,
  query = adipocytes,
  dims = 1:30
)

predictions.adipo <- TransferData(
  anchorset = anchors.adipo,
  refdata = ref.adipo$cell_type,
  dims = 1:30
)

adipocytes <- AddMetaData(adipocytes, metadata = predictions.adipo)
```

## Visualize Seurat transfer results

```{r}
DimPlot(adipocytes, group.by = "predicted.id", label = TRUE, repel = TRUE)

VlnPlot(
  adipocytes,
  features = "prediction.score.max",
  group.by = "predicted.id"
)
```

# ASPC mapping

This section mirrors the adipocyte mapping workflow for adipocyte progenitor
cells.

## Build the ASPC reference object with Symphony

```{r}
reference.aspc <- symphony::buildReference(
  ref.aspc@assays$RNA["counts"],
  ref.aspc@meta.data,
  vars = c("sample"),
  K = 100,
  verbose = TRUE,
  do_umap = TRUE,
  do_normalize = TRUE,
  vargenes_method = "vst",
  vargenes_groups = "cell_type",
  topn = 2000,
  theta = 2,
  d = 20,
  additional_genes = NULL,
  save_uwot_path = file.path(OUTPUT_DATA_DIR ,"./testing_uwot_model_aspc")
)
```

## Map ASPC query cells to the ASPC reference

```{r}
query.aspc <- mapQuery(
  aspc@assays$RNA["counts"],
  aspc@meta.data,
  reference.aspc,
  do_normalize = FALSE,
  do_umap = TRUE
)

query.aspc <- knnPredict(query.aspc, reference.aspc, reference.aspc$meta_data$cell_type, k = 5)

aspc$cell_type_pred_knn <- query.aspc$meta_data$cell_type_pred_knn
aspc$cell_type_pred_knn_prob <- query.aspc$meta_data$cell_type_pred_knn_prob

query.aspc.reduction <- CreateDimReducObject(
  embeddings = query.aspc$umap,
  key = "ref.umap",
  assay = DefaultAssay(aspc)
)

aspc[["ref_umap"]] <- query.aspc.reduction
```

## Visualize ASPC mapping results

```{r}
DimPlot(
  aspc,
  reduction = "ref_umap",
  group.by = "cell_type_pred_knn",
  label = TRUE,
  repel = TRUE
)
```
