---
title: "CellChat analysis of human neck adipose tissue"
output: html_document
date: "2024-12-02"
---

# Introduction

This notebook runs CellChat separately in Deep and Superficial neck
adipose depots, then merges the resulting CellChat objects for
downstream comparative analysis.

The workflow is:

1. Load the cleaned Seurat object.
2. Run CellChat independently for Deep and Superficial depots.
3. Save the depot-specific CellChat objects.
4. Merge the two CellChat objects and save the merged result.

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(CellChat)
library(Seurat)
library(tidyverse)
library(here)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------
INPUT_RDS <- here(
  "data",
  "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds"
)

OUTPUT_DATA_DIR <- here("output", "4.downstream_pipelines", "cellchat", "data")
dir.create(OUTPUT_DATA_DIR, showWarnings = FALSE, recursive = TRUE)

# Which population.size settings to run
POP_SIZE_VALUES <- c(FALSE, TRUE)

# Helper to convert TRUE/FALSE to filename-safe text
pop_size_label <- function(x) {
  paste0("pop.size.", tolower(as.character(x)))
}

# Define all output files at the beginning
OUTPUT_PATHS <- lapply(POP_SIZE_VALUES, function(pop_size) {
  label <- pop_size_label(pop_size)

  list(
    pop_size = pop_size,
    deep = file.path(
      OUTPUT_DATA_DIR,
      paste0("cellchat_deep_", label, "_cell_type_short.rds")
    ),
    superficial = file.path(
      OUTPUT_DATA_DIR,
      paste0("cellchat_superficial_", label, "_cell_type_short.rds")
    ),
    merged = file.path(
      OUTPUT_DATA_DIR,
      paste0("cellchat_merged_", label, "_cell_type_short.rds")
    )
  )
})

names(OUTPUT_PATHS) <- sapply(POP_SIZE_VALUES, pop_size_label)
```

# Helper function

```{r}
run_cellchat <- function(seurat_obj,
                         subset_ident = "neck_region",
                         idents,
                         group.by = "cell_type",
                         cellchat_db = CellChatDB.human,
                         min.cells = 10,
                         population.size = FALSE) {
  # Subset the Seurat object to the requested compartment
  Idents(seurat_obj) <- subset_ident
  seurat_obj <- subset(seurat_obj, idents = idents)

  # Initialize CellChat object
  cellchat <- createCellChat(object = seurat_obj, group.by = group.by)
  cellchat@DB <- cellchat_db

  # Standard CellChat preprocessing and inference workflow
  cellchat <- subsetData(cellchat)
  cellchat <- identifyOverExpressedGenes(cellchat)
  cellchat <- identifyOverExpressedInteractions(cellchat)
  cellchat <- computeCommunProb(cellchat, population.size = population.size)
  cellchat <- filterCommunication(cellchat, min.cells = min.cells)
  cellchat <- computeCommunProbPathway(cellchat)
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

  cellchat
}
```

# Load Seurat object

```{r}
seurat_obj <- readRDS(INPUT_RDS)

cellchat_sample_summary <- seurat_obj@meta.data %>%
  filter(neck_region %in% c("Deep", "Superficial")) %>%
  count(neck_region, sample, cell_type_short, name = "n_nuclei") %>%
  arrange(neck_region, sample, cell_type_short)

readr::write_csv(
  cellchat_sample_summary,
  file.path(OUTPUT_DATA_DIR, "cellchat_input_nuclei_by_sample_depot_celltype.csv")
)
```

# Run CellChat by depot

```{r}
# Create an empty list to store all CellChat results in memory.
# Each entry will correspond to one population.size setting:
#   - "pop.size.false"
#   - "pop.size.true"
cellchat_results <- list()

# Loop over each output configuration defined earlier in OUTPUT_PATHS.
for (nm in names(OUTPUT_PATHS)) {

  # Extract the configuration for the current iteration.
  cfg <- OUTPUT_PATHS[[nm]]

  # -----------------------------------------------------------------------
  # Run CellChat on the Deep neck_region subset
  # -----------------------------------------------------------------------
  deep_cellchat <- run_cellchat(
    seurat_obj = seurat_obj,
    subset_ident = "neck_region",
    idents = "Deep",
    group.by = "cell_type_short",
    population.size = cfg$pop_size
  )

  # -----------------------------------------------------------------------
  # Run CellChat on the Superficial neck_region subset
  # -----------------------------------------------------------------------
  # This repeats the same workflow as above, but for the Superficial depot.
  superficial_cellchat <- run_cellchat(
    seurat_obj = seurat_obj,
    subset_ident = "neck_region",
    idents = "Superficial",
    group.by = "cell_type_short",
    population.size = cfg$pop_size
  )

  # -----------------------------------------------------------------------
  # Save the depot-specific CellChat objects to disk
  # -----------------------------------------------------------------------
  saveRDS(deep_cellchat, cfg$deep)
  saveRDS(superficial_cellchat, cfg$superficial)

  # -----------------------------------------------------------------------
  # Combine the two depot-specific CellChat objects into a named list
  # -----------------------------------------------------------------------
  object_list <- list(
    Deep = deep_cellchat,
    Superficial = superficial_cellchat
  )

  # -----------------------------------------------------------------------
  # Merge Deep and Superficial CellChat objects for comparative analysis
  # -----------------------------------------------------------------------
  cellchat_merged <- mergeCellChat(
    object.list = object_list,
    add.names = names(object_list)
  )

  # Save the merged CellChat object to disk
  saveRDS(cellchat_merged, cfg$merged)

  # -----------------------------------------------------------------------
  # Store all results in memory as well
  # -----------------------------------------------------------------------
  cellchat_results[[nm]] <- list(
    deep = deep_cellchat,
    superficial = superficial_cellchat,
    merged = cellchat_merged
  )
}
```

## Session Info
```{r}
writeLines(
  capture.output(sessionInfo()),
  file.path(OUTPUT_DATA_DIR, "sessionInfo_cellchat.txt")
)
```

