---
title: "Descriptive_analysis_Shiyi_object"
author: "Henrique Camara"
date: "2023-11-14"
output: html_document
editor_options:
  markdown:
    wrap: 72
---

```{r}
# Purpose: Annotate cell types, add subject metadata, perform initial QC,
# and save a cleaned Seurat object for downstream analysis.
# Run after: Seurat object has been created from the AnnData conversion step.
# Inputs: Seurat RDS and subject metadata table.
# Outputs: cleaned Seurat object and QC/annotation plots.
# Must not change: clustering parameters, annotation assignments, filtering thresholds,
# and saved output object content.
# Cleanup level: annotation and readability only, with no intended effect on outputs.
```

```{r setup, include=FALSE}
getwd()
here::here()

# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(glmGamPoi)
library(Seurat)
library(DESeq2)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(openxlsx)
library(here)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------
INPUT_RDS <- here("data", "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18.rds")

INPUT_SUBJECT_INFO <- here("data", "2023-09-14_SubjectInformationTable.xlsx")

# Save output RDS in the same folder as the input object
sans_ext_rds <- tools::file_path_sans_ext(INPUT_RDS)
OUTPUT_RDS <- paste0(sans_ext_rds, "_cleaned.rds")

OUTPUT_FIG_DIR <- here("output", "1.processing", "figures")
OUTPUT_DATA_DIR <- here("output", "1.processing", "data")

dir.create(OUTPUT_FIG_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUTPUT_DATA_DIR, recursive = TRUE, showWarnings = FALSE)
```

# INTRODUCTION

The objective of this analysis is to compare the new objects generated by
the Berkeley team and determine QC metrics. This notebook combines
several utility steps used during object review, annotation, and cleaning.

## Setting the environment

```{r, echo=TRUE, message=FALSE}
# -------------------------------------------------------------------------
# Helper functions
# -------------------------------------------------------------------------

# Create directories with status messages
mkdir <- function(out_path) {
  if (dir.exists(out_path)) {
    print("Directory already exists.")
  } else {
    dir.create(out_path, recursive = TRUE)
    if (dir.exists(out_path)) {
      print("Directory created successfully.")
    } else {
      print("Failed to create the directory.")
    }
  }
}

# Reorder clusters in stacked VlnPlots based on average expression
VlnPlotOrdered <- function(s.object, features, identity = "leiden_scVI") {
  Idents(s.object) <- identity

  p <- VlnPlot(
    s.object,
    features = features,
    stack = TRUE,
    flip = TRUE
  ) +
    NoLegend() +
    theme(plot.margin = unit(c(0.1, 0.1, 0.1, 1.5), "cm"))

  df <- p$data

  # Average expression per gene per cluster
  df <- df %>%
    dplyr::group_by(feature, ident) %>%
    summarise(avg_exp = mean(expression), .groups = "drop") %>%
    arrange(desc(avg_exp))

  cluster_order <- df %>%
    dplyr::group_by(ident) %>%
    arrange(desc(avg_exp)) %>%
    dplyr::slice(1)

  # Initialize one container per gene
  feats <- list()
  for (gene in features) {
    feats[[gene]] <- character()
  }

  # Iteratively assign each cluster to the gene for which it is most enriched
  while (nrow(df) > 0) {
    gene <- df$feature[1]
    cluster <- as.character(df$ident[1])

    feats[[gene]] <- c(feats[[gene]], cluster)
    df <- df %>% filter(ident != cluster)
  }

  # Reassemble cluster order
  cluster_order <- vector()
  for (ord in feats) {
    cluster_order <- c(cluster_order, ord)
  }

  p + scale_x_discrete(limits = cluster_order)
}
```

## Workflow {.tabset}

### Downloading the object

```{r}
s.object <- readRDS(INPUT_RDS)

names(s.object@assays)
DefaultAssay(s.object) <- "log_10k_norm"

Idents(s.object) <- "leiden_scVI"
UMAPPlot(s.object)
```

### Adding metadata

```{r, echo=TRUE, message=FALSE}
meta <- s.object@meta.data

# Remove GEX suffix from sample names
meta$clean_sample <- gsub("_GEX", "", meta$sample)

# Extract subject and biopsy identifiers from sample names
meta$subject <- sub("^(.*)_(.*)$", "\\1", meta$clean_sample)
meta$biopsy <- sub("^(.*)_(.*)$", "\\2", meta$clean_sample)

head(meta$subject)
head(meta$biopsy)
unique(meta$biopsy)

# Biopsy groups used to define anatomical neck region
deep <- c("PV", "BAT", "CS", "DC")
inter <- c("SP")
sup <- c("SQ", "WAT", "SC", "PS", "PP")

meta$neck_region <- NA
meta$neck_region[meta$biopsy %in% deep] <- "Deep"
meta$neck_region[meta$biopsy %in% sup] <- "Superficial"
meta$neck_region[meta$biopsy %in% inter] <- "Intermediate"

# Match subject IDs to the clinical metadata table
meta <- meta %>%
  mutate(SubjectID = toupper(subject))

meta_df <- read.xlsx(INPUT_SUBJECT_INFO, sheet = 3)
meta_df <- meta_df[1:18, ]

all(meta$SubjectID %in% meta_df$SubjectID)

# Preserve cell barcodes before join, then restore rownames afterward
meta$cells <- rownames(meta)
meta <- meta %>%
  left_join(meta_df, by = "SubjectID", multiple = "all")
rownames(meta) <- meta$cells

# Add obesity categories
meta$obesity <- NA
meta$obesity[meta$BMI < 25] <- "<25"
meta$obesity[meta$BMI >= 25] <- ">=25"
meta$obesity[meta$BMI >= 30] <- ">=30"

s.object@meta.data <- meta
```

### Increasing cluster resolution in the current Seurat object

```{r}
DefaultAssay(s.object)

s.object <- FindNeighbors(
  object = s.object,
  dims = 1:30,
  reduction = "scVI",
  assay = "log_10k_norm"
)

# Using 30 dimensions improved clustering relative to earlier tests with 10
resolutions <- c(0.8, 1.6)

for (res in resolutions) {
  s.object <- FindClusters(
    object = s.object,
    resolution = res,
    graph.name = "RNA_snn"
  )
}

plots <- list()
for (res in resolutions) {
  plots[[as.character(res)]] <- UMAPPlot(
    s.object,
    label = TRUE,
    repel = TRUE,
    group.by = paste0("RNA_snn_res.", res)
  ) + NoLegend()
}

plots
```

### Assigning clusters

#### Plotting our set of markers

```{r}
Idents(s.object) <- "RNA_snn_res.1.6"

features <- c(
  "ADIPOQ", "PPARGC1A", "PPARG", "PDGFRA", "DCN", "MYOCD", "PKD2",
  "ADGRF5", "JAM2", "STEAP4", "PKHD1L1", "CDH19", "MPZ",
  "CARD11",
  "BANK1", "MS4A1", # CD20
  "CD2","CD6",
  "CD163", "CPA3", "CSF3R", "CNTNAP2", "CASR",
  "DNAH11", "PRM1", "PTH"
)

p <- VlnPlot(
  s.object,
  features = features,
  stack = TRUE,
  flip = TRUE
) +
  NoLegend() +
  theme(plot.margin = unit(c(0.1, 0.1, 0.1, 1.5), "cm"))

head(p$data)
summary(p$data %>% filter(feature == "ADIPOQ"))

p
```
```{r}
lymp_mrk_top <- lymp_mrk |> rownames_to_column("gene") |> mutate(direction = ifelse(avg_log2FC > 0, "Up", "Down")) |> group_by(direction) |> arrange(p_val_adj) |> slice_head(n = 10)
lymp_mrk_top


classic_b_markers <- c(
  "MS4A1",   # CD20
  "BANK1")

classic_t_markers <- c(
  "CD2",
  "CD6"
)

lymp_mrk |> rownames_to_column("gene") |> filter(gene %in% c(classic_b_markers, classic_t_markers))
```

#### Clustering - cell_type_res.1.6

```{r}
meta <- s.object@meta.data
meta$cell_type_res.1.6 <- NA

# Unknown clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "10"] <- "Unknown_1"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "13"] <- "Unknown_2"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "14"] <- "Unknown_3"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "15"] <- "Unknown_4"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "27"] <- "Unknown_5"

# ADIPOQ-related clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "0"] <- "White_adipocyte_1"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "5"] <- "White_adipocyte_2"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "11"] <- "PreAdipocyte"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "23"] <- "Macrophage_ADIPOQ_pos"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "36"] <- "Smooth_muscle_ADIPOQ_pos"

# PPARGC1A-related cluster
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "26"] <- "Brown_adipocyte"

# DCN-related clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "1"] <- "ASPC_1"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "4"] <- "ASPC_2"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "16"] <- "Mesenchymal"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "18"] <- "ASPC_3"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "24"] <- "ASPC_4"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "31"] <- "ASPC_5"

# MYOCD-related clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "6"] <- "Smooth_muscle_1"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "28"] <- "Smooth_muscle_2"

# Endothelial subtypes
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "7"] <- "Endothelial_venous"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "8"] <- "Endothelial_capillary"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "9"] <- "Endothelial_arterial"

# Pericytes
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "20"] <- "Pericyte"

# Lymphatic endothelial subtypes
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "17"] <- "Lymphatic_endothelial_1"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "19"] <- "Lymphatic_endothelial_2"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "29"] <- "Lymphatic_endothelial_3"

# Schwann cells
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "32"] <- "Schwann_non_myelinating"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "30"] <- "Schwann_myelinating"

# Lymphocytes
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "12"] <- "T_Lymphocyte"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "33"] <- "B_Lymphocyte"

# Macrophages
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "3"] <- "Macrophage"

# Mast cells
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "25"] <- "Mast"

# Neutrophils
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "21"] <- "Neutrophil"

# Neuron-associated clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "2"] <- "Neuron_associated_1"

# Merged neuron/parathyroid-associated clusters
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "22"] <- "Neuron_associated_2"
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "34"] <- "Neuron_associated_2"

# PRM1-positive cluster
meta$cell_type_res.1.6[meta$RNA_snn_res.1.6 == "35"] <- "PRM1_pos"

s.object@meta.data <- meta

UMAPPlot(s.object, label = TRUE, group.by = "RNA_snn_res.1.6") + NoLegend()
UMAPPlot(s.object, label = TRUE, group.by = "cell_type_res.1.6") + NoLegend()

ggsave(
  file.path(OUTPUT_FIG_DIR, "UMAPPlot_cell_type_res_1.6.png"),
  width = 12,
  height = 10,
  dpi = 300
)
```

#### Clustering - Cell type group

```{r}
meta <- s.object@meta.data

meta$cell_type_group[meta$RNA_snn_res.1.6 == "0"] <- "White_adipocyte"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "5"] <- "White_adipocyte"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "11"] <- "PreAdipocyte"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "23"] <- "Macrophage_ADIPOQ_pos"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "36"] <- "Smooth_muscle_ADIPOQ_pos"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "26"] <- "Brown_adipocyte"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "1"] <- "ASPC"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "4"] <- "ASPC"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "16"] <- "Mesenchymal"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "18"] <- "ASPC"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "24"] <- "ASPC"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "31"] <- "ASPC"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "6"] <- "Smooth_muscle"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "28"] <- "Smooth_muscle"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "7"] <- "Endothelial_venous"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "8"] <- "Endothelial_capillary"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "9"] <- "Endothelial_arterial"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "20"] <- "Pericyte"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "17"] <- "Lymphatic_endothelial"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "19"] <- "Lymphatic_endothelial"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "29"] <- "Lymphatic_endothelial"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "32"] <- "Schwann"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "30"] <- "Schwann"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "12"] <- "T_Lymphocyte"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "33"] <- "B_Lymphocyte"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "3"] <- "Macrophage"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "25"] <- "Mast"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "21"] <- "Neutrophil"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "2"] <- "Neuron_associated_1"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "22"] <- "Neuron_associated_2"
meta$cell_type_group[meta$RNA_snn_res.1.6 == "34"] <- "Neuron_associated_2"

meta$cell_type_group[meta$RNA_snn_res.1.6 == "35"] <- "PRM1_pos"

s.object@meta.data <- meta

UMAPPlot(s.object, label = TRUE, group.by = "RNA_snn_res.1.6") + NoLegend()
UMAPPlot(s.object, label = TRUE, group.by = "cell_type_group") + NoLegend()

ggsave(
  file.path(OUTPUT_FIG_DIR, "UMAPPlot_cell_type_group.png"),
  width = 12,
  height = 10,
  dpi = 300
)
```

#### Clustering - Cell Type SDK

```{r, fig.width=10, fig.height=8}
metadata <- s.object@meta.data
metadata$cell_type_sdk <- NA

metadata$cell_type_sdk[which(metadata$cell_type_group == "Brown_adipocyte")] <- "Brown Adipocytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "White_adipocyte")] <- "White Adipocytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Macrophage_ADIPOQ_pos")] <- "Lipid Associated Macrophages"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Smooth_muscle_ADIPOQ_pos")] <- "ADIPOQ+ Smooth Muscle Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "PreAdipocyte")] <- "Pre-adipocytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "ASPC")] <- "Adipocyte Progenitor Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Mesenchymal")] <- "Mesenchymal Stem Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Endothelial_arterial")] <- "Arterial Endothelial Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Endothelial_venous")] <- "Venous Endothelial Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Endothelial_capillary")] <- "Capillary Endothelial Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Lymphatic_endothelial")] <- "Lymphatic Endothelial Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Pericyte")] <- "Pericytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Smooth_muscle")] <- "Smooth Muscle Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Macrophage")] <- "Macrophages"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Mast")] <- "Mast Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "T_Lymphocyte")] <- "T Lymphocytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "B_Lymphocyte")] <- "B Lymphocytes"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Neutrophil")] <- "Neutrophils"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Schwann")] <- "Schwann Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Neuron_associated_1")] <- "Neuron Associated Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "Neuron_associated_2")] <- "Parathyroid Associated Cells"
metadata$cell_type_sdk[which(metadata$cell_type_group == "PRM1_pos")] <- "PRM1+ Cells"

s.object@meta.data <- metadata

DimPlot(
  s.object,
  reduction = "umap",
  group.by = "cell_type_sdk",
  label = TRUE,
  label.size = 5,
  repel = TRUE
) +
  labs(x = "UMAP 1", y = "UMAP 2") +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.title = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(size = 1, color = "black")
  ) +
  NoLegend()

unique(s.object$cell_type_sdk)

ggsave(
  file.path(OUTPUT_FIG_DIR, "UMAPPlot_cell_type_sdk.png"),
  width = 12,
  height = 10,
  dpi = 300
)
```

# Cell type short

This will be the reference annotation used for downstream analysis.

```{r}
cell_type_mapping <- c(
  "Brown Adipocytes" = "BAds",
  "White Adipocytes" = "WAds",
  "Lipid Associated Macrophages" = "PreAds",
  "ADIPOQ+ Smooth Muscle Cells" = "PreAds",
  "Pre-adipocytes" = "PreAds",
  "Adipocyte Progenitor Cells" = "ASPC",
  "Mesenchymal Stem Cells" = "MSCs",
  "Arterial Endothelial Cells" = "EndoCs",
  "Venous Endothelial Cells" = "EndoCs",
  "Capillary Endothelial Cells" = "EndoCs",
  "Lymphatic Endothelial Cells" = "LECs",
  "Pericytes" = "Pericytes",
  "Smooth Muscle Cells" = "SMCs",
  "Macrophages" = "Macro",
  "Mast Cells" = "MastCs",
  "T Lymphocytes" = "T Cells",
  "B Lymphocytes" = "B Cells",
  "Neutrophils" = "Neutrophils",
  "Schwann Cells" = "SchwannCs",
  "Neuron Associated Cells" = "NeuroCs",
  "Parathyroid Associated Cells" = "ParaThyCs",
  "PRM1+ Cells" = "PRM1+"
)

s.object@meta.data <- s.object@meta.data %>%
  mutate(
    cell_type_short = factor(
      cell_type_mapping[cell_type_sdk],
      levels = unique(cell_type_mapping)
    )
  )
```

### Filtering Low Quality Cells

```{r}
# Start metadata of # nuclei per sample
nuclei_per_sample <- s.object@meta.data |> count(sample, name = "initial_nuclei")
nuclei_per_sample
```



#### Check UMI count distribution

```{r}
names(s.object@meta.data)

VlnPlot(
  s.object,
  features = "nCount_log_10k_norm",
  group.by = "cell_type_res.1.6",
  fill.by = "ident",
  cols = palette.use,
  pt.size = 1,
  alpha = 0.1
) +
  theme(axis.ticks.x = element_line()) +
  NoLegend() +
  geom_hline(
    yintercept = c(500, 750, 1000, 4000),
    linetype = "dashed",
    color = c("red", "orange", "grey", "purple")
  )

ggsave(
  file.path(OUTPUT_FIG_DIR, "VlnPlot_UMI_count.png"),
  device = "png",
  height = 8,
  width = 16
)
```

#### Remove cells with low UMI counts (<500)

```{r}
s.object <- subset(x = s.object, subset = nCount_log_10k_norm > 500)

VlnPlot(s.object, features = "nCount_log_10k_norm", group.by = "cell_type_res.1.6") +
  NoLegend() +
  geom_hline(yintercept = 500, linetype = "dashed", color = "red") +
  scale_y_continuous(limits = c(0, 5000))

# Add info
nuclei_per_sample <- s.object@meta.data |> count(sample, name = "umi_500_or_more") |> left_join(nuclei_per_sample, by = "sample")
nuclei_per_sample
```

#### Gene marker identification analysis

```{r}
Idents(s.object) <- "cell_type_res.1.6"
DefaultAssay(s.object) <- "log_10k_norm"

mrk_unk_1 <- FindMarkers(s.object, ident.1 = "Unknown_1", only.pos = TRUE, test.use = "wilcox")
mrk_unk_2 <- FindMarkers(s.object, ident.1 = "Unknown_2", only.pos = TRUE, test.use = "wilcox")
mrk_unk_3 <- FindMarkers(s.object, ident.1 = "Unknown_3", only.pos = TRUE, test.use = "wilcox")
mrk_unk_4 <- FindMarkers(s.object, ident.1 = "Unknown_4", only.pos = TRUE, test.use = "wilcox")
mrk_unk_5 <- FindMarkers(s.object, ident.1 = "Unknown_5", only.pos = TRUE, test.use = "wilcox")

mrk_wad1 <- FindMarkers(s.object, ident.1 = "White_adipocyte_1", only.pos = TRUE, test.use = "wilcox")
mrk_wad2 <- FindMarkers(s.object, ident.1 = "White_adipocyte_2", only.pos = TRUE, test.use = "wilcox")
mrk_bad <- FindMarkers(s.object, ident.1 = "Brown_adipocyte", only.pos = TRUE, test.use = "wilcox")
```

Let`s DotPlot the top markers based on p-value, logFC, and pct_exp.

Starting with adjusted p-value.

```{r, fig.width=250,fig.height=250}
feats <- mrk_unk_1 %>%
  mutate(genes = rownames(mrk_unk_1)) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 3) %>%
  pull(genes)

feats <- c(
  feats,
  mrk_unk_2 %>% mutate(genes = rownames(mrk_unk_2)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)
feats <- c(
  feats,
  mrk_unk_3 %>% mutate(genes = rownames(mrk_unk_3)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)
feats <- c(
  feats,
  mrk_unk_4 %>% mutate(genes = rownames(mrk_unk_4)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)
feats <- c(
  feats,
  mrk_unk_5 %>% mutate(genes = rownames(mrk_unk_5)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)

feats <- unique(feats)

Idents(s.object) <- factor(Idents(s.object), levels = sort(levels(Idents(s.object))))

p <- DotPlot(s.object, features = feats) +
  scale_color_viridis_c(direction = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Unknwon markers genes - adj_P-val")

p

ggsave(
  file.path(OUTPUT_FIG_DIR, "Dot_plot_unknown_pval.png"),
  device = "png",
  plot = p,
  height = 10,
  width = 13
)

feats <- c(
  mrk_wad1 %>% mutate(genes = rownames(mrk_wad1)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)
feats <- c(
  feats,
  mrk_wad2 %>% mutate(genes = rownames(mrk_wad2)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)
feats <- c(
  feats,
  mrk_bad %>% mutate(genes = rownames(mrk_bad)) %>% arrange(p_val_adj) %>% slice_head(n = 3) %>% pull(genes)
)

feats <- unique(feats)

p <- DotPlot(s.object, features = feats) +
  scale_color_viridis_c(direction = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Positive Control Markers - adj_P-val")

p
```

#### Remove Unknowns, as they do not have clear markers

```{r}
s.object <- subset(
  x = s.object,
  idents = unique(grep("Unknown", s.object$cell_type_res.1.6, value = TRUE)),
  invert = TRUE
)

# Add info
nuclei_per_sample <- s.object@meta.data |> count(sample, name = "unknwon_removed") |> left_join(nuclei_per_sample |> select(-"unknwon_removed"), by = "sample")
nuclei_per_sample
```

### Plot the final object

```{r}
UMAPPlot(s.object, group.by = "cell_type_short", cols = palette.use, label = TRUE) +
  NoLegend()

ncol(s.object) # Original analysis resulted in 37596 cells left

meta <- s.object@meta.data
meta <- meta %>%
  mutate(
    cell_type = case_when(
      cell_type_short == "EndoCs" ~ "Endothelial Cells",
      cell_type_short == "PreAds" ~ "Pre-adipocytes",
      TRUE ~ cell_type_sdk
    )
  )

s.object@meta.data <- meta
```
### Find All Markers
```{r}
# Find markers expressed in at least 30% of cells with logfc > 0.5
all_markers <- FindAllMarkers(s.object, logfc.threshold = 0.5, only.pos = TRUE, min.pct = 0.3)
```

```{r}
write_csv(all_markers, file.path(OUTPUT_DATA_DIR, "cell_type_markers.csv"))
```

### Saving the file

```{r}
Idents(s.object) <- "cell_type"

saveRDS(s.object, OUTPUT_RDS)
```

### Saving cell counts
```{r}
cell_counts <- s.object@meta.data %>%
  count(neck_region, cell_type, cell_type_short, name = "n_nuclei")

write.csv(cell_counts, file.path(OUTPUT_DATA_DIR, "final_cell_counts_by_depot_and_cell_type.csv"), row.names = FALSE)
```

### Saving nuclei per sample
```{r}
nuclei_per_sample_final <- rbind(
  nuclei_per_sample,
  c(
    "Total", 
    colSums(nuclei_per_sample[-1])
    )
)

write.csv(nuclei_per_sample_final, file.path(OUTPUT_DATA_DIR, "final_cell_counts_per_sample.csv"), row.names = FALSE)

```

# Save cell_type summary meta
```{r}
cell_type_meta <- s.object@meta.data |> 
  select(clean_sample, RNA_snn_res.0.8, RNA_snn_res.1.6, cell_type_res.1.6, cell_type, cell_type_short) |> 
  rownames_to_column("barcode")
write.csv(cell_type_meta, file.path(OUTPUT_DATA_DIR, "cell_type_metadata.csv"), row.names = FALSE)

```


```{r}
writeLines(capture.output(sessionInfo()), file.path(OUTPUT_DATA_DIR, "sessionInfo_seurat_preprocessing.txt"))
```

