---
title: "GTEx Correlation Analysis"
output: html_document
date: "2025-02-07"
editor_options:
  markdown:
    wrap: 72
---

# Introduction

This notebook tests associations between GTEx adipose HEAT ssGSEA scores
and subject metadata.

It assumes the GTEx ssGSEA tables were generated upstream and focuses on
the downstream metadata association workflow.

## Updated screening strategy

The original version screened pairwise HEAT associations across the full
combined GTEx adipose dataset. Because subcutaneous and visceral adipose tissue
can have dramatically different HEAT score distributions, pooled screening can
mask within-stratum associations.

This updated version performs the initial pairwise screening separately within
each sex-by-tissue stratum, then identifies variables that are significant in
**any** stratum. Those variables can then be carried forward into the adjusted
stratified linear models.

The retained outputs are:

- combined HEAT metadata table
- filtered metadata table used for association testing
- pooled association results across all GTEx samples
- stratified pairwise association results
- summary table of variables significant in any stratum
- summary figure of FDR-ranked pairwise associations by stratum
- stratified adjusted linear models by sex and tissue

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(tidyverse)
library(openxlsx)
library(ggplot2)
library(ggpubr)
library(broom)
library(here)
library(RColorBrewer)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------
INPUT_GTEX_SUB_SSGSEA <- here(
  "output", "3.heat_signature", "5.gtex", "data",
  "gene_tpm_2022-06-06_v10_adipose_subcutaneous_ssGSEA.csv"
)

INPUT_GTEX_VIS_SSGSEA <- here(
  "output", "3.heat_signature", "5.gtex", "data",
  "gene_tpm_2022-06-06_v10_adipose_visceral_omentum_ssGSEA.csv"
)

INPUT_GTEX_SUBJECT_META <- here(
  "local_data", "gtex",
  "GTEx_Analysis_2017-06-05_v8_Annotations_SubjectPhenotypesDS.txt"
)

INPUT_GTEX_META_DICT <- here(
  "local_data", "gtex",
  "GTEx_Analysis_2017-06-05_v8_Annotations_GTEx_Analysis_2017-06-05_v8_Annotations_SubjectPhenotypesDD.xlsx"
)

OUTPUT_DIR <- here("output", "3.heat_signature", "5.gtex")
OUT_DATA_DIR <- here(OUTPUT_DIR, "data")
OUT_FIGURE_DIR <- here(OUTPUT_DIR, "figures")

dir.create(OUT_DATA_DIR, recursive = TRUE, showWarnings = FALSE)
dir.create(OUT_FIGURE_DIR, recursive = TRUE, showWarnings = FALSE)

required_inputs <- c(
  INPUT_GTEX_SUB_SSGSEA,
  INPUT_GTEX_VIS_SSGSEA,
  INPUT_GTEX_SUBJECT_META,
  INPUT_GTEX_META_DICT
)

missing_inputs <- required_inputs[!file.exists(required_inputs)]
if (length(missing_inputs) > 0) {
  stop(
    "Missing input file(s):\n",
    paste(missing_inputs, collapse = "\n")
  )
}

# -------------------------------------------------------------------------
# Analysis settings
# -------------------------------------------------------------------------
PAIRWISE_FDR_CUTOFF <- 0.05
MIN_PAIRWISE_N <- 5
MIN_MODEL_N <- 30
```

# Load ssGSEA results and GTEx metadata

```{r}
sc_data <- read_csv(INPUT_GTEX_SUB_SSGSEA, show_col_types = FALSE)
vis_data <- read_csv(INPUT_GTEX_VIS_SSGSEA, show_col_types = FALSE)

final_data <- bind_rows(sc_data, vis_data) %>%
  mutate(
    tissue = case_when(
      str_detect(tolower(tissue), "sub") ~ "Subcutaneous",
      str_detect(tolower(tissue), "vis") ~ "Visceral",
      TRUE ~ tissue
    )
  )

gtex_subj_meta <- read_delim(
  INPUT_GTEX_SUBJECT_META,
  delim = "\t",
  show_col_types = FALSE
) %>%
  mutate(SUBJID = str_remove(SUBJID, "GTEX-"))

gtex_subj_meta_annot <- read.xlsx(INPUT_GTEX_META_DICT)

heat_meta <- final_data %>%
  select(HEAT, subject, tissue) %>%
  left_join(gtex_subj_meta, by = c("subject" = "SUBJID"))

write_csv(heat_meta, file.path(OUT_DATA_DIR, "HEAT_metadata.csv"))

gtex_sample_summary <- heat_meta %>%
  count(tissue, SEX, name = "n_samples") %>%
  arrange(tissue, SEX)

write_csv(
  gtex_sample_summary,
  file.path(OUT_DATA_DIR, "gtex_sample_summary_by_tissue_sex.csv")
)
```

# Helper function for variable-wise association testing

```{r}
testAssociation <- function(var, data, outcome = "HEAT") {
  x <- data[[var]]
  y <- data[[outcome]]

  complete_cases <- !is.na(x) & !is.na(y)
  df <- data[complete_cases, ]
  x <- df[[var]]
  y <- df[[outcome]]

  n_total <- nrow(data)
  n_used <- nrow(df)
  n_removed <- n_total - n_used

  n_obs <- df %>%
    count(!!sym(var)) %>%
    mutate(label = paste(!!sym(var), "n =", n)) %>%
    pull(label) %>%
    paste0(collapse = ";")

  if (n_used < MIN_PAIRWISE_N) {
    return(tibble(
      variable = var,
      method = "unhandled",
      statistic = NA_real_,
      p_value = NA_real_,
      n_removed = n_removed,
      n_obs = n_obs,
      direction = NA_real_
    ))
  }

  if (is.numeric(x) && length(unique(x)) > 10) {
    res <- cor.test(y, x)
    return(tibble(
      variable = var,
      method = "pearson",
      statistic = unname(res$estimate),
      p_value = res$p.value,
      n_removed = n_removed,
      n_obs = "Numeric",
      direction = unname(res$estimate)
    ))
  }

  if (length(unique(x)) == 2) {
    if (is.factor(x) || is.character(x)) {
      x <- factor(x)
      df[[var]] <- x
    }

    res <- tryCatch(
      t.test(y ~ x, data = df),
      error = function(e) NULL
    )

    if (is.null(res)) {
      return(tibble(
        variable = var,
        method = "t.test (failed)",
        statistic = NA_real_,
        p_value = NA_real_,
        n_removed = n_removed,
        n_obs = n_obs,
        direction = NA_real_
      ))
    }

    return(tibble(
      variable = var,
      method = "t.test",
      statistic = unname(res$statistic),
      p_value = res$p.value,
      n_removed = n_removed,
      n_obs = n_obs,
      direction = NA_real_
    ))
  }

  if (length(unique(x)) > 2 && length(unique(x)) <= 10) {
    fit <- tryCatch(
      lm(y ~ x, data = df),
      error = function(e) NULL
    )

    if (is.null(fit)) {
      return(tibble(
        variable = var,
        method = "anova (failed)",
        statistic = NA_real_,
        p_value = NA_real_,
        n_removed = n_removed,
        n_obs = n_obs,
        direction = NA_real_
      ))
    }

    res <- anova(fit)
    return(tibble(
      variable = var,
      method = "anova",
      statistic = NA_real_,
      p_value = res$`Pr(>F)`[1],
      n_removed = n_removed,
      n_obs = n_obs,
      direction = NA_real_
    ))
  }

  tibble(
    variable = var,
    method = "unhandled",
    statistic = NA_real_,
    p_value = NA_real_,
    n_removed = n_removed,
    n_obs = n_obs,
    direction = NA_real_
  )
}
```

# Clean metadata for association testing

```{r}
multi_category_integer <- gtex_subj_meta_annot %>%
  filter(TYPE == "integer, encoded value") %>%
  filter(str_detect(VALUES, "1") | str_detect(X9, "2") | str_detect(X10, "2")) %>%
  pull(VARNAME)

binary_vars <- gtex_subj_meta_annot %>%
  filter(TYPE == "integer, encoded value", !VARNAME %in% multi_category_integer) %>%
  pull(VARNAME)

meta_filtered <- heat_meta %>%
  mutate(
    across(
      all_of(binary_vars),
      ~ case_when(
        .x > 90 ~ NA_real_,
        TRUE ~ as.numeric(.x)
      )
    )
  ) %>%
  select(where(~ n_distinct(na.omit(.)) >= 2)) %>%
  mutate(
    SEX = case_when(
      SEX == 1 ~ "Male",
      SEX == 2 ~ "Female",
      TRUE ~ "Unknown"
    )
  )

write_csv(meta_filtered, file.path(OUT_DATA_DIR, "HEAT_metadata_filtered.csv"))
```

# Pooled associations across all GTEx metadata

This pooled screen is retained for reference, but the main selection step for
follow-up modeling now comes from the stratified pairwise screen below.

```{r}
assoc_results_pooled <- map_dfr(
  setdiff(names(meta_filtered), c("HEAT", "subject")),
  ~ testAssociation(.x, meta_filtered, outcome = "HEAT")
) %>%
  mutate(FDR = p.adjust(p_value, method = "BH"))

write_csv(
  assoc_results_pooled,
  file.path(OUT_DATA_DIR, "gtex_heat_correlation_pooled.csv")
)
```

# Stratified pairwise HEAT associations by sex and tissue

```{r}
pairwise_vars <- setdiff(names(meta_filtered), c("HEAT", "subject", "SEX", "tissue"))

group_combinations <- meta_filtered %>%
  distinct(SEX, tissue) %>%
  drop_na()

assoc_results_stratified <- map_dfr(seq_len(nrow(group_combinations)), function(i) {
  sex_val <- group_combinations$SEX[i]
  tissue_val <- group_combinations$tissue[i]

  sub_data <- meta_filtered %>%
    filter(SEX == sex_val, tissue == tissue_val)

  map_dfr(pairwise_vars, ~ testAssociation(.x, sub_data, outcome = "HEAT")) %>%
    mutate(
      SEX = sex_val,
      tissue = tissue_val,
      n_stratum = nrow(sub_data)
    )
}) %>%
  group_by(SEX, tissue) %>%
  mutate(FDR = p.adjust(p_value, method = "BH")) %>%
  ungroup()

write_csv(
  assoc_results_stratified,
  file.path(OUT_DATA_DIR, "gtex_heat_correlation_stratified.csv")
)

pairwise_significant_any_stratum <- assoc_results_stratified %>%
  filter(!is.na(FDR), FDR < PAIRWISE_FDR_CUTOFF) %>%
  left_join(
    gtex_subj_meta_annot %>% select(VARDESC, VARNAME),
    by = c("variable" = "VARNAME")
  ) %>%
  mutate(
    VARDESC = case_when(
      is.na(VARDESC) ~ variable,
      TRUE ~ VARDESC
    )
  ) %>%
  arrange(variable, SEX, tissue)

write_csv(
  pairwise_significant_any_stratum,
  file.path(OUT_DATA_DIR, "significant_pairwise_any_stratum.csv")
)

# Also write a multi-sheet Excel workbook
excel_file <- file.path(OUT_DATA_DIR, "stratified_pariwise_assoc_summary.xlsx")

wb <- createWorkbook()

addWorksheet(wb, "gtex_heat_pooled")
writeData(wb, "gtex_heat_pooled", assoc_results_pooled)

addWorksheet(wb, "gtex_heat_strat")
writeData(wb, "gtex_heat_strat", assoc_results_stratified)

addWorksheet(wb, "pairwise_significant")
writeData(wb, "pairwise_significant", pairwise_significant_any_stratum)



saveWorkbook(wb, excel_file, overwrite = TRUE)
```

# Stratified adjusted linear models by sex and tissue

Variables are carried forward if they are significant in **any** pairwise
screening stratum.

```{r}
# Select variables related to Medical History to check for independent associations of HEAT-sginature score and metabolic status
medical_history_vars <- gtex_subj_meta_annot |> filter(DOCFILE =="General Medical History") |> pull(VARNAME)

vars_to_plot <- pairwise_significant_any_stratum %>%
  distinct(variable) %>%
  filter(variable %in% medical_history_vars) |> 
  pull(variable)

stratified_model_res <- list()
stratified_models_list <- list()
failed_model_fits <- list()

for (var in vars_to_plot) {
  for (i in seq_len(nrow(group_combinations))) {
    sex_val <- group_combinations$SEX[i]
    tissue_val <- group_combinations$tissue[i]

    sub_data <- meta_filtered %>%
      filter(SEX == sex_val, tissue == tissue_val)

    n_samples_stratum <- nrow(sub_data)

    model_vars <- c(var, "HEAT", "AGE", "BMI", "TRDNISCH", "COHORT") 

    sub_data <- sub_data %>%
      drop_na(all_of(model_vars))

    n_samples_model <- nrow(sub_data)
    n_samples_removed <- n_samples_stratum - n_samples_model

    # Skip if too few samples
    if (n_samples_model < MIN_MODEL_N) {
      failed_model_fits[[paste(var, sex_val, tissue_val, sep = "_")]] <- tibble(
        health_var = var,
        SEX = sex_val,
        tissue = tissue_val,
        n_samples_stratum = n_samples_stratum,
        n_samples_model = n_samples_model,
        n_samples_removed = n_samples_removed,
        failure_stage = "precheck",
        failure_reason = "Too few samples after filtering"
      )
      next
    }

    # Skip if main variable has fewer than 2 unique values
    if (length(unique(sub_data[[var]])) < 2) {
      failed_model_fits[[paste(var, sex_val, tissue_val, sep = "_")]] <- tibble(
        health_var = var,
        SEX = sex_val,
        tissue = tissue_val,
        n_samples_stratum = n_samples_stratum,
        n_samples_model = n_samples_model,
        n_samples_removed = n_samples_removed,
        failure_stage = "precheck",
        failure_reason = "Main variable has fewer than 2 unique values"
      )
      next
    }

    # Check covariates for invalid factor-like columns with only one level
    covariates_to_check <- c("AGE", "BMI", "TRDNISCH", "COHORT")
    bad_covariates <- covariates_to_check[
      sapply(covariates_to_check, function(x) length(unique(sub_data[[x]])) < 2)
    ]

    if (length(bad_covariates) > 0) {
      failed_model_fits[[paste(var, sex_val, tissue_val, sep = "_")]] <- tibble(
        health_var = var,
        SEX = sex_val,
        tissue = tissue_val,
        n_samples_stratum = n_samples_stratum,
        n_samples_model = n_samples_model,
        n_samples_removed = n_samples_removed,
        failure_stage = "precheck",
        failure_reason = paste(
          "Covariate(s) with fewer than 2 unique values:",
          paste(bad_covariates, collapse = ", ")
        )
      )
      next
    }
    # Variables were selected based on the pariwise correlation with HEAT signature. "TRDNISCH" is ischemic time and also represents many of the variables related to "DTH" variables
    formula <- as.formula(
      paste("HEAT ~", var, "+ AGE + BMI + TRDNISCH + COHORT") 
    )

    model_id <- paste(var, sex_val, tissue_val, sep = "_")

    model <- tryCatch(
      lm(formula, data = sub_data),
      error = function(e) e
    )

    if (inherits(model, "error")) {
      failed_model_fits[[model_id]] <- tibble(
        health_var = var,
        SEX = sex_val,
        tissue = tissue_val,
        n_samples_stratum = n_samples_stratum,
        n_samples_model = n_samples_model,
        n_samples_removed = n_samples_removed,
        failure_stage = "lm",
        failure_reason = conditionMessage(model)
      )
      next
    }

    model_tidy <- tryCatch(
      tidy(model) %>%
        mutate(
          health_var = var,
          SEX = sex_val,
          tissue = tissue_val,
          n_samples_stratum = n_samples_stratum,
          n_samples_model = n_samples_model,
          n_samples_removed = n_samples_removed
        ),
      error = function(e) e
    )

    if (inherits(model_tidy, "error")) {
      failed_model_fits[[model_id]] <- tibble(
        health_var = var,
        SEX = sex_val,
        tissue = tissue_val,
        n_samples_stratum = n_samples_stratum,
        n_samples_model = n_samples_model,
        n_samples_removed = n_samples_removed,
        failure_stage = "tidy",
        failure_reason = conditionMessage(model_tidy)
      )
      next
    }

    stratified_model_res[[model_id]] <- model_tidy
    stratified_models_list[[model_id]] <- model
  }
}



stratified_model_aov <- bind_rows(stratified_model_res, .id = "model_id")
# Add FDR correction
stratified_model_aov <- stratified_model_aov %>%
  mutate(
    tested_term = term %in% vars_to_plot
  ) %>%
  group_by(SEX, tissue) %>%
  mutate(
    FDR = case_when(
      tested_term ~ p.adjust(p.value, method = "BH"),
      TRUE ~ NA_real_
    )
  ) %>%
  ungroup()

failed_model_fits_df <- bind_rows(failed_model_fits, .id = "model_id")

# -----------------------------
#  Save files 
# -----------------------------

write_csv(
  stratified_model_aov,
  file = file.path(OUT_DATA_DIR, "stratified_aov_statistics.csv")
)

write_csv(
  failed_model_fits_df,
  file = file.path(OUT_DATA_DIR, "failed_stratified_aov_models.csv")
)

significant_stratified_aov <- stratified_model_aov %>%
  filter(FDR < 0.05, term %in% vars_to_plot)

write_csv(
  significant_stratified_aov,
  file = file.path(OUT_DATA_DIR, "significant_stratified_aov.csv")
)

# Also write a multi-sheet Excel workbook
excel_file <- file.path(OUT_DATA_DIR, "stratified_aov_summary.xlsx")

wb <- createWorkbook()

addWorksheet(wb, "significant_models")
writeData(wb, "significant_models", significant_stratified_aov)

addWorksheet(wb, "all_models")
writeData(wb, "all_models", stratified_model_aov)

addWorksheet(wb, "failed_models")
writeData(wb, "failed_models", failed_model_fits_df)



saveWorkbook(wb, excel_file, overwrite = TRUE)
```
