---
title: "Plots for Figure 1"
output: html_document
date: "2025-01-14"
---

# Purpose

This notebook assembles the plots used in Figure 1 of the manuscript.
It mainly reads the cleaned Seurat object and upstream summary tables, then
formats the specific figure panels for export.

# Inputs and outputs

Main inputs:
- cleaned Seurat object
- propeller cell proportion table
- propeller statistical summary table
- subject information table
- shared plotting themes and manuscript color palettes

Main outputs:
- Figure 1 panel PDFs

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(speckle)
library(ggnewscale)
library(here)
library(openxlsx)

# Shared plotting functions and manuscript palettes
source(here("scripts/0.environment_setup/snRNAseq_graphics_setup.R"))

# -------------------------------------------------------------------------
# Define inputs
# -------------------------------------------------------------------------
SEURAT_FILE <- here(
  "data",
  "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds"
)

SUBJECT_INFO_FILE <- here(
  "data",
  "2023-09-14_SubjectInformationTable.xlsx"
)

PROPELLER_PROPORTIONS <- here(
  "output", "2.sobject_characterization", "propeller", "data",
  "cellular_proportions_table.csv"
)

PROPELLER_STATS <- here(
  "output", "2.sobject_characterization", "propeller", "data",
  "proportions_stat_table.csv"
)

# -------------------------------------------------------------------------
# Define outputs
# -------------------------------------------------------------------------
OUTPUT_DIR <- here("output", "6.figure_preparation", "Figure_1")
dir.create(OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)
```

# Load the main Seurat object

The starting point for this figure script is the cleaned integrated Seurat
object generated upstream.

```{r}
s.object <- readRDS(SEURAT_FILE)
```

# Panel 1b. Demographics table

This chunk loads the subject information table used for the demographic
summary associated with Figure 1.

```{r}
subject_info <- openxlsx::read.xlsx(
  SUBJECT_INFO_FILE,
  sheet = "Labs",
  rows = 1:19
)

subject_info
summary(subject_info)
```

# Panel 1e. Violin plot of marker genes

This panel visualizes marker genes used to support cell type annotation in the
main dataset.

```{r}
features <- c(
  "ADIPOQ", "PPARG", "PPARGC1A", "PDGFRA", "DCN", "JAM2", "SHANK3",
  "PROX1", "STEAP4", "MYOCD", "CD163", "CPA3", "IL7R", "MS4A1",
  "CSF3R", "CDH19", "CNTNAP5", "PTH", "PRM1"
)

gg_violin <- VlnPlot(
  s.object,
  features = features,
  flip = TRUE,
  stack = TRUE,
  group.by = "cell_type_short",
  combine = TRUE
) +
  scale_fill_manual(values = gene_type_palette) +
  NoLegend() +
  theme(
    axis.text.x = element_text(size = 6, face = "plain"),
    axis.title.y = element_text(size = 6, face = "plain"),
    axis.text.y.right = element_text(size = 6, face = "plain"),
    strip.text = element_text(size = 6, face = "plain"),
    strip.background = element_rect(linewidth = 0.3),
    axis.line = element_line(color = "black", linewidth = 0.3),
    axis.ticks = element_line(color = "black", linewidth = 0.3),
    axis.title.x = element_blank(),
    axis.ticks.y = element_blank()
  )

# Preserve the thinner violin outline used in the original script
gg_violin$layers[[1]]$aes_params$size <- 0.3
```

```{r, fig.width = 2.755906, fig.height = 2.362205}
mm_to_in(c(70, 60))

gg_violin <- gg_violin +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "Expression of Cell Type Marker Genes") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5))

gg_violin

ggsave(
  file.path(OUTPUT_DIR, "Figure_1_D.pdf"),
  plot = gg_violin,
  width = 70,
  height = 60,
  units = "mm",
  bg = "white",
  dpi = 600
)
```

# Panel 1f. Comparison of cell type abundance

This panel reads the propeller-derived cellular abundance tables calculated
upstream and plots regional differences in cell type proportions.

```{r}
propeller_proportions <- read_csv(PROPELLER_PROPORTIONS)
propeller_stats <- read_csv(PROPELLER_STATS)
```

```{r}
# Adjust statistic color display
propeller_stats$colors <- ifelse(propeller_stats$p.val < 0.05, "red", "black")
```


```{r fig.height = 4, fig.width = 10}
p_value_y_position <- max(propeller_stats$y.position)

order <- propeller_proportions %>%
  group_by(clusters) %>%
  summarise(avg_pct = mean(n_pct), .groups = "drop") %>%
  arrange(desc(avg_pct)) %>%
  pull(clusters)

propeller_proportions <- propeller_proportions %>%
  mutate(
    neck_region = factor(
      neck_region,
      levels = c("Superficial", "Intermediate", "Deep")
    )
  )

gg_propeller <- ggplot(
  propeller_proportions,
  aes(x = clusters, y = n_pct, fill = neck_region)
) +
  geom_boxplot(
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    alpha = 0.75
  ) +
  geom_jitter(
    color = "black",
    shape = 21,
    alpha = 0.9,
    size = 1,
    position = position_jitterdodge(
      jitter.width = 0.1,
      dodge.width = 0.75
    )
  ) +
  scale_fill_manual(values = neck.color, name = "Neck Region") +
  scale_color_manual(values = neck.color, name = "Neck Region") +
  new_scale_color() +
  stat_pvalue_manual(
    data = propeller_stats,
    color = "colors",
    xmin = "xmin",
    xmax = "xmax",
    label = "p_display",
    y.position = "y.position",
    vjust = -0.2,
    tip.length = 0.006,
    inherit.aes = FALSE,
    size = 6 / .pt
  ) +
  scale_color_identity(guide = "none") +
  coord_cartesian(ylim = c(NA, p_value_y_position + 2)) +
  ggpubr::grids(axis = "y") +
  scale_x_discrete(limits = order) +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.x = element_blank(),
    legend.position = c(1, 0.80),
    legend.justification = c("right", "top"),
    line = element_line(linewidth = 0.1)
  ) +
  ylab("Proportion of Cells in Sample")
```

```{r, fig.width = 7.086614, fig.height = 1.574803}
mm_to_in(c(180, 40))

gg_propeller <- gg_propeller +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "Cell Type Abundance per Cervical Adipose Tissue Depot") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5)) +
  scale_y_continuous(expand = expansion(c(0.05, 0.1)))

gg_propeller

ggsave(
  file.path(OUTPUT_DIR, "Figure_1_E.pdf"),
  plot = gg_propeller,
  width = 180,
  height = 40,
  units = "mm",
  bg = "white",
  dpi = 600
)

pdf(
  file.path(OUTPUT_DIR, "Figure_1_E.pdf"),
  width = mm_to_in(180),
  height = mm_to_in(40)
)
gg_propeller
dev.off()
```

```{r}
# -----------------------------
#  Prepare subplots 
# -----------------------------

abundance_cutoff <- 3

high_label <- paste0("Mean abundance ≥ ", abundance_cutoff, "%")
low_label  <- paste0("Mean abundance < ", abundance_cutoff, "%")

cluster_abundance <- propeller_proportions %>%
  group_by(clusters) %>%
  summarise(avg_pct = mean(n_pct, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    abundance_group = if_else(
      avg_pct < abundance_cutoff,
      low_label,
      high_label
    )
  )

propeller_proportions_plot <- propeller_proportions %>%
  left_join(cluster_abundance, by = "clusters") %>%
  mutate(
    abundance_group = factor(
      abundance_group,
      levels = c(high_label, low_label)
    ),
    neck_region = factor(
      neck_region,
      levels = c("Superficial", "Intermediate", "Deep")
    )
  )

# -----------------------------
#  Create Inputs Function
# -----------------------------
prepare_propeller_plot_inputs <- function(
    plot_data,
    stats_data,
    cluster_col_stats = "cell_type_short",
    y_expand = 0.08,
    bracket_width = 0.20
) {
  
  cluster_order <- plot_data %>%
    group_by(clusters) %>%
    summarise(avg_pct = mean(n_pct, na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(avg_pct)) %>%
    pull(clusters)
  
  plot_data <- plot_data %>%
    mutate(
      clusters = factor(clusters, levels = cluster_order)
    )
  
  panel_range <- range(plot_data$n_pct, na.rm = TRUE)
  y_offset <- diff(panel_range) * y_expand
  
  if (!is.finite(y_offset) || y_offset == 0) {
    y_offset <- max(plot_data$n_pct, na.rm = TRUE) * y_expand
  }
  
  if (!is.finite(y_offset) || y_offset == 0) {
    y_offset <- 0.1
  }
  
  y_positions <- plot_data %>%
    group_by(clusters) %>%
    summarise(
      y.position = max(n_pct, na.rm = TRUE) + y_offset,
      .groups = "drop"
    ) %>%
    mutate(
      clusters_chr = as.character(clusters)
    )
  
  stats_data_plot <- stats_data %>%
    select(
      -any_of(c(
        "x_position",
        "xmin",
        "xmax",
        "y.position",
        "y.position.x",
        "y.position.y",
        "clusters"
      ))
    ) %>%
    filter(.data[[cluster_col_stats]] %in% cluster_order) %>%
    left_join(
      y_positions %>%
        select(clusters_chr, y.position),
      by = setNames("clusters_chr", cluster_col_stats)
    ) %>%
    mutate(
      x_position = match(.data[[cluster_col_stats]], cluster_order),
      xmin = x_position - bracket_width,
      xmax = x_position + bracket_width,
      clusters = factor(.data[[cluster_col_stats]], levels = cluster_order)
    )
  
  list(
    plot_data = plot_data,
    stats_data = stats_data_plot,
    cluster_order = cluster_order
  )
}

# -----------------------------
#  Plotting function 
# -----------------------------
make_propeller_plot <- function(
    plot_data,
    stats_data = NULL,
    cluster_order = NULL,
    plot_title = NULL,
    show_legend = TRUE
) {
  
  if (is.null(cluster_order)) {
    cluster_order <- plot_data %>%
      group_by(clusters) %>%
      summarise(avg_pct = mean(n_pct, na.rm = TRUE), .groups = "drop") %>%
      arrange(desc(avg_pct)) %>%
      pull(clusters)
  }
  
  p <- ggplot(
    plot_data,
    aes(x = clusters, y = n_pct, fill = neck_region)
  ) +
    geom_hline(yintercept = 40, linewidth = 0.3, color = 3, linetype = "dashed") + 
    geom_boxplot(
      outlier.shape = NA,
      position = position_dodge(width = 0.75),
      alpha = 0.75
    ) +
    geom_jitter(
      color = "black",
      shape = 21,
      alpha = 0.9,
      size = 1,
      position = position_jitterdodge(
        jitter.width = 0.1,
        dodge.width = 0.75
      )
    ) +
    scale_fill_manual(values = neck.color, name = "Neck Region") +
    scale_color_manual(values = neck.color, name = "Neck Region") +
    scale_x_discrete(limits = cluster_order) +
    ggpubr::grids(axis = "y") +
    theme_nature_metabolism() +
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      axis.title.x = element_blank(),
      legend.position = if (show_legend) c(1, 0.80) else "none",
      legend.justification = c("right", "top"),
      line = element_line(linewidth = 0.1)
    ) +
    ylab("Proportion of Cells in Sample") +
    ggtitle(plot_title)
  
  if (!is.null(stats_data) && nrow(stats_data) > 0) {
    
    y_range <- range(plot_data$n_pct, na.rm = TRUE)
    y_padding <- diff(y_range) * 0.15
    
    if (!is.finite(y_padding) || y_padding == 0) {
      y_padding <- max(plot_data$n_pct, na.rm = TRUE) * 0.15
    }
    
    if (!is.finite(y_padding) || y_padding == 0) {
      y_padding <- 0.1
    }
    
    p <- p +
      new_scale_color() +
      stat_pvalue_manual(
        data = stats_data,
        color = "colors",
        xmin = "xmin",
        xmax = "xmax",
        label = "p_display",
        y.position = "y.position",
        vjust = -0.2,
        tip.length = 0.006,
        inherit.aes = FALSE,
        size = 6 / .pt
      ) +
      scale_color_identity(guide = "none") +
      coord_cartesian(
        ylim = c(
          NA,
          max(stats_data$y.position, na.rm = TRUE) + y_padding
        )
      )
  }
  
  p
}

# -----------------------------
#  Create plots 
# -----------------------------
propeller_plot_inputs <- list()
propeller_plots <- list()
clusters_per_plot <- vector()

for (group_name in levels(propeller_proportions_plot$abundance_group)) {
  
  plot_data_group <- propeller_proportions_plot %>%
    filter(abundance_group == group_name)
  
  cluster_in_plot <- length(unique(plot_data_group$clusters))
  clusters_per_plot <- c(clusters_per_plot, cluster_in_plot)
  
  if (nrow(plot_data_group) == 0) {
    next
  }
  
  propeller_plot_inputs[[group_name]] <- prepare_propeller_plot_inputs(
    plot_data = plot_data_group,
    stats_data = propeller_stats,
    cluster_col_stats = "cell_type_short",
    y_expand = 0.08,
    bracket_width = 0.20
  )
  
  propeller_plots[[group_name]] <- make_propeller_plot(
    plot_data = propeller_plot_inputs[[group_name]]$plot_data,
    stats_data = propeller_plot_inputs[[group_name]]$stats_data,
    cluster_order = propeller_plot_inputs[[group_name]]$cluster_order,
    plot_title = group_name,
    show_legend = TRUE
  )
}

# -----------------------------
#  Display plots 
# -----------------------------
propeller_plots

```

```{r, fig.width = 7.086614, fig.height = 1.968504, warning=FALSE}
mm_to_in(c(180, 50))

plot_widths <- clusters_per_plot / sum(clusters_per_plot)

gg_propeller_v2 <- wrap_plots(
  propeller_plots,
  widths = plot_widths
)

gg_propeller_v2

ggsave(
  file.path(OUTPUT_DIR, "Figure_1_E.v2.pdf"),
  plot = gg_propeller_v2,
  width = 180,
  height = 50,
  units = "mm",
  bg = "white",
  dpi = 600
)
```

# Panel 1d. Initial UMAP plot

This panel shows the main UMAP colored by simplified cell type labels used in
the manuscript.

```{r, fig.width = 5, fig.height = 4}
scale_factor <- 6
gg_umap <- DimPlot(
  s.object,
  reduction = "umap",
  group.by = "cell_type_short",
  label = TRUE,
  label.size = 6 * scale_factor/ .pt,
  repel = TRUE
) +
  scale_color_manual(values = palette.use, breaks = order)+
  theme_nature_metabolism() +
  theme_void() +
  theme(
    plot.title = element_blank(),
    legend.text = element_text(size = 6 * scale_factor),
    legend.key.size = unit(2.5 * scale_factor, "mm")
  ) +
  guides(
    color = guide_legend(
      ncol = 1,
      override.aes = list(size = 1 * scale_factor)
    )
  ) 
```

```{r, fig.width = 2.755906 * 6, fig.height = 2.362205 * 6}
mm_to_in(c(70, 60))

gg_umap <- gg_umap +
  labs(title = "Cervical AT snRNAseq\nUMAP Plot") +
  theme(plot.title = element_text(size = 6 * scale_factor, face = "bold", hjust = 0.5))

gg_umap

ggsave(
  file.path(OUTPUT_DIR, "Figure_1_C.pdf"),
  plot = gg_umap,
  width = 70* scale_factor,
  height = 60* scale_factor,
  units = "mm",
  bg = "white",
  dpi = 600
)
```

# Supplementary marker dotplot. Beige, brown, and adipocyte marker programs

This section reorganizes adipocytes by depot and visualizes classical and
thermogenic marker programs across white and brown adipocyte groups.

```{r, fig.width = 13.779528, fig.height = 3.937008}
mm_to_in(c(100, 100))

# -------------------------------------------------------------------------
# Regroup adipocytes by depot
# -------------------------------------------------------------------------
meta <- s.object@meta.data

meta <- meta %>%
  mutate(
    adipo_region = case_when(
      str_detect(cell_type, " Adipocyte") ~ paste0(cell_type, " (", neck_region, ")"),
      TRUE ~ "Other"
    )
  ) %>%
  mutate(
    adipo_region = case_when(
      adipo_region == "Brown Adipocytes (Intermediate)" ~ "Other",
      adipo_region == "Brown Adipocytes (Superficial)" ~ "Other",
      TRUE ~ adipo_region
    )
  ) %>%
  mutate(
    adipo_region = factor(
      adipo_region,
      levels = c(
        sort(unique(adipo_region[str_detect(adipo_region, "^Brown Adipocytes")])),
        sort(unique(adipo_region[str_detect(adipo_region, "^White Adipocytes")])),
        "Other"
      )
    )
  )

s.object@meta.data <- meta

# -------------------------------------------------------------------------
# Define marker gene groups
# -------------------------------------------------------------------------
adipo_markers <- c("ADIPOQ", "PPARG")
white_markers <- c("LEP", "SLC7A10", "HOCX8", "HOXC9")
thermogenic_markers <- c("UCP1", "PPARGC1A", "PRDM16", "EBF2", "DIO2")
brown_markers <- c("EBF3", "FBXO31", "ZIC1", "MPZL2", "LHX8")
beige_markers <- c("TBX1", "TMEM26", "TNFRSF9", "SHOX2")
fcc_markers <- c("GAMT", "GATM", "SLC6A8", "ALPL", "CKB", "CKMT1A", "CKMT1B", "CKMT2")
calcium_markers <- c("ATP2A2", "RYR2", "ARLN", "ALN", "ITPR1", "ITPR2")
lipid_cyc_markers <- c("PNPLA2", "DGAT1")

feature_groups <- list(
  "General Adipocyte" = adipo_markers,
  "White Adipocyte" = white_markers,
  "Brown Adipocyte" = brown_markers,
  "Beige Adipocyte" = beige_markers,
  "Thermogenic Genes" = thermogenic_markers,
  "FCC Genes" = fcc_markers,
  "Calcium Cycling Genes" = calcium_markers,
  "Lipid Cycling Genes" = lipid_cyc_markers
)

feature_groups <- lapply(feature_groups, function(gene_list) {
  found_genes <- gene_list %in% rownames(s.object)
  gene_list[found_genes]
})

group_labels <- c(
  "General Adipocyte" = "Adipocyte",
  "White Adipocyte" = "White",
  "Brown Adipocyte" = "Brown",
  "Beige Adipocyte" = "Beige",
  "Thermogenic Genes" = "Classical Thermogenic",
  "FCC Genes" = "Futile Creatine Cycling",
  "Calcium Cycling Genes" = "Calcium Cycling",
  "Lipid Cycling Genes" = "Lipid Cycling"
)

all_features <- unlist(feature_groups) %>% toupper()
feature_order <- all_features

# -------------------------------------------------------------------------
# Create accessory plotting data for category brackets
# -------------------------------------------------------------------------
group_lengths <- lengths(feature_groups)
group_ends <- cumsum(group_lengths)
group_starts <- c(1, head(group_ends, -1) + 1)
group_mids <- (group_starts + group_ends) / 2

bracket_df <- data.frame(
  group = names(feature_groups),
  start = group_starts,
  end = group_ends,
  mid = group_mids,
  label = str_wrap(unname(group_labels[names(feature_groups)]), width = 80)
)

# -------------------------------------------------------------------------
# Alternating background panels by feature group
# -------------------------------------------------------------------------
panel_df <- bracket_df %>%
  mutate(
    xmin = start - 0.5,
    xmax = end + 0.5,
    fill = rep(c("white", "grey95"), length.out = n())
  )

# -------------------------------------------------------------------------
# Build the dot plot
# -------------------------------------------------------------------------
dp <- DotPlot(
  s.object,
  features = all_features,
  group.by = "adipo_region",
  cols = c("lightgrey", "blue")
)

dp$data$features.plot <- factor(dp$data$features.plot, levels = feature_order)


main_plot <- ggplot(dp$data, aes(x = features.plot, y = id)) +
  geom_rect(
    data = panel_df,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
    inherit.aes = FALSE,
    color = NA
  ) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_fill_identity() +
  scale_size_continuous(
    limits = c(0, 100),
    breaks = c(20, 40, 60, 80),
    range = c(0, 5),
    name = "% cells"
  ) +
  scale_color_gradient(
    low = "lightgrey",
    high = "blue",
    limits = c(-2, 2),
    breaks = seq(-2, 2, 1),
    name = "Avg. expression"
  ) +
  labs(x = NULL, y = NULL) +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10)
  )

bracket_plot <- ggplot(bracket_df) +
  geom_segment(
    aes(x = start - 0.2, xend = end + 0.2, y = 0.2, yend = 0.2),
    linewidth = 0.3
  ) +
  geom_text(
    aes(x = mid, y = 0.65, label = label),
    size = 5 / .pt,
    fontface = "bold"
  ) +
  scale_x_continuous(
    limits = c(0.5, length(feature_order) + 0.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t = 5, r = 10, b = 0, l = 10)
  )

combined_plot <- bracket_plot / main_plot +
  plot_layout(heights = c(2.2, 12))
```

```{r, fig.width = 7.086614, fig.height = 1.574803}
mm_to_in(c(180, 40))

combined_plot

ggsave(
  file.path(OUTPUT_DIR, "Figure_1G_Classical_markers.v2.pdf"),
  plot = combined_plot,
  width = 180,
  height = 40,
  units = "mm",
  bg = "white",
  dpi = 600
)
```
