---
title: "Plots for Figure 2 and 3"
output: html_document
date: "2025-01-14"
---

Plotting of Figures 2 and 3 were combined as they share the VISION scores plotting for most part of it

# --- Setup ---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(colorRamp2)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(speckle)
library(scales)
library(openxlsx)
library(ComplexHeatmap)
library(circlize)
library(grid)
library(openxlsx)
library(ggnewscale)
library(ggplotify)
library(ggvenn)
library(ggpmisc)
library(VISION)
library(viridis)
library(enrichR)
library(colorspace)
library(here)

#Conflict solving
conflicted::conflicts_prefer(pheatmap::pheatmap)

# OBS:  theme_nature_metabolism() and color palettes set on "scripts/0.environment_setup/snRNAseq_graphics_setup.R"
source(here("scripts/0.environment_setup/snRNAseq_graphics_setup.R"))

# --- Setup paths ---
## Inputs
### Seurat input
SEURAT_INPUT <- here(
  "..",
  "NeckAdipose/ocean_code_data/cervical_at_gex_seurat.rds")

SEURAT_INPUT <- here("data/adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds")

### HEAT signature generation
HEAT_SIGNATURE_RDATA <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/heat_signature.Rdata")
LYNES_MICROARRAY_CTS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Supplementary_Table_cervical_at_microarray.xlsx")
LYNES_CORRELATION_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Lynes_et_al_correlation.csv")
SALEJ_BULK_CTS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/logCPM_for_correlations.csv")
SALEJ_CORRELATION_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Salej_et_al_logCPM_correlation.csv")
HEAT_VENN_RDATA <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/genes_for_venn.Rdata")
HEAT_SIGNATURE_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/heat_signature.csv")


# HEAT Enrichr
ENRICHED_RDS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/enriched_heat.rds")

# HEAT threshold
HEAT_THRESHOLD_RDS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/wad_heat_threshold.rds")


# BULK RNAseq
GSEA_HEAT_BULK_DIR <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/ssGSEA")

PARAGANG_BULK_MTX_RDS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/GSE49795_Sondergaard_expression_matrix.rds") 
BULK_DIR <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Cleaned_Datasets")
BULK_META_DIR <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Experimental_Design")

A41_CTS_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/camaraH_a41_d18_fpkm.xlsx")
A38_CTS_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/camaraH_a38_d18_tpm.xlsx")
CERO_CTS_XLSX <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/RNA-seq data set Cero et al.xlsx")


# VISION
VISION_SIG_DIR <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/vision_signature_scores")

# Angueira dataset
ANGUEIRA_ADIPO_SUB_RDS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/AngueiraA_Dataset_adipo_recluster.rds")

ANGUEIRA_HEATvUCP1_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/Angueira_HEATvUCP1.csv")

# Mouse datasets
## Shamsi 2021
SHAMSI_VISON_SIG <- file.path(VISION_SIG_DIR, "ShamsiF_Dataset_VISION_sigScore.csv")
SHAMSI_RDS <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/ShamsiF_Dataset.rds")

### CIBERSORTx
CIBERSORT_PROPS_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/deconvolution_cellular_proportions_table.csv")

# GTEx
GTEX_META_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/gtex_metadata_filtered.csv")
GTEX_AOV_STATS_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/stratified_aov_statistics.csv")
GTEX_AOV_STATS_SIGNIF_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/significant_stratified_aov.csv")
GTEX_META_ANNOT_CSV <- here(
  "..",
  "NeckAdipose/ocean_code_data/figure_2_3_S2_S3/gtex_dictionary_filtered.csv")

# Outputs
OUTPUT_DIR <- here("..", "results", "Figure_2_3_S2_S3")

dir.create(OUTPUT_DIR, showWarnings = F, recursive = T)
```

# ---Load the HEAT signature

```{r}
load(HEAT_SIGNATURE_RDATA)
```

# --- Panel S2A. Heatmap of UCP1 correlated genes in Microarray ---
## ---Load count matrix ---
```{r}
# Load and remove spaces from column names
lynes_cts <- read.xlsx(LYNES_MICROARRAY_CTS, sheet = "All Data by Individual")

# Replace spaces and dots
names(lynes_cts) <- gsub("[ \\.]", "_", names(lynes_cts))
```

## ---Load correlation info ---
```{r}
expanded_lynes_cor_df <- read_csv(file = LYNES_CORRELATION_CSV)
```

## --- Format, filter matrix and plot ---
```{r}
# Select deep neck (CS) samples and create unique gene_probe names
data_lynes <- lynes_cts %>%
  dplyr::select(grep("CS", names(lynes_cts), value = TRUE), Gene_Symbol, Probe_Set_ID) %>%
  mutate(Gene_Probe = paste0(Gene_Symbol, "_(", Probe_Set_ID, ")")) %>%
  column_to_rownames("Gene_Probe") %>%
  select(-c(Gene_Symbol, Probe_Set_ID))

# Order samples according to UCP1 expression
samples <- names(data_lynes)

ranked_sample_ucp1 <- data_lynes %>%
  t() %>%
  as_tibble(rownames = "Sample") %>%
  arrange(`UCP1_(11732874_at)`) %>%
  pull(Sample)

data_lynes <- data_lynes %>%
  select(all_of(ranked_sample_ucp1))

# Keep all significantly positive correlated genes
strongest_corrs <- expanded_lynes_cor_df %>%
  filter(
    Gene_Probe %in% rownames(data_lynes),
    p_value < 0.05,
    FDR < 0.1,
    estimate > 0
  ) %>%
  arrange(desc(estimate))

# Put UCP1 first, then all positive genes
gene_order_ids <- c(
  "UCP1_(11732874_at)",
  setdiff(strongest_corrs$Gene_Probe, "UCP1_(11732874_at)")
)

gene_order_ids <- gene_order_ids[gene_order_ids %in% rownames(data_lynes)]

data_lynes <- data_lynes[gene_order_ids, , drop = FALSE]
data_lynes <- as.matrix(data_lynes)

# Keep internal rownames unique
internal_ids <- rownames(data_lynes)

# Gene symbols for display
gene_symbols <- gsub("_\\(.*\\)$", "", internal_ids)

# Scale all positively correlated genes
scaled_data <- t(scale(t(data_lynes)))

# Split UCP1 from others
gene_group <- ifelse(gene_symbols == "UCP1", "UCP1", "Others")
gene_group <- factor(gene_group, levels = c("UCP1", "Others"))

# Which genes to label
genes_to_label <- union("UCP1", HEAT_signature)

# Find rows to mark
rows_to_mark <- which(gene_symbols %in% genes_to_label)
labels_to_mark <- gene_symbols[rows_to_mark]


col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))

ht <- Heatmap(
  scaled_data,
  name = "Z-score",
  row_split = gene_group,
  row_title = NULL,
  col = col_fun,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_column_names = FALSE, 
  show_row_names = FALSE,
  column_names_gp = gpar(fontsize = 5),
  rect_gp = gpar(col = "black", lwd = 0),
  # column_title = "Positively UCP1 correlated genes\nThis Study Microarray",
  # column_title_gp = gpar(fontsize = 6, fontface = "bold"),
  show_heatmap_legend = FALSE,
  left_annotation = rowAnnotation(
    mark = anno_mark(
      at = rows_to_mark,
      labels = labels_to_mark,
      side = "left",
      labels_gp = gpar(
        fontsize = 5,
        lineheight = 0.8
      ),
      lines_gp = gpar(
        col = grDevices::adjustcolor("gray50", alpha.f = 0.6),
        lwd = 0.3
      ),
      link_width = unit(2, "mm"),
      link_height = unit(0.8, "mm"),
      padding = unit(0.2, "mm"),
      extend = unit(0, "mm")
    )
  )
)

```

## --- Save plot ---
```{r, fig.width=1.181102, fig.height=7.086614}
mm_to_in(c(30, 180))
ht

pdf(file.path(OUTPUT_DIR, "S2A_ucp1_heatmap_microarray.pdf"), width = mm_to_in(30), height = mm_to_in(180))
ht
dev.off()
```


# --- Panel S2B. Heatmap of UCP1 correlated genes in Salej Duran et al 2025 ---
## ---Load count matrix
```{r}
# Load TPM data
log2_cpm <- read_csv(SALEJ_BULK_CTS)

# Filter out NA genes
log2_cpm <- log2_cpm[!is.na(log2_cpm$gene), ]

# Store gene names
ensmbl_genes <- log2_cpm %>%
  select(X, gene) %>%
  rename("Ensembl" = "X")

# Add genes to rownames and keep Deep neck samples
log2_cpm_clean <- log2_cpm %>%
  # filter(gene %in% c("UCP1", HEAT_signature)) %>%
  mutate(gene = make.unique(gene)) %>%
  column_to_rownames("X") %>%
  select(contains("Deep"))

# Transpose
cpm_deep.t <- as.data.frame(t(log2_cpm_clean))

# Convert to numeric
cpm_deep.t <- data.frame(lapply(cpm_deep.t, function(x) as.numeric(as.character(x))))
```

## ---Load correlation info
```{r}
salej_cor_df <- read_csv(file = SALEJ_CORRELATION_CSV)
```

## --- Format, filter matrix and plot ---
```{r}
# Create object data_salej (Sample x Gene ensembls at this stage)
data_salej <- cpm_deep.t

# Order samples according to UCP1 expression
ucp1_ensembl <- ensmbl_genes %>%
  filter(gene == "UCP1") %>%
  pull(Ensembl)

data_salej <- data_salej %>%
  arrange(.data[[ucp1_ensembl]])

# Keep all significantly positive correlated genes
strongest_corrs <- salej_cor_df %>%
  filter(
    Ensembl %in% names(data_salej),
    PValue < 0.05,
    FDR < 0.1,
    estimate > 0
  ) %>%
  arrange(desc(estimate))

gene_order_ids <- c(
  ucp1_ensembl,
  setdiff(strongest_corrs$Ensembl, ucp1_ensembl)
)

gene_order_ids <- gene_order_ids[gene_order_ids %in% names(data_salej)]

data_salej <- data_salej[, gene_order_ids, drop = FALSE]

# Map Ensembl to gene symbol
gene_symbols <- ensmbl_genes$gene[match(gene_order_ids, ensmbl_genes$Ensembl)]

# Keep internal IDs unique
data_salej <- as.matrix(data_salej)
colnames(data_salej) <- gene_order_ids

# Convert to Genes x Samples
data_salej <- t(data_salej)

scaled_data <- t(scale(t(data_salej)))

gene_group <- ifelse(gene_symbols == "UCP1", "UCP1", "Others")
gene_group <- factor(gene_group, levels = c("UCP1", "Others"))

genes_to_label <- union("UCP1", HEAT_signature)

rows_to_mark <- which(gene_symbols %in% genes_to_label)
labels_to_mark <- gene_symbols[rows_to_mark]

col_fun <- colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))

ht <- Heatmap(
  scaled_data,
  name = "Z-score",
  row_split = gene_group,
  row_title = NULL,
  col = col_fun,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_column_names = FALSE, 
  show_row_names = FALSE,
  column_names_gp = gpar(fontsize = 5),
  rect_gp = gpar(col = "black", lwd = 0),
  # column_title = "Positively UCP1 correlated genes\nSalej et al. RNAseq",
  # column_title_gp = gpar(fontsize = 6, fontface = "bold"),
  show_heatmap_legend = FALSE,
  left_annotation = rowAnnotation(
    mark = anno_mark(
      at = rows_to_mark,
      labels = labels_to_mark,
      side = "left",
      labels_gp = gpar(
        fontsize = 5,
        lineheight = 0.8
      ),
      lines_gp = gpar(
        col = grDevices::adjustcolor("gray50", alpha.f = 0.6),
        lwd = 0.3
      ),
      link_width = unit(2, "mm"),
      link_height = unit(0.8, "mm"),
      padding = unit(0.2, "mm"),
      extend = unit(0, "mm")
    )
  )
)
```


## --- Save plot ---
```{r, fig.width=1.181102, fig.height=7.086614}
mm_to_in(c(30, 180))

ht

pdf(file.path(OUTPUT_DIR, "S2B_ucp1_heatmap_salej.pdf"), width = mm_to_in(30), height = mm_to_in(180))
ht
dev.off()
```


# --- Panel 2A. VennDiagram HEAT signature strategy ---
```{r, fig.width=1.181102, fig.height=1.181102}
mm_to_in(c(30,50))
# Load the list of genes_for_venn for HEAT signature plot analysis
load(file = HEAT_VENN_RDATA)
str(genes_for_venn)
# ---Plot Venn Diagram ---
myCol <- brewer.pal(length(genes_for_venn), "Pastel2")
venn <- ggvenn(
  text_size = 6/.pt,
  genes_for_venn,
  fill_color = myCol, 
  show_percentage = FALSE, 
  set_name_color = colorspace::darken(myCol, amount = 0.6),
  stroke_size = 0, 
  set_name_size = 6/.pt, 
) 
venn
# Save PDF
ggsave(file.path(OUTPUT_DIR, "2A.2_venn_heat.pdf"), venn, width = 30, height = 50, units = "mm", bg = "white",  dpi = 600)

# ---Plot Venn Diagram Subset ---
myCol <- brewer.pal(2, "Pastel2")
venn <- ggvenn(
  text_size = 6/.pt,
  genes_for_venn[1:2],
  fill_color = myCol[1:2], 
  show_percentage = FALSE, 
  set_name_color = colorspace::darken(myCol[1:2], amount = 0.6),
  stroke_size = 0, 
  set_name_size = 6/.pt, 
) 
venn
# Save PDF
ggsave(file.path(OUTPUT_DIR, "2A.1_venn_heat.pdf"), venn, width = 30, height = 50, units = "mm", bg = "white",  dpi = 600)
```


# --- Panel 2B. Heatmap of HEAT signature in cervical AT of CamaraH ---
## --- Load Seurat - CamaraH ---
```{r}
s.object <- readRDS(SEURAT_INPUT)
```


## --- Format data ---
```{r}
library(colorspace)

# Load HEAT signature
load(HEAT_SIGNATURE_RDATA)
data <- DotPlot(s.object, HEAT_signature, group.by = "cell_type_short")
data <- data$data

cell_order <- data |>
  group_by(id) |>
  summarise(mean.exp = mean(avg.exp.scaled)) |>
  arrange(mean.exp) |>
  pull(id)
gene_order <- data |>
  filter(id == "BAds") |>
  arrange(-avg.exp.scaled) |>
  pull(features.plot)
data <- data |> mutate(
  id = factor(id, levels = cell_order),
  features.plot = factor(features.plot, levels = gene_order)
)
```

## --- Plotting and save ---
```{r, fig.width=7.086614, fig.height=1.968504}
mm_to_in(c(180, 50))

gg_heat_heatmap_cervical <- ggplot(data, aes(features.plot, id)) +
  geom_tile(aes(fill = avg.exp.scaled)) +
  scale_fill_continuous_divergingx(
    palette = "RdGy", rev = T,
    mid = 0,
    name = "Scaled Expression"
  ) +
  theme_nature_metabolism(base_size = 6) +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    axis.line = element_blank()
  ) +
  labs(title = "HEAT Signature Gene Expression")

gg_heat_heatmap_cervical + theme(legend.position = "right")

ggsave(file.path(OUTPUT_DIR, "2B_heat_heatmap_cervical_at.pdf"), plot = gg_heat_heatmap_cervical, width = 180, height = 50, units = "mm", bg = "white",  dpi = 600)
```


# --- Plot VISION scores ---
## --- Read signature score files ---
```{r}
# List files to read
sigscores_df_files <- list.files(
  VISION_SIG_DIR,
  full.names = TRUE,
  pattern = "_sigScore.csv"
)

# Read forcing the right type for each column
sigscore_list <- lapply(
  sigscores_df_files,
  read_csv,
  col_types = "cddcccccdccc"
)

# Merge the files in a single one
sigscores_df <- bind_rows(sigscore_list)

# Remove in case there is duplicated signature loading
sigscores_df <- sigscores_df %>%
  distinct(barcode, Hallmark, dataset, .keep_all = T)

sigscores_df |> distinct(Hallmark) |> filter(str_detect(Hallmark, "OXIDATIVE"))
sigscores_df |> filter(Hallmark %in% c("HEAT", 
                                       "HALLMARK_FATTY_ACID_METABOLISM", 
                                       "HALLMARK_OXIDATIVE_PHOSPHORYLATION", 
                                       "BATLAS", 
                                       "GOBP_ADAPTIVE_THERMOGENESIS")) |> 
write_csv("~/Downloads/CamaraH_vision_subset.csv")

# Clean dataset name
sigscores_df$dataset <- str_remove(sigscores_df$dataset, "_Dataset")

sigscores_df |> distinct(Hallmark) |> filter(str_detect(Hallmark, "OXIDATIVE"))
sigscores_df |> filter(Hallmark %in% c("HEAT", 
                                       "HALLMARK_FATTY_ACID_METABOLISM", 
                                       "HALLMARK_OXIDATIVE_PHOSPHORYLATION", 
                                       "BATLAS", 
                                       "GOBP_ADAPTIVE_THERMOGENESIS")) |> 
write_csv("~/Downloads/CamaraH_vision_subset.csv")
```

## --- Add graphical information ---
```{r}
# Rename Efthymiou adipocytes
sigscores_df <- sigscores_df %>% mutate(
  cell_type = case_when(
    str_detect(dataset, "Eft") & str_detect(cell_type, "Adipo") ~ "Adipocytes",
    TRUE ~ cell_type
  )
)

# Rename Emont cell types to Title Case
sigscores_df <- sigscores_df %>% mutate(
  cell_type = case_when(
    str_detect(dataset, "Emont") ~ str_to_title(cell_type),
    TRUE ~ cell_type
  )
)

# Add some graph formatting information
sigscores_df <- sigscores_df %>%
  mutate(
    fill_color = case_when(
      str_detect(cell_type, "ASPC|[Pp]rogenitor") ~ "grey90",
      str_detect(cell_type, "Brown") ~ "#97600f",
      (region == "perivasc" & str_detect(cell_type, "Adipo")) ~ "#deb373",
      (region == "neck" & str_detect(cell_type, "Adipo")) ~ "#deb373",
      str_detect(cell_type, "Adipo") ~ "#ffe65a",
      TRUE ~ "grey90"
    ),
    region_label = case_when(
      region == "subc" ~ "Subcutaneous Abdominal",
      region == "visc" ~ "Visceral Abdominal",
      region == "perivasc" ~ "Perivascular",
      region == "Subc" ~ "Subcutaneous Cervical",
      region == "neck" ~ "Deep Cervical",
      region == "Deep" ~ "Deep Cervical",
      region == "Superficial" ~ "Subcutaneous Cervical",
      TRUE ~ NA
    )
  ) %>%
  mutate(cell_type_dataset = paste(cell_type, dataset, sep = "_"))

sigscores_df <- sigscores_df |> 
  mutate(publication = case_when(dataset == "CamaraH" ~ "This Study",
                                            dataset == "AngueiraA" ~ "Angueira et al. 2021",
                                            dataset == "SunW" ~ "Sun et al. 2020",
                                            dataset == "WangT" ~ "Wang et al. 2024",
                                            dataset == "EfthymiouV" ~ "Efthymiou et al. 2025",
                                            dataset =="LazarescuO" ~ "Lazarescu et al. 2025",
                                            dataset == "Emont" ~ "Emont et al. 2022",
                                            TRUE ~ "Unknown"),
          depot = case_when(dataset == "CamaraH" ~ "Cervical",
                                            dataset == "AngueiraA" ~ "Perivascular",
                                            dataset == "SunW" ~ "Cervical",
                                            dataset == "WangT" ~ "Cervical",
                                            dataset == "EfthymiouV" ~ "Abdominal",
                                            dataset =="LazarescuO" ~ "Abdominal",
                                            dataset == "Emont" ~ "Abdominal",
                                            TRUE ~ "Unknown"),
          cell_type_Region = interaction(cell_type, region_label, sep = " | ", drop = TRUE),
         region_extension = str_remove(region_label, " .*"),
         region_extension = ifelse(str_detect(region_extension, "vascular"), "", region_extension)) 



# Put CamaraH as the first dataset by factoring
dataset_levels <- sort(unique(sigscores_df$dataset))
idx <- match("CamaraH", dataset_levels)
dataset_levels <- c(dataset_levels[idx], dataset_levels[-idx])
sigscores_df$dataset <- factor(sigscores_df$dataset, levels = dataset_levels)

# Remove cells with less than 10
less_than_10_cells <- sigscores_df %>%
  distinct(cell_type_dataset, region, barcode) %>%
  count(cell_type_dataset, region) %>%
  filter(n <= 10) %>%
  mutate(filter_col = paste0(cell_type_dataset, region, sep = "_")) %>%
  pull(filter_col)

sigscores_df <- sigscores_df %>%
  mutate(filter_col = paste0(cell_type_dataset, region, sep = "_")) %>%
  filter(!filter_col %in% less_than_10_cells)

sigscores_df %>% filter(
  str_detect(dataset, "Eft") & str_detect(cell_type, "Adipo")
)

sigscores_df |> count(dataset, cell_type_dataset)
```

## --- Load Seurat from CamaraH  ---
```{r}
s.object <- readRDS(SEURAT_INPUT)
```

# --- Panel 2C. HEAT signature Feature Plot ---
```{r, fig.width=1.13189, fig.height=1.13189}
mm_to_in(28.75)

scale_factor <- 2

library(ggrastr)

meta_simplified <- s.object@meta.data %>%
  rownames_to_column("barcode") %>%
  select(barcode, cell_type_short)
graph.data <- sigscores_df %>%
  filter(dataset == "CamaraH", Hallmark == "HEAT") %>%
  left_join(meta_simplified, by = "barcode")

### Circle the brown adipocyte cluster
brown <- graph.data %>%
  distinct(barcode, .keep_all = TRUE) |> 
  filter(cell_type == "Brown Adipocytes") %>%
  select(umap_1, umap_2)

# robust center
cx <- median(brown$umap_1, na.rm = TRUE)
cy <- median(brown$umap_2, na.rm = TRUE)

# robust radius: e.g. 90% of points within the circle
d <- sqrt((brown$umap_1 - cx)^2 + (brown$umap_2 - cy)^2)
r <- quantile(d, 0.90, na.rm = TRUE) * 1.15  # +15% padding

theta <- pi/4
x_lab <- cx + 2 * r * cos(theta)
y_lab <- cy + r * sin(theta)

gg_heat_featplot_annot <- gg_heat_featplot +
  annotate("path",
           x = cx + r * cos(seq(0, 2*pi, length.out = 361)),
           y = cy + r * sin(seq(0, 2*pi, length.out = 361)),
           linewidth = 0.4,
           color = "black") +
  annotate("text",
           x = x_lab, y = y_lab,
           label = "BAds",
           hjust = 0, vjust = 0,          # anchor text from lower-left
           nudge_x = 0.1, nudge_y = 0.1,   # tweak for your UMAP scale
           size = 6* scale_factor/.pt)
```

## --- Plotting and save ---
```{r, fig.width=1.13189* scale_factor, fig.height=1.13189* scale_factor}
gg_heat_featplot_annot
# Save
ggsave(file.path(OUTPUT_DIR, "2C_heat_vision_featplot.pdf"), plot = gg_heat_featplot_annot, width = 100 * scale_factor, height = 50 * scale_factor, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel S2D. HEAT signature Violin Plot in CamaraH ---
## --- Read the Q99 HEAT percentile of White Adipocytes ---

```{r}
wad_HEAT_percentile <- readRDS(HEAT_THRESHOLD_RDS)
```

```{r}
# Violin in All Cells
gg_heat_vln_allcells <- ggplot(graph.data, aes(x = cell_type_short, y = sigScores, fill = cell_type_short)) +
  geom_hline(yintercept = wad_HEAT_percentile, color = "black", linetype = "dashed", linewidth = 0.2) +
  geom_jitter(aes(color = cell_type_short), width = 0.2, alpha = 0.5, size = 0.5, stroke = 0) +
  geom_violin(trim = FALSE, color = "black", linewidth = 0.3) + # <- black border here
theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "in")
  ) +
  labs(
    title = "HEAT Signature score",
    x = "Cell Type",
    y = paste("HEAT Signature score\n(VISION)")
  ) +
  scale_fill_manual(values = palette.use) +
  scale_color_manual(values = palette.use) +
  NoLegend()
```

## --- Plotting and save ---
```{r}
gg_heat_vln_allcells

scale <- 180
ggsave(file.path(OUTPUT_DIR, "S2D_heat_vln_allcells.pdf"), plot = gg_heat_vln_allcells, width = 60, height = 32, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel S2C. UCP1-correlated genes signature Feature Plot ---
```{r, fig.width=1.417323*4, fig.height=1.259843*4}
mm_to_in(c(36, 32))
graph.data <- sigscores_df %>%
  filter(dataset == "CamaraH", Hallmark == "UCP1 correlated") %>%
  left_join(meta_simplified, by = "barcode")

### Circle the brown adipocyte cluster
brown <- graph.data %>%
    distinct(barcode, .keep_all = TRUE) |> 
  filter(cell_type == "Brown Adipocytes") %>%
  select(umap_1, umap_2)

# robust center
cx <- median(brown$umap_1, na.rm = TRUE)
cy <- median(brown$umap_2, na.rm = TRUE)

# robust radius: e.g. 90% of points within the circle
d <- sqrt((brown$umap_1 - cx)^2 + (brown$umap_2 - cy)^2)
r <- quantile(d, 0.90, na.rm = TRUE) * 1.15  # +15% padding

theta <- pi/4
x_lab <- cx + 2 * r * cos(theta)
y_lab <- cy + r * sin(theta)

gg_ucp1_corr_featplot_annot <- gg_ucp1_corr_featplot +
  annotate("path",
           x = cx + r * cos(seq(0, 2*pi, length.out = 361)),
           y = cy + r * sin(seq(0, 2*pi, length.out = 361)),
           linewidth = 0.4,
           color = "black") +
  annotate("text",
           x = x_lab, y = y_lab,
           label = "BAds",
           hjust = 0, vjust = 0,          # anchor text from lower-left
           nudge_x = 0.1, nudge_y = 0.1,   # tweak for your UMAP scale
           size = 5*3/.pt)
```

## --- Plotting and save ---
```{r, fig.width=1.417323*4, fig.height=1.259843*4}
gg_ucp1_corr_featplot_annot

# Save
ggsave(file.path(OUTPUT_DIR, "S2C_ucp1_corr_vision_featplot.pdf"), plot = gg_ucp1_corr_featplot_annot, width = 36 * 3, height = 32 * 3, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel S2E-H: Enrichr HEAT signature analysis ---
```{r}
# -----------------------------
#  Load enriched paths 
# -----------------------------
enriched <- readRDS(ENRICHED_RDS)
```

```{r}
websiteLive <- getOption("enrichR.live")
```

## --- Set formatting helper functions ---
```{r}
dbs <- names(enriched)
# 1) One-hue per DB, light -> dark.
# Uses colorspace for good palette generation, but you can swap to RColorBrewer if you prefer.
db_base_cols <- setNames(
  colorspace::qualitative_hcl(length(dbs), palette = "Dark 3"),
  dbs
)

mono_fill_scale_for_db <- function(db_name) {
  base <- db_base_cols[[db_name]]
  low  <- colorspace::lighten(base, amount = 0.90)
  high <- colorspace::lighten(base,  amount = 0.45)

  # Most enrichment plots map fill to P.value where smaller is "more significant".
  # Reverse so smaller p gets darker.
  scale_fill_gradient(low = low, high = high, trans = "reverse")
}

# 2) Keep existing coord system, but disable clipping.
# If plot already uses coord_flip, we re-add coord_flip with clip='off' (same coord, no cropping).
add_no_clip_coord <- function(p) {
  if (inherits(p$coordinates, "CoordFlip")) {
    p + coord_flip(clip = "off")
  } else {
    p + coord_cartesian(clip = "off")
  }
}

# 3) Add extra space on the *continuous* axis, respecting coord_flip.
add_continuous_expand <- function(p, mult = c(0.02, 0.35)) {
  if (inherits(p$coordinates, "CoordFlip")) {
    # When flipped, the continuous axis is usually y (Count, -log10(p), etc.)
    p + scale_y_continuous(expand = expansion(mult = mult))
  } else {
    p + scale_x_continuous(expand = expansion(mult = mult))
  }
}

# 4) Put pathway names "inside" the bars (visually) and left-aligned to the axis origin.
# For coord_flip plots (common): x is Term (discrete), y is Count (continuous).
# Place text at y=0, so after flipping it lands at x=0.
add_inside_left_labels <- function(p, term_col = "Term", wrap_width = 25) {
  p +
    theme(
      axis.text.y  = element_blank(),
      axis.ticks.y = element_blank(),
      plot.margin  = margin(t = 5.5, r = 5.5, b = 5.5, l = 5.5)
    ) +
    geom_text(
      data = p$data,
      aes(
        x = .data[[term_col]],
        y = 0,
        label = stringr::str_wrap(.data[[term_col]], width = wrap_width)
      ),
      hjust = 0,
      vjust = 0.5,
      size = 5 / .pt,
      inherit.aes = FALSE,
      fontface = "bold"
    )
}
```

## --- Plot ---
```{r}
plot_list <- list()
base_size <- 6

if (websiteLive) {
  for (database in dbs) {

    plot_data <- enriched[[database]] %>%
      mutate(
        Term = str_remove_all(Term, "\\s*\\(.*"),
        Term = str_wrap(Term, width = 100)
      )

    p <- plotEnrich(
      plot_data,
      showTerms = 5,
      numChar = 500,
      y = "Ratio",              # will auto-fallback to Count if needed (your warning)
      orderBy = "P.value",
      title = str_wrap(str_replace_all(database, "_", " "), width = 30)
    ) +
      theme_nature_metabolism(base_size) +
      theme(legend.position = "right") +
      mono_fill_scale_for_db(database)   # will replace existing fill scale (fine)

    p <- p %>%
      add_inside_left_labels(term_col = "Term", wrap_width = 45) %>%
      add_no_clip_coord() %>%
      add_continuous_expand(mult = c(0.02, 0.35))

    print(p)
    plot_list[[database]] <- p
  }
}

```
## --- Plotting and save --- 
```{r, fig.width=2.834646*1.5, fig.height=2.519685}
mm_to_in(c(72, 64))
for(pathway in names(plot_list)){
  print(plot_list[[pathway]])
  ggsave(file.path(OUTPUT_DIR, paste0("S2E_H_heat_enrichr_", pathway,".pdf")), plot = plot_list[[pathway]], width = 60, height = 32, units = "mm", bg = "white",  dpi = 600)

}
```



# --- Panel 2D. Signature Score Violin Plot in Adipocytes - CamaraH ---
## --- Pivot signatures ---
```{r}
# Set signatures of interest. This is dictate plotting order
signatures_order <- c("HEAT", "FA METABOLISM", "OXPHOS", "BATLAS")

graph.data <- sigscores_df %>%
  filter(dataset == "CamaraH") %>%
  mutate(
    Hallmark = str_replace_all(Hallmark, "HALLMARK_", "H: "),
    Hallmark = str_replace_all(Hallmark, "_", " "),
    Hallmark = str_replace_all(Hallmark, "SIGNATURE", ""),
    Hallmark = case_when(
      str_detect(Hallmark, "OXIDATIVE PHOSPHORYLATION") ~ "H: OXPHOS",
      str_detect(Hallmark, "FATTY ACID") ~ "H: FA METABOLISM",
      str_detect(Hallmark, "G2M") ~ "H: G2M",
      str_detect(Hallmark, "REACTIVE OXYGEN") ~ "H: ROS",
      str_detect(Hallmark, "UNFOLDED PROTEIN") ~ "H: UPR",
      TRUE ~ Hallmark
    )
  ) %>%
  mutate(Hallmark = str_remove(Hallmark, "H: ")) %>% # Fix the introduced problem of adding the H: to identify Hallmark pathways
  filter(Hallmark %in% signatures_order)

graph.data |> distinct(Hallmark)
sigscores_df |> distinct(Hallmark)

```
## --- Create plot data ---

```{r, fig.width=2.834646, fig.height= 2.214567}
plot_data <- graph.data %>%
  filter(
    str_detect(cell_type, "Adipocytes"),
    !str_detect(cell_type, "Pre")
  ) %>%
  mutate(Hallmark = str_wrap(Hallmark, 5)) %>%
  left_join(meta_simplified, by = "barcode") %>%
  mutate(Hallmark = factor(Hallmark,
    level = str_wrap(
      signatures_order, 5
    )
  ))
```

## --- Create plot ---
```{r, fig.width=2.834646, fig.height= 2.214567}
# Violin in Adipocytes
gg_vision_adipo_vln <- ggplot(
  plot_data,
  aes(x = cell_type_short, y = sigScores, fill = cell_type_short)
) +
  geom_jitter(aes(color = cell_type_short), width = 0.2, alpha = 0.5, stroke = 0, size = 1) +
  geom_violin(trim = FALSE, color = "black", linewidth = 0.3) +
  stat_compare_means(
    method = "t.test",
    comparisons = list(c("BAds", "WAds")),
    label = "p.signif",
    symnum.args = list(
      cutpoints = c(0, 0.001, 0.05, 1),
      symbols   = c("<0.001", "<0.05", "ns")
    ),
    size = 5 / 2.81, 
    vjust = -0.3, 
  ) +
  theme_nature_metabolism() +
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_text(size = 6),
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "Signature score",
    fill = NULL,
    color = NULL
  ) +
  scale_fill_manual(values = palette.use) +
  scale_color_manual(values = palette.use) +
  scale_y_continuous(expand = expansion(c(0.05, 0.2))) +
  facet_wrap(vars(Hallmark), ncol = 5)
```

## --- Plotting and save ---
```{r, fig.width=3.149606, fig.height= 1.13189}
mm_to_in(c(80, 28.75))
gg_vision_adipo_vln

scale <- 180
ggsave(file.path(OUTPUT_DIR, "2D_vision_adipo_vln.pdf"), plot = gg_vision_adipo_vln, width = 80, height = 28.75, units = "mm", bg = "white",  dpi = 600)
```
# --- Supp_Table - Cohen's effect size (d) ---
```{r}
# -----------------------------
#  Calculate effect sizes 
# -----------------------------

library(dplyr)
library(effsize)

results_list <- list()

for (hallmark in unique(plot_data$Hallmark)) {

  df <- plot_data |>
    filter(Hallmark == hallmark) |>
    select(score = sigScores, cell_type = cell_type_short)

  # Means
  stats <- df %>%
    group_by(cell_type) %>%
    summarise(mean = mean(score), sd = sd(score), n = n(), .groups = "drop")

  mean_bad <- stats$mean[stats$cell_type == "BAds"]
  mean_wad <- stats$mean[stats$cell_type == "WAds"]
  
        n_bad <- stats$n[stats$cell_type == "BAds"]
  n_wad <- stats$n[stats$cell_type == "WAds"]


  # Effect size
  eff <- cohen.d(score ~ cell_type, data = df)

  results_list[[hallmark]] <- data.frame(
    feature = "gene_signature",
    feature_name = hallmark,    
    cohen_d = eff$estimate,
    CI_low = eff$conf.int[1],
    CI_high = eff$conf.int[2],
    mean_BAds = mean_bad,
    mean_WAds = mean_wad,
         n_BAds = n_bad,
    n_WAds = n_wad,
    delta_mean = mean_bad - mean_wad,
    level = "nuclei"

  )
  # -----------------------------
  #  Sample-level effect size 
  # -----------------------------
  
  df_sample_unfiltered <- plot_data |>
  filter(Hallmark == hallmark) |>
  group_by(sample, cell_type_short) |>
  summarise(score = mean(sigScores), .groups = "drop") |> 
  rename("cell_type" = "cell_type_short")
  
  # -----------------------------
#  Filter data 
# -----------------------------
  df_sample <- df_sample_unfiltered |>
    left_join(
      plot_data |>
          filter(Hallmark == hallmark) |>
        group_by(sample) |>
        count(cell_type_short),
      by = c("sample", "cell_type" = "cell_type_short")
    ) |>
    filter(n > 10)

  
  # Means
  stats <- df_sample %>%
    group_by(cell_type) %>%
    summarise(mean = mean(score), sd = sd(score), n = n(), .groups = "drop")

  mean_bad <- stats$mean[stats$cell_type == "BAds"]
  mean_wad <- stats$mean[stats$cell_type == "WAds"]
  
      n_bad <- stats$n[stats$cell_type == "BAds"]
  n_wad <- stats$n[stats$cell_type == "WAds"]

  # Sample effect size
  eff <- cohen.d(score ~ cell_type, data = df_sample)

  results_list[[paste0(hallmark,"_sample")]] <- data.frame(
    feature = "gene_signature",
    feature_name= hallmark,    
    cohen_d = eff$estimate,
    CI_low = eff$conf.int[1],
    CI_high = eff$conf.int[2],
    mean_BAds = mean_bad,
    mean_WAds = mean_wad,
         n_BAds = n_bad,
    n_WAds = n_wad,
    delta_mean = mean_bad - mean_wad,
    level = "sample"
  )
  
  
}

eff_size_table <- bind_rows(results_list)
eff_size_table <- eff_size_table %>%
  arrange(level, desc(cohen_d))

eff_size_table <- eff_size_table %>%
  mutate(effect_size_label = case_when(
    cohen_d < 0.5 ~ "small",
    cohen_d < 0.8 ~ "medium",
    TRUE ~ "large"
  ))

eff_size_table

write_csv(eff_size_table, file.path(OUTPUT_DIR, "effect_size_wad_bad.csv"), col_names = TRUE)
```


# --- Panel 2E. Thermogenic Genes Violin Plot in Adipocytes - CamaraH ---
## --- Calculate stats ---
```{r}
stats_df <- FindMarkers(s.object, ident.1 = "WAds", ident.2 = "BAds", group.by = "cell_type_short", test.use = "wilcox")
```

## --- Add label information ---
```{r}
plot_stats <- stats_df %>%
  rownames_to_column("feature") %>%
  mutate(
    group2 = "WAds",
    group1 = "BAds",
    y.position = 5.5,
    label = case_when(
      p_val_adj < 0.001 ~ "<0.001",
      TRUE ~ as.character(round(p_val_adj, 2))
    ),
    xmin = 0.5,
    xmax = 1.5,
    colors = ifelse(p_val_adj < 0.05, "red", "black"),
  ) %>%
  select(feature, group1, group2, y.position, label, colors)
```

## --- Create Plot ---
```{r}
### Define thermo genes
thermo_genes <- c("EBF2", "PPARGC1A", "PRDM16", "UCP1")

plot_stats_filtered <- plot_stats %>% filter(feature %in% thermo_genes)

p <- VlnPlot(s.object, features = thermo_genes, group.by = "cell_type_short", stack = TRUE)
graph.data <- p$data
graph.data <- graph.data %>% filter(str_detect(ident, "WAds|BAds"))

# Get unique values of 'ident' (group names)
ident_levels <- as.character(unique(graph.data$ident))

# Create all pairwise comparisons dynamically
comparisons_list <- combn(ident_levels, 2, simplify = FALSE)

# Plot Violin
gg_thermo_genes_adipo_vln <- ggplot(graph.data, aes(x = ident, y = expression, fill = ident)) +
  geom_jitter(aes(color = ident), width = 0.1, alpha = 0.5, stroke = 0, size = 1) +
  geom_violin(trim = TRUE, color = "black", linewidth = 0.3, width = 1) +
  theme_nature_metabolism() +
  theme(
    legend.position = "none",
    axis.line = element_line(linewidth = 0.3),
    strip.background = element_blank(),
  ) +
  labs(
    title = NULL,
    x = NULL,
    y = "Gene expression",
    fill = NULL,
    color = NULL
  ) +
  scale_fill_manual(values = palette.use) +
  scale_color_manual(values = palette.use) +
  facet_wrap(vars(feature), ncol = 5) +
    new_scale_color() +

  stat_pvalue_manual(
    plot_stats_filtered,
    xmin = "group1",
    xmax = "group2",
    label = "label",
    y.position = "y.position",
    size = 5 / ggplot2::.pt,
    color = "colors",
    inherit.aes = FALSE # <- key line to prevent inheriting fill/color=ident
  ) +
  scale_y_continuous(expand = expansion(c(0.05, 0.20)))
```

## --- Plotting and save ---

```{r, fig.width=2.805118, fig.height=1.131890}
mm_to_in(c(71.25, 28.75))
gg_thermo_genes_adipo_vln

scale <- 180
ggsave(file.path(OUTPUT_DIR, "2E_thermo_genes_adipo_vln.pdf"), plot = gg_thermo_genes_adipo_vln, width = 71.25, height = 28.75, units = "mm", bg = "white",  dpi = 600)
```
# --- Supp_Table - Cohen's effect size (d) - Continued ---
```{r}
# -----------------------------
#  Load data 
# -----------------------------

PSEUDO_CPM_CTS_RDS <- here("output/4.downstream_pipelines/de_analysis/data/CamaraH_pseudobulk_count_matrix_for_ggplot.rds")
pseudobulk_exp <- read_rds(PSEUDO_CPM_CTS_RDS)
```


```{r}
# -----------------------------
#  Filter data 
# -----------------------------

pseudobulk_exp_filtered <- pseudobulk_exp |> 
  filter(cell_type_short %in% c("BAds", "WAds"),
         gene %in% thermo_genes,
         n_cells > 10) |> # Filter to a minimum of cells per sample
  select(feature = gene, expression = cpm_log, ident = cell_type_short, sample)
```


```{r}
# -----------------------------
#  Calculate Cohen's d 
# -----------------------------

library(dplyr)
library(effsize)

if(!exists("results_list")){
results_list <- list()
}

for (gene in thermo_genes) {

  df <- graph.data |>
    filter(feature == gene) |>
    select(score = expression, cell_type = ident)

  # Means
  stats <- df %>%
    group_by(cell_type) %>%
    summarise(mean = mean(score), sd = sd(score), n = n(), .groups = "drop")

  mean_bad <- stats$mean[stats$cell_type == "BAds"]
  mean_wad <- stats$mean[stats$cell_type == "WAds"]
  
    n_bad <- stats$n[stats$cell_type == "BAds"]
  n_wad <- stats$n[stats$cell_type == "WAds"]

  # Effect size
  eff <- cohen.d(score ~ cell_type, data = df)

  results_list[[gene]] <- data.frame(
    feature = "gene",
    feature_name = gene,    
    cohen_d = eff$estimate,
    CI_low = eff$conf.int[1],
    CI_high = eff$conf.int[2],
    mean_BAds = mean_bad,
    mean_WAds = mean_wad,
         n_BAds = n_bad,
    n_WAds = n_wad,
    delta_mean = mean_bad - mean_wad,
    level = "nuclei"

  )
  
    # Sample effect size
    df_sample <- pseudobulk_exp_filtered |>
    filter(feature == gene) |>
    select(score = expression, cell_type = ident)

  # Means
  stats <- df_sample %>%
    group_by(cell_type) %>%
    summarise(mean = mean(score), sd = sd(score), n = n(), .groups = "drop")

  mean_bad <- stats$mean[stats$cell_type == "BAds"]
  mean_wad <- stats$mean[stats$cell_type == "WAds"]
  
  n_bad <- stats$n[stats$cell_type == "BAds"]
  n_wad <- stats$n[stats$cell_type == "WAds"]

  eff <- cohen.d(score ~ cell_type, data = df_sample)

  results_list[[paste0(gene,"_sample")]] <- data.frame(
    feature = "gene",
    feature_name = gene,    
    cohen_d = eff$estimate,
    CI_low = eff$conf.int[1],
    CI_high = eff$conf.int[2],
    mean_BAds = mean_bad,
    mean_WAds = mean_wad,
    n_BAds = n_bad,
    n_WAds = n_wad,
    delta_mean = mean_bad - mean_wad,
    level = "sample"

  )
}

eff_size_table <- bind_rows(results_list)
eff_size_table <- eff_size_table %>%
  arrange(level, desc(cohen_d))

eff_size_table <- eff_size_table %>%
  mutate(effect_size_label = case_when(
    cohen_d < 0.5 ~ "small",
    cohen_d < 0.8 ~ "medium",
    TRUE ~ "large"
  ))
```

## --- Save Supp Table - Cohen's d ---
```{r}
eff_size_table

write_csv(eff_size_table, file.path(OUTPUT_DIR, "effect_size_wad_bad.csv"), col_names = TRUE)

```
# --- Panel 3I-O. VISION HEAT ViolinPlot in different human snRNAseq ---
```{r}
# Read the Q99 percentile
wad_HEAT_percentile <- readRDS(HEAT_THRESHOLD_RDS)
```

## --- Create plot data ---
``` {r, fig.width = 7.086614, fig.height = 1.771654}
plot_data_sig_score <- sigscores_df %>%
  filter(
    Hallmark %in% c("HEAT"),
    str_detect(cell_type, "Adipo"),
    !str_detect(cell_type, "Adipoq"),
    !str_detect(tolower(cell_type), "pre|progenitor"),
    !str_detect(region, "Intermediate")
)
```

## --- Supp_Table - HEAT+ adipocyte percentage (%) ---
```{r}
## Calculate separating brown and white adipocytes
heat_positive_pct_per_cell <- plot_data_sig_score %>%
  count(HEAT_positive = sigScores > wad_HEAT_percentile, cell_type, region, dataset) %>%
  group_by(cell_type, region, dataset) %>%
  mutate(props = round(n / sum(n) * 100, 1)) %>%
  arrange(dataset, cell_type, region)

dir.create(here(OUTPUT_DIR, "..", "supplemental_tables"), showWarnings = F, recursive = T)
write_csv(heat_positive_pct_per_cell, 
          here(OUTPUT_DIR, "..", "supplemental_tables","heat_positive_pct_per_cell.csv")
          )

## Calculate aggregating all adipocytes
heat_positive_pct_per_sample <- plot_data_sig_score %>%
  count(HEAT_positive = sigScores > wad_HEAT_percentile, dataset) %>%
  group_by(dataset) %>%
  mutate(props = round(n / sum(n) * 100, 1)) %>%
  arrange(dataset)

write_csv(heat_positive_pct_per_cell, here(OUTPUT_DIR, "..", "supplemental_tables","heat_positive_pct_per_dataset.csv"))
```


## --- Format plot data ---
``` {r, fig.width = 7.086614, fig.height = 1.771654, warning = FALSE}
library(dplyr)
library(ggplot2)
library(purrr)
library(patchwork)
library(ggpubr)

plot_data_sig_score$dataset <- factor(plot_data_sig_score$dataset,
  levels = c("CamaraH","AngueiraA","SunW","WangT","EfthymiouV","LazarescuO","Emont")
)

# Add plotting information
plot_data_sig_score <- plot_data_sig_score |> 
  mutate(publication = case_when(dataset == "CamaraH" ~ "This Study",
                                            dataset == "AngueiraA" ~ "Angueira et al. 2021",
                                            dataset == "SunW" ~ "Sun et al. 2020",
                                            dataset == "WangT" ~ "Wang et al. 2024",
                                            dataset == "EfthymiouV" ~ "Efthymiou et al. 2025",
                                            dataset =="LazarescuO" ~ "Lazarescu et al. 2025",
                                            dataset == "Emont" ~ "Emont et al. 2022",
                                            TRUE ~ "Unknown"),
          depot = case_when(dataset == "CamaraH" ~ "Cervical",
                                            dataset == "AngueiraA" ~ "Perivascular",
                                            dataset == "SunW" ~ "Cervical",
                                            dataset == "WangT" ~ "Cervical",
                                            dataset == "EfthymiouV" ~ "Abdominal",
                                            dataset =="LazarescuO" ~ "Abdominal",
                                            dataset == "Emont" ~ "Abdominal",
                                            TRUE ~ "Unknown"),
          cell_type_Region = interaction(cell_type, region_label, sep = " | ", drop = TRUE),
         region_extension = str_remove(region_label, " .*"),
         cell_type = case_when(str_detect(cell_type, "White") ~ "WAds",
                               str_detect(cell_type, "Brown") ~ "BAds",
                               TRUE ~ cell_type)) 

# Create palette
fill_color_palette <- plot_data_sig_score %>%
  distinct(cell_type_dataset, fill_color) %>%
  drop_na(fill_color) %>%
  deframe()


# Global y limits so every plot matches
y_lim <- range(plot_data_sig_score$sigScores, na.rm = TRUE)
```

## --- Set helper function to merge plots  ---
``` {r, fig.width = 7.086614, fig.height = 1.771654, warning = FALSE}
make_one <- function(d) {
  df <- plot_data_sig_score %>% filter(dataset == d)
  title_txt <- paste(unique(df$publication), collapse = ", ")
  subtitle_txt <- paste(unique(df$depot), collapse = ", ")

  ggplot(df, aes(x = region_extension, y = sigScores, fill = cell_type_dataset)) +
    geom_rect(
      data = data.frame(ymin = wad_HEAT_percentile),
      aes(ymin = ymin),
      xmin = -Inf, xmax = Inf, ymax = Inf,
      fill = "indianred1", alpha = 0.1,
      inherit.aes = FALSE
    ) +
    geom_rect(
      data = data.frame(ymax = wad_HEAT_percentile),
      aes(ymax = ymax),
      xmin = -Inf, xmax = Inf, ymin = -Inf,
      fill = "grey", alpha = 0.1,
      inherit.aes = FALSE
    )  +
  geom_jitter(aes(color = cell_type_dataset),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.6),
    shape = 21,
    alpha = 0.5,
    stroke = 0.1,
    size = 1
  ) +
    geom_violin(trim = FALSE, color = "black", width = 0.8, linewidth = 0.3) +
    geom_hline(yintercept = wad_HEAT_percentile, color = "firebrick",
               linetype = "dashed", linewidth = 0.2) +
    stat_compare_means(
      method = "t.test",
      comparisons = list(c("Brown Adipocytes", "White Adipocytes")),
      label = "p.format", size = 4 / 2.81
    ) +
    facet_wrap(vars(cell_type), scales = "free_x") +
    theme_nature_metabolism() +
    theme(
      legend.position = "right",
      strip.background = element_blank(),
      strip.text = element_text(face = "plain", size = 4),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    ) +
    labs(
      title = subtitle_txt,
      subtitle = title_txt,
      x = NULL,
      y = "Signature score",
      fill = NULL,
      color = NULL
    ) +
    scale_fill_manual(values = fill_color_palette) +
    scale_color_manual(values = fill_color_palette) +
    scale_y_continuous(limits = y_lim, 
                       expand = expansion(c(0.25,0.25)), 
                       # breaks = round(c(-1, 0, as.numeric(wad_HEAT_percentile), 2, 3, 4), 2)
                       )
} + guides(color = "none", fill = "none")
```

## --- Plotting and save ---
``` {r, fig.width = 7.086614, fig.height = 1.771654, warning = FALSE}
mm_to_in(c(180,45))
plots <- levels(plot_data_sig_score$dataset) %>%
  set_names() %>%
  map(make_one)

x_counts <- plot_data_sig_score %>%
  filter(!is.na(cell_type_Region)) %>%
  group_by(dataset) %>%
  summarise(n_x = n_distinct(cell_type_Region), .groups = "drop")

plots <- levels(plot_data_sig_score$dataset) %>% map(make_one)

widths <- x_counts$n_x
widths <- c(4.8,2,2, 2.2, 4.5, 4.5,4.5)
gg_heat_vlnplot_alldataset <- wrap_plots(plots, nrow = 1, widths = widths) +
  theme(plot.tag = element_text(face = "bold", size = 6))

gg_heat_vlnplot_alldataset

ggsave(file.path(OUTPUT_DIR, "3I-O_heat_vlnplot_alldataset.pdf"), plot = gg_heat_vlnplot_alldataset, width = 180, height = 45, units = "mm", bg = "white",  dpi = 600)

```

# --- Panel S3D-I. VISION HEAT FeaturePlot in different human snRNAseq ---
```{r}
library(dplyr)
library(ggplot2)

publication_levels = c("Angueira et al. 2021", "Sun et al. 2020", "Wang et al. 2024","Efthymiou et al. 2025", "Lazarescu et al. 2025","Emont et al. 2022")

top_frac <- 0.005      # top 0.5% as "signature+"
rad_q    <- 0.90      # circle contains 90% of those signature+ cells
pad      <- 2.5      # inflate radius by 15%

df <- sigscores_df %>%
  filter(Hallmark == "HEAT", publication %in% publication_levels, str_detect(cell_type, "^Adipocyte"), str_detect(cell_type, "Progenitor", negate = T))

# Build circle parameters per dataset
circles <- df %>%
  group_by(publication) %>%
  mutate(thr = quantile(sigScores, 1 - top_frac, na.rm = TRUE)) %>%
  filter(sigScores >= thr) %>%
  summarise(
    cx = median(umap_1, na.rm = TRUE),
    cy = median(umap_2, na.rm = TRUE),
    r  = quantile(sqrt((umap_1 - median(umap_1, na.rm = TRUE))^2 +
                       (umap_2 - median(umap_2, na.rm = TRUE))^2),
                  rad_q, na.rm = TRUE) * pad,
    .groups = "drop"
  ) |> 
  filter(publication %in% c("Angueira et al. 2021", "Efthymiou et al. 2025", "Sun et al. 2020"))
  

# Turn circle params into drawable paths
circle_paths <- circles %>%
  tidyr::crossing(t = seq(0, 2*pi, length.out = 361)) %>%
  mutate(
    x = cx + r * cos(t),
    y = cy + r * sin(t)
  )

# Re-filter
df <- sigscores_df %>%
  filter(Hallmark == "HEAT", dataset != "CamaraH")  |> 
  mutate(publication = factor(publication, levels = ))

#Comment in if you want to have higher HEAT cells on top
df <- df |> arrange(sigScores) 

gg_heat_featplot_alldataset <- ggplot(df, aes(x = umap_1, y = umap_2, color = sigScores)) +
  geom_point(size = 0.5, alpha = 0.7, stroke = 0) +
  geom_path(
    data = circle_paths,
    aes(x = x, y = y, group = publication),
    inherit.aes = FALSE,
    linewidth = 0.4,
    color = "black"
  ) +
  scale_color_viridis(
    option = "viridis",
    limits = c(-0.3, 2.6)
  ) +
  labs(
    x = "UMAP Dimension 1",
    y = "UMAP Dimension 2",
    color = "HEAT Signature\nScore (VISION)"
  ) +
  theme_nature_metabolism() +
  theme_no_axis() +
  facet_wrap(vars(publication), ncol = length(sigscore_list), scales = "free")


gg_heat_featplot_alldataset
```
## --- Plotting and save ---
``` {r, fig.width = 6.377953, fig.height = 1.259843}
mm_to_in(c(162, 32))
gg_heat_featplot_alldataset

gg_heat_featplot_alldataset <- gg_heat_featplot_alldataset +
  scale_color_gradientn(
  colours = c("#F7F7F7", "#BDBDBD", "#6A51A3", "#D7301F"),
  limits = c(-0.3, 2.6)
)

gg_heat_featplot_alldataset

ggsave(file.path(OUTPUT_DIR, "S3D-I_heat_featplot_alldataset.pdf"), plot = gg_heat_featplot_alldataset, width = 180, height = 32, units = "mm", bg = "white",  dpi = 600)

```

# --- Panel S3J. In vitro HEAT singature expression ---
## Load heat signature
```{r}
load(HEAT_SIGNATURE_RDATA) #Creates HEAT_signature
```


## --- A38 Cells - Plotting and save ---
``` {r fig.height = 2, fig.width = 3.5}
library(openxlsx)
a38.cells <- read.xlsx(A38_CTS_CSV, rowNames = F) %>%
  dplyr::select(-1, -transcript_id.s.)
names(a38.cells) <- gsub("\\.", " ", names(a38.cells))

a38.de.data <- a38.cells[!duplicated(a38.cells$Symbol) & !duplicated(a38.cells$Symbol, fromLast = TRUE), ] %>%
  select(-1) %>%
  remove_rownames() %>%
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

a38.metadata <- data.frame(
  samples = colnames(a38.de.data),
  separate = colnames(a38.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-") %>%
  mutate(Type = ifelse(Type == "A38 W", "A38 WAT", Type))

heatmap <- as.matrix(a38.cells %>%
  dplyr::filter(Symbol %in% c("UCP1", HEAT_signature)) %>%
  select(
    Symbol, 
    "A38 WAT-D18-1", "A38 WAT-D18-2", "A38 WAT-D18-3",
    "A38 BAT-D18-1", "A38 BAT-D18-2", "A38 BAT-D18-3"
  ) %>%
  column_to_rownames(var = "Symbol"))

# Filter metadata to keep only selected samples
a38.metadata <- a38.metadata[a38.metadata$samples %in% colnames(heatmap),]

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S3J_A38_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))

pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 7,
  gaps_row = 1,
  gaps_col = c(3),
  legend = TRUE,
  annotation_col = a38.metadata %>%
    remove_rownames() %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "A38 WAT", "A38 hWAd", "A38 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D6", "D9", "D12", "D15", "D18"))),
  annotation_colors = list("Diff_Day" = c("D0" = "grey", "D18" = "black"),
                           "Type" = c("A38 hBAd" = "#b3575e", "A38 hWAd" = "#313594")),
  scale = "row"
)

dev.off()

grep("COB", a38.cells$Symbol, value = T)
a38.cells |>  filter(Symbol %in% custom.order) |> select("Symbol", contains("D18"))

a38.cells |> select(contains("A38")) |> colSums(na.rm = TRUE)
```

## --- A41 Cells - Plotting and save ---
``` {r}
a41.cells <- read.xlsx(A41_CTS_CSV, rowNames = FALSE) %>%
  dplyr::select(-gene_id, -transcript_ids, -Species, -Name, -Description)

a41.de.data <- a41.cells[!duplicated(a41.cells$Symbol) & !duplicated(a41.cells$Symbol, fromLast = TRUE), ] %>%
  remove_rownames() |> 
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

a41.metadata <- data.frame(
  samples = colnames(a41.de.data),
  separate = colnames(a41.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-")

heatmap <- a41.cells %>%
  dplyr::filter(Symbol %in% c("UCP1", HEAT_signature)) %>%
  select(Symbol, 
          "EV-D18-1_FPKM", "EV-D18-2_FPKM", "EV-D18-3_FPKM", 
          "BAT-D18-1_FPKM", "BAT-D18-2_FPKM", "BAT-D18-3_FPKM") %>% 
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S3J_A41_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))
pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 6,
  gaps_row = 1,
  gaps_col = c(3),
  legend = TRUE,
  annotation_col = a41.metadata %>%
    filter(samples %in% colnames(heatmap)) %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "EV", "A41 hWAd", "A41 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D9", "D12", "D18"))),
    annotation_colors = list("Diff_Day" = c("D9" = "grey", "D18" = "black"),
                           "Type" = c("A41 hBAd" = "#b3575e", "A41 hWAd" = "#313594")),

  scale = "row"
)
dev.off()
a41.cells |>  filter(Symbol %in% custom.order) |> select("Symbol", contains("D18")) |> select(!contains("OE"))

```

## --- Cero et al. 2023 - Plotting and save ---
```{r}
cero <- read.xlsx(CERO_CTS_XLSX)

cero_mtx <- cero %>%
  column_to_rownames(var = "Name") %>%
  select(contains("d30.CPM")) %>%
  as.matrix()

cero_metadata <- data.frame(
  samples = colnames(cero_mtx),
  separate = colnames(cero_mtx)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Count_Type", "Replicate"), sep = "\\.")

heatmap <- cero_mtx 

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S3J_cero_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))
pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 6,
  gaps_row = 1,
  gaps_col = c(4),
  legend = TRUE,
  annotation_col = cero_metadata %>%
    filter(samples %in% colnames(heatmap)) %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Diff_Day = factor(toupper(Diff_Day), levels = c("D0", "D30"))),
    annotation_colors = list("Diff_Day" = c("D0" = "grey", "D30" = "black"),
                           "Type" = c("hBA" = "#b3575e", "hWA" = "#313594")),

  scale = "row"
)
dev.off()
```

# --- Panel S3K-M. Angueira Adipocyte Subset ---
## --- Load files ---
```{r}
# Load Seurat
s.object.angueira.sub <- readRDS(ANGUEIRA_ADIPO_SUB_RDS)

# Load UCP1 correlation
ucp1_v_sigscores <- read_csv(ANGUEIRA_HEATvUCP1_CSV)
```

## --- Setup common plotting parameters ---
```{r, fig.width=5.708661, fig.height=1.265467}
mm_to_in(c(145, 225 / 7))
# Assing plot dimensions
width <- 30
height <- 25

# Rename adipocytes
meta <- s.object.angueira.sub@meta.data
meta <- meta |> mutate(
  adipo_cell_type = paste0("hAd", seurat_clusters),
  adipo_cell_type = factor(adipo_cell_type, levels = paste0("hAd", 0:4))
)
s.object.angueira.sub@meta.data <- meta

# Reassign idents
Idents(s.object.angueira.sub) <- "adipo_cell_type"
```

## ---- Set color palette ----
```{r}
# Select a monochromatic palette
pal <- brewer.pal(5, "Oranges")
```

# --- Panel S3K. Adipocyte UCP1 x HEAT correlation - Angueira et al 2021 ---
## --- Plotting and save ---
```{r, fig.width=1.417323, fig.height=1.968504}
mm_to_in(c(width, height*2))
# Filter df to run correlation only in cells expressing UCP1
df_filt <- ucp1_v_sigscores |> filter(UCP1 > 0)
pal <- brewer.pal(3, "Oranges")
gg_angueira_correlation <- ggplot() +
  geom_point(
    data = ucp1_v_sigscores,
    aes(UCP1, sigScores),
    size = 1, stroke = 0.3, alpha = 0.25, color = pal[1]
  ) +
  geom_smooth(
    data = df_filt,
    aes(UCP1, sigScores),
    method = "lm", se = TRUE, color = pal[3], fill = pal[3], linewidth = 0.3
  ) +
  stat_cor(
    data = df_filt,
    aes(UCP1, sigScores),
    method = "pearson",
    digits = 2,
    size = 6 / .pt,
    color = "black",
    label.x = -Inf, label.y = Inf, hjust = -0.1, vjust = 1.2,
    p.accuracy = 0.01
  ) +
  theme_nature_metabolism() +
  labs(
    title = "HEAT Score x UCP1",
    subtitle = "Angueira et al. 2021 - Adipocytes",
    y = str_wrap("HEAT Signature Score (VISION)", width = 20),
    x = "UCP1 expression"
  ) +
  scale_y_continuous(expand = expansion(c(0.05, 0.1)))

gg_angueira_correlation


ggsave(file.path(OUTPUT_DIR, "S3K_heat_ucp1_corr_anguieira.pdf"), bg = "white", width = width, height = height, units = "mm", dpi = 600)
```
## --- Register N  for correlations ---
```{r}
s.object.angueira.sub@meta.data |> count(adipo_cell_type)

corr_used_adipocytes <- ucp1_v_sigscores |> count(cell_type, UCP1 > 0)
corr_used_adipocytes
```

# --- Panel S3L. Adipocyte UMAP - Angueira et al 2021 ---
## --- Plotting and save ---
```{r, fig.width=1.417323 * 4, fig.height=0.984252 * 4}
mm_to_in(c(width, height))

scale_factor <- 4
gg_angueira_umap <- UMAPPlot(s.object.angueira.sub, cols = pal, pt.size = 0.1) + labs(title = "Angueira et al. 2021", subtitle = "Adipocyte subtypes") + theme_nature_metabolism(base_size = 6 * scale_factor) + theme_no_axis()
ggsave(file.path(OUTPUT_DIR, "S3L_angueira_adipo_subset_umap.pdf"), plot = gg_angueira_umap, bg = "white", width = width * scale_factor, height = height * scale_factor, units = "mm", dpi = 600)

gg_angueira_umap
```
# --- Panel S3M. Adipocyte ViolinPlot - Angueira et al 2021 ---
## --- Plotting and save ---
```{r, fig.width=1.417323, fig.height=0.984252}
gg_angueira_violin <- VlnPlot(s.object.angueira.sub, features = c("ADIPOQ", "UCP1", "HEAT"), flip = T, stack = T, fill.by = "ident", cols = pal) + theme_nature_metabolism() + theme(
  strip.text = element_text(size = 3),
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  axis.title.x = element_blank()
) +
  NoLegend()
ggsave(file.path(OUTPUT_DIR, "S3M_angueira_adipo_subset_feat_heat.pdf"), plot = gg_angueira_violin, bg = "white", width = width, height = height, units = "mm", dpi = 600)
gg_angueira_violin
```

## --- Supp_Table - HEAT+ adipocyte percentage (%) - Continued ---
```{r}
# Count proportions of heat-positive cells per subcluster
meta <- s.object.angueira.sub@meta.data
angueira_heat_positive_pct <- meta %>%
  count(heat_positive, adipo_cell_type) %>%
  group_by(adipo_cell_type) %>%
  mutate(props = round(n / sum(n) * 100, 1))

angueira_heat_positive_pct

# Harmonize to match with heat_positive_pct_per_cell (Calculated above)
colnames(heat_positive_pct_per_cell)
colnames(angueira_heat_positive_pct)

angueira_heat_positive_pct <- angueira_heat_positive_pct |>
  rename(
    "HEAT_positive" = "heat_positive",
    "cell_type" = "adipo_cell_type"
  ) |>
  mutate(region = "perivasc", dataset = "AngueiraA")

heat_positive_pct_per_cell_merged <- rbind(heat_positive_pct_per_cell, angueira_heat_positive_pct) |> arrange(dataset, region, cell_type)
heat_positive_pct_per_cell_merged

# Save
dir.create(here(OUTPUT_DIR, "..", "supplemental_tables"), showWarnings = F, recursive = T)
write_csv(heat_positive_pct_per_cell_merged, here(OUTPUT_DIR, "..", "supplemental_tables","heat_positive_pct_per_cell.csv"))

```
# --- Panel S3N-O. VISION in mouse dataset ---
## --- Load files ---
```{r}
# Load files
sigscore_shamsi <- read_csv(SHAMSI_VISON_SIG)
shamsi_rds <- readRDS(SHAMSI_RDS)

# Merge signature in the object
sigscore_shamsi <- sigscore_shamsi |> filter(Hallmark == "Mouse_HEAT")
if (all(sigscore_shamsi$barcode == rownames(shamsi_rds@meta.data))) {
  shamsi_rds$mouse_heat <- sigscore_shamsi$sigScores
}

ucp1_shamsi <- FetchData(shamsi_rds, vars = "Ucp1", layer = "data")
ucp1_v_sigscores <- sigscore_shamsi |> left_join(ucp1_shamsi |> rownames_to_column("barcode"), by = "barcode")
```
## --- Add non-adipocyte subgroup ---
```{r}
shamsi_rds@meta.data <- shamsi_rds@meta.data |> mutate(cell_type_custom = ifelse(cell_type =="Adipocytes", "Adipocytes", "Non-Adipocytes"))
```

## --- Define color palettes ---
```{r}
shamsi_palette <- c(
  "Pdgfra_APC" = "#8bbe9b",
  "Lymph_EC" = "#cc9fd7",
  "EC" = "#dfa9c1",
  "Macrophages" = "#0461cf",
  "Pericytes" = "#e0641c",
  "VSM" = "#fbb294",
  "NMSC" = "#f02943",
  "MSC" = "#f79399",
  "Adipocytes" = "#ffe65a",
  "Non-Adipocytes" = "grey75"
)

shamsi_palette_low  <- colorspace::lighten(shamsi_palette, amount = 0.60)
shamsi_palette_high <- colorspace::darken(shamsi_palette,  amount = 0.15)
```


# --- Panel S3N. Adipocyte UCP1 x HEAT correlation - Shamsi et al 2023 ---
## --- Plotting and save ---
```{r, fig.width=1.377953, fig.height=1.265467}
ucp1_v_sigscores
mm_to_in(c(35, 225 / 7))
# Filter df to run correlation only in cells expressing UCP1
df_filt <- ucp1_v_sigscores |> filter(Ucp1 > 0, cell_type %in% c("Brown adipocytes", "Adipocytes"), Hallmark == "Mouse_HEAT")
pal <- brewer.pal(3, "Oranges")
gg_shamsi_correlation <- ggplot() +
  geom_point(
    data = df_filt,
    aes(Ucp1, sigScores, ),
    size = 1, stroke = 0.3, alpha = 0.25,
    color = shamsi_palette_low[["Adipocytes"]]
  ) +
  geom_smooth(
    data = df_filt,
    aes(Ucp1, sigScores),
    method = "lm", se = TRUE, linewidth = 0.3, color = shamsi_palette[["Adipocytes"]], fill = shamsi_palette[["Adipocytes"]]
  ) +
  stat_cor(
    data = df_filt,
    aes(Ucp1, sigScores),
    method = "pearson",
    digits = 2,
    size = 6 / .pt,
    color = "black", 
    label.x = -Inf, label.y = Inf, hjust = -0.1, vjust = 1.2, p.accuracy = 0.01) +
  theme_nature_metabolism() +
  labs(
    title = "HEAT Score x UCP1",
    subtitle = "Shamsi et al. 2023 - Mouse Adipocytes",
    y = str_wrap("HEAT Signature Score (VISION)",width = 20),
    x = "UCP1 expression"
  ) +
  scale_y_continuous(expand = expansion(c(0.05, 0.1))) +
  NoLegend()

gg_shamsi_correlation


ggsave(file.path(OUTPUT_DIR, "S3N_heat_ucp1_corr_shamsi.pdf"), bg = "white", width = width, height = height, units = "mm", dpi = 600)
```
```{r}
shamsi_rds@meta.data |> count(cell_type_custom)

corr_used_adipocytes <- ucp1_v_sigscores |> filter(cell_type %in% c("Brown adipocytes", "Adipocytes"), Hallmark == "Mouse_HEAT") |> count(Ucp1 > 0)
corr_used_adipocytes
```

# --- Panel S3O. Adipocyte UMAP - Shamsi et al 2023 ---
## --- Plotting and save ---
```{r, fig.width=3.779528, fig.height=2.952756}
mm_to_in(c(width*3, height*3))
# rEassing identities
Idents(shamsi_rds) <-"cell_type_custom"

# Plot
a <- UMAPPlot(shamsi_rds, pt.size = 0.5, cols = shamsi_palette) + theme_nature_metabolism(base_size = 6 *3) + theme_no_axis() + labs(title = "Shamsi et al. 2023", subtitle = "Mouse iBAT") + theme(legend.position = "bottom", legend.text = element_text(size = 5 * 3))
a
# Save
ggsave(file.path(OUTPUT_DIR, "S3O_shamsi_umap.pdf"), plot = a, bg = "white", width = width*3, height = height*3, units = "mm", dpi = 600)
```


# --- Panel S3P. Adipocyte ViolinPlot - Shamsi et al 2023 ---
## --- Plotting and save ---
```{r, fig.width=1.417323, fig.height=1.259843}
mm_to_in(c(36,32))
if("mouse_heat" %in% colnames(shamsi_rds@meta.data)){
shamsi_rds@meta.data <- shamsi_rds@meta.data |> rename("mHEAT" = "mouse_heat")
}

Idents(shamsi_rds) <- factor(Idents(shamsi_rds), levels = c("Adipocytes", "Non-Adipocytes"))

d <- VlnPlot(shamsi_rds, features = c("Adipoq", "Ucp1", "mHEAT"), flip = T, stack = T, fill.by = "ident", cols = shamsi_palette) + theme_nature_metabolism() + theme(
  strip.text = element_text(size = 3),
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
  axis.title.x = element_blank()
) +
  NoLegend()
d

ggsave(file.path(OUTPUT_DIR, "S3P_shamsi_violin.pdf"), plot = d, bg = "white", width = width, height = height, units = "mm", dpi = 600)

```

# --- Panel 3A-D. ssGSEA vs UCP1 in bulkRNAseq ---
## --- Load the ssGSEA data
```{r}
#--- Determine files to loop over
# Results files
files <- list.files(GSEA_HEAT_BULK_DIR, full.names = TRUE)
SSGSEA_FILES <- grep("HEATvUCP1.csv", files, value = TRUE)


# Define dataset names
dataset_names <- c("Castella", "Din", "Gavrila", "Giroud", "Ohja")
```

```{r}
## Create data frame to map names
dataset_keys <- data.frame(
  dataset_name = c("Castella", "Din", "Gavrila", "Giroud", "Ohja"),
  full_name = c("Castella et al. 2023", "Din et al. 2018", "Salej Duran et al. 2025", "Giroud et al. 2023", "Ojha et al. 2016")
) |> column_to_rownames("dataset_name")
```

## --- Define Color Palettes ---
```{r}
neck.color
dev_age_color <- c("Child" = "orchid4", "Infant" = "orchid3", "Neonate" = "orchid1")
pheo_color <- c("control" = "grey25", "pheochromocytoma" = "darkgreen", "Control" = "grey25", "Pheo" = "darkgreen")

color_palette_gsea_bulk <- c(neck.color, dev_age_color, pheo_color)
```

## --- Plotting ---
```{r}
corrplot_list <- list()
plot_data_list <- list()

# Filter out Sondergaard
dataset_names <- grep("Sondergaard", dataset_names, ignore.case = TRUE, invert = TRUE, value = TRUE)
# Loop over each sample and plot data
for (dataset in dataset_names) {
  ucp1vHEAT <- read.csv(str_subset(SSGSEA_FILES, dataset))
  ucp1vHEAT <- ucp1vHEAT %>%
    rename("group" = "X.") %>%
    mutate(
      group = str_replace(group, "pheochromocytoma", "Pheo"),
      group = str_replace(group, "control", "Control")
    )
  if (dataset == "ojha") {
    ucp1vHEAT <- ucp1vHEAT %>%
      mutate(group = factor(group, levels = c("Neonate", "Infant", "Child")))
  }
  
  plot_data_list[[dataset]] <- ucp1vHEAT |> count(group)
  write_csv(plot_data_list[[dataset]], file.path(OUTPUT_DIR, paste0(dataset,"_bulk_ssgsea_heat_n.csv")))

  # --- Plotting HEAT vs UCP1 ---
  gg_2 <- ggplot(
  ucp1vHEAT,
  aes(x = HEAT, y = log10(UCP1 + 1), color = group)
) +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    linetype = "dashed",
    linewidth = 0.1,
    se = FALSE,
    show.legend = FALSE
  ) +
  stat_cor(
    method = "pearson",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
    size = 4 / 2.81
  ) +
  geom_point(size = 1) +
  theme_nature_metabolism() +
  theme(legend.position = "top") +
  scale_color_manual(values = color_palette_gsea_bulk) +
  labs(
    x = "HEAT Score (ssGSEA)",
    y = expression(log[10](UCP1 + 1)),
    title = dataset_keys[dataset,]
  )
  
  
  print(gg_2)

  corrplot_list[[dataset]] <- gg_2
}
```

## --- Save plot ---
```{r, fig.width=7.086614, fig.height= 1.771654, warning=FALSE}
# Wrap and save
gg_heat_bulk_corrplot <- wrap_plots(corrplot_list[c("Gavrila", "Din", "Ohja", "Castella")], ncol = 4)
gg_heat_bulk_corrplot
scale <- 180
ggsave(file.path(OUTPUT_DIR, "3A-D_heat_ucp1_bulk_corrplot.pdf"), plot = gg_heat_bulk_corrplot, width = 180, height = 225 / 5, units = "mm", bg = "white",  dpi = 600)

```
# --- Panel 3E-H. ssGSEA vs CKMT2 in bulkRNAseq ---
```{r}
boxplots_list <- list()
corrplot_list <- list()

# Filter out Sondergaard
dataset_names <- grep("Sondergaard", dataset_names, ignore.case = TRUE, invert = TRUE, value = TRUE)
# Loop over each sample and plot data
for (dataset in dataset_names) {
  ucp1vHEAT <- read.csv(str_subset(SSGSEA_FILES, dataset))
  ucp1vHEAT <- ucp1vHEAT %>%
    rename("group" = "X.") %>%
    mutate(
      group = str_replace(group, "pheochromocytoma", "Pheo"),
      group = str_replace(group, "control", "Control")
    )
  if (dataset == "ojha") {
    ucp1vHEAT <- ucp1vHEAT %>%
      mutate(group = factor(group, levels = c("Neonate", "Infant", "Child")))
  }
  # --- Plotting HEAT vs CKMT2 ---
    gg_2 <- ggplot(
  ucp1vHEAT,
  aes(x = HEAT, y = log10(CKMT2 + 1), color = group)
) +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    linetype = "dashed",
    linewidth = 0.1,
    se = FALSE,
    show.legend = FALSE
  ) +
  stat_cor(
    method = "pearson",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
    size = 4 / 2.81
  ) +
  geom_point(size = 1) +
  theme_nature_metabolism() +
  theme(legend.position = "top") +
  scale_color_manual(values = color_palette_gsea_bulk) +
  labs(
    x = "HEAT Score (ssGSEA)",
    y = expression(log[10](CKMT2 + 1)),
    title = dataset_keys[dataset,]
  )
  
  
  print(gg_2)

  corrplot_list[[dataset]] <- gg_2
}
```

## --- Save plot ---
```{r, fig.width=7.086614, fig.height= 1.771654, warning=FALSE}
# Wrap and save
gg_heat_bulk_corrplot <- wrap_plots(corrplot_list[c("Gavrila", "Din", "Ohja", "Castella")], ncol = 4)
gg_heat_bulk_corrplot
scale <- 180
ggsave(file.path(OUTPUT_DIR, "3E-H_heat_ckmt2_bulk_corrplot.pdf"), plot = gg_heat_bulk_corrplot, width = 180, height = 225 / 5, units = "mm", bg = "white",  dpi = 600)

```

# --- Panel S3A. Heatmap HEAT signature - Sondergaard et al 2014 ---
```{r}
expression_matrix <- read_rds(PARAGANG_BULK_MTX_RDS)

pdf(file.path(OUTPUT_DIR, "S3A_paragang_heatmap.pdf"), height = mm_to_in(180), width = mm_to_in(45))
# Create heatmap
pheatmap::pheatmap(expression_matrix,
  scale = "row", # Normalize rows (genes)
  clustering_distance_rows = "euclidean", # Clustering for rows
  cellwidth = 5,
  cellheight = 5,
  fontsize = 6,
  clustering_distance_cols = "euclidean", # Clustering for columns
  clustering_method = "complete", # Hierarchical clustering method
  treeheight_row = 0.5, treeheight_col = 0.5
)
dev.off()

pheatmap::pheatmap(expression_matrix,
  scale = "row", # Normalize rows (genes)
  clustering_distance_rows = "euclidean", # Clustering for rows
  cellwidth = 5,
  cellheight = 5,
  fontsize = 6,
  clustering_distance_cols = "euclidean", # Clustering for columns
  clustering_method = "complete", # Hierarchical clustering method
)
```

# --- Panel S3B. Heatmap HEAT signature - Castella et al 2023 ---
## --- Format dataset names ---
```{r}
## Create data frame to map names
dataset_keys <- data.frame(
  dataset_name = c("Castella", "Din", "Gavrila", "Giroud", "Ohja"),
  full_name = c("Castella et al. 2023", "Din et al. 2018", "Salej Duran et al. 2025", "Giroud et al. 2023", "Ojha et al. 2016")
) |> column_to_rownames("dataset_name")
```

## --- Load HEAT_signature ---
```{r}
# Load HEAT signature
load(HEAT_SIGNATURE_RDATA) # This creates object "HEAT_signature"
```

## --- List files ---
```{r}
#--- Determine files to loop over
# Expression files
exp_files <- list.files(BULK_DIR, full.names = TRUE)
cleaned_files <- grep("_Cleaned.csv$", exp_files, value = TRUE)

#Metadata files
meta_files <- list.files(BULK_META_DIR, full.names = TRUE)

#Define dataset names
dataset_names <- c("Castella")
```

## --- Plotting and save ---
```{r}
for(dataset in dataset_names){
 
  # --- Read in the data ---
  object <- read.csv(str_subset(cleaned_files, dataset), header = TRUE, row.names = 1)
  
  meta <- read.csv(str_subset(meta_files, dataset))
  
  meta$group <- factor(meta$group, levels = unique(meta$group[order(meta$order)]))
  rownames(meta) <- meta$sample
  
  meta <- meta %>% select(group) |> mutate(group = case_when(
    grepl("WAT|SC", rownames(meta)) ~ "Superficial",
    grepl("BAT|Deep", rownames(meta)) ~ "Deep",
    TRUE ~ group
  ))

  
  # Reorder
  UCP1_exp <- object["UCP1", ] |> 
    t() |> 
    as.data.frame( ) |> 
    rownames_to_column("sample") |> 
    left_join(meta |> rownames_to_column("sample"), by = "sample") 
  col_order <- UCP1_exp |> 
    arrange(group, UCP1) |> 
    pull(sample)

  
  heatmap <- object[,col_order] %>%
    as.matrix()
  
  #Colors
  dev_age_color <- c("Child" = "orchid4", "Infant" = "orchid3", "Neonate" = "orchid1")
  pheo_color <- c("control" = "grey25", "pheochromocytoma" = "darkgreen", "Control" = "grey25", "Pheo" = "darkgreen")
  annotation_colors <- list(group = c(neck.color, dev_age_color, pheo_color))
  
  annotation_colors$group <- annotation_colors$group[names(annotation_colors$group) %in% meta$group]
  
  custom.order <- c("UCP1", sort(HEAT_signature))
  custom.order <- custom.order[custom.order %in% rownames(heatmap)]
  
  
  # -----------------------------
  #  Save plot as pdf 
  # -----------------------------
  
  pdf(file.path(OUTPUT_DIR, paste("S3B", dataset, "HEAT_heatmap.pdf", sep= "_")), width = mm_to_in(50), height = mm_to_in(90))
  # # --- Plot heatmap ---
  mat <- heatmap[custom.order, , drop = FALSE]

  print(pheatmap(mat,
           cluster_cols = FALSE,
           cluster_rows = FALSE,
           cellwidth = 6,
           cellheight = 5,
           fontsize = 6,
           gaps_row = 1,
           gaps_col = 3,
           show_colnames = FALSE,
           legend = TRUE,  
           annotation_legend = FALSE,
           annotation_col = meta,
           annotation_colors = annotation_colors,
           annotation_names_col = FALSE,
           scale = "row",
          main = dataset_keys[dataset,]
))
  dev.off()
  
  # -----------------------------
  #  Print in file 
  # -----------------------------
  
  print(pheatmap(mat,
           cluster_cols = FALSE,
           cluster_rows = FALSE,
           cellwidth = 6,
           cellheight = 5,
           fontsize = 6,
           gaps_row = 1,
           gaps_col = first_break,
           show_colnames = FALSE,
           legend = TRUE,  
           annotation_legend = FALSE,
           annotation_col = meta,
           annotation_colors = annotation_colors,
           annotation_names_col = FALSE,
           scale = "row",
          main = dataset_keys[dataset,]
))


}

```


# --- Panel S3C. CIBERSORT ---
## --- Load CIBERSORT data ---
```{r}
# Load output from X.02.X.CIBERSORTx
combined_data_with_meta <- read.csv(CIBERSORT_PROPS_CSV)
```

## --- Create plots ---
```{r}
# Plotting according to the datasets
plot_list <- list()
for (dataset in unique(combined_data_with_meta$Dataset)) {
  # Filter the data
  plot_data <- combined_data_with_meta %>%
    dplyr::filter(
      Cell_type == "Brown_adipocyte",
      Dataset == dataset
    )


  ## --- Define Color Palettes ---
  neck.color
  dev_age_color <- c("Child" = "orchid4", "Infant" = "orchid3", "Neonate" = "orchid1")
  pheo_color <- c("control" = "grey25", "pheochromocytoma" = "darkgreen", "Control" = "grey25", "Pheo" = "darkgreen")

  color_palette_gsea_bulk <- c(neck.color, dev_age_color, pheo_color)

  ## --- Format the plot title ---
  plot_title <-dataset_keys[dataset, ]
  # plot_title <- str_replace(plot_title, "et", "\net") #Scape if needed

  # Generate the plot
  p <- ggplot(plot_data, aes(x = Depot, y = Proportions, fill = Depot)) +
    # In case you want barplot + sem
    #    stat_summary(geom = "col", fun = "mean", aes(fill = Depot, color = Depot),  alpha = 0.5, width = 0.5, linewidth = 0.3) +
    # stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.15, linewidth = 0.3, aes(color = Depot)) +

    # In case you want boxplot
    geom_boxplot(alpha = 0.4, outlier.shape = NA, color = "black", width = 0.4, linewidth = 0.3) +
    geom_jitter(position = position_jitterdodge(jitter.width = 0), size = 0.5, shape = 21, color = "black", stroke = 0.3) + # Jittered points
    theme_classic() +
    theme(
      axis.title.x = element_blank(),
      plot.title = element_text(hjust = 0.5)
    ) +
    labs(
      title = plot_title,
      y = "Estimated BAds"
    ) +
    theme_nature_metabolism() +
    theme(
      axis.title.x = element_blank(),
      # axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    ) +
    scale_fill_manual(values = color_palette_gsea_bulk) +
    scale_color_manual(values = color_palette_gsea_bulk) +
    coord_cartesian(clip = "off") +
    NoLegend() +
    scale_y_continuous(expand = expansion(c(0.05, 0.25)))

  # --- Add statistics ---

  
  stat.test <- compare_means(
  Proportions ~ Depot,
  data = plot_data,
  method = "t.test",
  p.adjust.method = "holm"
  ) %>%
  mutate(
    p.format = signif(p.adj, 2),
    label = p.format,
    fontface = ifelse(p.adj <= 0.05, "bold", "plain"),
    y.position = max(plot_data$Proportions, na.rm = TRUE) * 1.15
  )

  p <- p +
    stat_pvalue_manual(
      stat.test,
      label = "label",
      size = 5 / ggplot2::.pt,
      inherit.aes = FALSE,
      fontface = "fontface"
    )

  plot_list[[dataset]] <- p
}
```

## --- Save plots ---
```{r, fig.width=1.181102, fig.height= 0.984252, warning=FALSE}
mm_to_in(c(30, 25))

for(nm in names(plot_list)){
gg_bad_cibersort <- wrap_plots(plot_list[c(
  "Gavrila", 
  "Din",
  "Castella"
  )], ncol = 3) 

print(plot_list[[nm]])

ggsave(file.path(OUTPUT_DIR, paste0("S3C_brown_ad_cibersort_", nm, ".pdf")), plot = plot_list[[nm]], width = 30, height = 25, units = "mm", bg = "white", dpi = 600)
}
```

# --- GTEx Correlations ----

## --- Load data ---
```{r}
meta_filtered <- read.csv(GTEX_META_CSV)
```

# --- Panel 3P. HEAT GTEx x BMI correlation --- 
```{r, fig.width= 1.417323, fig.height=1.771654}
# make sure tissue has exactly 2 levels
tissue_levels <- unique(meta_filtered$tissue)

# monochromatic blue palette
tissue_cols <- c(
  lighten("blue", 0.3),
  darken("blue", 0.3)
)
names(tissue_cols) <- tissue_levels

# Initialize plot ---
gg_heat_bmi_corr <- ggplot(meta_filtered, aes(BMI, HEAT, color = tissue, fill = tissue)) +
  geom_point(shape = 21, alpha = 0.2, stroke = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.3) +
  stat_cor(
    aes(group = tissue, color = tissue),
    method = "pearson",
    size = 6 / .pt, 
    show.legend = FALSE, 
    label.y = c(6.8, 5.8)
  ) +
  scale_color_manual(values = tissue_cols, name = "Tissue") +
  scale_fill_manual(values = tissue_cols, name = "Tissue") +
  theme_nature_metabolism() +
  theme(legend.position = "bottom") +
  labs(
    title = "HEAT Score x BMI",
    y = "HEAT Score (ssGSEA)"
  ) +
  coord_cartesian(clip = "off")
```

## --- Save plot ---
```{r, fig.width= 1.574803, fig.height=1.771654}
mm_to_in(c(40, 45))

# Expand axis and add tag
gg_heat_bmi_corr <- gg_heat_bmi_corr +
  scale_x_continuous(expand = expansion(0.1)) +
  scale_y_continuous(expand = expansion(0.1))

gg_heat_bmi_corr

# Save and output
ggsave(file.path(OUTPUT_DIR, "3P_heat_bmi_corr.pdf"), plot = gg_heat_bmi_corr, width = 40, height = 45, units = "mm", bg = "white",  dpi = 600)

```

# --- Panel 3Q. HEAT GTEx x Age correlation --- 
```{r, fig.width= 1.417323, fig.height=1.771654}
# make sure tissue has exactly 2 levels
tissue_levels <- unique(meta_filtered$tissue)

# monochromatic blue palette
tissue_cols <- c(
  lighten("red", 0.1),
  darken("red", 0.3)
)
names(tissue_cols) <- tissue_levels

# Initialize plot ---
gg_heat_age_corr <- ggplot(meta_filtered, aes(AGE, HEAT, color = tissue, fill = tissue)) +
  geom_point(shape = 21, alpha = 0.2, stroke = 0.2, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 0.3) +
  stat_cor(
    aes(group = tissue, color = tissue),
    method = "pearson",
    size = 6 / .pt, 
    show.legend = FALSE, , 
    label.y = c(6.8, 5.8),
  ) +
  scale_color_manual(values = tissue_cols, name = "Tissue") +
  scale_fill_manual(values = tissue_cols, name = "Tissue") +
  theme_nature_metabolism() +
  theme(legend.position = "bottom") +
  labs(
    title = "HEAT Score x Age",
    y = "HEAT Score (ssGSEA)"
  ) +
  coord_cartesian(clip = "off")
```

## --- Save plot ---
```{r, fig.width= 1.574803, fig.height=1.771654}
mm_to_in(c(40, 45))

# Expand axis and add tag
gg_heat_age_corr <- gg_heat_age_corr +
  scale_x_continuous(expand = expansion(0.1)) +
  scale_y_continuous(expand = expansion(0.1))

gg_heat_age_corr

# Save and output
ggsave(file.path(OUTPUT_DIR, "3Q_heat_age_corr.pdf"), plot = gg_heat_age_corr, width = 40, height = 45, units = "mm", bg = "white",  dpi = 600)

```

# --- Panel 3R. HEAT GTEx x Sex Boxplot --- 
```{r, fig.width=2.165354, fig.height= 1.771654}
plot_data <- meta_filtered %>%
  mutate(
    tissue_short = case_when(
      tissue == "Subcutaneous" ~ "SubQ",
      tissue == "Visceral" ~ "Vis"
    ),
    sex_short = case_when(
      SEX == "Female" ~ "Fem",
      SEX == "Male" ~ "Male"
    )
  ) %>%
  mutate(x_label = paste(tissue_short, sex_short, sep = "\n"))

comparisons_list <- combn(unique(plot_data$x_label), 2, simplify = FALSE)
comparisons_list <- list(
  c("SubQ\nFem", "SubQ\nMale"),
  c("SubQ\nFem", "Vis\nFem"),
  c("Vis\nMale", "Vis\nFem"),
  c("SubQ\nMale", "Vis\nMale")
)
gg_heat_sex_tissue_boxplot <- ggplot(plot_data, aes(x = x_label, y = HEAT)) +
  geom_jitter(aes(alpha = tissue, fill = SEX, colour = SEX), shape = 21, size = 1, width = 0.05, stroke = 0.2) +
stat_compare_means(
  comparisons = comparisons_list,
  method = "wilcox.test",
  paired = FALSE,
  p.adjust.method = "none",
  aes(
    label = after_stat(ifelse(p < 0.001, "<0.001", sprintf("%.2f", p))),
    fontface = after_stat(ifelse(p < 0.05, "bold", "plain")),
    colour   = after_stat(ifelse(p < 0.05, "red", "black"))
  ),
  size = 5 / .pt
) +
  geom_boxplot(aes(color = SEX), outliers = FALSE, width = 0.3, fill = "white", linewidth = 0.3) +
  scale_fill_manual(values = c("Female" = "coral", "Male" = "darkslategray")) +
  scale_color_manual(values = c("Female" = "coral", "Male" = "darkslategray")) +
  scale_alpha_manual(values = c(0.4, 0.6)) +
  theme_nature_metabolism() +
  theme(legend.position = "none") +
  theme(
    plot.title = element_text(size = 6, hjust = 0.5),
    axis.line = element_line(linewidth = 0.3) # smaller = thinner lines
  ) +
  labs(
    title = "Tissue and Gender",
    subtitle = "GTEx - Adipose Tissue",
    y = "HEAT Score (ssGSEA)",
    x = NULL
  ) +
  coord_cartesian(clip = "off")
```


## --- Save the pplot ---
```{r, fig.width= 1.574803, fig.height=1.771654}
mm_to_in(c(40, 45))
gg_heat_sex_tissue_boxplot <- gg_heat_sex_tissue_boxplot +
  scale_y_continuous(expand = expansion(0.08))

gg_heat_sex_tissue_boxplot
ggsave(file.path(OUTPUT_DIR, "3R_heat_sex_tissue_boxplot.pdf"), plot = gg_heat_sex_tissue_boxplot, width = 40, height = 45, , units = "mm", bg = "white",  dpi = 600)
```

# --- Panel 3S. GTEx HEAT x Disease Association BoxPlots --- 
## --- Load the data ---
```{r}
# Load calculated significant correlations
stratified_model_aov <- read.csv(GTEX_AOV_STATS_CSV)
# significant_stratified_aov <- read.csv(GTEX_AOV_STATS_SIGNIF_CSV)

# Load file with HEAT score and metadata values
meta_filtered <- read.csv(GTEX_META_CSV)

# Load metadata explanation
gtex_subj_meta_annot <- read.csv(GTEX_META_ANNOT_CSV)
```

# --- Panel S3Q. GTEx HEAT x Disease Association BoxPlots - All comparisons --- 
```{r}
# Store significant associations ----
sig_vars <- c("MHHTN",
              "MHRNLFLR",
              "MHT2D"
              )

# Extract GTEx variables description ----
var_description <- gtex_subj_meta_annot %>%
  dplyr::filter(VARNAME %in% sig_vars) %>%
  select(VARDESC, VARNAME)


# Prepare plot data ----
plot_data <- meta_filtered %>%
  select(tissue, SEX, HEAT, COHORT, all_of(sig_vars)) %>%
  pivot_longer(cols = all_of(sig_vars), names_to = "disease", values_to = "disease_status") %>%
  mutate(disease_status = as.character(disease_status)) %>%
  mutate(disease_status = case_when(
    disease_status == "0" ~ "Non-Diseased",
    disease_status == "1" ~ "Diseased",
    TRUE ~ "Unknown"
  )) %>%
  mutate(disease_status = factor(disease_status, levels = c("Non-Diseased", "Diseased"))) %>%
  drop_na(disease_status) |> # Drop rows with any NA in the model variables
  mutate(tissue = case_when(
    str_detect(tolower(tissue), "sub") ~ "Subcutaneous",
    str_detect(tolower(tissue), "vis") ~ "Visceral",
    TRUE ~ "Unknown"
  )) %>%
  select(tissue, SEX, HEAT, COHORT, disease, disease_status) %>%
  mutate(
    tissue_short = case_when(
      tissue == "Subcutaneous" ~ "SubQ",
      tissue == "Visceral" ~ "Vis"
    ),
    sex_short = case_when(
      SEX == "Female" ~ "Fem",
      SEX == "Male" ~ "Male"
    ),
    x_var = paste(tissue, as.character(SEX), sep = "\n")
  ) %>%
  mutate(x_label = paste(tissue_short, sex_short, sep = "\n")) %>%
  left_join(var_description, by = c("disease" = "VARNAME")) %>%
  mutate(VARDESC = str_remove(VARDESC, "\\(NIDDM, adult onset diabetes\\)")) |>
  mutate(x_label = factor(x_label))

x_levels <- levels(plot_data$x_label)

# --- Add statistics ---
## Extract x variable names to left join
names_map <- plot_data %>% distinct(x_var, x_label)

## Filter pre-calculated statistics
stat_df <- stratified_model_aov %>%
  filter(term %in% sig_vars) %>%
  left_join(var_description, by = c("term" = "VARNAME")) %>%
  mutate(
    x = paste(tissue, SEX, sep = "\n"),
    group1 = "Non-Diseased",
    group2 = "Diseased",
    y.position = max(plot_data$HEAT) * 1.05,
    label = case_when(
      # FDR < 0.001 ~ "<0.001",
      TRUE ~ as.character(signif(FDR, 2))
    )
  ) %>%
  select(VARDESC, x, group1, group2, p.value, y.position, label) %>%
  left_join(names_map, by = c("x" = "x_var")) %>%
  mutate(VARDESC = str_remove(VARDESC, "\\(NIDDM, adult onset diabetes\\)"))

## Add aesthetic annotations
stat_df <- stat_df %>% mutate(
  xmin = as.numeric(x_label) - 0.25,
  xmax = as.numeric(x_label) + 0.25,
  colors = ifelse(p.value < 0.05, "red", "black"),
  fontface = ifelse(p.value < 0.05, "bold", "plain")
)


#  --- Create the plot ---
gg_gtex_ancov_boxplot_all <- ggplot(
  plot_data,
  aes(
    x = x_label,
    y = HEAT,
    color = disease_status
  )
) +
  geom_jitter(aes(colour = disease_status), alpha = 0.2, shape = 21, size = 0.5, stroke = 0.2, position = position_jitterdodge(dodge.width = 0.8)) +
  geom_boxplot(outliers = F, linewidth = 0.3) +
  labs(x = NULL, y = "HEAT Score (ssGSEA)") +
  theme_nature_metabolism() +
  theme(strip.background = element_blank()) +
  scale_color_manual(values = c("grey65", "deeppink4"), name = "Disease status") +
  stat_pvalue_manual(
    stat_df,
    label = "label",
    xmin = "xmin",
    xmax = "xmax",
    fontface = "fontface",
    y.position = "y.position", bracket.size = 0.2,
    size = 5 / .pt
  ) +
  facet_wrap(vars(VARDESC)) +
  coord_cartesian(clip = "off")
```

## --- Record N ---
```{r}
gtex_n <- plot_data |> count(tissue, SEX, disease, disease_status)
```

## --- Save plot ---
```{r, fig.width=7.086614, fig.height=1.265467}
mm_to_in(c(180, 225 / 7))
gg_gtex_ancov_boxplot_all <- gg_gtex_ancov_boxplot_all+
  scale_y_continuous(expand = expansion(mult = c(0.15)))

gg_gtex_ancov_boxplot_all

# Save and output
scale <- 180
ggsave(file.path(OUTPUT_DIR, "S3Q_gtex_ancov_boxplot_all.pdf"), plot = gg_gtex_ancov_boxplot_all, width = 180, height = 225 / 7, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel 3S. GTEx HEAT x Disease Association BoxPlots - Female SubQ --- 
```{r}
# Filter plot data, rename and assign levels
plot_data_2 <- plot_data |>
  filter(tissue == "Subcutaneous", SEX == "Female") |>
  mutate(
    VARDESC = case_when(
      str_detect(VARDESC, "Diabetes") ~ "T2D",
      TRUE ~ VARDESC
    ),
    VARDESC = factor(VARDESC)
  )

# Extract levels for x axis
x_levels <- levels(plot_data_2$VARDESC)

# Create statistics table with matching x axis variables
stat_df <- stratified_model_aov %>%
  filter(
    term %in% sig_vars,
    SEX == "Female",
    tissue == "Subcutaneous"
  ) %>%
  left_join(var_description, by = c("term" = "VARNAME")) %>%
  mutate(
    VARDESC = case_when(
      str_detect(VARDESC, "Diabetes") ~ "T2D",
      TRUE ~ VARDESC
    ),
    x = factor(VARDESC, levels = x_levels),
    group1 = "Non-Diseased",
    group2 = "Diseased",
    y.position = max(plot_data_2$HEAT) * 1.10,
    label = case_when(
      # FDR < 0.001 ~ "<0.001",
      TRUE ~ as.character(signif(FDR, 2))
    ),
    xmin = as.numeric(x) - 0.25,
    xmax = as.numeric(x) + 0.25,
    colors = ifelse(FDR < 0.05, "red", "black"),
    fontface = ifelse(FDR < 0.05, "bold", "plain")
  ) %>%
  select(
    VARDESC, x, xmin, xmax,
    group1, group2,
    p.value, y.position, label, fontface
  )


# Create the plot with ggpubr
gg_gtex_ancov_boxplot <- ggplot(
  plot_data_2,
  aes(
    x = VARDESC,
    y = HEAT,
    color = disease_status
  )
) +
  geom_jitter(aes(colour = disease_status), alpha = 0.5, shape = 21, size = 0.5, stroke = 0.2, position = position_jitterdodge(dodge.width = 0.8)) +
  geom_boxplot(outliers = F, linewidth = 0.3) +
  labs(x = NULL, y = "HEAT Score (ssGSEA)") +
  theme_nature_metabolism() +
  theme(strip.background = element_blank()) +
  # theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  scale_color_manual(values = c("grey65", "deeppink4"), name = "Disease status") +
  stat_pvalue_manual(
    stat_df,
    label = "label",
    xmin = "xmin",
    xmax = "xmax",
    fontface = "fontface",
    y.position = "y.position",
    bracket.size = 0.2,
    size = 5 / .pt
  ) +
  labs(
    title = "HEAT Score ",
    subtitle = "GTEx - Subcutaneous WAT"
  )+
  coord_cartesian(clip = "off")

gg_gtex_ancov_boxplot
```


## --- Save plot ---
```{r, fig.width=2.874016, fig.height=1.771654}
mm_to_in(c(73, 45))
gg_gtex_ancov_boxplot <- gg_gtex_ancov_boxplot +
  scale_y_continuous(expand = expansion(mult = c(0.15)))

gg_gtex_ancov_boxplot

# Save and output
ggsave(file.path(OUTPUT_DIR, "3S_gtex_ancov_boxplot.pdf"), plot = gg_gtex_ancov_boxplot, width = 73, height = 45, units = "mm", bg = "white",  dpi = 600)
```

