---
title: "3-FigurePlotting"
author: "Henrique"
date: "2025-07-17"
output: html_document
---
# --- Setup ---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

mm_to_in <- function(mm) mm / 25.4

# Load libraries
library(Seurat)
library(SeuratObject)
library(ggplot2)
library(RColorBrewer)
library(VISION)
library(Matrix)
library(dplyr)
library(CellChat)
library(circlize) ### A tutorial on how to create your own circlize can be found here: https://r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling.html
library(ComplexHeatmap)
library(NMF)
library(stringr)
library(readr)
library(viridis)
library(tidyverse)
library(patchwork)
library(igraph)
library(ggraph)
library(tidygraph)
library(ggnewscale)

conflicts_prefer(ggplot2::annotate)
conflicts_prefer(base::unname)


# Define short cell names
cell_type_mapping <- c(
  "Brown Adipocytes" = "BAds",
  "White Adipocytes" = "WAds",
  "Lipid Associated Macrophages" = "PreAds",
  "ADIPOQ+ Smooth Muscle Cells" = "PreAds",
  "Pre-adipocytes" = "PreAds",
  "Adipocyte Progenitor Cells" = "ASPC",
  "Mesenchymal Stem Cells" = "MSCs",
  "Arterial Endothelial Cells" = "EndoCs",
  "Venous Endothelial Cells" = "EndoCs",
  "Capillary Endothelial Cells" = "EndoCs",
  "Lymphatic Endothelial Cells" = "LECs",
  "Pericytes" = "Pericytes",
  "Smooth Muscle Cells" = "SMCs",
  "Macrophages" = "Macro",
  "Mast Cells" = "MastCs",
  "T Lymphocytes" = "T Cells",
  "B Lymphocytes" = "B Cells",
  "Neutrophils" = "Neutrophils",
  "Schwann Cells" = "SchwannCs",
  "Neuron Associated Cells" = "NeuroCs",
  "Parathyroid Associated Cells" = "ParaThyCs",
  "PRM1+ Cells" = "PRM1+"
)

# OBS:  theme_nature_metabolism() and color palettes set on "scripts/0.environment_setup/snRNAseq_graphics_setup.R"
source(here("scripts/0.environment_setup/snRNAseq_graphics_setup.R"))

# --- Setup paths ---

## Inputs
DESEQ_STATS_CSV <- here("output/4.downstream_pipelines/de_analysis/data/deseq2_stats.csv")
GSEA_HALLMARK <- here("output/4.downstream_pipelines/de_analysis/data//HallMark_GSEA.csv")
GSEA_GO <- here("output/4.downstream_pipelines/de_analysis/data//GO_GSEA.csv")
GSEA_REACTOME <- here("output/4.downstream_pipelines/de_analysis/data//Reactome_GSEA.csv")

CELLCHAT_DEEP <- here("output/4.downstream_pipelines/cellchat/data/cellchat_deep_pop.size.false_cell_type_short.rds")
CELLCHAT_SUPERFICIAL <- here("output/4.downstream_pipelines/cellchat/data/cellchat_superficial_pop.size.false_cell_type_short.rds")
CELLCHAT_MERGED <- here("output/4.downstream_pipelines/cellchat/data/cellchat_merged_pop.size.false_cell_type_short.rds")

SEURAT_INPUT <- here("data/adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds")

MEBOCOST_COMM_CSV <- here("output/4.downstream_pipelines/mebocost/data/communication_result_multi.csv")

# Inputs to update
HEATMAP_INPUT_RDS <- here("output/4.downstream_pipelines/de_analysis/data/heatmap_input.rds")

# Outputs
OUTPUT_DIR <- here("output", "5.figure_preparation", "Figure_3")

dir.create(OUTPUT_DIR, showWarnings = F, recursive = T)
```

### --- Load Seurat ---
```{r}
s.object <- readRDS(SEURAT_INPUT)
```

# --- Heatmap White Adipocytes DEG by depot ---
```{r plot-heatmap}
# Load heatmap input list (.rds) generated on #X.03.X_Heatmap_Processing.Rmd
pheatmap_data_all_deg <- readRDS(HEATMAP_INPUT_RDS)

# Remove unwanted annotations
pheatmap_data_all_deg$ann_df.sub <- pheatmap_data_all_deg$ann_df.sub |> select(-`Sample ID`)

# Save pheatmap as pdf specific size and font

multiplier <- 3
pdf(file.path(OUTPUT_DIR, "3A_heatmap_deg_wad.pdf"), width = mm_to_in(50 * multiplier), height = mm_to_in(180 * multiplier))



new_order <- c(
  grep("White Adipocytes_Superficial", colnames(pheatmap_data_all_deg$plot.data)),
  grep("White Adipocytes_Deep", colnames(pheatmap_data_all_deg$plot.data)),
  grep("Brown", colnames(pheatmap_data_all_deg$plot.data))
)

pheatmap::pheatmap(
  mat = log1p(pheatmap_data_all_deg$plot.data[, new_order]),
  col = rev(RColorBrewer::brewer.pal(11, "RdBu")),
  annotation_col = pheatmap_data_all_deg$ann_df.sub,
  annotation_colors = pheatmap_data_all_deg$ann_colors,
  legend_width = 10, legend = F,
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  show_colnames = FALSE,
  show_rownames = FALSE,
  scale = "row",
  gaps_col = c(14,25),
  gaps_row = pheatmap_data_all_deg$gaps_row,
  fontsize = 6 * multiplier, # Default font size
  fontsize_row = 6 * multiplier, # Row labels
  fontsize_col = 6 * multiplier # Column labels
)
dev.off()

colnames(pheatmap_data_all_deg$plot.data)

pheatmap_data_all_deg$ann_df.sub |> count(Cell, `Neck region`)
```


# --- Pseudobulk DESeq2 ---
## --- Read DESeq2 DE data.frame ---
```{r read-deseq-df, fig.width= 1.77, fig.height= 1.42}
final_DE <- read.csv(DESEQ_STATS_CSV)

final_DE_filt <- final_DE %>% dplyr::filter(padj <= 0.05, contrast.2 != "Intermediate", contrast.1 != "Intermediate")


# This adds a categorisation based on whether the log2FoldChange is up or down
final_DE_filt$direction <- ifelse(final_DE_filt$log2FoldChange > 0, "Up in Deep", "Up in Superficial")

final_DE_filt %>%
  filter(cell_type == "White Adipocytes") %>%
  count(direction)
```
## --- Volcano Plot DEG in White Adipocytes ---
### --- Set up function ---
```{r}
library(ggrepel)

PlotVolcano <- function(de_table, cluster, cont.1 = "Deep", cont.2 = "Superficial",
                        fc_cutoff = 1.5, p_cutoff = 0.05, show_genes = NULL,
                        n_top_show = 40, label_size = 6) {
  res <- de_table %>% dplyr::filter(cell_type == cluster)
  sub_res <- res %>% dplyr::filter(contrast.1 == cont.1, contrast.2 == cont.2)

  larger_cont_name <- max(nchar(cont.1), nchar(cont.2))
  contrast_name <- paste0(cluster, ": ", cont.1, " vs ", cont.2)
  axis_cont.1 <- sprintf("%-*s", larger_cont_name, as.character(cont.1))
  axis_cont.2 <- sprintf("%-*s", larger_cont_name, as.character(cont.2))

  volcano_df <- sub_res %>%
    dplyr::mutate(
      neg_log10_p = -log10(padj),
      direction = dplyr::case_when(
        padj < p_cutoff & log2FoldChange >= fc_cutoff ~ "Up in Deep",
        padj < p_cutoff & log2FoldChange <= -fc_cutoff ~ "Up in Superficial",
        TRUE ~ "Not Significant"
      )
    )

  maxFC <- max(abs(volcano_df$log2FoldChange), na.rm = TRUE)
  maxY <- max(volcano_df$neg_log10_p, na.rm = TRUE)

  top_genes <- volcano_df %>%
    dplyr::filter(direction != "Not Significant") %>%
    dplyr::arrange(pvalue) %>%
    dplyr::slice_head(n = n_top_show) %>%
    dplyr::pull(gene)

  label_genes <- unique(c(show_genes, top_genes))

  volcano_df <- volcano_df %>%
    dplyr::mutate(label = dplyr::if_else(gene %in% label_genes, as.character(gene), NA_character_))

  volcano_df$direction <- factor(
    volcano_df$direction,
    levels = c("Up in Superficial", "Not Significant", "Up in Deep")
  )

  gg_volcano <- ggplot(volcano_df, aes(x = log2FoldChange, y = neg_log10_p)) +
    geom_vline(
      xintercept = c(-fc_cutoff, 0, fc_cutoff),
      linetype = "dashed", color = "black", linewidth = 0.1
    ) +
    geom_hline(
      yintercept = -log10(p_cutoff),
      linetype = "dashed", color = "black", linewidth = 0.1
    ) +

    # Solid points, no outline, smaller size
    geom_point(aes(color = direction), shape = 16, size = 0.7, alpha = 0.15) +
    geom_point(
  data = volcano_df %>% dplyr::filter(gene %in% label_genes),
  aes(x = log2FoldChange, y = neg_log10_p, fill = direction),
  shape = 21,
  size = 0.9,
  stroke = 0.2,
  color = "black",
  alpha = 1,
  inherit.aes = FALSE
) + 
    scale_color_manual(values = c(
      "Up in Superficial" = neck.color[["Superficial"]],
      "Up in Deep"        = neck.color[["Deep"]],
      "Not Significant"   = "grey70"
    )) +
scale_fill_manual(values = c(
  "Up in Superficial" = colorspace::darken(neck.color[["Superficial"]], amount = 0.15),
  "Up in Deep"        = colorspace::darken(neck.color[["Deep"]], amount = 0.15),
  "Not Significant"   = "grey60"
)) + 
    labs(
      title = "Cervical Depot DGE",
      subtitle = cluster,
      x = paste0("Log2 fold-change"),
      y = "-log10(adjusted p-value)",
      color = NULL
    ) +
    theme(
      legend.position = "top",
      plot.subtitle = element_text(hjust = 0.5)
    ) +
    scale_x_continuous(limits = c(-maxFC, maxFC)) +
    scale_y_continuous(limits = c(0, maxY * 1.05))

  gg_volcano <- gg_volcano +

    # Add a text
    geom_text_repel(
      aes(label = label),
      max.overlaps = Inf,
      size = label_size / .pt,
      box.padding = 0.4,
      point.padding = 0.25,
      segment.color = "black",
      segment.size = 0.3,
      min.segment.length = 0,
      na.rm = TRUE
    ) +

    # Corner annotations with neck colors
    annotate(
      "text",
      x = -Inf, y = Inf, label = "Up in Superficial",
      hjust = -0.05, vjust = 1.2, color = neck.color["Superficial"], size = 5 / .pt
    ) +
    annotate(
      "text",
      x = Inf, y = Inf, label = "Up in Deep",
      hjust = 1.05, vjust = 1.2, color = neck.color["Deep"], size = 5 / .pt
    )

  gg_volcano
}
```

### --- Plot B: Volcano plot ---
```{r plot-volcano, fig.width= 1.968504, fig.height= 1.968504}
mm_to_in(c(50, 50))

# Choose genes to plot
deep_genes <- c("PDE1A", "PDE4D", "PDE3A", "EBF2", "PRDM16", "UCP1", "PPARGC1A", "MECOM")
sup_genes <- c("HSPA1A", "CRYAB", "ADAMTS4", "MMP17")

# Plot
gg_volcano <- PlotVolcano(de_table = final_DE, cluster = "White Adipocytes", show_genes = c(deep_genes, sup_genes), n_top_show = 0, cont.1 = "Deep", cont.2 = "Superficial", label_size = 5, fc_cutoff = 0.5) + theme_nature_metabolism() + theme(plot.subtitle = element_text(hjust = 0.5)) + NoLegend()
# gg_volcano <- gg_volcano + plot_annotation(tag_levels = list("b")) & theme(plot.tag = element_text(face = "bold", size = 6 * 1.3))
gg_volcano

# Save as png
ggsave(file.path(OUTPUT_DIR, "3B_Volcano.pdf"), gg_volcano, width = 50, height = 50, units = "mm", bg = "white",  dpi = 600)
```



## --- Plot Supp B: Volcano Plot DEG in ASPCs ---
```{r, fig.width=2.36, fig.height=1.77}
gg_volcano_aspc <- PlotVolcano(de_table = final_DE, cluster = "Adipocyte Progenitor Cells", n_top_show = 40, cont.1 = "Deep", cont.2 = "Superficial") + theme_nature_metabolism() + theme(plot.subtitle = element_text(hjust = 0.5)) + NoLegend() +
  plot_annotation(tag_level = list("Supp B"))

gg_volcano_aspc

# Save as pdf
ggsave(file.path(OUTPUT_DIR, "S3B_volcano_aspc.pdf"), gg_volcano_aspc, width = 70, height = 225 / 4, units = "mm", bg = "white",  dpi = 600)
```
# --- GSEA ---
## --- Load the data ---
```{r}
gsea_hallmark <- read.csv(file = GSEA_HALLMARK)
gsea_GO_BP <- read.csv(file = GSEA_GO)
gsea_Rctm <- read.csv(file = GSEA_REACTOME)
```

## --- Set up plotting function---
```{r}
GeneRatioDotPlot <- function(gsea_res, wrap = 80, text_size = 8, left_margin = 10, padj_cutoff = 0.1, size_by = "gene count", slice_top = FALSE, top_n = 5) {
  df <- as.data.frame(gsea_res)

  df <- df %>%
    mutate(
      gene_count = str_count(core_enrichment, pattern = "/") + 1,
      gene_ratio = gene_count / setSize,
      Description = str_replace_all(Description, "_", " "),
      Description = str_wrap(as.character(Description), width = wrap),
      Description = str_remove(Description, "GOBP |REACTOME "),
      direction = case_when(NES > 0 ~ "Superficial Enriched", NES < 0 ~ "Deep Enriched")
    ) %>%
    filter(p.adjust <= padj_cutoff)

  if (slice_top) {
    df <- df %>%
      group_by(direction) %>%
      arrange(p.adjust) %>%
      slice_head(n = top_n) %>%
      ungroup()
  }

  # Ensure factor levels are ordered per direction
  df <- df %>%
    group_by(direction) %>%
    mutate(Description = str_to_title(Description)) %>%
    mutate(Description = factor(Description, levels = unique(Description))) %>%
    ungroup()

  p <- ggplot(df, aes(x = gene_ratio, y = Description)) +
    geom_point(aes(fill = p.adjust, size = gene_count), shape = 21, color = "black") +
    scale_fill_viridis_c(option = "inferno", direction = -1, name = "p.adjust") +
    scale_size(range = c(1, 2), name = size_by) +
    facet_wrap(~direction, scales = "free_y") +
    theme_nature_metabolism() +
    theme(
      legend.position = "right",
      panel.grid.major = element_line(color = "grey80"),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3),
      strip.background = element_blank(), # <- removes grey background
      strip.text = element_text(face = "bold", size = text_size),
      axis.text.y = element_text(size = text_size, hjust = 1),
      axis.title.y = element_blank()
    ) +
    scale_y_discrete(limits = rev) +
    labs(x = "Gene Ratio", title = "GSEA Dotplot by Regulation Direction")

  return(p)
}
```

## --- Plot C: Gene Ontology BP ---
```{r}
GOBPDotPlot <- function(gsea_res, wrap = 80, text_size = 8, left_margin = 10, padj_cutoff = 0.1, size_by = "Gene ratio", slice_top = FALSE, top_n = 5) {
  df <- as.data.frame(gsea_res)

  df <- df %>%
    mutate(
      NES = NES,
      gene_count = str_count(core_enrichment, pattern = "/") + 1,
      gene_ratio = gene_count / setSize,
      Description = str_replace_all(Description, "_", " "),
      Description = str_wrap(as.character(Description), width = wrap),
      Description = str_remove(Description, "GOBP |REACTOME "),
      direction = case_when(NES > 0 ~ "Superficial Enriched", NES < 0 ~ "Deep Enriched")
    ) %>%
    filter(p.adjust <= padj_cutoff)

  if (slice_top) {
    df <- df %>%
      group_by(direction) %>%
      arrange(p.adjust) %>%
      slice_head(n = top_n) %>%
      ungroup()
  }

  # Ensure factor levels are ordered per direction
  df <- df %>%
    group_by(direction) %>%
    mutate(Description = str_to_title(Description)) %>%
    mutate(Description = factor(Description, levels = unique(Description))) %>%
    ungroup()

  p <- ggplot(df, aes(x = -log10(p.adjust), y = Description)) +
    geom_point(aes(fill = NES, size = gene_ratio), shape = 21, color = "black") +
    scale_fill_gradient2(name = "NES") +
    scale_size(range = c(1, 2), name = size_by, limits = c(0, 0.6), breaks = c(0, 0.3, 0.6), ) +
    scale_y_discrete(limits = rev) +
    facet_wrap(~direction, scales = "free_y") +
    theme_nature_metabolism() +
theme(
  legend.position = "right",
  panel.grid.major = element_line(color = "grey80"),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3),
  strip.background = element_blank(),
  strip.text = element_text(face = "bold", size = text_size),
  axis.text.y = element_text(size = text_size, hjust = 1),
  axis.title.y = element_blank(),
  plot.margin = margin(t = 5.5, r = 5.5, b = 5.5, l = left_margin)
) +
    labs(x = "-log10(p.adjusted)", title = "GSEA Dotplot by Regulation Direction")

  return(p)
}
```


```{r, fig.width=4.724409 , fig.height=1.968504}
mm_to_in(c(130, 50))
gg_gsea_gobp <- GOBPDotPlot(gsea_GO_BP, padj_cutoff = 0.1, wrap = 30, slice_top = TRUE, top_n = 8, text_size = 5, left_margin = 5.5) +
  labs(title = "GO - Biological Process\nWhite Adipocytes") 


gg_gsea_gobp 

# Save the file
ggsave(file.path(OUTPUT_DIR, "3C_dotplot_gsea_go_bp_nes.pdf"), gg_gsea_gobp, width = 130, height = 50, units = "mm", bg = "white",  dpi = 600)
```

# --- CellChat ---
## --- Load the objects and create CC object
```{r}
# --- Load the cellchat objects
deep_cellchat <- readRDS(CELLCHAT_DEEP)
superficial_cellchat <- readRDS(CELLCHAT_SUPERFICIAL)
cellchat_merged <- readRDS(CELLCHAT_MERGED)
```

## --- Create object_list
```{r}
object_list <- list(Deep = deep_cellchat, Superficial = superficial_cellchat)
```


## --- Extract signaling information
```{r}
# Extract signaling information
df.sup <- as.data.frame(subsetCommunication(superficial_cellchat, thresh = 1))
df.deep <- as.data.frame(subsetCommunication(deep_cellchat, thresh = 1))

# Merge and filter significant
df.merge <- rbind(
  df.sup |> mutate(depot = "Superficial"),
  df.deep |> mutate(depot = "Deep")
) |>
  filter(pval <= 0.05)  

annotation_order <- df.merge |> count(annotation) |> arrange(-n) |> pull(annotation)
df.merge <- df.merge |> mutate(annotation = factor(annotation, levels = annotation_order))
```

# --- Plotting
## --- Plot Supp C: CELLCHAT - Number of interactions per depot
```{r, fig.width= 1.968504, fig.height= 1.417323, warning=FALSE}
# Main Communication types
pd <- position_dodge(width = 0.9)

gg_comm_bar <- ggplot(df.merge, aes(x = annotation, fill = depot)) +
  geom_bar(position = pd) +
  geom_text(
    stat = "count",
    aes(label = after_stat(count), group = depot),
    position = pd,
    vjust = -0.15,
    size = 5 / .pt
  ) +
  labs(
    x = "Signaling Type", y = str_wrap("Number of inferred interactions", 25), fill = "Depot",
    title = "Communication Types"
  ) +
  theme_nature_metabolism() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = neck.color) +
  scale_y_continuous(expand = expansion(c(0.05, 0.20)))


# Main Communication types to Wads
gg_comm_to_wad <- ggplot(df.merge |> filter(target == "WAds"), aes(x = annotation, fill = depot)) +
  geom_bar(position = pd) +
  geom_text(
    stat = "count",
    aes(label = after_stat(count), group = depot),
    position = pd,
    vjust = -0.15,
    size = 5 / .pt
  ) +
  labs(x = "Signaling Type", y = str_wrap("Number of inferred interactions", 35), fill = "Depot", title = "Communication Types", subtitle = "White Adipocytes as Target") +
  theme_nature_metabolism() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = neck.color) +
  scale_y_continuous(expand = expansion(c(0.05, 0.15)))
```


```{r, fig.width= 1.968504, fig.height= 1.224409, warning=FALSE}
mm_to_in(c(50, 31.1))
gg_comm_bar <- gg_comm_bar +
  # plot_annotation(tag_levels = list("d")) + theme(plot.tag = element_text(face = "bold", size = 6)) +
  # NoLegend() +
  theme(axis.title.x = element_blank())
gg_comm_bar

ggsave(file.path(OUTPUT_DIR, "3Z_barplot_communication_types.pdf"), plot = gg_comm_bar, width = 50, height = 31.1, units = "mm", dpi = 600, bg = "white")
```


```{r, fig.width= 1.574803, fig.height= 1.771654, warning=FALSE}
mm_to_in(c(40,45))
gg_comm_to_wad
ggsave(file.path(OUTPUT_DIR, "3Z_barplot_communication_types_to_wads.pdf"), plot = gg_comm_to_wad, width = 40, height = 45, units = "mm", dpi = 600, bg = "white")
```


## --- Communication flow analysis ---
```{r, fig.width=1.574803, fig.height=1.417323, warning=FALSE}
mm_to_in(c(40, 36))

# Helper function to plot only one signaling type consistently
customRankNet <- function(cellchat = cellchat_merge, ccht_comm_prob_df = df.merge, comm_type){  
  pathway.show <- ccht_comm_prob_df |>
    filter(annotation == comm_type) |>
    pull(pathway_name)
  rankNet(cellchat_merged, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = "WAds", stacked = T, do.stat = TRUE, color.use = neck.color, signaling = pathway.show) + theme_nature_metabolism() + theme(legend.position = "top") +
    labs(title = comm_type, subtitle = "WAds as Receivers") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```


```{r, fig.width=1.614173, fig.height=1.771654, warning=FALSE}
mm_to_in(c(41,45))
customRankNet(cellchat_merge, df.merge, comm_type = "Secreted Signaling")
ggsave(file.path(OUTPUT_DIR, paste0("3Z_communication_flow_to_wads_", "Secreted Signaling", ".pdf")), width = 41, height = 45, units = "mm", dpi = 600, bg = "white")
```

```{r, fig.width=1.574803, fig.height=1.574803, warning=FALSE}
mm_to_in(c(40,40))
customRankNet(cellchat_merge, df.merge, comm_type = "ECM-Receptor")
ggsave(file.path(OUTPUT_DIR, paste0("3Z_communication_flow_to_wads_", "ECM-Receptor", ".pdf")), width = 40, height = 40, units = "mm", dpi = 600, bg = "white")

customRankNet(cellchat_merge, df.merge, comm_type = "Cell-Cell Contact")
ggsave(file.path(OUTPUT_DIR, paste0("3Z_communication_flow_to_wads_", "Cell-Cell Contact", ".pdf")), width = 40, height = 40, units = "mm", dpi = 600, bg = "white")

```


## --- 2D communication plot ---
```{r, warning=FALSE}
library(ggrepel)
# I have confirmed that the interaction strength is simply the sum of the communication probability within all the significant pathwys
# We can thus create our own plot

## Initialize empty list
gg_2d <- list()

# Start loop to plot for each signaling type
for (comm_type in unique(df.merge$annotation)) {
  df.filter <- df.merge |> filter(annotation == comm_type)

  # Calculate strength
  ## Output signal strength
  out_strength_df <- df.filter |>
    group_by(source, depot) |>
    summarise(out_prob = sum(prob)) |>
    rename("cell" = "source")

  ## Input signal strength
  in_strength_df <- df.filter |>
    group_by(target, depot) |>
    summarise(in_prob = sum(prob)) |>
    rename("cell" = "target")

  # Communication count per cell (counts rows where cell appears as source or target).
  communication_count_df <- df.filter %>%
    mutate(target = ifelse(source == target, "", target)) %>% # Remove cell if autocrine to count autocrine just once
    select(depot, source, target) %>%
    pivot_longer(cols = c(source, target), values_to = "cell") %>%
    count(depot, cell, name = "comm_count")

  # Merge data frames
  strengt_df <- left_join(out_strength_df, in_strength_df, by = c("cell", "depot"))

  strengt_df <- strengt_df %>%
    left_join(communication_count_df, by = c("depot", "cell")) %>%
    mutate(
      comm_count = coalesce(comm_count, 0L),
      cell_label = cell
    )
  
  # Map names
  meta_lookup <- s.object@meta.data |>
    dplyr::select(cell_type, cell_type_short) |>
    dplyr::distinct(cell_type, .keep_all = TRUE) |>
    mutate(cell_type_short = as.character(cell_type_short))
  strengt_df <- dplyr::left_join(strengt_df, meta_lookup, by = c("cell" = "cell_type_short"))

  # Filter cells of interest
  if (comm_type == "Secreted Signaling") {
    strengt_df <- strengt_df |> mutate(cell_label = case_when(cell_type %in% c(
      "White Adipocytes",
      "Brown Adipocytes",
      "Lymphatic Endothelial Cells",
      "Schwann Cells"
    ) ~ cell, TRUE ~ NA))
  } else if (comm_type == "ECM-Receptor") {
    strengt_df <- strengt_df |> mutate(cell_label = case_when(cell_type %in% c(
      "White Adipocytes",
      "Brown Adipocytes",
      "Adipocyte Progenitor Cells",
      "Smooth Muscle Cells",
      "Pericytes",
      "Endothelial Cells",
      "Lymphatic Endothelial Cells",
      "Schwann Cells"
    ) ~ cell, TRUE ~ NA))
  } else if (comm_type == "Cell-Cell Contact") {
    strengt_df <- strengt_df |> mutate(cell_label = case_when(cell_type %in% c(
      "White Adipocytes",
      "B Lymphocytes",
      "Macrophages",
      "Mast Cells",
      "Neutrophils",
      "T Lymphocytes",
      "Endothelial Cells",
      "Lymphatic Endothelial Cells",
      "Schwann Cells"
    ) ~ cell, TRUE ~ NA))
  }


  sup_max <- strengt_df %>%
    ungroup() |>
    dplyr::filter(depot == "Superficial") %>%
    dplyr::summarise(
      x = max(out_prob, na.rm = TRUE) * 1.05,
      y = max(in_prob, na.rm = TRUE) * 1.05
    )

  # Plot 2D plot

  gg_2d[[comm_type]] <- ggplot(strengt_df, aes(x = out_prob, y = in_prob)) +
    geom_rect(
      data = sup_max,
      aes(xmin = -Inf, ymin = -Inf, xmax = x * 1.15, ymax = y * 1.15),
      inherit.aes = FALSE,
      fill = "grey90",
      alpha = 0.3,
      color = "black",
      linewidth = 0.2
    ) +
    geom_point(aes(fill = cell, size = comm_count), shape = 21, color = "black", alpha = 0.75) +
    geom_text_repel(
      aes(label = cell_label, color = cell),
      size = 5 / .pt,
      max.overlaps = Inf,
      box.padding = 0.35,
      point.padding = 0.2,
      show.legend = FALSE, face = "bold"
    ) +
    facet_wrap(~depot) +
    scale_fill_manual(values = palette.use) +
    scale_color_manual(values = palette.use) +
    scale_size(range = c(1, 2))




  gg_2d[[comm_type]] <- gg_2d[[comm_type]] +
    theme_nature_metabolism() +
    NoLegend() +
    theme(
      panel.grid.major.x = element_line(linewidth = 0.1, color = "grey90"),
      panel.grid.major.y = element_line(linewidth = 0.1, color = "grey90"),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),
      panel.spacing.x = unit(1.8, "mm"),
      axis.line = element_blank()
    ) +
    scale_x_continuous(
      expand = expansion(c(0.05, 0.15)),
      labels = scales::label_number(accuracy = 0.1, trim = TRUE)
    ) +
    scale_y_continuous(expand = expansion(c(0.05, 0.20))) +
    labs(
      title = comm_type,
      subtitle = "All to All",
      y = "Incoming Signal Strength",
      x = "Outgoing Signal Strength"
    )

  ggsave(file.path(OUTPUT_DIR, paste0("3Z_2d_communication_plots_", comm_type, ".pdf")), plot = gg_2d[[comm_type]], width = 55, height = 33, units = "mm", dpi = 600, bg = "white")
  
  if(comm_type == "Cell-Cell Contact"){
      ggsave(file.path(OUTPUT_DIR, paste0("3Z_2d_communication_plots_", comm_type, ".pdf")), plot = gg_2d[[comm_type]], width = 55, height = 43, units = "mm", dpi = 600, bg = "white")
  }
}
```


```{r, fig.width= 2.165354, fig.height= 1.299213, warning=FALSE}
mm_to_in(c(55, 33))
gg_2d
```

## --- Plot Supp B: Stacked communications ---
```{r, fig.width=1.771654, fig.height=4.429134}
mm_to_in(c(45, 112.5))
# Stacked flow ---
gg_ccht_ranknet <- rankNet(cellchat_merged, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = "WAds", stacked = F, do.stat = TRUE, color.use = neck.color) +
  theme_nature_metabolism() +
  theme(legend.position = "top",
        panel.border = element_rect(color = "black", fill = NA),
) +
  labs(
    title = "CellChat\nInformation Flow",
    subtitle = "WAds as Receivers"
  )

gg_ccht_ranknet

# Save as png
ggsave(file.path(OUTPUT_DIR, "S3B_cellchat_ranknet.pdf"), gg_ccht_ranknet, width = 45, height = 112.5, units = "mm", bg = "white",  dpi = 600)

# Save as pdf
pdf(file.path(OUTPUT_DIR, "S3B_cellchat_ranknet.pdf"), width = 45, height = 112.5)
gg_ccht_ranknet
dev.off()
```

## --- Plot E: CELLCHAT - Circle Plot of selected communications
### --- Define color pallete 
```{r}
# Define the cells and filter the color pallete accordingly
cells <- levels(object_list[[1]]@idents)
color.use <- palette.use[cells]
color.use <- color.use[!is.na(color.use)]
color.use
```

```{r}
meta <- s.object@meta.data

celltype_map <- meta %>%
  dplyr::select(cell_type, cell_type_short) %>%
  dplyr::distinct(cell_type, .keep_all = TRUE) %>%
  tibble::deframe()

rename_celltypes <- function(x, map) {
  x_clean <- stringr::str_remove(x, ".*~ ")
  dplyr::coalesce(unname(map[x_clean]), x_clean)
}

cellchat_merged@idents["Macrophages"]


for (pathways.show in c("COLLAGEN", "LAMININ", "CD99", "BMP", "PTPRM", "FGF", "MPZ")) {
  # Setting up a loop to get around issues when a single depot has significant communication
  weight.max <- 0
  for (i in 1:length(object_list)) {
    print(i)
    tryCatch(
      {
        weight <- getMaxWeight(object_list[i], slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
        if (weight > weight.max) {
          weight.max <- weight
        }
      },
      error = function(e) {
        message(sprintf("Error at index %d: %s", i, e$message))
        # Continue to next iteration
      }
    )
  }

  # Plot in R
  par(mfrow = c(1, 2), xpd = TRUE)
  for (i in 1:length(object_list)) {
    tryCatch(
      {
        netVisual_aggregate(object_list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object_list)[i]), color.use = color.use)
      },
      error = function(e) {
        print(e$message)
      }
    )
  }

  file_name <- paste0("3E_cellchat_", tolower(pathways.show), "_circleplot.pdf")
  
  # Adjust dimension per plot
  if(pathways.show == "BMP"){
    plot.width = 40
    plot.heigth = 45
  } else if (pathways.show == "COLLAGEN"){
    plot.width = 70
    plot.heigth = 33
  } else {
    plot.width = 50
    plot.heigth = 40
    }
  # Save as pdf
  ## Note: This needs to be further edited using softwares like Inkscape or Illustrator
  pdf(file.path(OUTPUT_DIR, file_name), width = mm_to_in(plot.width*3), height = mm_to_in(plot.heigth*3))
  par(mfrow = c(1, 2), xpd = TRUE)
  for (i in 1:length(object_list)) {
    tryCatch(
      {
        netVisual_aggregate(object_list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object_list)[i]), color.use = color.use, )
      },
      error = function(e) {
        print(e$message)
      }
    )
  }

  dev.off()
  
  if(pathways.show %in% c("COLLAGEN", "CD99")){
      file_name <- paste0("3E_cellchat_", tolower(pathways.show), "_WAd_as_target_circleplot.pdf")
      plot.width = 50
      plot.heigth = 40


      pdf(file.path(OUTPUT_DIR, file_name), width = mm_to_in(plot.width*3), height = mm_to_in(plot.heigth*3))
  par(mfrow = c(1, 2), xpd = TRUE)
  for (i in 1:length(object_list)) {
    tryCatch(
      {
        netVisual_aggregate(object_list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object_list)[i]), color.use = color.use, targets.use = "WAds")
      },
      error = function(e) {
        print(e$message)
      }
    )
  }

  dev.off()
  }
}

```

### --- Plot F: Ligand-Receptor Violin Plots ---
```{r}
plot_list <- list()

Idents(s.object) <- "neck_region"
filtered_s.object <- subset(s.object, idents = c("Intermediate"), invert = TRUE)
Idents(filtered_s.object) <- "cell_type"
Idents(s.object) <- "cell_type"
```


```{r, fig.width=2.755906, fig.height=1.574803, warning=FALSE}
mm_to_in(c(70, 40))
Idents(filtered_s.object) <- "cell_type"
gg_ccht_bmp_vln <- VlnPlot(filtered_s.object, features = c("BMP5", "BMP8B", "ACVR1", "ACVR2A", "BMPR2", "BMPR1A"), idents = c("Smooth Muscle Cells", "Schwann Cells", "White Adipocytes"), split.by = "neck_region", flip = T, stack = T, pt.size = 0) + 
  scale_fill_manual(values = neck.color) + 
  theme_nature_metabolism() + 
  theme(axis.title = element_blank()) + 
  theme(legend.position = "top") + 
  # plot_annotation(tag_levels = list("f")) + theme(plot.tag = element_text(size = 6, face = "bold")) +
  scale_x_discrete(limits = c("Smooth Muscle Cells", "Schwann Cells", "White Adipocytes"))


gg_ccht_bmp_vln
# Save as png
ggsave(file.path(OUTPUT_DIR, "3F_cellchat_bmp_vln.pdf"), gg_ccht_bmp_vln, width = 70, height = 40, units = "mm", bg = "white",  dpi = 600)
```
```{r}
vln_plot_n <- gg_ccht_bmp_vln$data |> count(feature, ident, split)
vln_plot_n |> distinct(ident, split, n) |> arrange(n)
```


# --- MEBOCOST
## Load communication table
```{r}
# Load the communication table
mebocost_comm <- read.csv(MEBOCOST_COMM_CSV)

mebocost_comm %>% distinct(Condition, Sender)

# Filter for significant communications to White Adipocytes in Deep and Superficial
mebocost_comm_filtered <- mebocost_comm %>%
  rename("neck_region" = "Condition") %>%
  mutate(
    Sender = str_remove(Sender, "^.*~ "),
    Receiver = str_remove(Receiver, "^.*~ ")
  ) %>%
  filter(
    neck_region != "Intermediate",
    permutation_test_fdr <= 0.05,
    str_detect(Receiver, "White")
  )
```

## --- Plot Supp E: the most frequent metabolites to White Adipocytes ---
```{r}
# Most frequent metabolite to White Adipocytes per depot
graph.data <- mebocost_comm_filtered %>%
  count(Metabolite_Name, neck_region) %>%
  arrange(n) %>%
  mutate(Metabolite_Name = factor(Metabolite_Name, levels = unique(Metabolite_Name)))
graph.data

# Plotting
# Use position = position_dodge()
gg_mebocost_top_metab <- ggplot(graph.data, aes(x = Metabolite_Name, y = n)) +
  geom_col(aes(color = neck_region, fill = neck_region), position = position_dodge(0.8), width = 0.7) +
  # scale_y_continuous(limits = c(0, max.y*1.05)) +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(size = 10),
    legend.position = "top"
  ) +
  scale_color_manual(values = neck.color) +
  scale_fill_manual(values = neck.color) +
  labs(
    title = "Top metabolite signaling to White Adipocytes",
    y = "Number of communications",
    x = "Metabolite",
    fill = "Neck region",
    colour = "Neck region"
  ) +
  theme(plot.margin = unit(c(0, 0, 0, 0), units = "in"))

gg_mebocost_top_metab <- gg_mebocost_top_metab + coord_flip()
gg_mebocost_top_metab

ggsave(filename = file.path(OUTPUT_DIR, "S3E_mebocost_top_wad_metabolites.pdf"), plot = gg_mebocost_top_metab,  width = 50, height = 90, units = "mm", dpi = 300)
```
# --- Plot Suppp F: DotPlot of metabolites to White Adipocytes ---
```{r}
# Annotate sender region and link metabolite and sensor for plotting
graph.data <- mebocost_comm_filtered %>%
  mutate(
    Sender = paste0(Sender, " (", neck_region, ")"),
    Metabolite_Name = paste(Metabolite_Name, Sensor, sep = "->")
  ) |> 
  filter(str_detect(Metabolite_Name, 
                    paste0(c(
                      # "Liothyronine",
                      # "Thyroxine",
                      "Dehydroepiandrosterone"
                      ), collapse = "|")))

# unique x-axis categories in order
sender_levels <- levels(factor(graph.data$Sender))

# Dataframe to add background boxes
box_df <- data.frame(
  Sender = sender_levels,
  x_center = seq_along(sender_levels),
  x_min = seq_along(sender_levels) - 0.5,
  x_max = seq_along(sender_levels) + 0.5,
  neck_region = graph.data$neck_region[match(sender_levels, graph.data$Sender)]
)

# Define the changing points by comparing left_shifted and right_shifted x_axist vector
cell_type_from_sender <- sub(" \\(.*", "", sender_levels)
change_points <- which(cell_type_from_sender[-1] != cell_type_from_sender[-length(cell_type_from_sender)]) + 0.5

gg_mebocost_wad_comm_dotplot <- ggplot(graph.data, aes(x = Sender, y = Metabolite_Name)) +

  geom_vline(
    xintercept = change_points,
    linetype = "dashed",
    color = "black",
    linewidth = 0.3,
    alpha = 0.6
  ) +

  # --- Points ---
  geom_point(aes(fill = log2(Norm_Commu_Score), size = log2(Norm_Commu_Score)), shape = 21) +
  scale_fill_viridis_c(option = "H", name = "Communication\nScore (log2)") +

  # --- Theming ---
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    legend.position = "right"
  ) +
  labs(
    title = "MEBOCOST Communication Score",
    subtitle = "White Adipocytes as Receivers",
    y = NULL
  )
```

### --- Adjust and save ---
```{r, fig.width=3.543307, fig.height=1.574803}
mm_to_in(c(90, 40))
gg_mebocost_wad_comm_dotplot <- gg_mebocost_wad_comm_dotplot + theme(axis.text.x = element_text(angle = 30, hjust = 1, vjust = 1),
                                                              panel.border = element_rect(color = "black", fill = NA),
                                                            #  # legend.position = "none",
                                                            #  # plot.title = element_blank(),
                                                            #  # plot.subtitle = element_blank(),
                                                            # plot.margin = margin(30, 1, 1, 1)
                                                            )
gg_mebocost_wad_comm_dotplot
ggsave(filename = file.path(OUTPUT_DIR, "S3C_mebocost_white_adipocytes_comm_dotplot.pdf"), plot = gg_mebocost_wad_comm_dotplot,  width = 90, height = 40, units = "mm", dpi = 300)
```
## --- Plot G: Mebocost relative communication flow ---
```{r, fig.width=2.440945, fig.height= 4.429134}
# Calculate communication score by adding communication for each metabolite. by neck region
relative_comm_flow <- mebocost_comm_filtered %>%
  group_by(Metabolite_Name, neck_region) %>%
  summarise(sum_commu_score = sum(Commu_Score)) %>%
  group_by(Metabolite_Name) %>%
  mutate(relative_score = round(sum_commu_score / sum(sum_commu_score), 3))

# Order from most Deep enriched to most Superficial enriched
order <- relative_comm_flow %>%
  filter(str_detect(neck_region, "Deep")) %>%
  arrange(relative_score) %>%
  pull(Metabolite_Name)
relative_comm_flow <- relative_comm_flow %>% mutate(Metabolite_Name = factor(Metabolite_Name, levels = order))

# Plot stacked proportions
gg_mebo_comm_flow_stacked <- ggplot(relative_comm_flow, aes(x = Metabolite_Name, y = relative_score)) +
  geom_col(aes(fill = neck_region)) +
  geom_hline(yintercept = .5, linetype = "dashed", color = "grey50") +
  theme_nature_metabolism() +
  scale_fill_manual(values = neck.color, name = NULL) +
  labs(title = "Metabolite Signaling", subtitle = "WAd as Receiver", y = "Relative information flow", x = NULL) +
  theme(legend.position = "top", 
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

### --- Adjust and save ---
```{r, fig.width=2.165354, fig.height= 2.318898}
mm_to_in(c(55, 58.9))
gg_mebo_comm_flow_stacked <- gg_mebo_comm_flow_stacked 

gg_mebo_comm_flow_stacked

ggsave(filename = file.path(OUTPUT_DIR, "3G_mebocost_wad_relative_information_flow_stacked.pdf"), plot = gg_mebo_comm_flow_stacked,  width = 55, height = 58.9, units = "mm", dpi = 300)
```

## --- Plot Supp D: Mebocost relative communication flow (Raw) ---
```{r, fig.width=2.440945, fig.height= 4.429134}
# Order from strongest communication
order <- relative_comm_flow %>%
  group_by(Metabolite_Name) |> 
  summarise(comm_strengt_total = sum(sum_commu_score)) |> 
  arrange(comm_strengt_total) |> 
  pull(Metabolite_Name)

relative_comm_flow <- relative_comm_flow %>% mutate(Metabolite_Name = factor(Metabolite_Name, levels = order))


# Plot dodged raw communication scores
gg_mebo_comm_flow_dodged <- ggplot(
  relative_comm_flow |> mutate(Metabolite_Name = str_wrap(Metabolite_Name, width = 80)),
  aes(x = Metabolite_Name, y = sum_commu_score)
) +
  geom_col(aes(fill = neck_region), position = position_dodge()) +
  theme_nature_metabolism() +
  scale_fill_manual(values = neck.color, name = NULL) +
  labs(
    title = "MEBOCOST\nInformation Flow",
    subtitle = "WAds as Receivers",
    y = "Information flow", x = NULL
  ) +
  scale_x_discrete(limits = str_wrap(order, width = 80)) +
  theme(legend.position = "top") +
  coord_flip(clip = "off")
```

### --- Adjust and save ---
```{r, fig.width=1.574803, fig.height= 4.724409}
mm_to_in(c(40, 120))
gg_mebo_comm_flow_dodged <- gg_mebo_comm_flow_dodged + theme(axis.text.x = element_text(angle = 270, vjust = 0.5),
                                                             axis.text.y = element_text(angle = 315, hjust = 1, vjust = 1),
                                                              panel.border = element_rect(color = "black", fill = NA),
                                                             legend.position = "none",
                                                             plot.title = element_blank(),
                                                             plot.subtitle = element_blank(),
                                                            plot.margin = margin(30, 1, 1, 1)
                                                            )
gg_mebo_comm_flow_dodged

ggsave(filename = file.path(OUTPUT_DIR, "S3D_mebocost_wad_information_flow_dodged.pdf"), plot = gg_mebo_comm_flow_dodged,  width = 40, height = 120, units = "mm", dpi = 300)
```

# --- Plot I: CirclePlot of Mebocost ---


```{r, fig.width=2.165354, fig.height= 1.968504}
meta <- s.object@meta.data

celltype_map <- meta %>%
  dplyr::select(cell_type, cell_type_short) %>%
  dplyr::distinct(cell_type, .keep_all = TRUE) %>%
  tibble::deframe()

rename_celltypes <- function(x, map) {
  x_clean <- stringr::str_remove(x, ".*~ ")
  dplyr::coalesce(unname(map[x_clean]), x_clean)
}

plot_list <- list()

for(metabolite_name in c("Dehydroepiandrosterone", "Liothyronine", "Thyroxine")){
# --- Get global maximum edge weight across both depots ---
max_edge <- mebocost_comm_filtered %>%
  filter(
    str_detect(Metabolite_Name, metabolite_name),
    str_detect(neck_region, "Deep|Superficial"),
    str_detect(Receiver, "White Adipo")
  ) %>%
  group_by(Sender, Receiver, Metabolite_Name, neck_region) %>%
  summarise(edge_weight = sum(Commu_Score), .groups = "drop") %>%
  pull(edge_weight) %>%
  max()



for (depot in c("Deep", "Superficial")) {
  # Filter to create the edges
  edges <- mebocost_comm_filtered %>%
  filter(
    str_detect(Metabolite_Name, metabolite_name),
    str_detect(neck_region, depot),
    str_detect(Receiver, "White Adipo")
  ) %>%
  group_by(Sender, Receiver, Metabolite_Name) %>%
  summarise(edge_weight = sum(Commu_Score), .groups = "drop") %>%
  mutate(
    Sender = rename_celltypes(Sender, celltype_map),
    Receiver = rename_celltypes(Receiver, celltype_map)
  ) %>%
  select(Sender, Receiver, edge_weight)

  # Build graph with all nodes from unfiltered mebocost_comm
all_nodes <- unique(c(mebocost_comm$Sender, mebocost_comm$Receiver))
all_nodes <- rename_celltypes(all_nodes, celltype_map)

nodes_df <- tibble(name = unique(all_nodes))

  g <- graph_from_data_frame(edges, vertices = nodes_df, directed = TRUE)
  g_tbl <- as_tbl_graph(g)

  node_from <- g_tbl %>%
    as_tibble() %>%
    mutate(from = row_number())

  new_g_tbl <- g_tbl %>%
    activate(edges) %>%
    left_join(node_from)

  # filter palette
  mebo.palette <- palette.use[nodes_df$name]
  
  # plot
  mebo_graph <- ggraph(new_g_tbl, layout = "circle") +
    geom_edge_arc(aes(color = name, width = edge_weight),
      arrow = arrow(length = unit(2.5, "mm"), type = "closed"),
      strength = 0.1, show.legend = FALSE
    ) +
    geom_edge_loop(aes(color = name, width = edge_weight),
      arrow = arrow(length = unit(2.5, "mm"), type = "closed"),
      strength = 0.8, show.legend = FALSE
    ) +
    geom_node_point(aes(fill = name), shape = 21, color = "black", size = 2) +
    geom_node_text(aes(label = name), size = 6 *2/ .pt, repel = TRUE) +

    # Use global max_edge for comparability
    scale_edge_width(
      range = c(0.1, 1.5),
      limits = c(0, max_edge),
      name = "Communication\nScore"
    ) +

    scale_edge_colour_manual(values = mebo.palette, guide = "none") +
    scale_fill_manual(values = mebo.palette, guide = "none") +  
    coord_equal(xlim = c(-1.5, 1.5), ylim = c(-1.5, 1.5), clip = "off") +
    theme_void() +
    # labs(title = paste("Dehydroepiandrosterone", depot, sep = "\n")) +
    theme(plot.title = element_text(hjust = 0.5, size = 6 / .pt), plot.title.position = "plot") 

  plot_list[[paste(metabolite_name, depot)]] <- mebo_graph
  ggsave(file.path(OUTPUT_DIR, paste0("3I_mebocost_", tolower(metabolite_name), "_", depot, ".pdf")), mebo_graph, width = 25, height = 25, units = "mm", dpi = 600)
}
}
```

```{r}
## Helper function
make_balanced_label_df <- function(lay, label_radius = 1.22, y_pad = 0.92) {
  
  max_y <- label_radius * y_pad
  
  label_df <- lay %>%
    dplyr::mutate(
      side = dplyr::if_else(x >= 0, "right", "left")
    ) %>%
    dplyr::group_by(side) %>%
    dplyr::arrange(dplyr::desc(y), .by_group = TRUE) %>%
    dplyr::mutate(
      y_lab = seq(from = max_y, to = -max_y, length.out = dplyr::n()),
      x_lab = dplyr::if_else(
        side == "right",
        sqrt(pmax(label_radius^2 - y_lab^2, 0)),
        -sqrt(pmax(label_radius^2 - y_lab^2, 0))
      ),
      hjust = dplyr::if_else(side == "right", 0, 1)
    ) %>%
    dplyr::ungroup()
  
  label_df
}
```

```{r}
meta <- s.object@meta.data

celltype_map <- meta %>%
  dplyr::select(cell_type, cell_type_short) %>%
  dplyr::distinct(cell_type, .keep_all = TRUE) %>%
  tibble::deframe()

rename_celltypes <- function(x, map) {
  x_clean <- stringr::str_remove(x, ".*~ ")
  dplyr::coalesce(unname(map[x_clean]), x_clean)
}

plot_list <- list()

for (metabolite_name in c("Dehydroepiandrosterone", "Liothyronine", "Thyroxine")) {

  ## Global max edge weight across depots for this metabolite
  max_edge <- mebocost_comm_filtered %>%
    dplyr::filter(
      stringr::str_detect(Metabolite_Name, metabolite_name),
      stringr::str_detect(neck_region, "Deep|Superficial"),
      stringr::str_detect(Receiver, "White Adipo")
    ) %>%
    dplyr::group_by(Sender, Receiver, Metabolite_Name, neck_region) %>%
    dplyr::summarise(edge_weight = sum(Commu_Score), .groups = "drop") %>%
    dplyr::pull(edge_weight) %>%
    max(na.rm = TRUE)

  for (depot in c("Deep", "Superficial")) {

    ## Edges for this panel
    edges <- mebocost_comm_filtered %>%
      dplyr::filter(
        stringr::str_detect(Metabolite_Name, metabolite_name),
        stringr::str_detect(neck_region, depot),
        stringr::str_detect(Receiver, "White Adipo")
      ) %>%
      dplyr::group_by(Sender, Receiver, Metabolite_Name) %>%
      dplyr::summarise(edge_weight = sum(Commu_Score), .groups = "drop") %>%
      dplyr::mutate(
        Sender = rename_celltypes(Sender, celltype_map),
        Receiver = rename_celltypes(Receiver, celltype_map)
      ) %>%
      dplyr::select(Sender, Receiver, edge_weight)
    
    edges <- mebocost_comm_filtered %>%
  filter(
    str_detect(Metabolite_Name, metabolite_name),
    str_detect(neck_region, depot),
    str_detect(Receiver, "White Adipo")
  ) %>%
  group_by(Sender, Receiver, Metabolite_Name) %>%
  summarise(edge_weight = sum(Commu_Score), .groups = "drop") %>%
  mutate(
    Sender = rename_celltypes(Sender, celltype_map),
    Receiver = rename_celltypes(Receiver, celltype_map)
  ) %>%
  select(Sender, Receiver, edge_weight) %>%
  arrange(desc(edge_weight))

    ## All nodes, consistently renamed
    all_nodes <- unique(c(mebocost_comm$Sender, mebocost_comm$Receiver))
    all_nodes <- rename_celltypes(all_nodes, celltype_map)

    nodes_df <- tibble::tibble(name = unique(all_nodes))

    ## Graph
    g <- igraph::graph_from_data_frame(edges, vertices = nodes_df, directed = TRUE)
    g_tbl <- tidygraph::as_tbl_graph(g)

    node_from <- g_tbl %>%
      tibble::as_tibble() %>%
      dplyr::mutate(from = dplyr::row_number())

    new_g_tbl <- g_tbl %>%
      tidygraph::activate(edges) %>%
      dplyr::left_join(node_from, by = "from")

    ## Palette aligned to actual node order
    mebo.palette <- palette.use[nodes_df$name]

    ## Build layout once and extract coordinates
    lay <- ggraph::create_layout(new_g_tbl, layout = "circle")

    ## Fixed radial label placement
    label_df <- make_balanced_label_df(lay, label_radius = 1.22, y_pad = 0.999)

    ## Plot
    mebo_graph <- ggraph::ggraph(lay) +
      ggraph::geom_edge_arc(
        aes(color = name, width = edge_weight),
        strength = 0.1,
        show.legend = FALSE
      ) +
      ggraph::geom_edge_loop(
        aes(color = name, width = edge_weight),
        strength = 0.8,
        show.legend = FALSE
      ) +
      ggraph::geom_node_point(
        aes(fill = name),
        shape = 21,
        color = "black",
        size = 2
      ) +
      ggplot2::geom_text(
        data = label_df,
        mapping = ggplot2::aes(
          x = x_lab,
          y = y_lab,
          label = name,
          hjust = hjust
        ),
        size = 6 / .pt,
        inherit.aes = FALSE
      ) +
      ggraph::scale_edge_width(
        range = c(0.1, 1.5),
        limits = c(0, max_edge),
        name = "Communication\nScore"
      ) +
      ggraph::scale_edge_colour_manual(values = mebo.palette, guide = "none") +
      ggplot2::scale_fill_manual(values = mebo.palette, guide = "none") +
      coord_equal(
        xlim = c(-1.55, 1.55),
        ylim = c(-1.55, 1.55),
        clip = "off"
      ) + theme_void() + theme(
  plot.margin = margin(1, 12, 1, 19))

    plot_list[[paste(metabolite_name, depot)]] <- mebo_graph

    ggplot2::ggsave(
      filename = file.path(
        OUTPUT_DIR,
        paste0("3I_mebocost_", tolower(metabolite_name), "_", depot, ".pdf")
      ),
      plot = mebo_graph,
      width = 40,
      height = 40,
      units = "mm",
      dpi = 600
    )
  }
}
```


```{r, fig.width=1.574803, fig.height= 1.574803}
mm_to_in(40)
plot_list
```



