---
title: "3-FigurePlotting"
author: "Henrique"
date: "2025-07-17"
output: html_document
---
# ---- Setup ----
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(Seurat)
library(SeuratObject)
library(ggplot2)
library(RColorBrewer)
library(VISION)
library(Matrix)
library(dplyr)
library(CellChat)
library(circlize) ### A tutorial on how to create your own circlize can be found here: https://r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling.html
library(ComplexHeatmap)
library(NMF)
library(stringr)
library(readr)
library(viridis)
library(tidyverse)
library(patchwork)

# OBS:  theme_nature_metabolism() and color palettes set on "scripts/0.environment_setup/snRNAseq_graphics_setup.R"
source(here("scripts/0.environment_setup/snRNAseq_graphics_setup.R"))

# --- Resolve conflicts ---
conflicts_prefer(pheatmap::pheatmap)


# --- Setup paths ---
## Inputs
DESEQ_STATS_CSV <- here("output/4.downstream_pipelines/de_analysis/data/deseq2_stats.csv")
SEURAT_INPUT <- here("data/adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds")
VISION_OBJECTS_LIST <- list(
  here("output/4.downstream_pipelines/vision/data/vision_objects/CamaraH_Dataset_VISION.rds")
  )


# Inputs to update
HEATMAP_LOG_CPM_RDS <- here("output/4.downstream_pipelines/de_analysis/data/CamaraH_pseudobulk_log_CPM_count_matrix_for_heatmap.rds")
HEATMAP_METADATA_RDS <- here("output/4.downstream_pipelines/de_analysis/data/CamaraH_metadata_for_heatmap.rds")
HEATMAP_CTS_RDS <- here("output/4.downstream_pipelines/de_analysis/data/CamaraH_pseudobulk_count_matrix_for_ggplot.rds")

A41_CTS_CSV <- here("data/6.internal_datasets/Tseng_Datasets/Supplementary_Table_a41_fpkm.csv")
A38_CTS_CSV <- here("data/6.internal_datasets/Tseng_Datasets/Supplementary_Table_a38_tpm.xlsx")
CERO_CTS_XLSX <- here("data/5.published_datasets/Public_BulkRNAseq/Cero 2023/RNA-seq data set Cero et al.xlsx")


BULK_DIR <- here("output/3.heat_signature/3.public_bulkrnaseq/data/Cleaned_Datasets")
BULK_META_DIR <- here("output/3.heat_signature/3.public_bulkrnaseq/data/Experimental_Design")


# Outputs
OUTPUT_DIR <- here("output", "5.figure_preparation", "Figure_4")

dir.create(OUTPUT_DIR, showWarnings = F, recursive = T)
```


# ---- Heatmap Plots ----
## ---- Load Pseudobulk expression matrix for heatmap ----
```{r}
mtx <- readRDS(HEATMAP_LOG_CPM_RDS)
```
## ---- load meadata ----
```{r}
hm_meta <- readRDS(HEATMAP_METADATA_RDS)
```
## --- Helper function to filter Matrix ---
```{r def-filt-func}
  # Define function to filter the matrix
  filterHmMatrix <- function(matrix, genes, sample_pattern, negate_pattern = "none") {
    #If more then 1 term is passed as argument, collapse for search
    if(length(sample_pattern) > 1){
      sample_pattern <- paste(sample_pattern, collapse = "|")
    }
    # Filter the matrix
    matrix_filter <- matrix[genes, (str_detect(colnames(matrix), sample_pattern) &
                                      !str_detect(colnames(matrix), negate_pattern)
                                    )
                            ]
    return(matrix_filter)
  }
```

```{r}
hm_data <- filterHmMatrix(mtx, c(
    "PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", 
    "HOXA3", "HOXA5", "HOXA6"
  ), sample_pattern = c("White", "Progenitor", "Endothelial", "Smooth Muscle"), 
  negate_pattern = "Intermediate")
```

## ---- Plot Supp D: Hox Genes HeatMap: White Adipocytes and Adipocyte Progenitors ----
```{r}
cell_type_of_interest <-  c("White", "Progenitor", "Endothelial", "Smooth Muscle")
# Determine cell type patterns to filter for
cell_type_of_interest <- c("White", "Progenitor")

# Filter expression matrix
  hm_data <- filterHmMatrix(mtx, 
                            c(
                              "PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", 
                              "HOXA3", "HOXA4", "HOXA5", "HOXA6"
                              ),
                            sample_pattern = cell_type_of_interest,
                            negate_pattern = "Intermediate")
  
  
  # Create metadata color mapping
  ## Indicate the columns in hm_meta that you want colors for
  cols_for_colors <- c("cell_type", "neck_region")

  ## Filter metadata 
  cell_type_filter <- paste(cell_type_of_interest, collapse = "|")
  annotation_col <- hm_meta |> 
    filter(str_detect(cell_type, cell_type_filter),
           neck_region != "Intermediate") |> 
    select(all_of(cols_for_colors))

  # Bulit annotation_color list
  annotation_colors <- lapply(cols_for_colors, function(col) {
    vals <- unique(annotation_col[[col]])
    setNames(
      ifelse(vals %in% names(c(palette.use, neck.color)), c(palette.use, neck.color)[vals], "grey"),
      vals
    )
  })
  names(annotation_colors) <- cols_for_colors
  
    #Set up size for saving
  ncol <- ncol(hm_data)
  nrow <- nrow(hm_data)

  target_width_mm <- 180
  target_height_mm <- 35

  cellwidth <- max(4, target_width_mm / ncol)
  cellheight <- max(5, (target_height_mm - 6) / nrow)    # subtract a bit for annotations
  
  
    pdf(file.path(OUTPUT_DIR,paste0("S4D_heatmap_hox_wad_aspcs.pdf")), width = mm_to_in(180), height = mm_to_in(35))
    pheatmap::pheatmap(
    mat = hm_data,
    annotation_col = annotation_col,
    annotation_colors = annotation_colors,
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    scale = "row", 
    cellwidth = cellwidth, cellheight = cellheight,
    fontsize = 6,
    gaps_row = 5,
    gaps_col = annotation_col |> count(cell_type, neck_region) |> pull(n) |> cumsum(),
    show_colnames = F
  )
  dev.off()
  
  # Plot in the markdown file
    pheatmap::pheatmap(
    mat = hm_data,
    annotation_col = annotation_col,
    annotation_colors = annotation_colors,
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    scale = "row", 
    cellwidth = cellwidth, cellheight = cellheight,
    fontsize = 6,
    gaps_row = 5,
    gaps_col = annotation_col |> count(cell_type, neck_region) |> pull(n) |> cumsum(),
    show_colnames = F
  )
```
```{r}
heatmap_aspc_adipo_n <- annotation_col |> count(cell_type, neck_region)
heatmap_aspc_adipo_n
```

## ---- Plot Supp E: Hox Genes HeatMap: Endothelial and Smooth Muscle Cells ----
```{r}
# Determine cell type patterns to filter for
cell_type_of_interest <-  c("Endothelial", "Smooth")

# Filter expression matrix
  hm_data <- filterHmMatrix(mtx, 
                            c(
                              "PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", 
                              "HOXA3", "HOXA4", "HOXA5", "HOXA6"
                              ),
                            sample_pattern = cell_type_of_interest,
                            negate_pattern = "Intermediate")
  
  
  # Create metadata color mapping
  ## Indicate the columns in hm_meta that you want colors for
  cols_for_colors <- c("cell_type", "neck_region")

  ## Filter metadata 
  cell_type_filter <- paste(cell_type_of_interest, collapse = "|")
  annotation_col <- hm_meta |> 
    filter(str_detect(cell_type, cell_type_filter),
           neck_region != "Intermediate") |> 
    select(all_of(cols_for_colors))

  # Bulit annotation_color list
  annotation_colors <- lapply(cols_for_colors, function(col) {
    vals <- unique(annotation_col[[col]])
    setNames(
      ifelse(vals %in% names(c(palette.use, neck.color)), c(palette.use, neck.color)[vals], "grey"),
      vals
    )
  })
  names(annotation_colors) <- cols_for_colors
  
  #Set up size for saving
  ncol <- ncol(hm_data)
  nrow <- nrow(hm_data)

  target_width_mm <- 180
  target_height_mm <- 35

  cellwidth <- max(4, target_width_mm / ncol)
  cellheight <- max(5, (target_height_mm - 6) / nrow)    # subtract a bit for annotations

  
  
    pdf(file.path(OUTPUT_DIR,paste0("S4E_heatmap_hox_endo_smc_aspcs.pdf")), width = mm_to_in(180), height = mm_to_in(35))
    pheatmap::pheatmap(
    mat = hm_data,
    annotation_col = annotation_col,
    annotation_colors = annotation_colors,
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    scale = "row", 
    cellwidth = cellwidth, cellheight = cellheight,
    fontsize = 6,
    gaps_row = 5,
    gaps_col = annotation_col |> count(cell_type, neck_region) |> pull(n) |> cumsum(),
    show_colnames = F
  )
  dev.off()
  
  # Plot in the markdown file
    pheatmap::pheatmap(
    mat = hm_data,
    annotation_col = annotation_col,
    annotation_colors = annotation_colors,
    cluster_cols = FALSE,
    cluster_rows = FALSE,
    scale = "row", 
    cellwidth = cellwidth, cellheight = cellheight,
    fontsize = 6,
    gaps_row = 5,
    gaps_col = annotation_col |> count(cell_type, neck_region) |> pull(n) |> cumsum(),
    show_colnames = F
  )
```
```{r}
heatmap_endo_smc_n <- annotation_col |> count(cell_type, neck_region)
heatmap_endo_smc_n
```

## ---- Plot A: HOX Volcanos ---
## --- Read DESeq2 DE data.frame ---
```{r read-deseq-df, fig.width= 1.77, fig.height= 1.42}
final_DE <- read.csv(DESEQ_STATS_CSV)

final_DE_filt <- final_DE %>% dplyr::filter(padj <= 0.05, contrast.2 != "Intermediate", contrast.1 != "Intermediate")


# This adds a categorisation based on whether the log2FoldChange is up or down
final_DE_filt$direction <- ifelse(final_DE_filt$log2FoldChange > 0, "Up in Deep", "Up in Superficial")

final_DE %>% distinct(cell_type)
```

## --- Loop and create plots  ---
```{r}
library(ggrepel)

plot_list <-  list()
for(filter_cell in c("Adipocyte Progenitor Cell", "White Adipocytes")){

df <- final_DE %>% filter(str_detect(cell_type, filter_cell),
                    contrast.1 == "Deep", 
                    contrast.2 == "Superficial")

fc_cutoff = 0.5
p_cutoff = 0.05

  # Categorize points
  volcano_df <- df %>%
    mutate(
      log2FoldChange = -log2FoldChange,
      label = ifelse(str_detect(gene, "^HOX|^IRX|^PAX") & padj < 0.05, gene, NA),
      dev_gen_cat = case_when(str_detect(gene, "^HOX") ~ "HOX",
                              str_detect(gene, "^IRX") ~ "IRX",
                              str_detect(gene, "^PAX") ~ "PAX",
                              TRUE ~ NA),
      transparent = is.na(dev_gen_cat),
      label_color = ifelse(padj < 0.05 & log2FoldChange < 0, neck.color["Deep"], neck.color["Superficial"])
      ) %>%
    arrange(desc(is.na(dev_gen_cat)), dev_gen_cat)
  
  # Extract FC for symetric x-axis
  maxFC <- max(abs(volcano_df$log2FoldChange))

  # Determine color pallete for the HOX genes  
  fill.color = setNames(c(brewer.pal(3, name = "Set2"), "grey"), c("HOX", "PAX", "IRX", NA))

  # Volcano plot
  gg_volcano <- ggplot(volcano_df, aes(x = log2FoldChange, y = -log10(padj))) +
    geom_vline(xintercept = c(-fc_cutoff, fc_cutoff), linetype = "dashed", color = "black", linewidth = 0.1) +
    geom_hline(yintercept = -log10(p_cutoff), linetype = "dashed", color = "black", linewidth = 0.1) +
    geom_point(aes(fill = dev_gen_cat, alpha = transparent) , size = 1, color = "black", stroke = 0.1, shape = 21) +
    annotate("text", x = -Inf, y = Inf, label = "Up in Deep", hjust = -0.1, vjust = 2, size = 5/.pt, color = neck.color["Deep"]) +
    annotate("text", x =  Inf, y = Inf, label = "Up in Superficial",  hjust = 1, vjust = 2, size = 5/.pt, color = neck.color["Superficial"]) +
    scale_alpha_manual(values = c(1, 0.1), guide = "none") +
    labs(
      title = paste0("Developmental Patterning Genes"),
      subtitle = paste0(filter_cell,": Deep vs Superficial"),
      x = paste0("Log2 fold-change"),
      y = "-log10 (adjusted p-value)"
    ) +
    scale_x_continuous(limits = c(-maxFC, maxFC)) +
    scale_fill_manual(values = fill.color, name = "Gene Cateogry") + 
    theme_nature_metabolism() +
    theme(legend.position = "right")
  
    # Add gene labels
  gg_volcano <- gg_volcano +
    geom_text_repel(
      aes(label = label),
      max.overlaps = 30,
      size = 4.5 / .pt,
      box.padding = 0.4,
      point.padding = 0.3,
      segment.color = "grey50", 
      segment.size = 0.1,
      segment.alpha = 1,   # force all segments to be drawn
  force_pull = 0  ,     # prevents shortening to zero
      na.rm = TRUE
    ) 
  
  plot_list[[filter_cell]] <-  gg_volcano
}
```
## ---- Plot White Adipocytes HOX ---
```{r, fig.width=2.165354, fig.height=2.519685}
mm_to_in(c(55, 64))
  gg_volcano <-  plot_list[["White Adipocytes"]]
  gg_volcano <- gg_volcano + plot_annotation(tag_levels = list("d")) + theme(plot.tag = element_text(size = 6, face = "bold"))
  gg_volcano   
  
  # Save as png
ggsave(file.path(OUTPUT_DIR, "4D_volcano_white_adipocyte_hox.pdf"), gg_volcano, width = 55, height = 64, units = "mm", bg = "white",  dpi = 600)

  # Save as pdf
  output_file <- paste0("4D_volcano_white_adipocyte_hox.pdf")
  pdf(file.path(OUTPUT_DIR, "4D_volcano_white_adipocyte_hox.pdf"), width = mm_to_in(55), height = mm_to_in(64))
  gg_volcano
  dev.off()
```
## ---- Plot Adipocyte Progenitors HOX ---
```{r, fig.width=2.657480, fig.height=1.181102}
mm_to_in(c(67.5, 30))
  gg_volcano <-  plot_list[["Adipocyte Progenitor Cell"]]
  gg_volcano <- gg_volcano + 
    plot_annotation(tag_levels = list(" ")) + 
    scale_y_continuous(expand = expansion(c(0.05,0.55))) 
  gg_volcano   
  
  # Save as png
  output_file <- paste0("S4B_volcano_aspc_hox.pdf")
ggsave(file.path(OUTPUT_DIR, output_file), gg_volcano, width = 67.5, height = 30, units = "mm", bg = "white",  dpi = 600)
```

# ---- Load Pseudobulk expression table for ggplot (long format) ---- 
```{r}
counts_table <- read_rds(HEATMAP_CTS_RDS)
```



## ---- Define genes of interest ---- 
```{r}
genes_of_interest <- c(
  "MECOM", "EBF2", "PRDM16", "COBL", "PPARGC1A",
  "LEP", "SLC7A10", "ADIPOQ", "PLIN1", "PLIN4", "FABP4",
  # "CKMT2", "GK", "CA12", "BDH1",
  "PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", 
  "HOXA3", "HOXA4", "HOXA5", "HOXA6",
  "ALPL", "CKMT1A", "CKMT1B", "CKMT2", "CKB",
  "UCP1"
)

all(genes_of_interest %in% counts_table$gene)
```
## --- Cheat sheet l1p conversions---
```{r}
log1p_df <- data.frame("log1p_val" = c(-1,0,0.5, 1:15)) %>%
  mutate(orig_val = round(expm1(log1p_val),2))

log1p_df
```
## --- Load DESeq2 comparisons ---
```{r}
pseudobulk_deseq_dge <- read.csv(DESEQ_STATS_CSV)
head(pseudobulk_deseq_dge)
```


## --- Plot Supp A: Plot Gene Expression per subject/cell_type pair ----
```{r}
plot_list <- list()

for (plot_cells in c("ASPC", "adipo")) {
  negate_search <- plot_cells != "ASPC"

  # ensure bucket exists
  if (is.null(plot_list[[plot_cells]])) plot_list[[plot_cells]] <- list()

  for (goi in genes_of_interest) {

    plot_data <- counts_table %>%
      filter(gene == goi) %>%
      filter(
        neck_region != "Intermediate",
        str_detect(cell_type, "Progenitor") |
          str_detect(cell_type, "White") |
          (str_detect(cell_type, "Brown") & (neck_region == "Deep")),
        n_cells >= 10
      ) %>%
      mutate(
        neck_region_short = str_replace(neck_region, "Deep", "DN"),
        neck_region_short = str_replace(neck_region_short, "Superficial", "SC"),
        cell_type_short = factor(cell_type_short, levels = c("ASPC", "WAds", "BAds")),
        neck_region_short = factor(neck_region_short, levels = c("SC", "DN")),
        region_cell_type_short = paste(neck_region_short, cell_type_short, sep = "\n")
      )

    max.y.pos <- max(plot_data$cpm_log, na.rm = TRUE) * 1.15

    stat_df <- pseudobulk_deseq_dge %>%
      filter(
        gene == goi,
        contrast.1 == "Deep",
        contrast.2 == "Superficial",
        str_detect(cell_type, "White") |
          str_detect(cell_type, "Progenitor") |
          str_detect(cell_type, "Brown")
      ) %>%
      mutate(
        padj = if_else(is.na(padj), 1.0, padj),
        p.format = if_else(padj < 0.01, "<0.01", as.character(round(padj, 2))),
        fontface = if_else(padj <= 0.05, "bold", "plain"),
        group1 = str_replace(contrast.1, "Deep", "DN"),
        group2 = str_replace(contrast.2, "Superficial", "SC"),
        y.position = max.y.pos
      ) %>%
      left_join(plot_data %>% distinct(cell_type, cell_type_short),
                by = "cell_type", multiple = "first")
    
    
    # Remove ASPCS from thermogenic genes plot
    if (goi %in% c("MECOM","EBF2","PRDM16","COBL","PPARGC1A",
                   "ADIPOQ","PLIN1","PLIN4","FABP4", "LEP", "SLC7A10",
                   "ALPL","CKMT1A","CKMT1B","CKMT2","CKB","UCP1")) {

      plot_data <- plot_data %>%
        filter(str_detect(cell_type_short, "ASPC", negate = negate_search))

      stat_df <- stat_df %>%
        filter(str_detect(cell_type, "Progenitor", negate = negate_search))
    }

    p <- ggplot(plot_data, aes(x = neck_region_short, y = cpm_log)) +
      geom_line(aes(group = subject), alpha = 0.5, linewidth = 0.1) +
      geom_point(aes(fill = cell_type), size = 1, shape = 21, color = "black") +
      facet_grid(. ~ cell_type_short, scales = "free_x", space = "free_x") +
      scale_x_discrete(drop = TRUE) +
      theme_nature_metabolism(base_size = 6) +
      theme(
        strip.background = element_blank(),
        strip.text = element_text(size = 5, face = "bold")
      ) +
      scale_fill_manual(values = palette.use, guide = NULL) +
      labs(title = goi, y = "Mean expression\n(log1p CPM)", x = NULL) +
      coord_cartesian(clip = FALSE)

    # only add p-value brackets if they exist and are usable
    stat_df_ok <- nrow(stat_df) > 0 &&
      all(c("group1","group2","y.position","p.format","fontface") %in% names(stat_df)) &&
      !all(is.na(stat_df$group1)) &&
      !all(is.na(stat_df$group2))

    if (stat_df_ok) {
      p <- p +
        stat_pvalue_manual(
          data = stat_df,
          xmin = "group1",
          xmax = "group2",
          inherit.aes = FALSE,
          label = "p.format",
          fontface = "fontface",
          size = 5 / ggplot2::.pt,
          y.position = "y.position",
          bracket.size = 0.2,
          tip.length = 0
        )
    }

    plot_list[[plot_cells]][[goi]] <- p
  }
}

```


## --- Plot A: Thermogenic genes ---
```{r, warning=F, message=FALSE, fig.width=4.049492, fig.height=1.181102}
mm_to_in(c(102.8571, 30))
for(plot_cells in c("ASPC", "adipo")){
# Plot and make small adjustments
gg_thermo_lineplot <- wrap_plots(plot_list[[plot_cells]][c("MECOM", "EBF2", "PRDM16", "PPARGC1A")], ncol = 4) + 
  plot_annotation(
                  title = "Thermogenic Transcriptional Regulators", 
                  theme = theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, margin = margin(b =0)))
                  )  & 
  scale_y_continuous(expand = expansion(c(0.05,0.15))) & 
  theme(axis.text.x = element_text(size = 5)) +
  theme(plot.tag = element_text(size = 6, face = "bold"))

print(gg_thermo_lineplot )

if(plot_cells == "ASPC"){
ggsave(file.path(OUTPUT_DIR,"S4A_thermo_lineplot_aspc.pdf"), gg_thermo_lineplot, width = 102.8571, height = 30, units = "mm", bg = "white", dpi = 600)
} else if(tolower(plot_cells) == "adipo"){
    ggsave(file.path(OUTPUT_DIR,"4A_thermo_lineplot_adipo.pdf"), gg_thermo_lineplot, width = 102.8571, height = 30, units = "mm", bg = "white", dpi = 600)
}
}
```

```{r}
for(plot_cells in c("ASPC", "adipo")){
plot_n <- plot_list[[plot_cells]]$HOXA3$data |> count(neck_region, cell_type_short)
print(plot_n)
}
```

## --- Plot B: White Adipocyte genes ---
```{r, warning=F, message=FALSE, fig.width=3.037122, fig.height=1.181102}
mm_to_in(c(77.1429, 30))

gg_adipo_lineplot <- wrap_plots(plot_list[["adipo"]][c("PLIN4", "LEP", "SLC7A10")], ncol = 3) + 
  plot_annotation(title = "White Adipocyte-Enriched Genes", 
                  theme = theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, margin = margin(b =0)))
                  )  & 
  scale_y_continuous(expand = expansion(c(0.05,0.15))) & 
  theme(axis.text.x = element_text(size = 5))

gg_adipo_lineplot
ggsave(file.path(OUTPUT_DIR,"4C_adipo_lineplot.pdf"), gg_adipo_lineplot, width = 77.1429, height = 30, units = "mm", bg = "white",  dpi = 600)
```

## --- Plot E: Deep HOX genes ---
```{r, warning=F, message=FALSE, fig.width=4.921260, fig.height=1.181102}
mm_to_in(c(125, 30))

gg_hox_deep_lineplot <- wrap_plots(plot_list[["adipo"]][c("HOXA3", "HOXA4", "HOXA5", "HOXA6")], ncol = 4) + 
  plot_annotation(title = "Deep-Enriched Developmental Patterning Genes", 
                  theme = theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, margin = margin(b =0)))
                  )  & 
  scale_y_continuous(expand = expansion(c(0.05,0.15))) & 
  theme(axis.text.x = element_text(size = 5), strip.text = element_text(size = 5)) + theme(plot.tag = element_text(size = 6, face = "bold"))
gg_hox_deep_lineplot

ggsave(file.path(OUTPUT_DIR,"4E_hox_deep_lineplot.pdf"), gg_hox_deep_lineplot, width = 125, height = 30, units = "mm", bg = "white",  dpi = 600)
```

## --- Plot F: Superficial HOX genes ---
```{r, warning=F, message=FALSE, fig.width=7.086614, fig.height=1.181102}
mm_to_in(c(180, 30))
gg_hox_sup_lineplot <- wrap_plots(plot_list[["adipo"]][c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5")], ncol = 5) + 
  plot_annotation(title = "Superficial-Enriched Developmental Patterning Genes", 
                  theme = theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, margin = margin(b =0)))
                  )  & 
  scale_y_continuous(expand = expansion(c(0.05,0.15))) & 
  theme(axis.text.x = element_text(size = 5), strip.text = element_text(size = 5)) + theme(plot.tag = element_text(size = 6, face = "bold"))
gg_hox_sup_lineplot

ggsave(file.path(OUTPUT_DIR,"4F_hox_sup_lineplot.pdf"), gg_hox_sup_lineplot, width = 180, height = 30, units = "mm", bg = "white",  dpi = 600)
```


# --- VISION in White Adipocytes
## --- Load seurat
```{r}
s.object <- readRDS(SEURAT_INPUT)
```
## --- Laod VISION

``` {r}
# Load VISIOn files
vision_obj_list <- lapply(VISION_OBJECTS_LIST, readRDS)
```

## --- Extract signatures to generate graph data

``` {r}
# Extract umap form the first VISION object
UMAP <- as.data.frame(getProjections(vision_obj_list[[1]])[["Seurat_umap"]]) %>% rownames_to_column("barcode")

# Load objects and extract signatures
signatures <- lapply(vision_obj_list, getSignatureScores)
signatures <- bind_cols(signatures)
names(signatures) <- gsub("\\.+", "_", names(signatures))
names(signatures) <- gsub(" ", "_", names(signatures))

# Extract metadata
meta <- s.object@meta.data %>%
  select(cell_type, cell_type_short, sample, neck_region, subject) %>%
  rownames_to_column("barcode")

# generate combined final data.frame
graph.data <- bind_cols(UMAP, signatures) %>% left_join(meta)
```

## ---Pivot signatures
```{r}
sig_names <- names(signatures)

graph.data.long <- pivot_longer(graph.data, cols = all_of(sig_names), names_to = "Hallmark", values_to = "sigScores")
```

## --- Rename signatures
```{r}
graph.data.renamed <- graph.data.long %>%
  mutate(
    Hallmark = str_replace_all(Hallmark, "_", " "),
    Hallmark = str_replace_all(Hallmark, "SIGNATURE", ""),
    Hallmark = str_remove_all(Hallmark, "GOBP "),
    Hallmark = str_remove_all(Hallmark, "HALLMARK "),
    Hallmark = case_when(
      str_detect(Hallmark, "OXIDATIVE PHOSPHORYLATION") ~ "OXPHOS",
      str_detect(Hallmark, "FATTY ACID") ~ "FA METABOLISM",
      str_detect(Hallmark, "G2M") ~ "G2M",
      str_detect(Hallmark, "REACTIVE OXYGEN") ~ "ROS",
      str_detect(Hallmark, "UNFOLDED PROTEIN") ~ "UPR",
      TRUE ~ Hallmark
    )
  )
```


## --- Plot Supp B: UMAP signature
```{r, fig.width=7.09, fig.height=1.77}
round(digits = 2, mm_to_in(c(180, 180 / 4)))

# Determine paths to plot
distinct(graph.data.renamed, Hallmark) %>% arrange(Hallmark)

signatures_of_interest <- c(
  # "CELLULAR RESPIRATION",
  #                   "AEROBIC RESPIRATION",
                    "ADAPTIVE THERMOGENESIS",
                    "BATLAS",
                    "HEAT")

# Filter to plot
graph.data.filter <- graph.data.renamed %>%
  filter(Hallmark %in% signatures_of_interest, cell_type_short %in% c("BAds", "WAds")) |> 
  mutate(Hallmark =factor(Hallmark, levels = signatures_of_interest))

```


```{r, fig.width=2.362205* 1.7, fig.height=0.7874016* 1.7}
mm_to_in(c(60, 20))
scale_factor <- 1.7


hallmark_umap_plot_list <- list()
for(hallmark in unique(graph.data.filter$Hallmark)){
gg_sup_a <- ggplot(graph.data.filter |> 
                     filter(Hallmark == hallmark, neck_region != "Intermediate") |> 
                     mutate(Hallmark = str_wrap(Hallmark, width = 8))) +
  aes(x = umap_1, y = umap_2, color = sigScores) +
  geom_point(size = 0.5, alpha = 0.7) +
  scale_color_viridis(
    option = "viridis",
    limits = c(-0.3, 2.6)
  ) +
  labs(
    x = "UMAP Dimension 1",
    y = "UMAP Dimension 2",
    color = "Signature Scores"
  ) +
  theme_void() +
  theme_nature_metabolism(base_size = 6* scale_factor) +
 theme_no_axis()+
  # remove strip
  theme(
    legend.position = "right",
    strip.text = element_text(size = 5* scale_factor, face = "plain"),
    strip.background = element_rect(fill = "white", colour = "white"), plot = element_blank()
  ) +
  facet_wrap(vars(neck_region)) +
  labs(title = hallmark) +
  coord_cartesian(clip = "off")

hallmark_umap_plot_list[[hallmark]] <- gg_sup_a

ggsave(file.path(OUTPUT_DIR,paste0("S4C_vision_umapplot_", hallmark, ".pdf")), gg_sup_a, width = 60 * scale_factor, height = 20 * scale_factor, units = "mm", bg = "white",  dpi = 600)
}

hallmark_umap_plot_list
```

## --- Plot B: Plot mean signature in WAds per subject
```{r, fig.width=7.09, fig.height=1.77}
# Filter for White Adipocytes only
mean_sig_data <- graph.data.filter %>%
  filter(
    (
      cell_type == "White Adipocytes" & neck_region != "Intermediate" | cell_type == "Brown Adipocytes" & neck_region == "Deep"
    )
  ) %>%
  group_by(sample, cell_type, Hallmark, subject, neck_region) %>%
  summarise(mean_score = mean(sigScores), .groups = "drop")


# Filter out samples with less than 10 cells
cells_to_keep <- meta %>% count(sample, cell_type) %>% filter(n >=10) %>% mutate(sample_cell_type = paste(sample, cell_type, sep = "_")) %>% pull(sample_cell_type)

mean_sig_data %>% count(sample, cell_type)
mean_sig_data <- mean_sig_data %>% filter(paste(sample, cell_type, sep = "_") %in% cells_to_keep)


plot_list <- list()
for (sig in unique(mean_sig_data$Hallmark)) {
  plot_data <- mean_sig_data %>% filter(Hallmark == sig)
  plot_data <- plot_data %>% mutate(region_cell_type = case_when(
    cell_type == "White Adipocytes" ~ paste(neck_region, "WAds", sep = "\n"),
    cell_type == "Brown Adipocytes" ~ paste(neck_region, "BAds", sep = "\n")
  ))
  plot_data <- plot_data %>% mutate(Hallmark = str_wrap(as.character(Hallmark), width = 5))
  max.y.pos <- max(plot_data$mean_score)
  min.y.pos <- min(plot_data$mean_score)
  
  # ---- Calculate statistics ----
  # Define the comparisons explicitly
comparisons_df <- tibble::tribble(
  ~group1,              ~group2,
  "Deep\nWAds",         "Superficial\nWAds",
  "Deep\nWAds",         "Deep\nBAds"
)

# Compute t tests manually
stat_df <- comparisons_df %>%
  rowwise() %>%
  mutate(
    p.value = t.test(
      plot_data$mean_score[plot_data$region_cell_type == group1],
      plot_data$mean_score[plot_data$region_cell_type == group2]
    )$p.value
  ) %>%
  ungroup() %>%
  mutate(
    label = case_when(
      p.value < 0.01 ~ "<0.01",
      TRUE ~ formatC(p.value, digits = 2, format = "f")
    ),
    y.position = max.y.pos * 1.10,
    colors = ifelse(p.value < 0.05, "red", "black"),
    fontface = ifelse(p.value < 0.05, "bold", "plain")
  )

  p <- ggplot(plot_data, aes(x = region_cell_type, y = mean_score)) +
    geom_line(aes(group = subject), alpha = 0.5) +
    geom_point(size = 1, shape = 21, aes(fill = cell_type), color = "black") +
    facet_wrap(vars(Hallmark), scales = "free") +
    scale_y_continuous(limits = c(min.y.pos * 0.98, max.y.pos * 1.15)) +
    scale_x_discrete(limits = c("Superficial\nWAds", "Deep\nWAds", "Deep\nBAds"),
                     labels = c("SC\nWAds", "DN\nWAds", "DN\nBAds")) +
    scale_fill_manual(values = palette.use) +
    theme_nature_metabolism(base_size = 6) +
    theme(
      strip.background = element_blank(), # <- removes grey background
      strip.text = element_text(size = 5),
      legend.position = "none"
      # axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
    ) +
    labs(
      y = "Mean Signature Score\n(VISION)",
      x = NULL
    )
  plot_list[[sig]] <- p +
    stat_pvalue_manual(
      stat_df,
      label = "label",
      fontface = "fontface",
      y.position = "y.position",
      bracket.size = 0.2,
      size = 5 / ggplot2::.pt
    ) 
}
```

# --- Adjust and save ---
```{r, warning=F, message=FALSE, fig.width=3.937008, fig.height=1.181102}
mm_to_in(c(100, 30))
plotting_order <- signatures_of_interest

plot_list

# Combine with patchwork (e.g., 2 columns)
gg_b <- wrap_plots(plot_list[plotting_order], ncol = length(plotting_order))
gg_b <- gg_b + plot_annotation(title = "Thermogenesis-Associated Signature Score", 
                  theme = theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, margin = margin(b =0)))
                  )  & 
  scale_y_continuous(expand = expansion(c(0.05,0.15))) & 
  theme(axis.text.x = element_text(size = 5), strip.text = element_text(size = 5)) + theme(plot.tag = element_text(size = 6, face = "bold"))
gg_b

scale <- length(plotting_order)
ggsave(file.path(OUTPUT_DIR,"4B_adipo_vision_lineplot.pdf"), gg_b, width = 100, height = 30, units = "mm", bg = "white",  dpi = 600)
```

```{r}
# Retained sample N 
vision_vln_n <- plot_list[[1]]$data |> count(region_cell_type)
vision_vln_n

kept_samples <- plot_list[[1]]$data |> distinct(sample) |> pull(sample)

# Original nuclei N
vision_vln_nuclei_n <- graph.data.filter %>%
  distinct(barcode, .keep_all = TRUE) |> 
  filter(
    (
      cell_type == "White Adipocytes" & neck_region != "Intermediate" | cell_type == "Brown Adipocytes" & neck_region == "Deep"
    )
  ) %>% count(cell_type_short, sample, neck_region) |> 
  filter(n > 10) |>
  group_by(cell_type_short, neck_region) |> 
  summarise(sum(n))
vision_vln_nuclei_n
```

# --- HOX genes in other datasets ---
## A38 cells
``` {r fig.height = 2, fig.width = 3.5}
library(openxlsx)
a38.cells <- read.xlsx(A38_CTS_CSV, rowNames = F) %>%
  dplyr::select(-1, -transcript_id.s.)
names(a38.cells) <- gsub("\\.", " ", names(a38.cells))

a38.de.data <- a38.cells[!duplicated(a38.cells$Symbol) & !duplicated(a38.cells$Symbol, fromLast = TRUE), ] %>%
  select(-1) %>%
  remove_rownames() %>%
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

a38.metadata <- data.frame(
  samples = colnames(a38.de.data),
  separate = colnames(a38.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-") %>%
  mutate(Type = ifelse(Type == "A38 W", "A38 WAT", Type))

heatmap <- as.matrix(a38.cells %>%
  dplyr::filter(Symbol %in% c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", "HOXA3", "HOXA4", "HOXA5", "HOXA6")) %>%
  select(
    Symbol, "A38 WAT-D0-1", "A38 WAT-D0-2", "A38 WAT-D0-3",
    "A38 WAT-D18-1", "A38 WAT-D18-2", "A38 WAT-D18-3",
    "A38 BAT-D0-1", "A38 BAT-D0-2", "A38 BAT-D0-3",
    "A38 BAT-D18-1", "A38 BAT-D18-2", "A38 BAT-D18-3"
  ) %>%
  column_to_rownames(var = "Symbol"))

# Filter metadata to keep only selected samples
a38.metadata <- a38.metadata[a38.metadata$samples %in% colnames(heatmap),]

custom.order <- c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

pdf(file.path(OUTPUT_DIR,"S4Z_A38_hox_heatmap.pdf"), width = mm_to_in(180), height = mm_to_in(60))

pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 7,
  gaps_row = 5,
  gaps_col = c(6),
  legend = FALSE,
  annotation_col = a38.metadata %>%
    remove_rownames() %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "A38 WAT", "A38 hWAd", "A38 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D6", "D9", "D12", "D15", "D18"))),
  annotation_colors = list("Diff_Day" = c("D0" = "grey", "D18" = "black"),
                           "Type" = c("A38 hBAd" = "#b3575e", "A38 hWAd" = "#313594")),
  scale = "row"
)

dev.off()
```

## --- Plot Supp D: A41 cells heatmap ---
``` {r}
a41.cells <- read_csv(file = A41_CTS_CSV) %>%
  dplyr::select(-gene_id, -transcript_ids, -Species, -Name, -Description)

a41.de.data <- a41.cells[!duplicated(a41.cells$Symbol) & !duplicated(a41.cells$Symbol, fromLast = TRUE), ] %>%
  column_to_rownames(var = "Symbol") %>%
  select(
    -"EV-D0-4_FPKM", -"EV-D0-5_FPKM", -"EV-D0-6_FPKM",
    -"BAT-D0-4_FPKM", -"BAT-D0-5_FPKM", -"BAT-D0-6_FPKM",
    -"OE-D0-4_FPKM", -"OE-D0-5_FPKM", -"OE-D0-6_FPKM",
    -"OE-D0-4_FPKM", -"OE-D0-5_FPKM", -"OE-D0-6_FPKM",
    -"OE-D9-1_FPKM", -"OE-D9-2_FPKM", -"OE-D9-3_FPKM",
    -"OE-D12-1_FPKM", -"OE-D12-2_FPKM", -"OE-D12-3_FPKM",
    -"OE-D18-1_FPKM", -"OE-D18-2_FPKM", -"OE-D18-3_FPKM"
  ) %>%
  as.matrix()

a41.metadata <- data.frame(
  samples = colnames(a41.de.data),
  separate = colnames(a41.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-")

heatmap <- a41.cells %>%
  dplyr::filter(Symbol %in% c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", "HOXA3", "HOXA4", "HOXA5", "HOXA6")) %>%
  select(Symbol, 
          "EV-D9-1_FPKM", "EV-D9-4_FPKM", "EV-D9-3_FPKM", 
          "EV-D18-1_FPKM", "EV-D18-2_FPKM", "EV-D18-3_FPKM", 
          "BAT-D9-1_FPKM", "BAT-D9-2_FPKM", "BAT-D9-3_FPKM", 
          "BAT-D18-1_FPKM", "BAT-D18-2_FPKM", "BAT-D18-3_FPKM") %>% 
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

custom.order <- c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", "HOXA3", "HOXA4", "HOXA5", "HOXA6")

pdf(file.path(OUTPUT_DIR,"S4Z_A41_hox_heatmap.pdf"), width = mm_to_in(180), height = mm_to_in(60))
pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 6,
  gaps_row = 5,
  gaps_col = c(6),
  legend = FALSE,
  annotation_col = a41.metadata %>%
    filter(samples %in% colnames(heatmap)) %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "EV", "A41 hWAd", "A41 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D9", "D12", "D18"))),
    annotation_colors = list("Diff_Day" = c("D9" = "grey", "D18" = "black"),
                           "Type" = c("A41 hBAd" = "#b3575e", "A41 hWAd" = "#313594")),

  scale = "row"
)
dev.off()
```


## Public BulkRNAseq datasets
```{r}
## Create data frame to map names
dataset_keys <- data.frame(
  dataset_name = c("Castella", "Din", "Gavrila", "Giroud", "Ohja"),
  full_name = c("Castella et al. 2023", "Din et al. 2018", "Salej et al. 2025", "Giroud et al. 2023", "Ojha et al. 2016")
) |> column_to_rownames("dataset_name")
```


```{r}
#--- Determine files to loop over
# Expression files
exp_files <- list.files(BULK_DIR, full.names = TRUE)
cleaned_files <- grep("_Cleaned", exp_files, value = TRUE)

#Metadata files
meta_files <- list.files(BULK_META_DIR, full.names = TRUE)

#Define dataset names
dataset_names <-  str_remove(basename(cleaned_files), "_Cleaned.csv")


for(dataset in dataset_names){
 
  # --- Read in the data ---
  object <- read.csv(str_subset(cleaned_files, dataset), header = TRUE, row.names = 1)
  
  meta <- read.csv(str_subset(meta_files, dataset))
  
  meta$group <- factor(meta$group, levels = unique(meta$group[order(meta$order)]))
  rownames(meta) <- meta$sample
  
  if(dataset == "Ohja"){
    idx <- match(names(object), meta$sample)
    names(object) <- meta$group[idx]
    names(object) <- make.unique(names(object))    
    rownames(meta)[idx] <-  names(object)

  }
  
  meta <- meta %>% select(group) |> mutate(group = case_when(
    grepl("WAT|SC", rownames(meta)) ~ "Superficial",
    grepl("BAT|Deep", rownames(meta)) ~ "Deep",
    TRUE ~ group
  ))

  
  # Reorder
  col_order <- c(
    grep("WAT|SC|C", names(object), value = TRUE),
    grep("BAT|Deep|F", names(object), value = TRUE),
    grep("Neonate", names(object), value = TRUE),
    grep("Infant", names(object), value = TRUE),
    grep("Child", names(object), value = TRUE)
  )
  
  heatmap <- object[,col_order] %>%
    as.matrix()
  
  if(dataset == "Din"){
  first_break <- 15
  } else if(dataset == "Gavrila"){
    first_break <- 14
  } else {
    first_break <- 0
  }
  
  #Colors
  dev_age_color <- c("Child" = "orchid4", "Infant" = "orchid3", "Neonate" = "orchid1")
  pheo_color <- c("control" = "grey25", "pheochromocytoma" = "darkgreen", "Control" = "grey25", "Pheo" = "darkgreen")
  annotation_colors <- list(group = c(neck.color, dev_age_color, pheo_color))
  
  annotation_colors$group <- annotation_colors$group[names(annotation_colors$group) %in% meta$group]
  
  custom.order <- c("PAX3", "HOXD3", "HOXD4", "IRX2", "IRX5", "HOXA3", "HOXA4", "HOXA5", "HOXA6")
  pdf(file.path(OUTPUT_DIR, paste("S4F-G", dataset, "hox_heatmap.pdf", sep= "_")), width = mm_to_in(180), height = mm_to_in(60))
  # # --- Plot heatmap ---
  mat <- heatmap[custom.order, , drop = FALSE]

  print(pheatmap(mat,
           cluster_cols = FALSE,
           cluster_rows = FALSE, 
           show_colnames = FALSE,
           cellwidth = 5,
           cellheight = 5.5,
           fontsize = 6, 
           gaps_row = 5,
           gaps_col = first_break,
           legend = TRUE,
           annotation_col = meta,
           annotation_colors = annotation_colors,
           scale = "row",
          main = dataset_keys[dataset,]
))
  dev.off()


}

```

# --- Multiome analysis ---
```{r}
library(Signac)
MULTIOME_SEURAT_RDS <- here("data_onedrive/9.multiome_data/multiome_seurat.rds")
query_obj <- readRDS(MULTIOME_SEURAT_RDS)
```

```{r}
# ---- Plot coverage of HOX genes in White Adipocytes from deep vs superficial ----
stopifnot("peaks" %in% Assays(query_obj))
DefaultAssay(query_obj) <- "peaks"
ann <- Annotation(query_obj)
seqlevels(ann) <- paste0("chr", gsub("^(chr)+", "", seqlevels(ann)))
Annotation(query_obj[["peaks"]]) <- ann


# save current identities
old_idents <- Idents(query_obj)

# Create a new identity. This is a workaround on defective subsetting of Seurat v5 for multiome
query_obj$cell_type_neck_region <- paste(query_obj$cell_type_short, query_obj$neck_region, sep = "_")

# Assing and select cells to plot
Idents(query_obj) <- "cell_type_neck_region"

idents_query <- paste0(c("WAds"),
                       collapse = "|")

idents_to_plot <- grep(idents_query, unique(query_obj$cell_type_neck_region), value = T)
idents_to_plot <- grep(pattern = "Intermediate", idents_to_plot, value = T, invert = T)

# Select genes
features <- c("HOXA3", "HOXA4", "HOXA5", "HOXA6",
              "PAX3", "IRX2", "IRX5", "HOXD3", "HOXD4")

# Set palette
id_levels <- levels(Idents(query_obj))
neck_region_from_ident <- sub("^.*_", "", id_levels)
ident_colors <- setNames(neck.color[neck_region_from_ident], id_levels)
ident_colors

for (goi in features){

  scale_factor <- 2

# plot only the WAds-derived identity groups

  p <- CoveragePlot(
    query_obj,
    region = goi,
    features = goi,
    assay = "peaks",
    expression.assay = "SCT",
    peaks = TRUE,
    idents = idents_to_plot
  ) &
    scale_fill_manual(values = ident_colors) &
    theme(
      text = element_text(size = 5 * scale_factor),
      axis.title = element_text(size = 5 * scale_factor),
      axis.text = element_text(size = 5 * scale_factor),
      strip.text = element_text(size = 5 * scale_factor),
      legend.title = element_text(size = 5 * scale_factor),
      legend.text = element_text(size = 5 * scale_factor),
      axis.line = element_line(linewidth = 0.3 * scale_factor),
      axis.ticks = element_line(linewidth = 0.3 * scale_factor),
      # panel.border = element_rect(fill = NA, linewidth = 0.3 * scale_factor)
    )

ggsave(
  filename = file.path(OUTPUT_DIR, paste0("4G_coverage_", goi, "_", idents_query,"_by_neck.region.pdf")),
  plot = p,
  width = 65 * scale_factor,
  height = 50 * scale_factor,
  units = "mm",
  bg = "white"
)

}
# restore original identities afterward
Idents(query_obj) <- old_idents
```




# --- HEAT genes in other datasets ---
## Load heat signature
```{r}
load(here("output/3.heat_signature/data/heat_signature.Rdata")) #Creates HEAT_signature
```


## A38 cells
``` {r fig.height = 2, fig.width = 3.5}
library(openxlsx)
a38.cells <- read.xlsx(A38_CTS_CSV, rowNames = F) %>%
  dplyr::select(-1, -transcript_id.s.)
names(a38.cells) <- gsub("\\.", " ", names(a38.cells))

a38.de.data <- a38.cells[!duplicated(a38.cells$Symbol) & !duplicated(a38.cells$Symbol, fromLast = TRUE), ] %>%
  select(-1) %>%
  remove_rownames() %>%
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

a38.metadata <- data.frame(
  samples = colnames(a38.de.data),
  separate = colnames(a38.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-") %>%
  mutate(Type = ifelse(Type == "A38 W", "A38 WAT", Type))

heatmap <- as.matrix(a38.cells %>%
  dplyr::filter(Symbol %in% c("UCP1", HEAT_signature)) %>%
  select(
    Symbol, 
    # "A38 WAT-D0-1", "A38 WAT-D0-2", "A38 WAT-D0-3",
    "A38 WAT-D18-1", "A38 WAT-D18-2", "A38 WAT-D18-3",
    # "A38 BAT-D0-1", "A38 BAT-D0-2", "A38 BAT-D0-3",
    "A38 BAT-D18-1", "A38 BAT-D18-2", "A38 BAT-D18-3"
  ) %>%
  column_to_rownames(var = "Symbol"))

# Filter metadata to keep only selected samples
a38.metadata <- a38.metadata[a38.metadata$samples %in% colnames(heatmap),]

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S4Z_A38_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))

pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 7,
  gaps_row = 1,
  gaps_col = c(3),
  legend = TRUE,
  annotation_col = a38.metadata %>%
    remove_rownames() %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "A38 WAT", "A38 hWAd", "A38 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D6", "D9", "D12", "D15", "D18"))),
  annotation_colors = list("Diff_Day" = c("D0" = "grey", "D18" = "black"),
                           "Type" = c("A38 hBAd" = "#b3575e", "A38 hWAd" = "#313594")),
  scale = "row"
)

dev.off()

grep("COB", a38.cells$Symbol, value = T)
a38.cells |>  filter(Symbol %in% custom.order) |> select("Symbol", contains("D18"))

a38.cells |> select(contains("A38")) |> colSums(na.rm = TRUE)
```

## --- Plot Supp D: A41 cells heatmap ---
``` {r}
a41.cells <- read_csv(file = A41_CTS_CSV) %>%
  dplyr::select(-gene_id, -transcript_ids, -Species, -Name, -Description)

a41.de.data <- a41.cells[!duplicated(a41.cells$Symbol) & !duplicated(a41.cells$Symbol, fromLast = TRUE), ] %>%
  column_to_rownames(var = "Symbol") %>%
  select(
    -"EV-D0-4_FPKM", -"EV-D0-5_FPKM", -"EV-D0-6_FPKM",
    -"BAT-D0-4_FPKM", -"BAT-D0-5_FPKM", -"BAT-D0-6_FPKM",
    -"OE-D0-4_FPKM", -"OE-D0-5_FPKM", -"OE-D0-6_FPKM",
    -"OE-D0-4_FPKM", -"OE-D0-5_FPKM", -"OE-D0-6_FPKM",
    -"OE-D9-1_FPKM", -"OE-D9-2_FPKM", -"OE-D9-3_FPKM",
    -"OE-D12-1_FPKM", -"OE-D12-2_FPKM", -"OE-D12-3_FPKM",
    -"OE-D18-1_FPKM", -"OE-D18-2_FPKM", -"OE-D18-3_FPKM"
  ) %>%
  as.matrix()

a41.metadata <- data.frame(
  samples = colnames(a41.de.data),
  separate = colnames(a41.de.data)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Sample"), sep = "-")

heatmap <- a41.cells %>%
  dplyr::filter(Symbol %in% c("UCP1", HEAT_signature)) %>%
  select(Symbol, 
          # "EV-D9-1_FPKM", "EV-D9-4_FPKM", "EV-D9-3_FPKM", 
          "EV-D18-1_FPKM", "EV-D18-2_FPKM", "EV-D18-3_FPKM", 
          # "BAT-D9-1_FPKM", "BAT-D9-2_FPKM", "BAT-D9-3_FPKM", 
          "BAT-D18-1_FPKM", "BAT-D18-2_FPKM", "BAT-D18-3_FPKM") %>% 
  column_to_rownames(var = "Symbol") %>%
  as.matrix()

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S4Z_A41_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))
pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 6,
  gaps_row = 1,
  gaps_col = c(3),
  legend = TRUE,
  annotation_col = a41.metadata %>%
    filter(samples %in% colnames(heatmap)) %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Type = ifelse(Type == "EV", "A41 hWAd", "A41 hBAd")) %>%
    mutate(Diff_Day = factor(Diff_Day, levels = c("D0", "D9", "D12", "D18"))),
    annotation_colors = list("Diff_Day" = c("D9" = "grey", "D18" = "black"),
                           "Type" = c("A41 hBAd" = "#b3575e", "A41 hWAd" = "#313594")),

  scale = "row"
)
dev.off()
a41.cells |>  filter(Symbol %in% custom.order) |> select("Symbol", contains("D18")) |> select(!contains("OE"))

```

## --- Cero et al. 2023 ---
```{r}
cero <- read.xlsx(CERO_CTS_XLSX)

cero_mtx <- cero %>%
  column_to_rownames(var = "Name") %>%
  select(contains("d30.CPM")) %>%
  as.matrix()

cero_metadata <- data.frame(
  samples = colnames(cero_mtx),
  separate = colnames(cero_mtx)
) %>%
  separate(col = separate, into = c("Type", "Diff_Day", "Count_Type", "Replicate"), sep = "\\.")

heatmap <- cero_mtx 

custom.order <- c("UCP1", sort(HEAT_signature))

pdf(file.path(OUTPUT_DIR,"S4Z_cero_heat_heatmap.pdf"), width = mm_to_in(25), height = mm_to_in(90))
pheatmap(heatmap[custom.order, , drop = FALSE],
  cluster_cols = FALSE,
  cluster_rows = FALSE,
  cellwidth = 12,
  cellheight = 7.2,
  fontsize = 6,
  gaps_row = 1,
  gaps_col = c(4),
  legend = TRUE,
  annotation_col = cero_metadata %>%
    filter(samples %in% colnames(heatmap)) %>%
    column_to_rownames(var = "samples") %>%
    select(1:2) %>%
    mutate(Diff_Day = factor(toupper(Diff_Day), levels = c("D0", "D30"))),
    annotation_colors = list("Diff_Day" = c("D0" = "grey", "D30" = "black"),
                           "Type" = c("hBA" = "#b3575e", "hWA" = "#313594")),

  scale = "row"
)
dev.off()
```

## Public BulkRNAseq datasets
```{r}
## Create data frame to map names
dataset_keys <- data.frame(
  dataset_name = c("Castella", "Din", "Gavrila", "Giroud", "Ohja"),
  full_name = c("Castella et al. 2023", "Din et al. 2018", "Salej et al. 2025", "Giroud et al. 2023", "Ojha et al. 2016")
) |> column_to_rownames("dataset_name")
```


```{r}
#--- Determine files to loop over
# Expression files
exp_files <- list.files(BULK_DIR, full.names = TRUE)
cleaned_files <- grep("_Cleaned", exp_files, value = TRUE)

#Metadata files
meta_files <- list.files(BULK_META_DIR, full.names = TRUE)

#Define dataset names
dataset_names <-  str_remove(basename(cleaned_files), "_Cleaned.csv")


for(dataset in dataset_names){
 
  # --- Read in the data ---
  object <- read.csv(str_subset(cleaned_files, dataset), header = TRUE, row.names = 1)
  
  meta <- read.csv(str_subset(meta_files, dataset))
  
  meta$group <- factor(meta$group, levels = unique(meta$group[order(meta$order)]))
  rownames(meta) <- meta$sample
  
  if(dataset == "Ohja"){
    idx <- match(names(object), meta$sample)
    names(object) <- meta$group[idx]
    names(object) <- make.unique(names(object))    
    rownames(meta)[idx] <-  names(object)

  }
  
  meta <- meta %>% select(group) |> mutate(group = case_when(
    grepl("WAT|SC", rownames(meta)) ~ "Superficial",
    grepl("BAT|Deep", rownames(meta)) ~ "Deep",
    TRUE ~ group
  ))

  
  # Reorder
  UCP1_exp <- object["UCP1", ] |> 
    t() |> 
    as.data.frame( ) |> 
    rownames_to_column("sample") |> 
    left_join(meta |> rownames_to_column("sample"), by = "sample")
  if(dataset %in% c("Din", "Gavrila", "Giroud")){
  col_order <- UCP1_exp |> 
    arrange(rev(group), UCP1) |> 
    pull(sample)
  } else {
      col_order <- UCP1_exp |> 
    arrange(group, UCP1) |> 
    pull(sample)
  }
  
  # col_order <- c(
  #   grep("WAT|SC|C", names(object), value = TRUE),
  #   grep("BAT|Deep|F", names(object), value = TRUE),
  #   grep("Neonate", names(object), value = TRUE),
  #   grep("Infant", names(object), value = TRUE),
  #   grep("Child", names(object), value = TRUE)
  # )
  
  heatmap <- object[,col_order] %>%
    as.matrix()
  
  if(dataset == "Din"){
  first_break <- 15
  } else if(dataset == "Gavrila"){
    first_break <- 14
  } else if(dataset == "Castella"){
    first_break <- 3
  } else if(dataset == "Giroud"){
    first_break <- 5
  } else if(dataset == "Ohja"){
    first_break <- c(7,35)
  } else {
    first_break <- 0
  } 
  
  print(col_order)
  #Colors
  dev_age_color <- c("Child" = "orchid4", "Infant" = "orchid3", "Neonate" = "orchid1")
  pheo_color <- c("control" = "grey25", "pheochromocytoma" = "darkgreen", "Control" = "grey25", "Pheo" = "darkgreen")
  annotation_colors <- list(group = c(neck.color, dev_age_color, pheo_color))
  
  annotation_colors$group <- annotation_colors$group[names(annotation_colors$group) %in% meta$group]
  
  custom.order <- c("UCP1", sort(HEAT_signature))
  custom.order <- custom.order[custom.order %in% rownames(heatmap)]
  
  pdf(file.path(OUTPUT_DIR, paste("S4Z", dataset, "HEAT_heatmap.pdf", sep= "_")), width = mm_to_in(200), height = mm_to_in(90))
  # # --- Plot heatmap ---
  mat <- heatmap[custom.order, , drop = FALSE]

  print(pheatmap(mat,
           cluster_cols = FALSE,
           cluster_rows = FALSE,
           cellwidth = 6,
           cellheight = 8,
           fontsize = 6,
           gaps_row = 1,
           gaps_col = first_break,show_colnames = FALSE,
           legend = FALSE,  
           annotation_legend = FALSE,
           annotation_col = meta,
           annotation_colors = annotation_colors,
           annotation_names_col = FALSE,
           scale = "row",
          main = dataset_keys[dataset,]
))
  dev.off()


}

```
# --- Supp Plot G: HEAT signature paraganglioma ---
```{r}
PARAGANG_BULK_MTX_RDS <- here("output/3.heat_signature/3.public_bulkrnaseq/data/GSE49795_Sondergaard_expression_matrix.rds")

expression_matrix <- read_rds(PARAGANG_BULK_MTX_RDS)

pdf(file.path(OUTPUT_DIR, "S4Z_paragang_heatmap.pdf"), height = mm_to_in(50), width = mm_to_in(50))
# Create heatmap
pheatmap::pheatmap(expression_matrix,
  scale = "row", # Normalize rows (genes)
  clustering_distance_rows = "euclidean", # Clustering for rows
  cellwidth = 5,
  cellheight = 5,
  fontsize = 6,
  clustering_distance_cols = "euclidean", # Clustering for columns
  clustering_method = "complete", # Hierarchical clustering method
  treeheight_row = 0, treeheight_col = 0
)
dev.off()

pheatmap::pheatmap(expression_matrix,
  scale = "row", # Normalize rows (genes)
  # clustering_distance_rows = "euclidean", # Clustering for rows
  cellwidth = 5,
  cellheight = 5,
  fontsize = 6,
  # clustering_distance_cols = "euclidean", # Clustering for columns
  # clustering_method = "complete", # Hierarchical clustering method
  treeheight_row = 0, treeheight_col = 0
)
```
