---
title: "Plots for Supplementary Figure 1"
output: html_document
date: "2025-01-14"
---

#  --- Setup ---
```{r setup, include=FALSE}
library(Seurat)
library(ggplot2)
library(ggpubr)

library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(forcats)
library(tibble)

library(openxlsx)
library(speckle)
library(ggnewscale)
library(patchwork)
library(here)
library(grid)



theme_no_axis <- function() {
  theme(
    axis.title   = element_blank(),
    axis.text    = element_blank(),
    axis.ticks   = element_blank(),
    axis.line    = element_blank()
  )
}

# OBS:  theme_nature_metabolism() and color palettes set on "scripts/0.environment_setup/snRNAseq_graphics_setup.R"
source(here("0.environment_setup/snRNAseq_graphics_setup.R"))


# --- Setup paths ---
## Inputs
SEURAT_INPUT <- here(
  "..",
  "data/cervical_at_gex_seurat.rds")

GAVRILA_TPM <- here(
  "..",
  "data/Figure_S1/TPM_values.csv")
DIN_CTS <- here(
  "..",
  "data/Figure_S1/Din2018_countmatrix.txt")
XUE_CTS <- here(
  "..",
  "data/Figure_S1/Supplementary_Table_cervical_at_microarray.xlsx")
REF_MAP_APSC <- here(
  "..",
  "data/Figure_S1/aspc_combined.rds") 
REF_MAP_ADIPO <- here(
  "..",
  "data/Figure_S1/adipo_combined.rds") 

PROPELLER_PROPORTIONS <- here(
  "..",
  "data/Figure_S1/propeller_cellular_proportions_table.csv" )
PROPELLER_STATS <- here(
  "..",
  "data/Figure_S1/proportions_stat_table.csv")


# Outputs
OUTPUT_DIR <- here("..", "results", "Figure_Supp_1")

dir.create(OUTPUT_DIR, showWarnings = F, recursive = T)
```

# --- Loading Seurat ---

This first part is loading the neck dataset, adds the short cell labels, and adds color palettes used in the below figures.

``` {r}
s.object <- readRDS(SEURAT_INPUT)
```

# --- Panel S1A. DotPlot Markers ---
This section reorganizes adipocytes by depot and visualizes classical and
thermogenic marker programs across white and brown adipocyte groups.

```{r, fig.width = 13.779528, fig.height = 3.937008}
mm_to_in(c(100, 100))

# -------------------------------------------------------------------------
# Regroup adipocytes by depot
# -------------------------------------------------------------------------
meta <- s.object@meta.data

meta <- meta %>%
  mutate(
    adipo_region = case_when(
      str_detect(cell_type, " Adipocyte") ~ paste0(cell_type, " (", neck_region, ")"),
      TRUE ~ "Other"
    )
  ) %>%
  mutate(
    adipo_region = case_when(
      adipo_region == "Brown Adipocytes (Intermediate)" ~ "Other",
      adipo_region == "Brown Adipocytes (Superficial)" ~ "Other",
      TRUE ~ adipo_region
    )
  ) %>%
  mutate(
    adipo_region = factor(
      adipo_region,
      levels = c(
        sort(unique(adipo_region[str_detect(adipo_region, "^Brown Adipocytes")])),
        sort(unique(adipo_region[str_detect(adipo_region, "^White Adipocytes")])),
        "Other"
      )
    )
  )

s.object@meta.data <- meta

# -------------------------------------------------------------------------
# Define marker gene groups
# -------------------------------------------------------------------------
adipo_markers <- c("ADIPOQ", "PPARG")
white_markers <- c("LEP", "SLC7A10", "HOCX8", "HOXC9")
thermogenic_markers <- c("UCP1", "PPARGC1A", "PRDM16", "EBF2", "DIO2")
brown_markers <- c("EBF3", "FBXO31", "ZIC1", "MPZL2", "LHX8")
beige_markers <- c("TBX1", "TMEM26", "TNFRSF9", "SHOX2")
fcc_markers <- c("GAMT", "GATM", "SLC6A8", "ALPL", "CKB", "CKMT1A", "CKMT1B", "CKMT2")
calcium_markers <- c("ATP2A2", "RYR2", "ARLN", "ALN", "ITPR1", "ITPR2")
lipid_cyc_markers <- c("PNPLA2", "DGAT1")

feature_groups <- list(
  "General Adipocyte" = adipo_markers,
  "White Adipocyte" = white_markers,
  "Brown Adipocyte" = brown_markers,
  "Beige Adipocyte" = beige_markers,
  "Thermogenic Genes" = thermogenic_markers,
  "FCC Genes" = fcc_markers,
  "Calcium Cycling Genes" = calcium_markers,
  "Lipid Cycling Genes" = lipid_cyc_markers
)

feature_groups <- lapply(feature_groups, function(gene_list) {
  found_genes <- gene_list %in% rownames(s.object)
  gene_list[found_genes]
})

group_labels <- c(
  "General Adipocyte" = "Adipocyte",
  "White Adipocyte" = "White",
  "Brown Adipocyte" = "Brown",
  "Beige Adipocyte" = "Beige",
  "Thermogenic Genes" = "Classical Thermogenic",
  "FCC Genes" = "Futile Creatine Cycling",
  "Calcium Cycling Genes" = "Calcium Cycling",
  "Lipid Cycling Genes" = "Lipid Cycling"
)

all_features <- unlist(feature_groups) %>% toupper()
feature_order <- all_features

# -------------------------------------------------------------------------
# Create accessory plotting data for category brackets
# -------------------------------------------------------------------------
group_lengths <- lengths(feature_groups)
group_ends <- cumsum(group_lengths)
group_starts <- c(1, head(group_ends, -1) + 1)
group_mids <- (group_starts + group_ends) / 2

bracket_df <- data.frame(
  group = names(feature_groups),
  start = group_starts,
  end = group_ends,
  mid = group_mids,
  label = str_wrap(unname(group_labels[names(feature_groups)]), width = 80)
)

# -------------------------------------------------------------------------
# Alternating background panels by feature group
# -------------------------------------------------------------------------
panel_df <- bracket_df %>%
  mutate(
    xmin = start - 0.5,
    xmax = end + 0.5,
    fill = rep(c("white", "grey95"), length.out = n())
  )

# -------------------------------------------------------------------------
# Build the dot plot
# -------------------------------------------------------------------------
dp <- DotPlot(
  s.object,
  features = all_features,
  group.by = "adipo_region",
  cols = c("lightgrey", "blue")
)

dp$data$features.plot <- factor(dp$data$features.plot, levels = feature_order)


main_plot <- ggplot(dp$data, aes(x = features.plot, y = id)) +
  geom_rect(
    data = panel_df,
    aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
    inherit.aes = FALSE,
    color = NA
  ) +
  geom_point(aes(size = pct.exp, color = avg.exp.scaled)) +
  scale_fill_identity() +
  scale_size_continuous(
    limits = c(0, 100),
    breaks = c(20, 40, 60, 80),
    range = c(0, 5),
    name = "% cells"
  ) +
  scale_color_gradient(
    low = "lightgrey",
    high = "blue",
    limits = c(-2, 2),
    breaks = seq(-2, 2, 1),
    name = "Avg. expression"
  ) +
  labs(x = NULL, y = NULL) +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 0, r = 10, b = 10, l = 10)
  )

bracket_plot <- ggplot(bracket_df) +
  geom_segment(
    aes(x = start - 0.2, xend = end + 0.2, y = 0.2, yend = 0.2),
    linewidth = 0.3
  ) +
  geom_text(
    aes(x = mid, y = 0.65, label = label),
    size = 5 / .pt,
    fontface = "bold"
  ) +
  scale_x_continuous(
    limits = c(0.5, length(feature_order) + 0.5),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0)
  ) +
  coord_cartesian(clip = "off") +
  theme_void() +
  theme(
    plot.margin = margin(t = 5, r = 10, b = 0, l = 10)
  )

combined_plot <- bracket_plot / main_plot +
  plot_layout(heights = c(2.2, 12))
```

```{r, fig.width = 7.086614, fig.height = 1.574803}
mm_to_in(c(180, 40))

combined_plot

ggsave(
  file.path(OUTPUT_DIR, "S1A_dotplot_adipocyte_markers.pdf"),
  plot = combined_plot,
  width = 180,
  height = 40,
  units = "mm",
  bg = "white",
  dpi = 600
)
```

This below portion gets the cell type proportions by sample.

```{r}
propeller_list <- getTransformedProps(
  clusters = s.object$cell_type,
  sample = s.object$sample
)

props_meta <- as_tibble(propeller_list$Proportions) %>%
  dplyr::left_join(
    s.object@meta.data %>%
      select(sample, neck_region) %>%
      distinct(),
    by = "sample"
  ) %>%
  mutate(
    n_pct = n * 100,
    neck_region = factor(neck_region, levels = c("Superficial", "Deep"))
  )
```

# --- Panel S1B. Brown Adipocyte Abundance ---

```{r}
propeller_proportions <- read_csv(PROPELLER_PROPORTIONS)
propeller_stats <- read_csv(PROPELLER_STATS)
```
## --- Calculate cellular proportions ---
```{r}
propeller_list <- getTransformedProps(
  clusters = s.object$cell_type_short,
  sample = s.object$sample
)

propeller_proportions <- as_tibble(propeller_list$Proportions) %>%
  dplyr::left_join(
    s.object@meta.data %>%
      select(sample, neck_region) %>%
      distinct(),
    by = "sample"
  ) %>%
  mutate(
    n_pct = n * 100,
    neck_region = factor(neck_region, levels = c("Superficial", "Intermediate", "Deep"))
  ) |> 
  left_join(s.object@meta.data |> select(cell_type, cell_type_short) |> distinct(cell_type_short,.keep_all = TRUE), by = c("clusters"= "cell_type_short"))
```


```{r fig.height = 4, fig.width = 10}
bad_plot_data <- propeller_proportions %>%
    dplyr::filter(clusters == "BAds")
bad_plot_stats <- propeller_stats |> filter(cell_type_short == "BAds") |> mutate(xmin = 1, xmax = 3, y.position = 18)


p_value_y_position <- max(bad_plot_stats$y.position)


gg_propeller<- ggplot(
  bad_plot_data,
  aes(
    x = neck_region,
    y = n_pct,
    fill = neck_region
  )
) +

  # Add the boxplot layer
  geom_boxplot(
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    alpha = 0.75
  ) +

  # Add the jitter layer with dodging
  geom_jitter(
    color = "black",
    shape = 21,
    alpha = 0.9,
    size = 1,
    position = position_jitterdodge(
      jitter.width = 0.1,
      dodge.width = 0.75
    )
  ) +
  scale_fill_manual(values = neck.color, name = "Neck Region") +
  scale_color_manual(values = neck.color, name = "Neck Region") +
  new_scale_color() +
  stat_pvalue_manual(
    data = bad_plot_stats |> filter(cell_type_short == "BAds"),
    color = "colors",
    xmin = "xmin",
    xmax = "xmax",
    # x = "cell_type_short",
    label = "p_display",
    y.position = "y.position",
    tip.length = 0.006,
    inherit.aes = FALSE,
    size = 6 / .pt,
  ) +
  scale_color_identity(guide = "none") +
  coord_cartesian(ylim = c(NA, p_value_y_position + 2)) + #
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      vjust = 1,
      hjust = 1
    ),
    axis.title.x = element_blank(),
    legend.position = c(0.95, 0.60),
    legend.justification = c("right", "top"),
  ) +
  ylab("Proportion of Cells in Sample")

gg_propeller
```
## --- Plotting ---
```{r, fig.width = 0.7874016, fig.height = 1.3779528}
mm_to_in(c(20,35))
gg_propeller<- gg_propeller+ 
  theme(plot.tag = element_text(size = 6, face = "plain")) + labs(title = str_wrap("BAds Abundance", width = 5)) + 
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5),
        legend.position = "none") +
  guides(color = guide_legend(ncol = 1),
         fill = guide_legend(ncol = 1)) +
  labs(y = str_wrap("Proportion in Sample", width = 15)) +
  scale_y_continuous(expand = expansion(c(0.05,0.15)))
gg_propeller

# Save
ggsave(file.path(OUTPUT_DIR, "S1B_brown_adipocyte_abundance_zoom.pdf"), plot = gg_propeller, width = 20, height = 35, units = "mm", bg = "white", dpi = 600)
```


# --- Panel S1C. Thermogenic Singal Normalized Score (NS) ---
This section is intended to compare the number of samples that have a high number of brown adipocytes in this dataset compared to the number of samples that have high UCP1 expression in bulk transcriptomic datasets.

## --- Pulling data from snRNAseq brown adipocytes  ---

This first part pulls and graphs just the percentage of brown adipocyte nuclei in each sample.

```{r}
ba_freq <- props_meta %>%
  dplyr::filter(clusters == "Brown Adipocytes") %>%
  dplyr::filter(neck_region == "Deep") %>%
  arrange(desc(n_pct)) %>%
  dplyr::select(sample, n_pct) %>%
  mutate(
    Normalized = n_pct / max(n_pct),
    dataset = "SingleCell"
  ) %>%
  rename(
    Value = n_pct,
    Sample = sample
  )

ba_freq$highlow <- seq(1, by = 1, to = nrow(ba_freq))

ba_freq <- ba_freq %>%
  mutate(normalized_sample = highlow / max(highlow))
```

## --- Pulling data from the Gavrila dataset ---

This second part pulls and graphs the number of high UCP1 samples in the Gavrila dataset.

```{r}
gavrila <- read.csv(file = GAVRILA_TPM)

normalized_gavrila <- gavrila %>%
  filter(Gene.name == "UCP1") %>%
  gather(2:28, key = "Sample", value = "Value") %>%
  mutate(
    TissueType = str_extract(Sample, "^(Deep|SC)"),
    Individual = as.factor(str_extract(Sample, "\\d+$")),
    Normalized = Value / max(Value),
    dataset = "Gavrila"
  ) %>%
  filter(TissueType == "Deep") %>%
  arrange(desc(Value)) %>%
  select(-ID, -Gene.name, -Individual, -TissueType)

normalized_gavrila$highlow <- seq(1, by = 1, to = nrow(normalized_gavrila))

normalized_gavrila <- normalized_gavrila %>%
  mutate(normalized_sample = highlow / max(highlow))
```

## --- Pulling data from the Din dataset ---

This third part pulls and graphs the number of high UCP1 samples in the Din dataset.

```{r}
din <- read.delim(DIN_CTS)

normalized_din <- din %>%
  filter(row.names(.) == "UCP1") %>%
  gather(key = "Sample", value = "Value") %>%
  mutate(
    TissueType = str_extract(Sample, "BAT|WAT"),
    Normalized = Value / max(Value),
    dataset = "Din"
  ) %>%
  filter(TissueType == "BAT") %>%
  arrange(desc(Value)) %>%
  select(-TissueType)

normalized_din$highlow <- seq(1, by = 1, to = nrow(normalized_din))

normalized_din <- normalized_din %>%
  mutate(normalized_sample = highlow / max(highlow))
```

## --- Pulling data from the microarray dataset ---

This fourth part pulls and graphs the number of high UCP1 samples in the Xue dataset.

``` {r}
normalized_xue <- read.xlsx(XUE_CTS, sheet = 2, check.names = F) %>%
  rename_with(~ gsub("\\.", " ", .)) %>%
  dplyr::filter(`Gene Symbol` == "UCP1") %>%
  select(2:21) %>%
  gather(1:20, key = "Sample", value = "Value") %>%
  mutate(
    TissueType = str_extract(Sample, "SQ|CS"),
    Value = (Value)^2,
    Normalized = Value / max(Value),
    dataset = "Xue"
  ) %>%
  dplyr::filter(TissueType == "CS") %>%
  arrange(desc(Value)) %>%
  select(-TissueType)

normalized_xue$highlow <- seq(1, by = 1, to = nrow(normalized_xue))

normalized_xue <- normalized_xue %>%
  mutate(normalized_sample = highlow / max(highlow))
```

## --- Plotting ---

This final part pulls all of the above datasets into a single plot for comparing between datasets the comparable number of high and low thermogenic cells.


```{r, fig.height = 4, fig.width = 5}
total <- rbind(ba_freq, normalized_gavrila, normalized_din, normalized_xue) %>%
  mutate(dataset = factor(dataset, levels = c("SingleCell", "Gavrila", "Din", "Xue"))) %>%
  mutate(dataset_renamed = fct_recode(
    dataset,
    "This study (snRNAseq)" = "SingleCell",
    "This study (microarray)" = "Xue",
    "Salej et al. 2025 (bulk RNA-seq)" = "Gavrila",
    "Din et al. 2018 (bulk RNA-seq)" = "Din"
  ))

levels(total$dataset)
levels(total$dataset_renamed)

c <- ggplot(total, aes(x = normalized_sample * 100, y = Normalized, color = dataset_renamed)) +
  geom_hline(yintercept = c(0.2, 0.8), linetype = "dashed", linewidth = 0.3) +
  geom_line() +
  geom_point() +
  labs(
    x = "Population Percentage",
    y = str_wrap("Normalized Thermogenic Signal (NS)", 25),
    color = "Dataset"
  ) +
  theme_nature_metabolism() +
  coord_cartesian(clip = "off") +
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.95, 0.98), # top-right
    legend.justification = c(0.5, 1), # anchor legend's top-right corner
    legend.background = element_rect(fill = "white", color = NA),
    legend.box.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(5.5, 35, 5.5, 5.5) # extra right margin so long text isn't cut
  ) +
  guides(
    color = guide_legend(
      override.aes = list(size = 0.5)  # legend dot size
    ))
```

```{r}
total_counts <- total |>
  mutate(
    NS_group = case_when(
      Normalized > 0.8 ~ "Strong: NS > 0.8",
      Normalized > 0.2 ~ "Intermediate: 0.2 < NS ≤ 0.8",
      TRUE ~ "Low: NS ≤ 0.2"
    )
  ) |>
  count(dataset_renamed, NS_group)
total_counts

total |>
  mutate(
    NS_group = case_when(
      Normalized > 0.8 ~ "Strong: NS > 0.8",
      Normalized > 0.2 ~ "Intermediate: 0.2 < NS ≤ 0.8",
      TRUE ~ "Low: NS ≤ 0.2"
    )
  ) |>
  count(NS_group)
```
## --- Plotting ---
``` {r fig.width = 2.952756, fig.height = 1.377953}
total |> count(dataset_renamed, Normalized > 0.8,Normalized > 0.20) |> group_by(`Normalized > 0.8`, `Normalized > 0.2`) |> summarise(sum(n))
nrow(total)

mm_to_in(c(75, 35))
c <- c + 
  theme(plot.tag = element_text(size = 6, face = "plain")) + labs(title = "Distribution of Thermogenic Samples\nin Cervical AT Datasets") + theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 3)) + 
  theme(
  legend.key.width  = unit(0, "pt"),
) 
c

ggsave(file.path(OUTPUT_DIR, "S1C_thermogenic_sample_distribution.pdf"), plot = c, width = 75, height = 35, units = "mm", bg = "white",  dpi = 600)



```


# --- Panel S1D. Proportions of BAd by donor ---

This plot is a stacked barplot with the relative contribution of each sample to the total of brown adipocytes in the dataset. This is intended to show that most of these cells are from 2 samples and to show which samples are adding the key contributions.

```{r fig.height = 4, fig.width = 2.5}
bad_counts <- s.object@meta.data %>%
  dplyr::filter(cell_type_short == "BAds") %>%
  group_by(Age, Gender) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  arrange(desc(count)) %>%
  mutate(
    relative = round(100 * count / sum(count), 0),
    subject_char = paste(Age, ", ", Gender, sep = ""),
    subject_char = factor(subject_char, levels = subject_char)
  )


d <- ggplot(bad_counts, aes(x = "BAds", y = relative, fill = subject_char)) +
  geom_col() +
  theme_pubr() +
  scale_fill_brewer(palette = "Set3", name = "Subjects") +
  theme_nature_metabolism() +
  theme(
    axis.text.x = element_text(),
    axis.title.x = element_blank(),
    legend.position = "right"
  ) +
  labs(
    y = "Contribution to Total (%)"
  )
```

## --- Plotting ---
``` {r fig.width = 1.181102, fig.height = 1.377953}
mm_to_in(c(30, 35))
d <- d + 
  # plot_annotation(tag_levels = list("C")) + 
  theme(plot.tag = element_text(size = 6, face = "plain")) + labs(title = "BAd Cluster") + theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2))
d

ggsave(file.path(OUTPUT_DIR, "S1D_brown_adipocyte_subject_composition.pdf"), plot = d, width = 30, height = 35, units = "mm", bg = "white",  dpi = 600)
```




# --- Panel S1E-F. Reference Mapping ASPCs - Subclustering ---
This section is intended to show the subclustering of adipocyte stem/progenitor cells (ASPCs) based on reference mapping with the Emont dataset. It is broken up in three parts: (1) the distribution of subtypes from the Emont dataset on the new embedding generated by SCANVI, (2) the cells from neck adipose imposed on this embedding, and (3) the proportions of each subtype in the neck adipose samples.

``` {r}
aspc <- readRDS(REF_MAP_APSC)
```

``` {r}
aspc@meta.data <- aspc@meta.data %>%
  mutate(predictions_scanvi = case_when(
    batch == "Query" ~ predictions_scanvi,
    batch == "Reference" ~ "Reference",
    TRUE ~ NA_character_
  ))

palette <- c(
  "Reference" = "grey",
  "hASPC1" = "#8DD3C7",
  "hASPC2" = "#FFFFB3",
  "hASPC3" = "#B3DE69",
  "hASPC4" = "#FB8072",
  "hASPC5" = "#80B1D3",
  "hASPC6" = "#FDB462"
)

Idents(aspc) <- "batch"

e <- DimPlot(subset(aspc, idents = "Reference"),
  group.by = "cell_type",
  order = c(
    "hASPC1", "hASPC2", "hASPC3",
    "hASPC4", "hASPC5", "hASPC6"
  ),
  cols = palette
) +
  theme_nature_metabolism() +
  theme_no_axis() +
  NoLegend()

f <- DimPlot(aspc,
  group.by = "predictions_scanvi",
  order = c(
    "hASPC1", "hASPC2", "hASPC3",
    "hASPC4", "hASPC5", "hASPC6",
    "Reference"
  ),
  cols = palette
) +
  theme_nature_metabolism() +
  theme_no_axis() +
  guides(
    color = guide_legend(
      ncol = 1,
      # byrow = TRUE,
      override.aes = list(size = 1)
    )
  )
```

```{r}
aspc_n <- aspc@meta.data |> count(predictions_scanvi, batch, cell_type)
aspc_n
```


## --- Plotting ---
``` {r fig.width = 1.574803, fig.height = 1.377953}
mm_to_in(c(40, 35))

# Collect legend for saving
legend <- get_legend(
  f + theme(legend.position = "right")
)
ggsave(file.path(OUTPUT_DIR, "S1E_F_legend.pdf"), plot = legend, width = 20, height = 35, units = "mm", bg = "white",  dpi = 600)



e <- e + 
  # plot_annotation(tag_levels = list("D")) +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "Emont et al. 2022 - ASPC", subtitle = "Recalculated UMAP") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2))

e
ggsave(file.path(OUTPUT_DIR, "S1E_aspc_reference.pdf"), plot = e, width = 40, height = 35, units = "mm", bg = "white",  dpi = 600)

f
f_nolegend <- f + theme(legend.position = "none")
f_nolegend <- f_nolegend +
  # plot_annotation(tag_levels = list("E")) +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "This Study - ASPC", subtitle = "Recalculated UMAP") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2))

f_nolegend

ggsave(file.path(OUTPUT_DIR, "S1F_aspc_mapped.pdf"), plot = f_nolegend, width = 40, height = 35, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel S1G. Abundance of mapped ASPC  ---
``` {r}
Idents(aspc) <- "batch"
query.aspc <- subset(aspc, idents = "Query")

propeller_list <- getTransformedProps(
  clusters = query.aspc$predictions_scanvi,
  sample = query.aspc$sample
)

props_meta <- as_tibble(propeller_list$Proportions) %>%
  dplyr::left_join(
    s.object@meta.data %>%
      select(sample, neck_region) %>%
      distinct(),
    by = "sample"
  ) %>%
  dplyr::filter(neck_region %in% c("Superficial", "Deep")) %>%
  mutate(
    n_pct = n * 100,
    neck_region = factor(neck_region, levels = c("Superficial", "Deep"))
  )
```

``` {r}
query.aspc@meta.data <- query.aspc@meta.data %>%
  mutate(
    neck_region = case_when(
      sample %in% c("Axiom001_SP", "Creek003_SP", "ENTRY005_SP") ~ "Intermediate",
      sample %in% c(
        "Axiom001_PV", "Creek003_PV", "Dsubject3_BAT", "ENTRY005_PV", "FREYA006_PV",
        "HAPPY008_CS", "ICING009_DC", "KARMA011_DC", "LEAFY012_DC", "MAPLE013_CS",
        "QUAIL017_PV", "RUGBY018_PV", "TRUTH020_BAT"
      ) ~ "Deep",
      sample %in% c(
        "Axiom001_SQ", "Creek003_SQ", "Dsubject3_WAT", "ENTRY005_SQ", "FREYA006_SQ",
        "HAPPY008_SC", "ICING009_SC", "JOYES010_PS", "KARMA011_SC", "LEAFY012_PS",
        "MAPLE013_SC", "Nevis014_PP", "QUAIL017_SC", "RUGBY018_PP"
      ) ~ "Superficial",
      TRUE ~ NA_character_
    )
  )

Idents(query.aspc) <- "neck_region"
query.aspc <- subset(query.aspc, idents = c("Deep", "Superficial"))

prop_results <- propeller(
  clusters = query.aspc$predictions_scanvi,
  sample = query.aspc$sample,
  group = query.aspc$neck_region
) %>%
  rename(
    predictions_scanvi = BaselineProp.clusters,
    p.val = P.Value
  ) %>%
  mutate(
    p_display = case_when(
      p.val < 0.01 ~ "<0.01",
      TRUE ~ paste(round(p.val, 2))
    )
  ) %>%
  arrange(rownames(.))

prop_results$group1 <- "Superficial"
prop_results$group2 <- "Deep"
prop_results$y.position <- 100
prop_results$colors <- ifelse(prop_results$p.val < 0.05, "red", "black")

prop_results <- prop_results %>% mutate(
  predictions_scanvi = factor(predictions_scanvi),
  xmin = as.numeric(predictions_scanvi) - 0.25,
  xmax = as.numeric(predictions_scanvi) + 0.25
)
```

``` {r}
p_value_y_position <- max(prop_results$y.position)

g <- ggplot(
  props_meta %>%
    dplyr::filter(neck_region != "Intermediate"),
  aes(
    x = clusters,
    y = n_pct,
    fill = neck_region
  )
) +

  # Add the boxplot layer
  geom_boxplot(
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    size = 0.3,
    alpha = 0.75
  ) +

  # Add the jitter layer with dodging
  geom_jitter(
    color = "black",
    shape = 21,
    alpha = 0.9,
    size = 1,
    position = position_jitterdodge(
      jitter.width = 0.1,
      dodge.width = 0.75
    )
  ) +
  scale_fill_manual(values = neck.color, name = "Neck Region") +
  scale_color_manual(values = neck.color, name = "Neck Region") +
  new_scale_color() +
  stat_pvalue_manual(
    data = prop_results,
    color = "colors",
    xmin = "xmin",
    xmax = "xmax",
    tip.length = 0.01,
    label = "p_display",
    y.position = "y.position",
    remove.bracket = FALSE,
    inherit.aes = FALSE,
    size = 5 / ggplot2::.pt
  ) +
  scale_color_identity(guide = "none") +
  coord_cartesian(ylim = c(NA, p_value_y_position + 2)) + #
  # Add your other layers and themes
  ggpubr::grids(axis = "y") +
  # scale_x_discrete(limits = order) +
  theme_nature_metabolism() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "right"
  ) +
  ylab("Proportion of \n Cells in Sample")
```
```{r}
  props_meta %>%
    dplyr::filter(neck_region != "Intermediate") |> 
  distinct(sample, .keep_all = TRUE) |> 
  count(neck_region)
```

## --- Plotting ---
``` {r fig.width = 3.149606, fig.height = 1.377953}
mm_to_in(c(80, 35))
g <- g + 
  # plot_annotation(tag_levels = list("F")) +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "ASPC Subtype Abundance per Cervical Region") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2)) +
  scale_y_continuous(expand = expansion(c(0.05, 0.10)))
g
ggsave(file.path(OUTPUT_DIR, "S1G_aspc_abundance.pdf"), plot = g, width = 80, height = 35, units = "mm", bg = "white",  dpi = 600)
```

# --- Panel S1H-I. Reference Mapping Adipocytes - Subclustering ---
This section is intended to show the subclustering of adipocytes based on reference mapping with the Emont dataset. It is broken up in three parts: (1) the distribution of subtypes from the Emont dataset on the new embedding generated by SCANVI, (2) the cells from neck adipose imposed on this embedding, and (3) the proportions of each subtype in the neck adipose samples.

``` {r}
adipo <- readRDS(REF_MAP_ADIPO)

adipo@meta.data <- adipo@meta.data %>%
  mutate(predictions_scanvi = case_when(
    batch == "Query" ~ predictions_scanvi,
    batch == "Reference" ~ "Reference",
    TRUE ~ NA_character_
  ))

palette <- c(
  "Reference" = "grey",
  "hAd1" = "#8DD3C7",
  "hAd2" = "#FFFFB3",
  "hAd3" = "#B3DE69",
  "hAd4" = "#FB8072",
  "hAd5" = "#80B1D3",
  "hAd6" = "#FDB462",
  "hAd7" = "#FCCDE5"
)

Idents(adipo) <- "batch"

h <- DimPlot(subset(adipo, idents = "Reference"),
  group.by = "cell_type",
  order = c(
    "hAd1", "hAd2", "hAd3",
    "hAd4", "hAd5", "hAd6", "hAd7"
  ),
  cols = palette
) +
  theme_nature_metabolism() +
  theme_no_axis() +
  NoLegend()

i <- DimPlot(adipo,
  group.by = "predictions_scanvi",
  order = c(
    "hAd1", "hAd2", "hAd3",
    "hAd4", "hAd5", "hAd6", "hAd7",
    "Reference"
  ),
  cols = palette
) +
  theme_nature_metabolism() +
  theme_no_axis() +
  guides(
    color = guide_legend(
      ncol = 1,
      # byrow = TRUE,
      override.aes = list(size = 1)
    )
  )
```

```{r}
adipo_n <- adipo@meta.data |> count(predictions_scanvi, batch, cell_type)
adipo_n
```

## --- Plotting ---
``` {r fig.width = 1.574803, fig.height = 1.377953}
mm_to_in(c(40, 35))

# Collect legend for saving
legend <- get_legend(
  i + theme(legend.position = "right")
)
ggsave(file.path(OUTPUT_DIR, "S1H_I_legend.pdf"), plot = legend, width = 20, height = 35, units = "mm", bg = "white",  dpi = 600)



h <- h + 
  # plot_annotation(tag_levels = list("G")) + 
  theme(plot.tag = element_text(size = 6, face = "plain")) + labs(title = "Emont et al. 2022 - Adipocytes", subtitle = "Recalculated UMAP") + theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2))
h
ggsave(file.path(OUTPUT_DIR, "S1H_adipocyte_reference.pdf"), plot = h, width = 40, height = 35, units = "mm", bg = "white",  dpi = 600)

i
i_nolegend <- i + theme(legend.position = "none")
i_nolegend <- i_nolegend + 
  # plot_annotation(tag_levels = list("H")) + 
  theme(plot.tag = element_text(size = 6, face = "plain")) + labs(title = "This Study - Adipocytes", subtitle = "Recalculated UMAP") + theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2))
i_nolegend

ggsave(file.path(OUTPUT_DIR, "S1I_adipocyte_mapped.pdf"), plot = i_nolegend, width = 40, height = 35, units = "mm", bg = "white",  dpi = 600)
```


# --- Panel S1J. Abundance of mapped Adipocytes  ---
``` {r}
Idents(adipo) <- "batch"
query.adipo <- subset(adipo, idents = "Query")

propeller_list <- getTransformedProps(
  clusters = query.adipo$predictions_scanvi,
  sample = query.adipo$sample
)

props_meta <- as_tibble(propeller_list$Proportions) %>%
  dplyr::left_join(
    s.object@meta.data %>%
      select(sample, neck_region) %>%
      distinct(),
    by = "sample"
  ) %>%
  dplyr::filter(neck_region %in% c("Superficial", "Deep")) %>%
  mutate(
    n_pct = n * 100,
    neck_region = factor(neck_region, levels = c("Superficial", "Deep"))
  )
```

``` {r}
query.adipo@meta.data <- query.adipo@meta.data %>%
  mutate(
    neck_region = case_when(
      sample %in% c("Axiom001_SP", "Creek003_SP", "ENTRY005_SP") ~ "Intermediate",
      sample %in% c(
        "Axiom001_PV", "Creek003_PV", "Dsubject3_BAT", "ENTRY005_PV", "FREYA006_PV",
        "HAPPY008_CS", "ICING009_DC", "KARMA011_DC", "LEAFY012_DC", "MAPLE013_CS",
        "QUAIL017_PV", "RUGBY018_PV", "TRUTH020_BAT"
      ) ~ "Deep",
      sample %in% c(
        "Axiom001_SQ", "Creek003_SQ", "Dsubject3_WAT", "ENTRY005_SQ", "FREYA006_SQ",
        "HAPPY008_SC", "ICING009_SC", "JOYES010_PS", "KARMA011_SC", "LEAFY012_PS",
        "MAPLE013_SC", "Nevis014_PP", "QUAIL017_SC", "RUGBY018_PP"
      ) ~ "Superficial",
      TRUE ~ NA_character_
    )
  )

Idents(query.adipo) <- "neck_region"
query.adipo <- subset(query.adipo, idents = c("Deep", "Superficial"))

prop_results <- propeller(
  clusters = query.adipo$predictions_scanvi,
  sample = query.adipo$sample,
  group = query.adipo$neck_region
) %>%
  rename(
    predictions_scanvi = BaselineProp.clusters,
    p.val = P.Value
  ) %>%
  mutate(
    p_display = case_when(
      p.val < 0.01 ~ "<0.01",
      TRUE ~ paste(round(p.val, 2))
    )
  ) %>%
  arrange(rownames(.))

prop_results$group1 <- "Superficial"
prop_results$group2 <- "Deep"
prop_results$y.position <- 80
prop_results$colors <- ifelse(prop_results$p.val < 0.05, "red", "black")

prop_results <- prop_results %>% mutate(
  predictions_scanvi = factor(predictions_scanvi),
  xmin = as.numeric(predictions_scanvi) - 0.25,
  xmax = as.numeric(predictions_scanvi) + 0.25
)
```

``` {r}
p_value_y_position <- max(prop_results$y.position)

j <- ggplot(
  props_meta %>%
    dplyr::filter(neck_region != "Intermediate"),
  aes(
    x = clusters,
    y = n_pct,
    fill = neck_region
  )
) +

  # Add the boxplot layer
  geom_boxplot(
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    size = 0.3,
    alpha = 0.75
  ) +

  # Add the jitter layer with dodging
  geom_jitter(
    color = "black",
    shape = 21,
    alpha = 0.9,
    size = 1,
    position = position_jitterdodge(
      jitter.width = 0.1,
      dodge.width = 0.75
    )
  ) +
  scale_fill_manual(values = neck.color, name = "Neck Region") +
  scale_color_manual(values = neck.color, name = "Neck Region") +
  new_scale_color() +
  stat_pvalue_manual(
    data = prop_results,
    color = "colors",
    xmin = "xmin",
    xmax = "xmax",
    tip.length = 0.01,
    label = "p_display",
    y.position = "y.position",
    remove.bracket = FALSE,
    inherit.aes = FALSE,
    size = 5 / ggplot2::.pt
  ) +
  scale_color_identity(guide = "none") +
  coord_cartesian(ylim = c(NA, p_value_y_position + 2)) + #
  # Add your other layers and themes
  ggpubr::grids(axis = "y") +
  # scale_x_discrete(limits = order) +
  theme_nature_metabolism() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "right"
  ) +
  ylab("Proportion of \n Cells in Sample")
```

## --- Plotting ---
``` {r fig.width = 3.149606, fig.height = 1.377953}
mm_to_in(c(80, 35))
j <- j + 
  # plot_annotation(tag_levels = list("I")) +
  theme(plot.tag = element_text(size = 6, face = "plain")) +
  labs(title = "Adipocyte Subtype Abundance per Cervical Region") +
  theme(plot.title = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 2)) +
  scale_y_continuous(expand = expansion(c(0.05, 0.10)))
j
ggsave(file.path(OUTPUT_DIR, "S1J_adipo_abundance.pdf"), plot = j, width = 80, height = 35, units = "mm", bg = "white",  dpi = 600)
```




