---
title: "Sondergaard_Analysis"
output: html_document
date: "2024-12-16"
---

# Introduction

This notebook processes the raw Affymetrix CEL files from the
Sondergaard paraganglioma dataset, extracts HEAT-signature genes,
saves an expression matrix for downstream use, and computes ssGSEA
HEAT scores together with UCP1 expression for inspection.

The main saved output retained from the original workflow is:

- `GSE49795_Ohja_expression_matrix.rds`

```{r setup, include=FALSE}
# -------------------------------------------------------------------------
# Load required packages
# -------------------------------------------------------------------------
library(affy)   # Older Affymetrix arrays
library(oligo)  # Newer/high-density Affymetrix arrays
library(hugene10sttranscriptcluster.db)
library(AnnotationDbi)
library(GSVA)
library(GSEABase)
library(pheatmap)
library(tidyverse)
library(here)

# -------------------------------------------------------------------------
# Resolve package conflicts
# -------------------------------------------------------------------------
conflicts_prefer(affy::list.celfiles)
conflicts_prefer(oligo::rma)

# -------------------------------------------------------------------------
# Define input and output paths
# -------------------------------------------------------------------------
# Inputs
CEL_FILES <- here("data/5.published_datasets/Public_BulkRNAseq/Sondergaard 2014 - Paraganglioma patient")
HEAT_CSV <- here("output/3.heat_signature/data/heat_signature.csv")

# Outputs
OUTPUT_DIR <- here("output/3.heat_signature/3.public_bulkrnaseq/data/")
dir.create(OUTPUT_DIR, showWarnings = FALSE, recursive = TRUE)

OUTPUT_SSGSEA_DIR <- here("output/3.heat_signature/3.public_bulkrnaseq/data/ssGSEA")
dir.create(OUTPUT_SSGSEA_DIR, showWarnings = FALSE, recursive = TRUE)

OUTPUT_EXP_MTX <- file.path(OUTPUT_DIR, "GSE49795_Sondergaard_expression_matrix.rds")
```

# Load CEL files

The raw data are distributed as Affymetrix CEL files and are read with
`oligo`.

```{r}
oligo_data <- read.celfiles(list.celfiles(CEL_FILES, full.names = TRUE))

annotation(oligo_data)
```

# Background correction and normalization

The `rma()` workflow performs background correction, quantile
normalization, and summarization.

```{r}
eset <- rma(oligo_data)
```

# Extract expression matrix

```{r}
expr_matrix <- exprs(eset)
```

# Map probe IDs to gene symbols

The Affymetrix probe IDs are mapped to gene symbols using the
`hugene10sttranscriptcluster.db` annotation package.

```{r}
transcript_ids <- rownames(expr_matrix)

gene_symbols <- mapIds(
  hugene10sttranscriptcluster.db,
  keys = transcript_ids,
  column = "SYMBOL",
  keytype = "PROBEID"
)

expr_matrix <- cbind(expr_matrix, GeneSymbol = gene_symbols)
```

# Remove probes without gene symbols

```{r}
sum(is.na(gene_symbols))

expr_matrix_filtered <- expr_matrix[!is.na(gene_symbols), ]
```

# Extract HEAT genes

This section restricts the dataset to genes present in the HEAT
signature and reshapes the data into a small expression matrix.

```{r}
HEAT <- read.csv(HEAT_CSV, header = TRUE)

HEAT_data <- as.data.frame(expr_matrix_filtered) %>%
  dplyr::filter(GeneSymbol %in% HEAT$x) %>%
  rownames_to_column(var = "probeid") %>%
  gather(., 2:4, key = "sample", value = "value") %>%
  mutate(value = as.numeric(value)) %>%
  spread(sample, value) %>%
  dplyr::select(-probeid)

expression_matrix <- as.matrix(HEAT_data[, 2:4])
rownames(expression_matrix) <- HEAT_data$GeneSymbol
```

# Clean duplicated genes and zero-variance rows

Multiple probes can map to the same gene. The original workflow made
row names unique and removed genes with zero variance before saving.

```{r}
rownames(expression_matrix) <- make.unique(rownames(expression_matrix))

duplicated_rows <- duplicated(rownames(expression_matrix))
rownames(expression_matrix)[duplicated_rows]

zero_sd_rows <- apply(expression_matrix, 1, sd) == 0
expression_matrix <- expression_matrix[!zero_sd_rows, ]

expression_matrix <- expression_matrix[order(rownames(expression_matrix)), ]

colnames(expression_matrix) <- c(
  "Patient Subcutaneous",
  "Patient Visceral",
  "Control Visceral"
)
```

# Save expression matrix

```{r}
write_rds(expression_matrix, OUTPUT_EXP_MTX)
```

# Plot heatmap

This heatmap was kept because it is part of the original script flow,
even though it is not saved to disk.

```{r}
pheatmap::pheatmap(
  expression_matrix,
  scale = "row",
  clustering_distance_rows = "euclidean",
  cellwidth = 5,
  cellheight = 5,
  fontsize = 6,
  clustering_distance_cols = "euclidean",
  clustering_method = "complete",
  main = "Gene Expression Heatmap"
)
```

# Run ssGSEA using the HEAT signature

This section computes ssGSEA HEAT scores from the filtered expression
matrix.

```{r}
pheo <- as.data.frame(expr_matrix_filtered)
rownames(pheo) <- NULL

pheo <- pheo[!duplicated(pheo$GeneSymbol), ]
rownames(pheo) <- NULL

pheo <- pheo %>%
  rename(
    sample1 = "GSM1207166_M242-01.CEL",
    sample2 = "GSM1207167_M242-02.CEL",
    sample3 = "GSM1207168_M242-03.CEL"
  ) %>%
  column_to_rownames(var = "GeneSymbol") %>%
  mutate(
    sample1 = as.numeric(sample1),
    sample2 = as.numeric(sample2),
    sample3 = as.numeric(sample3)
  ) %>%
  as.matrix()

heat_collection <- GeneSetCollection(
  GeneSet(
    unique(HEAT_data$GeneSymbol),
    setName = "HEAT"
  )
)

expr_set <- ExpressionSet(assayData = pheo)

pheo.gsea <- gsva(ssgseaParam(expr_set, heat_collection))

t(pheo.gsea@assayData$exprs)

# Sample 1 is Patient Subcutaneous
# Sample 2 is Patient Visceral
# Sample 3 is Control Visceral
```

# Add UCP1 expression to ssGSEA output

```{r}
goi_idx <- grep("UCP1|CKMT2", expr_matrix_filtered[, 4])
colnames(expr_matrix_filtered)

goi_exp <- expr_matrix_filtered[goi_idx, 1:4] |>
  as_tibble() |>
  column_to_rownames("GeneSymbol") |>
  t() |>
  as.data.frame() |>
  mutate(across(everything(), as.numeric))

pheo.gsea.ucp <- t(pheo.gsea@assayData$exprs)
pheo.gsea.ucp <- cbind(pheo.gsea.ucp, goi_exp) |> rownames_to_column("sample")
pheo.gsea.ucp$group <- c("Patient Subcutaneous", "Patient Visceral", "Control Visceral")


# Save file
write_csv(pheo.gsea.ucp, file.path(OUTPUT_SSGSEA_DIR, "Sondergaard_HEATvUCP1.csv"))
```
