---
title: "Description of human neck composition"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: sentence
---

# Introduction {.tabset}

The purpose of this file is to explore the cellular composition of human
neck adipose tissue biopsies using Propeller.

## Setting up the environment

```{r setup}
library(speckle)
library(Seurat)
library(limma)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(ComplexUpset)
library(here)

conflicts_prefer(ggplot2::vars)

INPUT_RDS <- here("data", "adata_integrated_soupXoutput_with_infer_forR_temp_Dec18_cleaned.rds")

OUTPUT_DATA_DIR <- here("output", "2.sobject_characterization", "propeller", "data")
OUTPUT_FIG_DIR <- here("output", "2.sobject_characterization", "propeller", "figures")

dir.create(OUTPUT_DATA_DIR, showWarnings = FALSE, recursive = TRUE)
dir.create(OUTPUT_FIG_DIR, showWarnings = FALSE, recursive = TRUE)
```

# Loading the Seurat object

```{r, message=FALSE}
s.object <- readRDS(INPUT_RDS)

# Replace hyphens in metadata column names to avoid downstream issues
names(s.object@meta.data) <- gsub("-", "_", names(s.object@meta.data))
```

# Workflow

This section prepares the data for plotting and significance testing.
For the main comparison, only Deep and Superficial neck samples are used.

```{r}
Idents(s.object) <- "neck_region"
small.s.object <- subset(s.object, idents = c("Deep", "Superficial"))
propeller_input <- small.s.object@meta.data
```

## Cell type proportions by sample

```{r}
propeller_list <- getTransformedProps(
  clusters = s.object$cell_type_short,
  sample = s.object$sample
)

props_meta <- as_tibble(propeller_list$Proportions) %>%
  left_join(
    s.object@meta.data %>%
      select(sample, neck_region) %>%
      distinct(),
    by = "sample"
  ) %>%
  mutate(
    n_pct = n * 100,
    neck_region = factor(neck_region, levels = c("Superficial", "Intermediate", "Deep"))
  ) %>%
  left_join(
    s.object@meta.data %>%
      select(cell_type, cell_type_short) %>%
      distinct(cell_type_short, .keep_all = TRUE),
    by = c("clusters" = "cell_type_short")
  )

write_csv(props_meta, file.path(OUTPUT_DATA_DIR, "cellular_proportions_table.csv"))
```

## Average proportions

```{r}
avg_prop <- props_meta %>%
  group_by(clusters, cell_type) %>%
  summarise(
    n = sum(!is.na(n_pct)),
    mean_pct = mean(n_pct, na.rm = TRUE),
    sem_pct = sd(n_pct, na.rm = TRUE) / sqrt(n),
    mean_sem = paste0(round(mean_pct, 1), " ± ", round(sem_pct, 1), "%"),
    .groups = "drop"
  ) %>%
  mutate(text_description = paste0(cell_type, " (", clusters, ", ", mean_sem, ")"))

write_csv(avg_prop, file.path(OUTPUT_DATA_DIR, "cellular_average_proportions_table.csv"))
```

## Helper functions

```{r}
filter_propeller_samples <- function(df, min_cell_threshold = 0) {
  sample_count <- df %>% count(sample)
  samples_keep <- sample_count %>%
    filter(n > min_cell_threshold) %>%
    pull(sample)

  df %>% filter(sample %in% samples_keep)
}

run_propeller_test <- function(df, group_col) {
  res <- propeller(
    clusters = df$cell_type_short,
    sample = df$sample,
    group = df[[group_col]]
  )

  if ("BaselineProp.clusters" %in% colnames(res)) {
    res <- res %>%
      rename(cell_type_short = BaselineProp.clusters)
  } else {
    res <- res %>%
      rownames_to_column("cell_type_short")
  }

  res %>%
    rename(p.val = P.Value) %>%
    mutate(
      p_display = sprintf("%.3f", p.val),
      fdr_display = sprintf("%.3f", FDR)
    )
}


```

## Main Propeller test. Deep vs Superficial

```{r}
min_cell_threshold <- 0
propeller_input_filtered <- filter_propeller_samples(
  propeller_input,
  min_cell_threshold = min_cell_threshold
)

prop_results <- run_propeller_test(
  propeller_input_filtered,
  group_col = "neck_region"
) %>%
  mutate(
    min_cell_filter = min_cell_threshold,
    group1 = "Superficial",
    group2 = "Deep",
    y.position = 75,
    colors = ifelse(p.val < 0.05, "red", "black")
  )
```

```{r}
# Bracket coordinates for later annotation
bracket_length <- 0.2

cluster_order <- props_meta %>%
  group_by(clusters) %>%
  summarise(avg_abundance = mean(n_pct), .groups = "drop") %>%
  arrange(desc(avg_abundance)) %>%
  pull(clusters)

prop_results <- prop_results %>%
  mutate(
    cell_type_short = factor(cell_type_short, levels = cluster_order),
    xmin = as.numeric(cell_type_short) - bracket_length,
    xmax = as.numeric(cell_type_short) + bracket_length
  )

max_prop <- props_meta %>%
  group_by(clusters) %>%
  summarise(y.pos = max(n_pct), .groups = "drop") %>%
  mutate(y.pos = 1.1 * y.pos)

prop_results <- prop_results %>%
  left_join(max_prop, by = c("cell_type_short" = "clusters"))

write_csv(prop_results, file.path(OUTPUT_DATA_DIR, "proportions_stat_table.csv"))
```

```{r}
# -----------------------------
#  Export sample level counts 
# -----------------------------

sample_summary <- small.s.object@meta.data %>%
  count(sample, neck_region, name = "n_nuclei") %>%
  arrange(neck_region, sample)

write_csv(
  sample_summary,
  file.path(OUTPUT_DATA_DIR, "sample_nuclei_counts_deep_superficial.csv")
)

```

```{r}
# -----------------------------
#  Export celltype-counts by sample 
# -----------------------------

cell_type_sample_counts <- small.s.object@meta.data %>%
  count(sample, neck_region, cell_type_short, cell_type, name = "n_nuclei") %>%
  arrange(neck_region, sample, cell_type_short)

write_csv(
  cell_type_sample_counts,
  file.path(OUTPUT_DATA_DIR, "cell_type_counts_by_sample_deep_superficial.csv")
)
```

```{r}
writeLines(
  capture.output(sessionInfo()),
  file.path(OUTPUT_DATA_DIR, "sessionInfo_propeller_composition.txt")
)
```

